### 2017-01-16 2017-GECCO-Modeling-User-Interactions-in-IEC/sample-sigconf.tex
GECCO New ACM Format

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GECCO '17]{the Genetic and Evolutionary Computation Conference 2017}{July 15--19, 2017}{Berlin, Germany}
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
\title{SIG Proceedings Paper in LaTeX Format}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

\author{Ben Trovato}
\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{trovato@corporation.com}

\author{G.K.M. Tobin}
\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{webmaster@marysville-ohio.com}

\author{Lars Th{\o}rv{\"a}ld}
\authornote{This author is the
  one who did all the really hard work.}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla} 
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Lawrence P. Leipuner}
\affiliation{
  \institution{Brookhaven Laboratories}
  \streetaddress{P.O. Box 5000}}
\email{lleipuner@researchlabs.org}

\author{Sean Fogarty}
\affiliation{%
  \institution{NASA Ames Research Center}
  \city{Moffett Field}
  \state{California} 
  \postcode{94035}}
\email{fogartys@amesres.org}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas} 
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
\email{jsmith@affiliation.org}

\author{Julius P.~Kumquat}
\affiliation{\institution{The Kumquat Consortium}}
\email{jpkumquat@consortium.net}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{B. Trovato et. al.}

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings\footnote{This is an abstract footnote}. 
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

% We no longer use \terms command
%\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\input{samplebody-conf}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc}

\end{document}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
abstract
@jj how about this for starters

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{A graph-based approach for interactive 
evolutionary computation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents an alternative data model for the development of
Interactive Evolutionay Computation (IEC) algorithms. A graph database is
proposed to model the overall system which consists of: a social network of
volunteers, a population of individuals in the evolutionary algorithm,
the system\'s functionality and their interactions. This graph model can then be used
to tackle some of the issues associated with IEC; such as, 
user fatigue, lack of volunteer participation, fitness estimation and novelty. 
The proposed graph database could be used to model IEC systems that 
follow a user-centered approach. A case study is presented providing both 
conceptual and implementation details using the EvoSpace-Interactive platform.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

\subsection{Subsection Heading Here}
Subsection text here.
\cite{Sims:1991}

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
The conclusion goes here.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\bibliographystyle{IEEEtran}
\bibliography{evospace-i}
%\bibliography{languages}
\end{document}

% that's all folks
\end{document}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Starting Intro

@@
 \section{Introduction}
-% no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
+Interactive Evolutionary Algorithms (IEAs): are standard EAs whose
+fitness evaluations are performed by persons with in an interactive 
+system.  Thus, the main loop of the EA need a human intervention for
+the quality assessment of each proposed solution. These systems have an
+inherent drawback araising from the very nature of the model, namely,
+the human fatigue caused by the interaction. Some authors have already
+tried to address this basic problem \cite{Frade:2010:EvoGAMES}.
+nevertheless, IEAs have demonstrated their ability for effectively
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},.
+as well as other domains \cite{ie1}.
 
-\hfill mds
-
-\hfill August 26, 2015
-
+,
 \subsection{Subsection Heading Here}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
A little more Intro

@@
 fitness evaluations are performed by persons with in an interactive 
-system.  Thus, the main loop of the EA need a human intervention for
-the quality assessment of each proposed solution. These systems have an
-inherent drawback araising from the very nature of the model, namely,
-the human fatigue caused by the interaction. Some authors have already
-tried to address this basic problem \cite{Frade:2010:EvoGAMES}.
+system.  Thus, the main loop of the EA needs the intervention of a
+human for the quality assessment of each proposed solution. These
+systems have an inherent drawback araising from the very nature of 
+the algorithm, namely, the human fatigue caused by the interaction. 
+Some authors have already tried to address this basic problem \cite{Frade:2010:EvoGAMES}.
 nevertheless, IEAs have demonstrated their ability for effectively
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},.
-as well as other domains \cite{ie1}.
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
+as well as other products in many other domains \cite{ie1}. 
+Several IEC have been developed as web based systems \cite{picbreeder,Musart},
+depending on volunteers who visit the system to interactibly help
+with the seach.
 
-,
 \subsection{Subsection Heading Here}
-Subsection text here.
+Subsection text here.,
 \cite{Sims:1991}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Intro

@@
 as well as other products in many other domains \cite{ie1}. 
-Several IEC have been developed as web based systems \cite{picbreeder,Musart},
-depending on volunteers who visit the system to interactibly help
-with the seach.
+Several IEC have been developed as web based systems \cite{picbreeder},
+who depend on volunteers who visit the system to interactively help
+with the search, using both anonymous and registered users. In particular
+the EvoSpace-Interactive (ES-I) platform employs a collaborative technique,
+where several registered users assign a quality assessment to a single
+individual and then an aggregated fitness value is calculated. The fitness
+assigned to each individual depends on the taste of each particular user, 
+resulting in a many-to-many relationship between users and individuals. 
+Many systems query this user-individual relation to extract relevant
+knowledge about the process and the population, for instance showing the
+most popular individual, or the the user with more participation \cite{picbreeder}.
+In order to do this, individuals about individuals must be stored, even
+if they are no longer in the active population. In this work a graph database
+is proposed to model the interaction between users and individuals, with
+the following considerations:
+
+\begin{enumerate}
+  \item  The relationship between individuals is naturally expressed as a
+  tree, or a graph when both ancestors and descendancy is considered.
+
+  \item Relationship between users can also be modeled as a social network,
+  with well stablished algorithms and metrics (cite).
+
+  \item Fitness evaluation can also be extracted from different kind of relationships
+  between a user and an individual, using common actions? found in social networks,
+  for instance share, like, or add to collection.
+
+\end{enumerate}
 

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Intro enumeration

@@
 is proposed to model the interaction between users and individuals, with
-the following considerations:
+the following advantages:
 
@@
 
-  \item Fitness evaluation can also be extracted from different kind of relationships
-  between a user and an individual, using common actions? found in social networks,
+  \item Many kinds of relationships between users and individuals,
+  can be expressed using common actions found in social networks;
   for instance share, like, or add to collection.
 
+  \item The same graph can be used to have a user model.
 \end{enumerate}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Intro cont
Planned Sections

@@
   can be expressed using common actions found in social networks;
-  for instance share, like, or add to collection.
+  for instance share, like, or add to collection. [ cite Activity Streams 2.0
+  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]
 
-  \item The same graph can be used to have a user model.
+  \item The same graph can be used to immplement a user model, or it can
+  be extracted from a user\'s public profile.
+
 \end{enumerate}
 
-\subsection{Subsection Heading Here}
-Subsection text here.,
-\cite{Sims:1991}
+In this work..
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+The remainder of this paper proceeds as follows.
+Section \ref{sec:interactive} presents related work on the topic 
+of Interactive Evolution data models.
+Then, Section \ref{sec:evospace-i} presents the computational platform on which the main proposal of this work is developed. The proposed graph model is presented in Section \ref{sec:graph}.
+The experimental work and results are presented in Section \ref{sec:experiments}.
+Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
+
+\subsection{Related Work}
+\label{sec:interactive}
+
+\subsection{EvoSpace-Interactive}
+\label{sec:evospace-i}
+
+\subsection{IEC Graph based model}
+\label{sec:graph}
+
+\subsection{Experiments}
+\label{sec:experiments}
+
+\subsection{Conclusions}
+\label{sec:conclusions}
 

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Graph Model
Preliminar writings from Christian work added

@@
 % Do not put math or special symbols in the title.
-\title{A graph-based approach for interactive 
+\title{A graph-based data model for interactive 
 evolutionary computation}
@@
 
-\subsection{Related Work}
+\section{Related Work}
 \label{sec:interactive}
 
-\subsection{EvoSpace-Interactive}
+\section{EvoSpace-Interactive}
 \label{sec:evospace-i}
 
-\subsection{IEC Graph based model}
+\section{IEC Graph based model}
 \label{sec:graph}
+
+% Begins Christian´s writings
+% Needs TLC
+
+\subsection{Users}
+
+Users are a central part of this proposed framework as it aims to study their
+behavior when interacting with individuals within interactive evolutionary
+algorithms and other tasks that may exist within IEC systems. Thus users in
+this proposal are entities interacting with an evolutionary computation that
+has the fundamental purpose of evaluating individuals of a population
+replacing the fitness function according to their preferences and mood, among
+others []. This form of evaluate individual is given a subjectively way.
+
+On the other hand in order to capture the attention of users and possibly
+increase their participation, currently there exist some basic actions that
+users can perform in a given system, going from to be able to access a system
+through login mechanism, once the user has logged-in in to the system the
+users can interact with different actions, for example the inviting action is
+when users can be invited through social networks or maybe of person to
+person. Another example is the sharing action, that can occur when users want
+to share something of their interest through their social networks. In this sense the action of posting is when
+users want to put something on their wall who want others to know through
+their social networks. Likewise the storing action occurs when users want to
+retain permanently information that are of their interest. Finally the
+browsing action occurs when users are exploring content for their needs. All
+these actions go beyond only evaluating individuals.
+
+\begin{figure*}
+  \captionsetup{justification=centering,margin=2cm}
+  \centering
+  \setlength\fboxsep{0pt}
+  \setlength\fboxrule{0.7pt}
+  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/users.png}}
+  \caption{Users actions on interactive evolutionary systems.}
+  \label{fig:users}       
+\end{figure*}
+
+In order to start the task of evaluating individuals is necessary that users
+access the system through a login mechanism. This mechanism consists of
+providing a username and a password in order to grant access to the system. All
+users accessing in this way they are considered active users within an IEC
+system.
+
+For the evaluating action is proposed  that users evaluate individuals
+indirectly, this means users can evaluate accepting indirect recommendations of
+friends that are currently active in the system. These recommendations can be
+store individuals in the system of users who know each other.
+
+Also the  browsing action is proposed within IEC systems,  this means that users
+can explore information that other active users in the system are generating.
+
+Finally activity participation is proposed. This can be divided into four
+different actions as fallows:
+
+\begin{itemize} 
+\item The Inviting action occurs when a user invites another to the system through social networks or from person to person.  
+\item The sharing action occurs when users share their individual creations with others active user in the system.
+\item The posting action occurs when published what they are doing within the system.
+\item The storing action occurs when users keep individuals in a collection, the collections concept will be discussed later in this chapter.
+\end{itemize}
+
+All the mention above is shown in figure \ref{fig:users}
+
+\subsection{Individual}
+
+Individuals are entities that form the population for the evolutionary
+algorithm. In particular for  this work individuals are animated digital
+paintings, every individual in the population is defined by a chromosome [].
+This chromosome defines the behavior of the individual, so that it consists of a
+vector of real numbers of fifteen genes, where each gen define a particular
+behavior in painting. This individuals combined each other to generated new
+individual in the population following the genetic operators such as selection,
+crossover and mutation[]. An example of a individual in this case it is
+illustrated in figure \ref{fig:individual}
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/individual.png}}
+\caption{Individual representation.}
+\label{fig:graph}       
+\end{figure*}
+
+\subsection{User Model} Now that both users an individuals have been defined we can
+model the users behavior have with respect to individuals in a graph-based user
+model.
+
+The reason for using a graph to model the user-individual is because it could
+can express in a simple way the behavior of this two entities having the
+necessary information  to represent the knowledge,  for example a vertex user
+and individual vertex connected through a edge that  represents the
+relationship. this relationship can be the semantic  between these two entities
+as figure \ref{fig:User-Individual} show. Each of these vertices contain properties as well as the
+edges, these properties will be explained as fallows.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_individual.png}}
+\caption{User-Individual.}
+\label{fig:User-Individual}       
+\end{figure*}
+
+In this proposed graph-based user-individual model the vertices are represented
+by set of nodes of USERS, INDIVIDUALS and COLLECTIONS. The edges are define by a
+set of relationships as LIKES, KNOWS, PARENT, HAS that represents the
+relationships between the vertices as seen in figure \ref{fig:User-Individual}.
+
+Already explained that represent USERS and INDIVIDUALS, now an explanation is
+given of what concerns for the concept of COLLECTIONS.
+
+In this proposal a collection is defined as a  container where users can store
+selected individuals laid within the collection. A collection is created when
+the user wants to store permanently individuals within the IEC system.
+
+Now for the edges in this proposal are explained as follows:
+
+\begin{itemize} 
+\item The  LIKES edge represents a preference relationship that a user has with respect to an individual.  
+\item Now the  PARENT edge represents an ancestor of the individual, in other words where the individual comes from.
+\item Likewise the  HAS edge represents that users have a collection where can contain individuals.
+\item Finally the KNOWS edge represents a relationship of friendship between users.
+\end{itemize}
+
+In figure \ref{fig:Nodes_Edges} illustrated the above mentioned, about nodes and edges.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/user_individual_collections.png}}
+\caption{Nodes And Edges Representation.}
+\label{fig:Nodes_Edges}       
+\end{figure*}
+
+In each vertex is necessary to store knowledge about the users, individuals and
+also for the collections, this is what is known as property.
+
+It is worth to mention that all vertices have a property called “element\_type”
+among others. This property is used to identify which type of vertex is, for
+example if vertex corresponds to a user then this property label the vertex as a
+“person”. Likewise if the vertex corresponds to an individual this property
+label it as an “individual” and is the same for any vertex we want to add to the
+model. This property also has the purpose of being able to do operations on the
+graph according to the vertex type.
+
+The properties for the vertex “u” are the following:
+
+\begin{itemize} 
+\item id.  
+\item created.
+\item name.
+\item element\_type.
+\end{itemize}
+
+The “id” property defines that the vertex u has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The property
+“created” defines the creation date of the vertex $u$. The property “name”
+defines the name of the vertex u. The property “element\_type” defines the
+element type that will be the vertex $u$.
+
+In figure \ref{fig:User_node} the vertex $u$ properties are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_node.png}}
+\caption{User Properties.}
+\label{fig:User_node}       
+\end{figure*}
+
+The properties for the vertex $i$ are the following:
+\begin{itemize} 
+\item id. 
+\item created. 
+\item element\_type.
+\item chromosome.
+\item views.
+\end{itemize}
+
+The “id” property defines that the vertex $i$ has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The
+“chromosome” property defines the chromosome of the individual representing the
+vertex $i$. which as defined above in this section. The  property “views” defines
+the amount that users have seen this vertex $i$ The property “element\_type”
+defines the element type that will be the vertex $i$. The property “created”
+defines the creation date of the vertex $i$.
+
+In figure \ref{fig:Individual_node} the vertex $i$ properties are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/individual_node.png}}
+\caption{Individual Properties.}
+\label{fig:Individual_node}       
+\end{figure*}
+
+The properties for the vertex “c” are the following:
+
+\begin{itemize} 
+\item id. 
+\item element\_type.
+\item created. 
+\item name.
+\end{itemize}
+
+The “id” property defines that the vertex $c$ has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The property
+“element\_type” defines the element type that will be the vertex $c$. The
+property “name” defines the name of the vertex $c$. The property “created”
+defines the creation date of the vertex $c$.
+
+In figure \ref{fig:Collection_node} the vertex $c$ properties are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/collection_node.png}}
+\caption{Collection Properties.}
+\label{fig:Collection_node}       
+\end{figure*}
+
+In the same way that the vertices have properties also the edges, these
+properties are defined as follows.
+
+The “LIKES” edge has the following properties:
+
+\begin{itemize} 
+\item id. 
+\item created. 
+\item rate.
+\end{itemize}
+
+The “id” property defines that the “LIKES” edge  has a unique identifier in the
+set of edges, which means that there will not be duplicate edges. The property
+“created” defines the date of creation of the this edge. The property “rate”
+defines the rate that the users give to the individual store in this edge.
+
+In figure \ref{fig:Likes_edge} the edge “LIKES” properties and the relationships with the vertices $u$ and $i$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_likes.png}}
+\caption{LIKES edge Properties.}
+\label{fig:Likes_edge}       
+\end{figure*}
+
+The “KNOWS” edge has the following properties.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “KNOWS” has a unique identifier in the
+set of edges, which means that there will not be duplicate edges. The property
+“created” defines the date of creation of the edge.
+
+In figure \ref{fig:Knows_edge} the edge “KNOWS” properties and the relationships between vertex “u”
+are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_knows.png}}
+\caption{KNOWS edge Properties.}
+\label{fig:Knows_edge}       
+\end{figure*}
+
+The “PARENT” edge has the following properties, and represent the parents of a new individual.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “PARENT” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
+The property “created” defines the date of creation of the edge.
+
+In figure \ref{fig:Parent_edge} the edge “PARENT” properties and the relationship between vertex $i$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_parent.png}}
+\caption{PARENT edge Properties.}
+\label{fig:Parent_edge}       
+\end{figure*}
+
+The edge “HAS” has the following properties, and represent the parents of a new individual.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “HAS” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
+The property “created” defines the date of creation of the edge.
+
+In figure \ref{fig:Has_edge} the edge “HAS” properties and the relationship between vertices $u$, $i$ and $c$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
+\caption{HAS edge Properties.}
+\label{fig:Has_edge}       
+\end{figure*}
+
+\begin{equation*}\label{eq:graphRelDef} 
+\displaystyle 
+\begin{split} 
+V &= \{[u_1,u_2,u_3,...,u_n],[i_1,i_2,i_3,...,i_n],[c_1,c_2,c_3,...,c_n]\},\\ 
+E&= \{[l_1,l_2,l_3,..,l_n],[p_1,p_2,p_3,...,p_n],[h_1,h_2,h_3,...,h_n],[k_1,k_2,k_3,...,k_n]\}\\ 
+\end{split} 
+\end{equation*}
+
+Already defined the vertices and edges a graph-based user-individual model is
+proposed where vertices users are related to the vertices individuals, likewise
+the vertices individuals are related to themselves when evolution exist between
+them. In that sense the vertices users and individuals are related to the
+vertices collections as a vertex user can be related to a vertex collections and
+a vertex collections collections is related to $n$ individuals vertex. In Figure
+\ref{fig:user_moder} shows an example of the graph-based user-individual model.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=8cm,height=8cm,keepaspectratio]{img/model_representation.png}}
+\caption{Graph-based user-individual model.}
+\label{fig:user_moder}       
+\end{figure*}
 

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Starting model

@@
 
-% Begins Christian´s writings
-% Needs TLC
+Users are a central part of an IEC system, because not only their
+explicit evaluation is involved, the overall context of interaction
+affects the system as a whole as users interact not only with the graphical
+interface, and each individual\'s phenotype, there is also the interaction
+with other users, their previous evaluations and their own history of activities
+and past experiences. Thus users in this proposal are dinamic entities 
+interacting with an evolutionary computation and
+have the fundamental purpose of evaluating individuals according to their preferences,
+both explicitly or implicitly. Their degree of participation is affected by 
+the context of use. With the porpouse of modeling the system, the contextual
+factors are abstracted using a graph. This graph can then be used to 
+enhance the system to capture the attention of users and possibly
+increase their participation and mitigate the fatige.
+
+The current model is tailored to the EvoSpace-I functionality and capabilities
+but it can be modified to handle other platforms.
 
@@
 
-Users are a central part of this proposed framework as it aims to study their
-behavior when interacting with individuals within interactive evolutionary
-algorithms and other tasks that may exist within IEC systems. Thus users in
-this proposal are entities interacting with an evolutionary computation that
-has the fundamental purpose of evaluating individuals of a population
-replacing the fitness function according to their preferences and mood, among
-others []. This form of evaluate individual is given a subjectively way.
+Currently there exist some basic actions that
+users can perform in the system, in order to interact with EvoSpace-I platform
+users must login to the system through their facebook account. 
+Once a user has logged-in to the system, he or she can do several actions.
+Example of actions are the following:
 
-On the other hand in order to capture the attention of users and possibly
-increase their participation, currently there exist some basic actions that
-users can perform in a given system, going from to be able to access a system
-through login mechanism, once the user has logged-in in to the system the
-users can interact with different actions, for example the inviting action is
-when users can be invited through social networks or maybe of person to
-person. Another example is the sharing action, that can occur when users want
-to share something of their interest through their social networks. In this sense the action of posting is when
+\begin{enumerate}
+  \item  Users can invite their friends to participate through social networks .
+
+  \item Relationship between users can also be modeled as a social network,
+  with well stablished algorithms and metrics (cite).
+
+  \item Many kinds of relationships between users and individuals,
+  can be expressed using common actions found in social networks;
+  for instance share, like, or add to collection. [ cite Activity Streams 2.0
+  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]
+
+  \item The same graph can be used to immplement a user model, or it can
+  be extracted from a user\'s public profile.
+
+\end{enumerate}
+
+Another example is the sharing action, that can occur when In this sense the action of posting is when
 users want to put something on their wall who want others to know through

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Corrected packages and found a few errors
Abstract reviewed.
Needs some restructuring and to address some issues /cc @mariosky.

@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
   % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
+  \graphicspath{{../pdf/}{../jpeg/}}
   % and their extensions so you won't have to specify these with
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
@@
   % driver is specified.
-  % \usepackage[dvips]{graphicx}
+  \usepackage[dvips]{graphicx}
   % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
+  \graphicspath{{../eps/}}
   % and their extensions so you won't have to specify these with
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
+  \DeclareGraphicsExtensions{.eps}
 \fi
@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
 % *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
+\ifCLASSOPTIONcompsoc
+  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\else
+  \usepackage[caption=false,font=footnotesize]{subfig}
+\fi
 % subfig.sty, written by Steven Douglas Cochran, is the modern replacement
@@
 This paper presents an alternative data model for the development of
-Interactive Evolutionay Computation (IEC) algorithms. A graph database is
+Interactive Evolutionay Computation (IEC) algorithms. A graph %specify
+                                %the elements of the graph - JJ
+stored for efficiency in a graph database, is
 proposed to model the overall system which consists of: a social network of
-volunteers, a population of individuals in the evolutionary algorithm,
+volunteers participating in the interactive evolutionary experiment, a
+population of individuals in the evolutionary algorithm, 
 the system\'s functionality and their interactions. This graph model can then be used
-to tackle some of the issues associated with IEC; such as, 
-user fatigue, lack of volunteer participation, fitness estimation and novelty. 
+to tackle some of the issues associated with IEC such as, 
+user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
+%Maybe the abstract should start with the issues we want to tackle as
+%motivation. Motivation always goes first - JJ
+and novelty. % These two last issues are not too clear, maybe they
+             % could be explained in that motivation. 
 The proposed graph database could be used to model IEC systems that 
-follow a user-centered approach. A case study is presented providing both 
-conceptual and implementation details using the EvoSpace-Interactive platform.
+follow a user-centered approach. % Graph is the model, graph database
+                                % the implementation. Two different
+                                % things. 
+A case study is presented providing both 
+conceptual and implementation details using the EvoSpace-Interactive
+platform.
+% It must prove that the issues have been addressed - JJ
 \end{abstract}

### 2017-01-17 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
A bit of revision
But more grammar revision and adding the rest of the paper is needed

@@
 \section{Introduction}
-Interactive Evolutionary Algorithms (IEAs): are standard EAs whose
-fitness evaluations are performed by persons with in an interactive 
+% Maybe an introduction on why and where to use these types of
+% algorithms? - jj
+Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
+fitness evaluations are performed by persons within an interactive 
 system.  Thus, the main loop of the EA needs the intervention of a
 human for the quality assessment of each proposed solution. These
-systems have an inherent drawback araising from the very nature of 
-the algorithm, namely, the human fatigue caused by the interaction. 
-Some authors have already tried to address this basic problem \cite{Frade:2010:EvoGAMES}.
-nevertheless, IEAs have demonstrated their ability for effectively
+systems have an inherent drawback arising from the very nature of 
+the algorithm, namely, the human fatigue caused by the interaction. % Can it be quantified? Maybe a reference to NodIO? - JJ
+Some authors have already tried to address this basic problem
+\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
+                            %going to do here? - jj
+
+% Maybe this claim should go to the beginning. THIS is why we want to
+% use them - JJ
+Nevertheless, IEAs have demonstrated their ability for effectively
 producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
 as well as other products in many other domains \cite{ie1}. 
-Several IEC have been developed as web based systems \cite{picbreeder},
+Several IECs have been developed as web based systems \cite{picbreeder},
 who depend on volunteers who visit the system to interactively help
 with the search, using both anonymous and registered users. In particular
-the EvoSpace-Interactive (ES-I) platform employs a collaborative technique,
+the EvoSpace-Interactive (ES-I) %citation
+platform employs a collaborative technique,
 where several registered users assign a quality assessment to a single
@@
 most popular individual, or the the user with more participation \cite{picbreeder}.
-In order to do this, individuals about individuals must be stored, even
+In order to do this, individuals about individuals %what? Maybe data
+                                %about individuals? - JJ
+must be stored, even
 if they are no longer in the active population. In this work a graph database
@@
 
-The remainder of this paper proceeds as follows.
+The remainder of this paper is organized  as follows.
 Section \ref{sec:interactive} presents related work on the topic 
@@
 
-Users are a central part of an IEC system, because not only their
-explicit evaluation is involved, the overall context of interaction
+Users are a central part of an IEC system,  not only because their
+explicit evaluation is essential, but also because the overall context of interaction
 affects the system as a whole as users interact not only with the graphical
@@
 The current model is tailored to the EvoSpace-I functionality and capabilities
-but it can be modified to handle other platforms.
+but it can be modified to handle other platforms. %A hint on how
 
@@
 
-Currently there exist some basic actions that
+Currently there are some basic actions that
 users can perform in the system, in order to interact with EvoSpace-I platform
@@
 
-Another example is the sharing action, that can occur when In this sense the action of posting is when
+Another example is the sharing action, that can occur when %Something
+% missing here - JJ
+
+In this sense the action of posting is when
 users want to put something on their wall who want others to know through
@@
 
-In figure \ref{fig:User_node} the vertex $u$ properties are shown.
+In figure \ref{fig:User_node} the properties of vertex $u$  are shown.
 
@@
 defines the element type that will be the vertex $i$. The property “created”
-defines the creation date of the vertex $i$.
+defines the creation date of the vertex $i$. % Is there a point in
+                                % describing these properties? - JJ
 
@@
 
-Already defined the vertices and edges a graph-based user-individual model is
-proposed where vertices users are related to the vertices individuals, likewise
+Once  the vertices and edges have been defined a graph-based user-individual model is
+proposed where  user vertices are related to the individual vertices, likewise
 the vertices individuals are related to themselves when evolution exist between
-them. In that sense the vertices users and individuals are related to the
+them. % Grammar has to be improved - JJ
+In that sense the vertices users and individuals are related to the
 vertices collections as a vertex user can be related to a vertex collections and

### 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Working on the Introduction a bit

@@
 This paper presents an alternative data model for the development of
-Interactive Evolutionay Computation (IEC) algorithms. A graph %specify
+Interactive Evolutionary Computation (IEC) algorithms. A graph %specify
                                 %the elements of the graph - JJ
@@
 % algorithms? - jj
+% Done -Mario
 Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
@@
 system.  Thus, the main loop of the EA needs the intervention of a
-human for the quality assessment of each proposed solution. These
-systems have an inherent drawback arising from the very nature of 
-the algorithm, namely, the human fatigue caused by the interaction. % Can it be quantified? Maybe a reference to NodIO? - JJ
+human for the quality assessment of each proposed solution.
+IEAs have demonstrated their ability for effectively
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
+as well as other products in many other domains \cite{ie1}. 
+Several IECs have been developed as web based systems \cite{picbreeder},
+who depend on volunteers who visit the system to interactively help
+with the search, using both anonymous and registered users.
+These systems have several inherent drawbacks arising from the very nature of 
+the algorithm, namely, the human fatigue caused by the interaction,
+boredom or lack of interest when evaluating a large number of individuals,
+many of which are not interesting or are very similar to each other.
+IEAs share some of the issues commonly found in volunteer based systems
+\cite{sarmenta2001volunteer,web:BOINC} such as volunteer\'s lack of accountability,
+the need to build trust between participants and application providers. 
+There is also the difficulty of establishing the amount of time and resources
+a volunteer is willing to spend on the system, and how they decide if they
+participate or not \cite{JJ:2016}.
+
+ % Can it be quantified? Maybe a reference to NodIO? - JJ
 Some authors have already tried to address this basic problem
@@
 % use them - JJ
-Nevertheless, IEAs have demonstrated their ability for effectively
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
-as well as other products in many other domains \cite{ie1}. 
-Several IECs have been developed as web based systems \cite{picbreeder},
-who depend on volunteers who visit the system to interactively help
-with the search, using both anonymous and registered users. In particular
+ In particular
 the EvoSpace-Interactive (ES-I) %citation
@@
 of Interactive Evolution data models.
-Then, Section \ref{sec:evospace-i} presents the computational platform on which the main proposal of this work is developed. The proposed graph model is presented in Section \ref{sec:graph}.
+Then, Section \ref{sec:evospace-i} presents the computational platform on which 
+the main proposal of this work is developed. The proposed graph model is 
+presented in Section \ref{sec:graph}.
 The experimental work and results are presented in Section \ref{sec:experiments}.
@@
 with other users, their previous evaluations and their own history of activities
-and past experiences. Thus users in this proposal are dinamic entities 
+and past experiences. Thus users in this proposal are dynamic entities 
 interacting with an evolutionary computation and
@@
 both explicitly or implicitly. Their degree of participation is affected by 
-the context of use. With the porpouse of modeling the system, the contextual
+the context of use. With the purpose of modeling the system, the contextual
 factors are abstracted using a graph. This graph can then be used to 
 enhance the system to capture the attention of users and possibly
-increase their participation and mitigate the fatige.
+increase their participation and mitigate the fatigue.
 
@@
 users can perform in the system, in order to interact with EvoSpace-I platform
-users must login to the system through their facebook account. 
+users must login to the system through their Facebook account. 
 Once a user has logged-in to the system, he or she can do several actions.

### 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Working just on the Intro

@@
 These systems have several inherent drawbacks arising from the very nature of 
-the algorithm, namely, the human fatigue caused by the interaction,
-boredom or lack of interest when evaluating a large number of individuals,
+the algorithms, namely, the human fatigue caused by the interaction, and
+the boredom arising when users evaluate a large number of individuals 
 many of which are not interesting or are very similar to each other.
-IEAs share some of the issues commonly found in volunteer based systems
-\cite{sarmenta2001volunteer,web:BOINC} such as volunteer\'s lack of accountability,
-the need to build trust between participants and application providers. 
+IEAs also share some of the issues commonly found in volunteer based systems
+\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
+and the need to build trust between participants and application providers. 
 There is also the difficulty of establishing the amount of time and resources
 a volunteer is willing to spend on the system, and how they decide if they
-participate or not \cite{JJ:2016}.
+participate or not \cite{JJ:2016}. In order to increase participation and 
+engagement, and to tackle the issues mentioned above,  
+IEAs must follow a user centered design, giving extensive attention 
+to volunteer users, not only because their
+explicit evaluation is essential, but also because the context of the 
+interaction affects the system as a whole. Users interact not only with the graphical
+interface, and each individual\'s phenotype, there is also the interaction
+with other users, their previous evaluations and their own stream of activities
+and past experiences. Thus users in this proposal are defined as dynamic entities 
+interacting with an evolutionary computation having the fundamental purpose 
+of evaluating individuals according to their preferences,
+both explicitly or implicitly. In this work we propose giving the same
+importance to users and their interactions as the population of 
+individuals have in a traditional EA, even to the extend as using this
+knowledge as an integral part of the algorithm it self. We argue also
+that a graph model is a viable representation for this purpose.
+
+% Moving to related work - Mario
+%\cite{Performance for the masses and Christian's previous work},
 
  % Can it be quantified? Maybe a reference to NodIO? - JJ
-Some authors have already tried to address this basic problem
-\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
+%Some authors have already tried to address this basic problem
+%\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
                             %going to do here? - jj
+%% Moving this to Related Work Section -Mario
 
@@
 % use them - JJ
- In particular
+% Done - Mario
+
+As a proof of concept, a case study is presented providing both 
+conceptual and implementation details using the EvoSpace-Interactive
+platform.
+
 the EvoSpace-Interactive (ES-I) %citation

### 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Intro and ES-I Sections

@@
 and the need to build trust between participants and application providers. 
-There is also the difficulty of establishing the amount of time and resources
+For the provider there is also the difficulty of establishing 
+the amount of time and resources
 a volunteer is willing to spend on the system, and how they decide if they
@@
 knowledge as an integral part of the algorithm it self. We argue also
-that a graph model is a viable representation for this purpose.
+that a graph model is a viable representation for this purpose,  
+presenting a natural way of maping the relationships found in the system. 
+Most IEAs can have the following many-to-many relations:
+
+\begin{itemize}
+  \item { \tt Individual-Individual} The relationship between individuals 
+  is naturally expressed as a tree, or a graph when both ancestors 
+  and descendants is considered. Other semantic relations can also be considered
+  for instance: similarity or composition.
+
+  \item {\tt User-Individual}
+  Many kinds of relationships between users and individuals,
+  can be expressed using common actions found in social networks;
+  for instance share, like, rate or add to collection [cite Activity Streams 2.0
+  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
+  relationships are used to assigned a single or a collection of
+  fitness values to individuals [cite ES-I].
+
+  \item {\tt User-User}
+  Relationship between users can also be modeled 
+  as a social network, with well stablished algorithms and metrics (cite).
+  Identifyng leaders of opinion and the similarity between user\'s 
+  preferences is of special importance for IEAs, as they could be used
+  to recommend to users only those individuals they will probably like more.
+
+\end{itemize}
+
+These relations are naturaly maped to graphs, while other actions and 
+properties can also be modeled as a semantic network using
+basic subject–-predicate-–object expressions [cite Semantic triples].
+Another reason for using a graph model is the tools available for
+implementing and query the model, and the increased performance of
+native graph database systems [cite Christian].
 
@@
 
-As a proof of concept, a case study is presented providing both 
-conceptual and implementation details using the EvoSpace-Interactive
-platform.
-
-the EvoSpace-Interactive (ES-I) %citation
-platform employs a collaborative technique,
-where several registered users assign a quality assessment to a single
-individual and then an aggregated fitness value is calculated. The fitness
-assigned to each individual depends on the taste of each particular user, 
-resulting in a many-to-many relationship between users and individuals. 
-Many systems query this user-individual relation to extract relevant
-knowledge about the process and the population, for instance showing the
-most popular individual, or the the user with more participation \cite{picbreeder}.
-In order to do this, individuals about individuals %what? Maybe data
-                                %about individuals? - JJ
-must be stored, even
-if they are no longer in the active population. In this work a graph database
-is proposed to model the interaction between users and individuals, with
-the following advantages:
-
-\begin{enumerate}
-  \item  The relationship between individuals is naturally expressed as a
-  tree, or a graph when both ancestors and descendancy is considered.
-
-  \item Relationship between users can also be modeled as a social network,
-  with well stablished algorithms and metrics (cite).
-
-  \item Many kinds of relationships between users and individuals,
-  can be expressed using common actions found in social networks;
-  for instance share, like, or add to collection. [ cite Activity Streams 2.0
-  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]
-
-  \item The same graph can be used to immplement a user model, or it can
-  be extracted from a user\'s public profile.
-
-\end{enumerate}
-
-In this work..
+A case study is presented as a proof of concept, providing both 
+conceptual and implementation details of a graph data model for
+the implementation of an IEA in the EvoSpace-Interactive platform. 
+The graph database is then used to measure the degree of relationship
+between users and their participation.
 
@@
 \label{sec:evospace-i} 
+As a case study, a IEA was implented using the 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
+A brief description of the aplication is presented next, focusing
+on the data elements that where ported to the graph model.
+
+\subsection{Fitness Assignment}
+\label{sec:col}
+The ES-I platform employs a collaborative technique,
+where several registered users assign a quality assessment to a single
+individual and then an aggregated fitness value is calculated. The fitness
+assigned to each individual depends on the taste of each particular user, 
+resulting in a many-to-many relationship between users and individuals. 
+Many systems query this user-individual relation to extract relevant
+knowledge about the process and the population, for instance showing the
+most popular individual, or the the user with more participation \cite{picbreeder}.
+In order to do this, data about individuals %what? Maybe data
+                                %about individuals? - JJ
+must be permanently stored, even
+if they are no longer in the active population.
+
+\subsection{Collaboration}
+\label{sec:col}
+After entering the web application by using their Facebook account,
+users can collaborate with their Facebook friends, 
+sharing those individuals they like, or by taking individuals
+from their friend\'s collections by using the web interface depicted 
+in Figure \ref{fig:web}
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
+\caption{User interface ED01.}
+\label{fig:web}       
+\end{figure*}
+
+At the top left corner list of Facebook friends is presented, 
+to encourage users to interact with the system. In the central 
+\emph{ Wall } area, an individual sampled from the population 
+is shown to the user.
+Here, the user can interact with the system in two ways.
+He can assign a quality assessment to the individual using
+a five star rating measure.
+Additionally, a user can choose to add an image to one of their \emph{Collections}.
+A collection is a special directory to store individuals a user prefers and wishes
+to save. After the user finishes interacting the individual
+on the Wall, he can choose to retrieve a new one from population.
+At the left side, is the \emph{Collections} section.
+The user can create several collections, to group and organize his favorite 
+artifacts. Moreover, a user can browse the content of each collection and from
+there share images through the social network.
+When a user browses over an individual a detail pane shows how many users have
+liked the individual. The pane also includes a link to the individual's 
+details, the parents, genetic operators that created it, and genealogy information.
 
@@
 \label{sec:graph}
+The fundamental goal of this research is to develop a user-centered framework
+for interactive evolutionary computation (IEC) in order to increase users’
+participation and also to minimize the amount of evaluations needed for the
+evolutionary process in given Web-based IEC application.
 
-Users are a central part of an IEC system,  not only because their
-explicit evaluation is essential, but also because the overall context of interaction
-affects the system as a whole as users interact not only with the graphical
-interface, and each individual\'s phenotype, there is also the interaction
-with other users, their previous evaluations and their own history of activities
-and past experiences. Thus users in this proposal are dynamic entities 
-interacting with an evolutionary computation and
-have the fundamental purpose of evaluating individuals according to their preferences,
-both explicitly or implicitly. Their degree of participation is affected by 
-the context of use. With the purpose of modeling the system, the contextual
-factors are abstracted using a graph. This graph can then be used to 
-enhance the system to capture the attention of users and possibly
-increase their participation and mitigate the fatigue.
+In this chapter an explanation of the proposed framework is described. The
+different techniques used, such as user modeling, fuzzy logic, and human-
+interaction.
+
+This framework is presented in figure \ref{fig:uc_framework}. Each of the
+entities of this framework will be explained in detail in following sections.
 
 The current model is tailored to the EvoSpace-I functionality and capabilities
-but it can be modified to handle other platforms. %A hint on how
+but it can be modified to handle other platforms, just by adding new clases 
+of nodes.
+
+\begin{figure*}
+  \captionsetup{justification=centering,margin=2cm}
+  \centering
+  \setlength\fboxsep{0pt}
+  \setlength\fboxrule{0.7pt}
+  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
+  \caption{User-Centerd Framework.}
+  \label{fig:uc_framework}       
+\end{figure*}
+
+%A hint on how
+% done -Mario
 
 \subsection{Users}
-
-Currently there are some basic actions that
-users can perform in the system, in order to interact with EvoSpace-I platform
-users must login to the system through their Facebook account. 
-Once a user has logged-in to the system, he or she can do several actions.
-Example of actions are the following:
-
-\begin{enumerate}
-  \item  Users can invite their friends to participate through social networks .
-
-  \item Relationship between users can also be modeled as a social network,
-  with well stablished algorithms and metrics (cite).
-
-  \item Many kinds of relationships between users and individuals,
-  can be expressed using common actions found in social networks;
-  for instance share, like, or add to collection. [ cite Activity Streams 2.0
-  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]
-
-  \item The same graph can be used to immplement a user model, or it can
-  be extracted from a user\'s public profile.
-
-\end{enumerate}
-
-Another example is the sharing action, that can occur when %Something
-% missing here - JJ
 
@@
 individual in the population following the genetic operators such as selection,
-crossover and mutation[]. An example of a individual in this case it is
-illustrated in figure \ref{fig:individual}
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/individual.png}}
-\caption{Individual representation.}
-\label{fig:graph}       
-\end{figure*}
+crossover and mutation[].
 

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
About to start explaining the model

@@
 \begin{abstract}
-This paper presents an alternative data model for the development of
-Interactive Evolutionary Computation (IEC) algorithms. A graph %specify
-                                %the elements of the graph - JJ
-stored for efficiency in a graph database, is
-proposed to model the overall system which consists of: a social network of
-volunteers participating in the interactive evolutionary experiment, a
-population of individuals in the evolutionary algorithm, 
-the system\'s functionality and their interactions. This graph model can then be used
-to tackle some of the issues associated with IEC such as, 
-user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
+Interactive evolution is the leading technique employed by 
+Evolutionary Art researchers, where users' preferences guide the search. A common
+practice is to develop Interactive Evolutionary Computation (IEC) algorithms
+as web applications in order to give users easy access the application as they
+often depend on volunteers who visit the system to interactively help with the search. 
+There are several drawbacks that limit user participation 
+such as human fatigue and boredom resulting from evaluating 
+a large number of individuals. 
+To tackle this issues, we propose the design of IEC systems using a user-centered 
+approach, with a framework consisting of a social network of 
+volunteers, interacting with a population consisting again of a network 
+of individuals. In this work we propose a graph model as a viable 
+and efficient tool for mapping in a natural way the relationships 
+found in the system. A case study is presented as a proof of concept, 
+providing both conceptual and implementation details of a graph based model as it is 
+applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
+As a result, the graph model was then succesfully applied to measure the degree
+of relationship between users with the porpuse of quantifing 
+their participation.
+
+% what is this issue?
 %Maybe the abstract should start with the issues we want to tackle as
 %motivation. Motivation always goes first - JJ
-and novelty. % These two last issues are not too clear, maybe they
+%and novelty. % These two last issues are not too clear, maybe they
              % could be explained in that motivation. 
-The proposed graph database could be used to model IEC systems that 
-follow a user-centered approach. % Graph is the model, graph database
+%The proposed graph database could be used to model IEC systems that 
+%follow a user-centered approach. % Graph is the model, graph database
                                 % the implementation. Two different
                                 % things.
-A case study is presented providing both 
-conceptual and implementation details using the EvoSpace-Interactive
-platform.
+
 % It must prove that the issues have been addressed - JJ
@@
 individuals have in a traditional EA, even to the extend as using this
-knowledge as an integral part of the algorithm it self. We argue also
+knowledge as an integral part of the evolutionary algorithm. We argue also
 that a graph model is a viable representation for this purpose,  
 presenting a natural way of maping the relationships found in the system. 
-Most IEAs can have the following many-to-many relations:
+Most IEAs have the following many-to-many relations:
 
 \begin{itemize}
-  \item { \tt Individual-Individual} The relationship between individuals 
-  is naturally expressed as a tree, or a graph when both ancestors 
-  and descendants is considered. Other semantic relations can also be considered
-  for instance: similarity or composition.
+  \item { \tt Individual-Individual} The relationships between individuals in the
+  population are naturally expressed as a tree, or a graph when both ancestors 
+  and descendants are considered. Other semantic relations can also be of interest
+  for instance similarity or composition.
 
   \item {\tt User-Individual}
-  Many kinds of relationships between users and individuals,
+  Many kind of relationships between users and individuals,
   can be expressed using common actions found in social networks;
-  for instance share, like, rate or add to collection [cite Activity Streams 2.0
+  for instance: share, like, rate or add to collection [cite Activity Streams 2.0
   syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
-  relationships are used to assigned a single or a collection of
+  relationships are used to assign a single or a collection of
   fitness values to individuals [cite ES-I].
@@
   Relationship between users can also be modeled 
-  as a social network, with well stablished algorithms and metrics (cite).
-  Identifyng leaders of opinion and the similarity between user\'s 
-  preferences is of special importance for IEAs, as they could be used
-  to recommend to users only those individuals they will probably like more.
+  as a social network, with well stablished semantics, algorithms and metrics (cite).
+  A graph model could eneable researchers to find other ways of identifyng leaders of 
+  opinion or measuring the similarity between user\'s preferences. 
+  These measures can then be used by recommender algorithms selecting 
+  individuals according users\' preferences.
 
@@
 A case study is presented as a proof of concept, providing both
-conceptual and implementation details of a graph data model for
-the implementation of an IEA in the EvoSpace-Interactive platform. 
-The graph database is then used to measure the degree of relationship
-between users and their participation.
+conceptual and implementation details of a graph data model applied in the 
+implementation of an IEA. The experiment was implmented using
+the EvoSpace-Interactive framework, and consisted of a web application
+for evolving artistic drawings. The graph database is then used to
+measure the degree of relationship between users and their participation.
 
@@
 \label{sec:interactive}
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
+\caption{User interface of an EvoSpace-Interactive application.}
+\label{fig:web}       
+\end{figure*}
 
@@
 in Figure \ref{fig:web}
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
-\caption{User interface ED01.}
-\label{fig:web}       
-\end{figure*}
 
@@
 participation and also to minimize the amount of evaluations needed for the
-evolutionary process in given Web-based IEC application.
-
-In this chapter an explanation of the proposed framework is described. The
-different techniques used, such as user modeling, fuzzy logic, and human-
-interaction.
-
-This framework is presented in figure \ref{fig:uc_framework}. Each of the
-entities of this framework will be explained in detail in following sections.
-
-The current model is tailored to the EvoSpace-I functionality and capabilities
-but it can be modified to handle other platforms, just by adding new clases 
-of nodes.
+evolutionary process in given Web-based IEC application.  This framework is 
+presented in figure \ref{fig:uc_framework}. The current model is tailored 
+to the EvoSpace-I functionality and capabilities but it can be modified to 
+handle other platforms, just by adding new clases of nodes.
 
@@
 
+The user centered framework consists of three high level components:
+
+\begin{itemize}
+  \item { \tt Interactive Evolutionary Computational system.}  
+  This is the real world system that we are going to model, it consist of
+  human users and their interaction with one or more individuals from
+  the population, this could be through a device, user interface or
+  by directly interacting with real world objects generated as part of the algorithm.
+
+  \item {\tt IEC Graph}
+  This is the model proposed in this work: A graph database modeling
+  the IEC system. The  main components are the social network of users and the population. This database is the back-end of the system, is used by the EA and
+  the user interface.
+
+  \item {\tt Evolutionary Algorithm} 
+  This algorithm queries the graph model, and then updates those nodes that 
+  are part of the population. 
+\end{itemize}
+
 %A hint on how
@@
 
-\subsection{Users}
+\subsection{IEC Graph Model}
+
+In this section the Graph Model is presented,
 
@@
 \end{figure*}
-
-In order to start the task of evaluating individuals is necessary that users
-access the system through a login mechanism. This mechanism consists of
-providing a username and a password in order to grant access to the system. All
-users accessing in this way they are considered active users within an IEC
-system.
 

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Model Section moved to temporal file

@@
   These measures can then be used by recommender algorithms selecting 
-  individuals according users\' preferences.
+  individuals according to users
+In this sense the action of posting is when
+users want to put something on their wall who want others to know through
+their social networks. Likewise the storing action occurs when users want to
+retain permanently information that are of their interest. Finally the
+browsing action occurs when users are exploring content for their needs. All
+these actions go beyond only evaluating individuals.
+
+\begin{figure*}
+  \captionsetup{justification=centering,margin=2cm}
+  \centering
+  \setlength\fboxsep{0pt}
+  \setlength\fboxrule{0.7pt}
+  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/users.png}}
+  \caption{Users actions on interactive evolutionary systems.}
+  \label{fig:users}       
+\end{figure*}
+
+For the evaluating action is proposed  that users evaluate individuals
+indirectly, this means users can evaluate accepting indirect recommendations of
+friends that are currently active in the system. These recommendations can be
+store individuals in the system of users who know each other.
+
+Also the  browsing action is proposed within IEC systems,  this means that users
+can explore information that other active users in the system are generating.
+
+Finally activity participation is proposed. This can be divided into four
+different actions as fallows:
+
+\begin{itemize} 
+\item The Inviting action occurs when a user invites another to the system through social networks or from person to person.  
+\item The sharing action occurs when users share their individual creations with others active user in the system.
+\item The posting action occurs when published what they are doing within the system.
+\item The storing action occurs when users keep individuals in a collection, the collections concept will be discussed later in this chapter.
+\end{itemize}
+
+All the mention above is shown in figure \ref{fig:users}
+
+\subsection{Individual}
+
+Individuals are entities that form the population for the evolutionary
+algorithm. In particular for  this work individuals are animated digital
+paintings, every individual in the population is defined by a chromosome [].
+This chromosome defines the behavior of the individual, so that it consists of a
+vector of real numbers of fifteen genes, where each gen define a particular
+behavior in painting. This individuals combined each other to generated new
+individual in the population following the genetic operators such as selection,
+crossover and mutation[].
+
+\subsection{User Model} Now that both users an individuals have been defined we can
+model the users behavior have with respect to individuals in a graph-based user
+model.
+
+The reason for using a graph to model the user-individual is because it could
+can express in a simple way the behavior of this two entities having the
+necessary information  to represent the knowledge,  for example a vertex user
+and individual vertex connected through a edge that  represents the
+relationship. this relationship can be the semantic  between these two entities
+as figure \ref{fig:User-Individual} show. Each of these vertices contain properties as well as the
+edges, these properties will be explained as fallows.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_individual.png}}
+\caption{User-Individual.}
+\label{fig:User-Individual}       
+\end{figure*}
+
+In this proposed graph-based user-individual model the vertices are represented
+by set of nodes of USERS, INDIVIDUALS and COLLECTIONS. The edges are define by a
+set of relationships as LIKES, KNOWS, PARENT, HAS that represents the
+relationships between the vertices as seen in figure \ref{fig:User-Individual}.
+
+Already explained that represent USERS and INDIVIDUALS, now an explanation is
+given of what concerns for the concept of COLLECTIONS.
+
+In this proposal a collection is defined as a  container where users can store
+selected individuals laid within the collection. A collection is created when
+the user wants to store permanently individuals within the IEC system.
+
+Now for the edges in this proposal are explained as follows:
+
+\begin{itemize} 
+\item The  LIKES edge represents a preference relationship that a user has with respect to an individual.  
+\item Now the  PARENT edge represents an ancestor of the individual, in other words where the individual comes from.
+\item Likewise the  HAS edge represents that users have a collection where can contain individuals.
+\item Finally the KNOWS edge represents a relationship of friendship between users.
+\end{itemize}
+
+In figure \ref{fig:Nodes_Edges} illustrated the above mentioned, about nodes and edges.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/user_individual_collections.png}}
+\caption{Nodes And Edges Representation.}
+\label{fig:Nodes_Edges}       
+\end{figure*}
+
+In each vertex is necessary to store knowledge about the users, individuals and
+also for the collections, this is what is known as property.
+
+It is worth to mention that all vertices have a property called “element\_type”
+among others. This property is used to identify which type of vertex is, for
+example if vertex corresponds to a user then this property label the vertex as a
+“person”. Likewise if the vertex corresponds to an individual this property
+label it as an “individual” and is the same for any vertex we want to add to the
+model. This property also has the purpose of being able to do operations on the
+graph according to the vertex type.
+
+The properties for the vertex “u” are the following:
+
+\begin{itemize} 
+\item id.  
+\item created.
+\item name.
+\item element\_type.
+\end{itemize}
+
+The “id” property defines that the vertex u has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The property
+“created” defines the creation date of the vertex $u$. The property “name”
+defines the name of the vertex u. The property “element\_type” defines the
+element type that will be the vertex $u$.
+
+In figure \ref{fig:User_node} the properties of vertex $u$  are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_node.png}}
+\caption{User Properties.}
+\label{fig:User_node}       
+\end{figure*}
+
+The properties for the vertex $i$ are the following:
+\begin{itemize} 
+\item id. 
+\item created. 
+\item element\_type.
+\item chromosome.
+\item views.
+\end{itemize}
+
+The “id” property defines that the vertex $i$ has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The
+“chromosome” property defines the chromosome of the individual representing the
+vertex $i$. which as defined above in this section. The  property “views” defines
+the amount that users have seen this vertex $i$ The property “element\_type”
+defines the element type that will be the vertex $i$. The property “created”
+defines the creation date of the vertex $i$. % Is there a point in
+                                % describing these properties? - JJ
+
+In figure \ref{fig:Individual_node} the vertex $i$ properties are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/individual_node.png}}
+\caption{Individual Properties.}
+\label{fig:Individual_node}       
+\end{figure*}
+
+The properties for the vertex “c” are the following:
+
+\begin{itemize} 
+\item id. 
+\item element\_type.
+\item created. 
+\item name.
+\end{itemize}
+
+The “id” property defines that the vertex $c$ has a unique identifier in the set
+of vertices, which means that there will not be duplicate vertices. The property
+“element\_type” defines the element type that will be the vertex $c$. The
+property “name” defines the name of the vertex $c$. The property “created”
+defines the creation date of the vertex $c$.
+
+In figure \ref{fig:Collection_node} the vertex $c$ properties are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/collection_node.png}}
+\caption{Collection Properties.}
+\label{fig:Collection_node}       
+\end{figure*}
+
+In the same way that the vertices have properties also the edges, these
+properties are defined as follows.
+
+The “LIKES” edge has the following properties:
+
+\begin{itemize} 
+\item id. 
+\item created. 
+\item rate.
+\end{itemize}
+
+The “id” property defines that the “LIKES” edge  has a unique identifier in the
+set of edges, which means that there will not be duplicate edges. The property
+“created” defines the date of creation of the this edge. The property “rate”
+defines the rate that the users give to the individual store in this edge.
+
+In figure \ref{fig:Likes_edge} the edge “LIKES” properties and the relationships with the vertices $u$ and $i$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_likes.png}}
+\caption{LIKES edge Properties.}
+\label{fig:Likes_edge}       
+\end{figure*}
+
+The “KNOWS” edge has the following properties.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “KNOWS” has a unique identifier in the
+set of edges, which means that there will not be duplicate edges. The property
+“created” defines the date of creation of the edge.
+
+In figure \ref{fig:Knows_edge} the edge “KNOWS” properties and the relationships between vertex “u”
+are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_knows.png}}
+\caption{KNOWS edge Properties.}
+\label{fig:Knows_edge}       
+\end{figure*}
+
+The “PARENT” edge has the following properties, and represent the parents of a new individual.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “PARENT” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
+The property “created” defines the date of creation of the edge.
+
+In figure \ref{fig:Parent_edge} the edge “PARENT” properties and the relationship between vertex $i$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_parent.png}}
+\caption{PARENT edge Properties.}
+\label{fig:Parent_edge}       
+\end{figure*}
+
+The edge “HAS” has the following properties, and represent the parents of a new individual.
+\begin{itemize} 
+\item id. 
+\item created. 
+\end{itemize}
+
+The “id” property defines that the edge “HAS” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
+The property “created” defines the date of creation of the edge.
+
+In figure \ref{fig:Has_edge} the edge “HAS” properties and the relationship between vertices $u$, $i$ and $c$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
+\caption{HAS edge Properties.}
+\label{fig:Has_edge}       
+\end{figure*}
+
+\begin{equation*}\label{eq:graphRelDef} 
+\displaystyle 
+\begin{split} 
+V &= \{[u_1,u_2,u_3,...,u_n],[i_1,i_2,i_3,...,i_n],[c_1,c_2,c_3,...,c_n]\},\\ 
+E&= \{[l_1,l_2,l_3,..,l_n],[p_1,p_2,p_3,...,p_n],[h_1,h_2,h_3,...,h_n],[k_1,k_2,k_3,...,k_n]\}\\ 
+\end{split} 
+\end{equation*}
+
+Once  the vertices and edges have been defined a graph-based user-individual model is
+proposed where  user vertices are related to the individual vertices, likewise
+the vertices individuals are related to themselves when evolution exist between
+them. % Grammar has to be improved - JJ
+In that sense the vertices users and individuals are related to the
+vertices collections as a vertex user can be related to a vertex collections and
+a vertex collections collections is related to $n$ individuals vertex. In Figure
+\ref{fig:user_moder} shows an example of the graph-based user-individual model.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=8cm,height=8cm,keepaspectratio]{img/model_representation.png}}
+\caption{Graph-based user-individual model.}
+\label{fig:user_moder}       
+\end{figure*}
+'  preferences.
 
@@
 \begin{itemize}
-  \item { \tt Interactive Evolutionary Computational system.}  
+  \item Interactive Evolutionary Computational system.  
   This is the real world system that we are going to model, it consist of
@@
 
-  \item {\tt IEC Graph}
+  \item IEC Graph.
   This is the model proposed in this work: A graph database modeling
@@
 
-  \item {\tt Evolutionary Algorithm} 
+  \item Evolutionary Algorithm. 
   This algorithm queries the graph model, and then updates those nodes that 
@@
 In this section the Graph Model is presented,
-
-In this sense the action of posting is when
-users want to put something on their wall who want others to know through
-their social networks. Likewise the storing action occurs when users want to
-retain permanently information that are of their interest. Finally the
-browsing action occurs when users are exploring content for their needs. All
-these actions go beyond only evaluating individuals.
-
-\begin{figure*}
-  \captionsetup{justification=centering,margin=2cm}
-  \centering
-  \setlength\fboxsep{0pt}
-  \setlength\fboxrule{0.7pt}
-  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/users.png}}
-  \caption{Users actions on interactive evolutionary systems.}
-  \label{fig:users}       
-\end{figure*}
-
-For the evaluating action is proposed  that users evaluate individuals
-indirectly, this means users can evaluate accepting indirect recommendations of
-friends that are currently active in the system. These recommendations can be
-store individuals in the system of users who know each other.
-
-Also the  browsing action is proposed within IEC systems,  this means that users
-can explore information that other active users in the system are generating.
-
-Finally activity participation is proposed. This can be divided into four
-different actions as fallows:
-
-\begin{itemize} 
-\item The Inviting action occurs when a user invites another to the system through social networks or from person to person.  
-\item The sharing action occurs when users share their individual creations with others active user in the system.
-\item The posting action occurs when published what they are doing within the system.
-\item The storing action occurs when users keep individuals in a collection, the collections concept will be discussed later in this chapter.
-\end{itemize}
-
-All the mention above is shown in figure \ref{fig:users}
-
-\subsection{Individual}
-
-Individuals are entities that form the population for the evolutionary
-algorithm. In particular for  this work individuals are animated digital
-paintings, every individual in the population is defined by a chromosome [].
-This chromosome defines the behavior of the individual, so that it consists of a
-vector of real numbers of fifteen genes, where each gen define a particular
-behavior in painting. This individuals combined each other to generated new
-individual in the population following the genetic operators such as selection,
-crossover and mutation[].
-
-\subsection{User Model} Now that both users an individuals have been defined we can
-model the users behavior have with respect to individuals in a graph-based user
-model.
-
-The reason for using a graph to model the user-individual is because it could
-can express in a simple way the behavior of this two entities having the
-necessary information  to represent the knowledge,  for example a vertex user
-and individual vertex connected through a edge that  represents the
-relationship. this relationship can be the semantic  between these two entities
-as figure \ref{fig:User-Individual} show. Each of these vertices contain properties as well as the
-edges, these properties will be explained as fallows.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_individual.png}}
-\caption{User-Individual.}
-\label{fig:User-Individual}       
-\end{figure*}
-
-In this proposed graph-based user-individual model the vertices are represented
-by set of nodes of USERS, INDIVIDUALS and COLLECTIONS. The edges are define by a
-set of relationships as LIKES, KNOWS, PARENT, HAS that represents the
-relationships between the vertices as seen in figure \ref{fig:User-Individual}.
-
-Already explained that represent USERS and INDIVIDUALS, now an explanation is
-given of what concerns for the concept of COLLECTIONS.
-
-In this proposal a collection is defined as a  container where users can store
-selected individuals laid within the collection. A collection is created when
-the user wants to store permanently individuals within the IEC system.
-
-Now for the edges in this proposal are explained as follows:
-
-\begin{itemize} 
-\item The  LIKES edge represents a preference relationship that a user has with respect to an individual.  
-\item Now the  PARENT edge represents an ancestor of the individual, in other words where the individual comes from.
-\item Likewise the  HAS edge represents that users have a collection where can contain individuals.
-\item Finally the KNOWS edge represents a relationship of friendship between users.
-\end{itemize}
-
-In figure \ref{fig:Nodes_Edges} illustrated the above mentioned, about nodes and edges.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/user_individual_collections.png}}
-\caption{Nodes And Edges Representation.}
-\label{fig:Nodes_Edges}       
-\end{figure*}
-
-In each vertex is necessary to store knowledge about the users, individuals and
-also for the collections, this is what is known as property.
-
-It is worth to mention that all vertices have a property called “element\_type”
-among others. This property is used to identify which type of vertex is, for
-example if vertex corresponds to a user then this property label the vertex as a
-“person”. Likewise if the vertex corresponds to an individual this property
-label it as an “individual” and is the same for any vertex we want to add to the
-model. This property also has the purpose of being able to do operations on the
-graph according to the vertex type.
-
-The properties for the vertex “u” are the following:
-
-\begin{itemize} 
-\item id.  
-\item created.
-\item name.
-\item element\_type.
-\end{itemize}
-
-The “id” property defines that the vertex u has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The property
-“created” defines the creation date of the vertex $u$. The property “name”
-defines the name of the vertex u. The property “element\_type” defines the
-element type that will be the vertex $u$.
-
-In figure \ref{fig:User_node} the properties of vertex $u$  are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_node.png}}
-\caption{User Properties.}
-\label{fig:User_node}       
-\end{figure*}
-
-The properties for the vertex $i$ are the following:
-\begin{itemize} 
-\item id. 
-\item created. 
-\item element\_type.
-\item chromosome.
-\item views.
-\end{itemize}
-
-The “id” property defines that the vertex $i$ has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The
-“chromosome” property defines the chromosome of the individual representing the
-vertex $i$. which as defined above in this section. The  property “views” defines
-the amount that users have seen this vertex $i$ The property “element\_type”
-defines the element type that will be the vertex $i$. The property “created”
-defines the creation date of the vertex $i$. % Is there a point in
-                                % describing these properties? - JJ
-
-In figure \ref{fig:Individual_node} the vertex $i$ properties are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/individual_node.png}}
-\caption{Individual Properties.}
-\label{fig:Individual_node}       
-\end{figure*}
-
-The properties for the vertex “c” are the following:
-
-\begin{itemize} 
-\item id. 
-\item element\_type.
-\item created. 
-\item name.
-\end{itemize}
-
-The “id” property defines that the vertex $c$ has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The property
-“element\_type” defines the element type that will be the vertex $c$. The
-property “name” defines the name of the vertex $c$. The property “created”
-defines the creation date of the vertex $c$.
-
-In figure \ref{fig:Collection_node} the vertex $c$ properties are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/collection_node.png}}
-\caption{Collection Properties.}
-\label{fig:Collection_node}       
-\end{figure*}
-
-In the same way that the vertices have properties also the edges, these
-properties are defined as follows.
-
-The “LIKES” edge has the following properties:
-
-\begin{itemize} 
-\item id. 
-\item created. 
-\item rate.
-\end{itemize}
-
-The “id” property defines that the “LIKES” edge  has a unique identifier in the
-set of edges, which means that there will not be duplicate edges. The property
-“created” defines the date of creation of the this edge. The property “rate”
-defines the rate that the users give to the individual store in this edge.
-
-In figure \ref{fig:Likes_edge} the edge “LIKES” properties and the relationships with the vertices $u$ and $i$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_likes.png}}
-\caption{LIKES edge Properties.}
-\label{fig:Likes_edge}       
-\end{figure*}
-
-The “KNOWS” edge has the following properties.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “KNOWS” has a unique identifier in the
-set of edges, which means that there will not be duplicate edges. The property
-“created” defines the date of creation of the edge.
-
-In figure \ref{fig:Knows_edge} the edge “KNOWS” properties and the relationships between vertex “u”
-are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_knows.png}}
-\caption{KNOWS edge Properties.}
-\label{fig:Knows_edge}       
-\end{figure*}
-
-The “PARENT” edge has the following properties, and represent the parents of a new individual.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “PARENT” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
-The property “created” defines the date of creation of the edge.
-
-In figure \ref{fig:Parent_edge} the edge “PARENT” properties and the relationship between vertex $i$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_parent.png}}
-\caption{PARENT edge Properties.}
-\label{fig:Parent_edge}       
-\end{figure*}
-
-The edge “HAS” has the following properties, and represent the parents of a new individual.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “HAS” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
-The property “created” defines the date of creation of the edge.
-
-In figure \ref{fig:Has_edge} the edge “HAS” properties and the relationship between vertices $u$, $i$ and $c$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
-\caption{HAS edge Properties.}
-\label{fig:Has_edge}       
-\end{figure*}
-
-\begin{equation*}\label{eq:graphRelDef} 
-\displaystyle 
-\begin{split} 
-V &= \{[u_1,u_2,u_3,...,u_n],[i_1,i_2,i_3,...,i_n],[c_1,c_2,c_3,...,c_n]\},\\ 
-E&= \{[l_1,l_2,l_3,..,l_n],[p_1,p_2,p_3,...,p_n],[h_1,h_2,h_3,...,h_n],[k_1,k_2,k_3,...,k_n]\}\\ 
-\end{split} 
-\end{equation*}
-
-Once  the vertices and edges have been defined a graph-based user-individual model is
-proposed where  user vertices are related to the individual vertices, likewise
-the vertices individuals are related to themselves when evolution exist between
-them. % Grammar has to be improved - JJ
-In that sense the vertices users and individuals are related to the
-vertices collections as a vertex user can be related to a vertex collections and
-a vertex collections collections is related to $n$ individuals vertex. In Figure
-\ref{fig:user_moder} shows an example of the graph-based user-individual model.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=8cm,height=8cm,keepaspectratio]{img/model_representation.png}}
-\caption{Graph-based user-individual model.}
-\label{fig:user_moder}       
-\end{figure*}
 

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/temporal.tex
Model Section moved to temporal file

--- 2017-01-18 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-\ifCLASSOPTIONcompsoc
-  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\else
-  \usepackage[caption=false,font=footnotesize]{subfig}
-\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{}
-
-\begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{A graph-based data model for interactive 
-evolutionary computation}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Michael Shell}
-\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
-Email: http://www.michaelshell.org/contact.html}
-\and
-\IEEEauthorblockN{Homer Simpson}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
-\and
-\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678--2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-This paper presents an alternative data model for the development of
-Interactive Evolutionary Computation (IEC) algorithms. A graph %specify
-                                %the elements of the graph - JJ
-stored for efficiency in a graph database, is
-proposed to model the overall system which consists of: a social network of
-volunteers participating in the interactive evolutionary experiment, a
-population of individuals in the evolutionary algorithm, 
-the system\'s functionality and their interactions. This graph model can then be used
-to tackle some of the issues associated with IEC such as, 
-user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
-%Maybe the abstract should start with the issues we want to tackle as
-%motivation. Motivation always goes first - JJ
-and novelty. % These two last issues are not too clear, maybe they
-             % could be explained in that motivation. 
-The proposed graph database could be used to model IEC systems that 
-follow a user-centered approach. % Graph is the model, graph database
-                                % the implementation. Two different
-                                % things. 
-A case study is presented providing both 
-conceptual and implementation details using the EvoSpace-Interactive
-platform.
-% It must prove that the issues have been addressed - JJ
-\end{abstract}
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-% Maybe an introduction on why and where to use these types of
-% algorithms? - jj
-% Done -Mario
-Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
-fitness evaluations are performed by persons within an interactive 
-system.  Thus, the main loop of the EA needs the intervention of a
-human for the quality assessment of each proposed solution.
-IEAs have demonstrated their ability for effectively
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
-as well as other products in many other domains \cite{ie1}. 
-Several IECs have been developed as web based systems \cite{picbreeder},
-who depend on volunteers who visit the system to interactively help
-with the search, using both anonymous and registered users.
-These systems have several inherent drawbacks arising from the very nature of 
-the algorithms, namely, the human fatigue caused by the interaction, and
-the boredom arising when users evaluate a large number of individuals 
-many of which are not interesting or are very similar to each other.
-IEAs also share some of the issues commonly found in volunteer based systems
-\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
-and the need to build trust between participants and application providers. 
-For the provider there is also the difficulty of establishing 
-the amount of time and resources
-a volunteer is willing to spend on the system, and how they decide if they
-participate or not \cite{JJ:2016}. In order to increase participation and 
-engagement, and to tackle the issues mentioned above,  
-IEAs must follow a user centered design, giving extensive attention 
-to volunteer users, not only because their
-explicit evaluation is essential, but also because the context of the 
-interaction affects the system as a whole. Users interact not only with the graphical
-interface, and each individual\'s phenotype, there is also the interaction
-with other users, their previous evaluations and their own stream of activities
-and past experiences. Thus users in this proposal are defined as dynamic entities 
-interacting with an evolutionary computation having the fundamental purpose 
-of evaluating individuals according to their preferences,
-both explicitly or implicitly. In this work we propose giving the same
-importance to users and their interactions as the population of 
-individuals have in a traditional EA, even to the extend as using this
-knowledge as an integral part of the algorithm it self. We argue also
-that a graph model is a viable representation for this purpose,  
-presenting a natural way of maping the relationships found in the system. 
-Most IEAs can have the following many-to-many relations:
-
-\begin{itemize}
-  \item { \tt Individual-Individual} The relationship between individuals 
-  is naturally expressed as a tree, or a graph when both ancestors 
-  and descendants is considered. Other semantic relations can also be considered
-  for instance: similarity or composition.
-
-  \item {\tt User-Individual}
-  Many kinds of relationships between users and individuals,
-  can be expressed using common actions found in social networks;
-  for instance share, like, rate or add to collection [cite Activity Streams 2.0
-  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
-  relationships are used to assigned a single or a collection of
-  fitness values to individuals [cite ES-I].
-
-  \item {\tt User-User}
-  Relationship between users can also be modeled 
-  as a social network, with well stablished algorithms and metrics (cite).
-  Identifyng leaders of opinion and the similarity between user\'s 
-  preferences is of special importance for IEAs, as they could be used
-  to recommend to users only those individuals they will probably like more.
-
-\end{itemize}
-
-These relations are naturaly maped to graphs, while other actions and 
-properties can also be modeled as a semantic network using
-basic subject–-predicate-–object expressions [cite Semantic triples].
-Another reason for using a graph model is the tools available for
-implementing and query the model, and the increased performance of
-native graph database systems [cite Christian].
-
-% Moving to related work - Mario
-%\cite{Performance for the masses and Christian's previous work},
-
- % Can it be quantified? Maybe a reference to NodIO? - JJ
-%Some authors have already tried to address this basic problem
-%\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
-                            %going to do here? - jj
-%% Moving this to Related Work Section -Mario
-
-% Maybe this claim should go to the beginning. THIS is why we want to
-% use them - JJ
-% Done - Mario
-
-A case study is presented as a proof of concept, providing both 
-conceptual and implementation details of a graph data model for
-the implementation of an IEA in the EvoSpace-Interactive platform. 
-The graph database is then used to measure the degree of relationship
-between users and their participation.
-
-The remainder of this paper is organized  as follows.
-Section \ref{sec:interactive} presents related work on the topic 
-of Interactive Evolution data models.
-Then, Section \ref{sec:evospace-i} presents the computational platform on which 
-the main proposal of this work is developed. The proposed graph model is 
-presented in Section \ref{sec:graph}.
-The experimental work and results are presented in Section \ref{sec:experiments}.
-Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
-
-\section{Related Work}
-\label{sec:interactive}
-
-\section{EvoSpace-Interactive}
-\label{sec:evospace-i} 
-As a case study, a IEA was implented using the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
-A brief description of the aplication is presented next, focusing
-on the data elements that where ported to the graph model.
-
-\subsection{Fitness Assignment}
-\label{sec:col}
-The ES-I platform employs a collaborative technique,
-where several registered users assign a quality assessment to a single
-individual and then an aggregated fitness value is calculated. The fitness
-assigned to each individual depends on the taste of each particular user, 
-resulting in a many-to-many relationship between users and individuals. 
-Many systems query this user-individual relation to extract relevant
-knowledge about the process and the population, for instance showing the
-most popular individual, or the the user with more participation \cite{picbreeder}.
-In order to do this, data about individuals %what? Maybe data
-                                %about individuals? - JJ
-must be permanently stored, even
-if they are no longer in the active population.
-
-\subsection{Collaboration}
-\label{sec:col}
-After entering the web application by using their Facebook account,
-users can collaborate with their Facebook friends, 
-sharing those individuals they like, or by taking individuals
-from their friend\'s collections by using the web interface depicted 
-in Figure \ref{fig:web}
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
-\caption{User interface ED01.}
-\label{fig:web}       
-\end{figure*}
-
-At the top left corner list of Facebook friends is presented, 
-to encourage users to interact with the system. In the central 
-\emph{ Wall } area, an individual sampled from the population 
-is shown to the user.
-Here, the user can interact with the system in two ways.
-He can assign a quality assessment to the individual using
-a five star rating measure.
-Additionally, a user can choose to add an image to one of their \emph{Collections}.
-A collection is a special directory to store individuals a user prefers and wishes
-to save. After the user finishes interacting the individual
-on the Wall, he can choose to retrieve a new one from population.
-At the left side, is the \emph{Collections} section.
-The user can create several collections, to group and organize his favorite 
-artifacts. Moreover, a user can browse the content of each collection and from
-there share images through the social network.
-When a user browses over an individual a detail pane shows how many users have
-liked the individual. The pane also includes a link to the individual's 
-details, the parents, genetic operators that created it, and genealogy information.
-
-\section{IEC Graph based model}
-\label{sec:graph}
-The fundamental goal of this research is to develop a user-centered framework
-for interactive evolutionary computation (IEC) in order to increase users’
-participation and also to minimize the amount of evaluations needed for the
-evolutionary process in given Web-based IEC application.
-
-In this chapter an explanation of the proposed framework is described. The
-different techniques used, such as user modeling, fuzzy logic, and human-
-interaction.
-
-This framework is presented in figure \ref{fig:uc_framework}. Each of the
-entities of this framework will be explained in detail in following sections.
-
-The current model is tailored to the EvoSpace-I functionality and capabilities
-but it can be modified to handle other platforms, just by adding new clases 
-of nodes.
-
-\begin{figure*}
-  \captionsetup{justification=centering,margin=2cm}
-  \centering
-  \setlength\fboxsep{0pt}
-  \setlength\fboxrule{0.7pt}
-  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
-  \caption{User-Centerd Framework.}
-  \label{fig:uc_framework}       
-\end{figure*}
-
-%A hint on how
-% done -Mario
-
-\subsection{Users}
-
 In this sense the action of posting is when
@@
 \end{figure*}
-
-\subsection{Experiments}
-\label{sec:experiments}
-
-\subsection{Conclusions}
-\label{sec:conclusions}
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-\section{Conclusion}
-The conclusion goes here.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\bibliographystyle{IEEEtran}
-\bibliography{evospace-i}
-%\bibliography{languages}
-\end{document}
-
-% that's all folks
-\end{document}

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Moving section to temporal file

--- 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 
-In this sense the action of posting is when
-users want to put something on their wall who want others to know through
-their social networks. Likewise the storing action occurs when users want to
-retain permanently information that are of their interest. Finally the
-browsing action occurs when users are exploring content for their needs. All
-these actions go beyond only evaluating individuals.
-
-\begin{figure*}
-  \captionsetup{justification=centering,margin=2cm}
-  \centering
-  \setlength\fboxsep{0pt}
-  \setlength\fboxrule{0.7pt}
-  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/users.png}}
-  \caption{Users actions on interactive evolutionary systems.}
-  \label{fig:users}       
-\end{figure*}
-
-For the evaluating action is proposed  that users evaluate individuals
-indirectly, this means users can evaluate accepting indirect recommendations of
-friends that are currently active in the system. These recommendations can be
-store individuals in the system of users who know each other.
-
-Also the  browsing action is proposed within IEC systems,  this means that users
-can explore information that other active users in the system are generating.
-
-Finally activity participation is proposed. This can be divided into four
-different actions as fallows:
-
-\begin{itemize} 
-\item The Inviting action occurs when a user invites another to the system through social networks or from person to person.  
-\item The sharing action occurs when users share their individual creations with others active user in the system.
-\item The posting action occurs when published what they are doing within the system.
-\item The storing action occurs when users keep individuals in a collection, the collections concept will be discussed later in this chapter.
-\end{itemize}
-
-All the mention above is shown in figure \ref{fig:users}
-
-\subsection{Individual}
-
-Individuals are entities that form the population for the evolutionary
-algorithm. In particular for  this work individuals are animated digital
-paintings, every individual in the population is defined by a chromosome [].
-This chromosome defines the behavior of the individual, so that it consists of a
-vector of real numbers of fifteen genes, where each gen define a particular
-behavior in painting. This individuals combined each other to generated new
-individual in the population following the genetic operators such as selection,
-crossover and mutation[].
-
-\subsection{User Model} Now that both users an individuals have been defined we can
-model the users behavior have with respect to individuals in a graph-based user
-model.
-
-The reason for using a graph to model the user-individual is because it could
-can express in a simple way the behavior of this two entities having the
-necessary information  to represent the knowledge,  for example a vertex user
-and individual vertex connected through a edge that  represents the
-relationship. this relationship can be the semantic  between these two entities
-as figure \ref{fig:User-Individual} show. Each of these vertices contain properties as well as the
-edges, these properties will be explained as fallows.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_individual.png}}
-\caption{User-Individual.}
-\label{fig:User-Individual}       
-\end{figure*}
-
-In this proposed graph-based user-individual model the vertices are represented
-by set of nodes of USERS, INDIVIDUALS and COLLECTIONS. The edges are define by a
-set of relationships as LIKES, KNOWS, PARENT, HAS that represents the
-relationships between the vertices as seen in figure \ref{fig:User-Individual}.
-
-Already explained that represent USERS and INDIVIDUALS, now an explanation is
-given of what concerns for the concept of COLLECTIONS.
-
-In this proposal a collection is defined as a  container where users can store
-selected individuals laid within the collection. A collection is created when
-the user wants to store permanently individuals within the IEC system.
-
-Now for the edges in this proposal are explained as follows:
-
-\begin{itemize} 
-\item The  LIKES edge represents a preference relationship that a user has with respect to an individual.  
-\item Now the  PARENT edge represents an ancestor of the individual, in other words where the individual comes from.
-\item Likewise the  HAS edge represents that users have a collection where can contain individuals.
-\item Finally the KNOWS edge represents a relationship of friendship between users.
-\end{itemize}
-
-In figure \ref{fig:Nodes_Edges} illustrated the above mentioned, about nodes and edges.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/user_individual_collections.png}}
-\caption{Nodes And Edges Representation.}
-\label{fig:Nodes_Edges}       
-\end{figure*}
-
-In each vertex is necessary to store knowledge about the users, individuals and
-also for the collections, this is what is known as property.
-
-It is worth to mention that all vertices have a property called “element\_type”
-among others. This property is used to identify which type of vertex is, for
-example if vertex corresponds to a user then this property label the vertex as a
-“person”. Likewise if the vertex corresponds to an individual this property
-label it as an “individual” and is the same for any vertex we want to add to the
-model. This property also has the purpose of being able to do operations on the
-graph according to the vertex type.
-
-The properties for the vertex “u” are the following:
-
-\begin{itemize} 
-\item id.  
-\item created.
-\item name.
-\item element\_type.
-\end{itemize}
-
-The “id” property defines that the vertex u has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The property
-“created” defines the creation date of the vertex $u$. The property “name”
-defines the name of the vertex u. The property “element\_type” defines the
-element type that will be the vertex $u$.
-
-In figure \ref{fig:User_node} the properties of vertex $u$  are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/user_node.png}}
-\caption{User Properties.}
-\label{fig:User_node}       
-\end{figure*}
-
-The properties for the vertex $i$ are the following:
-\begin{itemize} 
-\item id. 
-\item created. 
-\item element\_type.
-\item chromosome.
-\item views.
-\end{itemize}
-
-The “id” property defines that the vertex $i$ has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The
-“chromosome” property defines the chromosome of the individual representing the
-vertex $i$. which as defined above in this section. The  property “views” defines
-the amount that users have seen this vertex $i$ The property “element\_type”
-defines the element type that will be the vertex $i$. The property “created”
-defines the creation date of the vertex $i$. % Is there a point in
-                                % describing these properties? - JJ
-
-In figure \ref{fig:Individual_node} the vertex $i$ properties are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/individual_node.png}}
-\caption{Individual Properties.}
-\label{fig:Individual_node}       
-\end{figure*}
-
-The properties for the vertex “c” are the following:
-
-\begin{itemize} 
-\item id. 
-\item element\_type.
-\item created. 
-\item name.
-\end{itemize}
-
-The “id” property defines that the vertex $c$ has a unique identifier in the set
-of vertices, which means that there will not be duplicate vertices. The property
-“element\_type” defines the element type that will be the vertex $c$. The
-property “name” defines the name of the vertex $c$. The property “created”
-defines the creation date of the vertex $c$.
-
-In figure \ref{fig:Collection_node} the vertex $c$ properties are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/collection_node.png}}
-\caption{Collection Properties.}
-\label{fig:Collection_node}       
-\end{figure*}
-
-In the same way that the vertices have properties also the edges, these
-properties are defined as follows.
-
-The “LIKES” edge has the following properties:
-
-\begin{itemize} 
-\item id. 
-\item created. 
-\item rate.
-\end{itemize}
-
-The “id” property defines that the “LIKES” edge  has a unique identifier in the
-set of edges, which means that there will not be duplicate edges. The property
-“created” defines the date of creation of the this edge. The property “rate”
-defines the rate that the users give to the individual store in this edge.
-
-In figure \ref{fig:Likes_edge} the edge “LIKES” properties and the relationships with the vertices $u$ and $i$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_likes.png}}
-\caption{LIKES edge Properties.}
-\label{fig:Likes_edge}       
-\end{figure*}
-
-The “KNOWS” edge has the following properties.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “KNOWS” has a unique identifier in the
-set of edges, which means that there will not be duplicate edges. The property
-“created” defines the date of creation of the edge.
-
-In figure \ref{fig:Knows_edge} the edge “KNOWS” properties and the relationships between vertex “u”
-are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_knows.png}}
-\caption{KNOWS edge Properties.}
-\label{fig:Knows_edge}       
-\end{figure*}
-
-The “PARENT” edge has the following properties, and represent the parents of a new individual.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “PARENT” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
-The property “created” defines the date of creation of the edge.
-
-In figure \ref{fig:Parent_edge} the edge “PARENT” properties and the relationship between vertex $i$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_parent.png}}
-\caption{PARENT edge Properties.}
-\label{fig:Parent_edge}       
-\end{figure*}
-
-The edge “HAS” has the following properties, and represent the parents of a new individual.
-\begin{itemize} 
-\item id. 
-\item created. 
-\end{itemize}
-
-The “id” property defines that the edge “HAS” has a unique identifier in the set of edges, which means that there will not be duplicate edges.
-The property “created” defines the date of creation of the edge.
-
-In figure \ref{fig:Has_edge} the edge “HAS” properties and the relationship between vertices $u$, $i$ and $c$ are shown.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
-\caption{HAS edge Properties.}
-\label{fig:Has_edge}       
-\end{figure*}
-
-\begin{equation*}\label{eq:graphRelDef} 
-\displaystyle 
-\begin{split} 
-V &= \{[u_1,u_2,u_3,...,u_n],[i_1,i_2,i_3,...,i_n],[c_1,c_2,c_3,...,c_n]\},\\ 
-E&= \{[l_1,l_2,l_3,..,l_n],[p_1,p_2,p_3,...,p_n],[h_1,h_2,h_3,...,h_n],[k_1,k_2,k_3,...,k_n]\}\\ 
-\end{split} 
-\end{equation*}
-
-Once  the vertices and edges have been defined a graph-based user-individual model is
-proposed where  user vertices are related to the individual vertices, likewise
-the vertices individuals are related to themselves when evolution exist between
-them. % Grammar has to be improved - JJ
-In that sense the vertices users and individuals are related to the
-vertices collections as a vertex user can be related to a vertex collections and
-a vertex collections collections is related to $n$ individuals vertex. In Figure
-\ref{fig:user_moder} shows an example of the graph-based user-individual model.
-
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
-\centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=8cm,height=8cm,keepaspectratio]{img/model_representation.png}}
-\caption{Graph-based user-individual model.}
-\label{fig:user_moder}       
-\end{figure*}
-
 \subsection{Experiments}

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Human centered conceptual framework

@@
 
-\section{IEC Graph based model}
+\section{IEC Human-centered framework}
 \label{sec:graph}
-The fundamental goal of this research is to develop a user-centered framework
-for interactive evolutionary computation (IEC) in order to increase users’
+The general goal of this research is to develop a human-centered [cite] framework
+for interactive evolutionary computation (IEC) in order to increase
 participation and also to minimize the amount of evaluations needed for the
-evolutionary process in given Web-based IEC application.  This framework is 
-presented in figure \ref{fig:uc_framework}. The current model is tailored 
-to the EvoSpace-I functionality and capabilities but it can be modified to 
-handle other platforms, just by adding new clases of nodes.
+evolutionary process in given IEC application. We deliberately use the more general
+term Human instead of User, because humans are not always playing the role of
+users in software engineering sense, they could just be interacting 
+naturaly with their environment and with out knowing they are implicitly
+evaluating a possible solution. In other cases the evaluation could be done
+by the natural interactions of a group of humans in a collaborative fashion
+or not. For instance the IEC algorithm could by trying to optimize 
+the design of an exit sign in an airport, then the quality of this sign could be
+determined by the average the time it takes a sample of people to 
+find the exit. In this example there could be certain collaboration
+between users if they are not traveling alone, of course this is only one of
+the considerations designers could make.
+
+%I will keep it general for this section -Mario
+%The current model is tailored 
+%to the EvoSpace-I functionality and capabilities but it can be modified to 
+%handle other platforms, just by adding new clases of nodes.
 
@@
 
-The user centered framework consists of three high level components:
+The human centered framework consists of three high level components depicted
+in figure \ref{fig:uc_framework}:
 
@@
   This is the real world system that we are going to model, it consist of
-  human users and their interaction with one or more individuals from
-  the population, this could be through a device, user interface or
-  by directly interacting with real world objects generated as part of the algorithm.
+  human users and their interactions with one or more individuals from
+  the population. There are many ways in which humans could interact 
+  with indivuals, this could be through a mobile device, a user interface 
+  or by interacting with real world objects generated as part of the algorithm.
+  There is also the posibility that fitness is or part of the search is done
+  in devices lended by the human. The information gathered through the 
+  interaction will guide the search.
+
+% We could reference Nod
 

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Framework described

@@
   \item { \tt Interactive Evolutionary Computational system.}  
-  This is the real world system that we are going to model, it consist of
-  human users and their interactions with one or more individuals from
+  This is the real world system that we are going to represent in the model, 
+  it consist of human users and their interactions with one or more individuals from
   the population. There are many ways in which humans could interact 
   with indivuals, this could be through a mobile device, a user interface 
-  or by interacting with real world objects generated as part of the algorithm.
-  There is also the posibility that fitness is or part of the search is done
-  in devices lended by the human. The information gathered through the 
-  interaction will guide the search.
+  or by interacting with real world objects generated as part of the algorithm 
+  [cite evodraw o unplugged].
+  There is also the posibility that fitness or even part of the search 
+  to be done by devices lended by the human [cite nodio].
+  The information gathered through the interaction is the primary concern
+  as it will guide the search.
 
-% We could reference Nod
-
-  \item {\tt IEC Graph}
-  This is the model proposed in this work: A graph database modeling
-  the IEC system. The  main components are the social network of users and the population. This database is the back-end of the system, is used by the EA and
-  the user interface.
+  \item {\tt IEC Data Model}
+  A data model is used to describe the IEC system prior to a physical 
+  implementation. 
+  %% This block is maybe not needed but in our ES-I implementation
+  %% we use Relational (PosgreSQL), Objects (JSON-Redis) and Graphs (Neo4J)
+  Depending on the domain several techniques can be used.
+  Parts of the system could be better described by an Entity relationship 
+  modeling to be used in a relational database or a Class Digram for a 
+  key-value store. 
+  %%
+  %%
+  A graph is proposed for modelling the social network
+  of users and their interaction with the population. This database 
+  is the back-end of the system, is used by the EA and the user interface.
 
   \item {\tt Evolutionary Algorithm} 
-  This algorithm queries the graph model, and then updates those nodes that 
-  are part of the population. 
+  The EA algorithm queries the graph model, and then updates those nodes that 
+  are part of the population. The EA could also be done with the help of
+  humans, for instance in the XY project [cite] artists physicaly created
+  the new population.  
 \end{itemize}
@@
 
-\subsection{IEC Graph Model}
+\subsection{IEC Graph Conceptual Model}
 

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Deleting comments and addressed issues

@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
 % *** Authors should verify (and, if needed, correct) their LaTeX system  ***
@@
 \section{Introduction}
-% Maybe an introduction on why and where to use these types of
-% algorithms? - jj
-% Done -Mario
 Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
@@
 
-\subsection{IEC Graph Conceptual Model}
+\section{IEC Graph Conceptual Model}
 
@@
 
-\subsection{Experiments}
+\section{Experiments}
 \label{sec:experiments}
 
-\subsection{Conclusions}
+\subsection{Conclusion}
 \label{sec:conclusions}
@@
 
-\section{Conclusion}
-The conclusion goes here.
-
 % conference papers do not normally have an appendix

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
New bib files
Making some small corrections and changes
It's more important that @mariosky addresses some concerns in the
paper.

@@
 \begin{itemize}
-  \item { \tt Individual-Individual} The relationships between individuals in the
+  \item {\bf Individual-Individual} The relationships between individuals in the
   population are naturally expressed as a tree, or a graph when both ancestors 
@@
 
-  \item {\tt User-Individual}
+  \item {\bf User-Individual}
   Many kind of relationships between users and individuals,
@@
 
-  \item {\tt User-User}
+  \item {\bf User-User}
   Relationship between users can also be modeled 
@@
 %% Moving this to Related Work Section -Mario
-
-% Maybe this claim should go to the beginning. THIS is why we want to
-% use them - JJ
-% Done - Mario
+% That section is not there yet. Forgot to paste? - JJ
 
@@
 
-\begin{figure*}
+\begin{figure*}[h!tbp]
 \captionsetup{justification=centering,margin=2cm}
@@
 \subsection{Fitness Assignment}
-\label{sec:col}
+\label{sec:assignment}
 The ES-I platform employs a collaborative technique,
@@
 
-At the top left corner list of Facebook friends is presented, 
+At the top left corner a list of Facebook friends is presented
 to encourage users to interact with the system. In the central 
-\emph{ Wall } area, an individual sampled from the population 
+\emph{ Wall } area, an individual sampled from the population that is
+being evolved via the evolutionary algorithm 
 is shown to the user.
@@
 
-\begin{figure*}
+\begin{figure*}[h!tbp]
   \captionsetup{justification=centering,margin=2cm}
@@
   \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
-  \caption{User-Centerd Framework.}
+  \caption{User-Centered Framework.} %Maybe a bit of more explanation,
+                                %showing the relationships and what
+                                %they mean.
   \label{fig:uc_framework}       
 \end{figure*}
-
+%
 The human centered framework consists of three high level components depicted
@@
 \bibliographystyle{IEEEtran}
-\bibliography{evospace-i}
+\bibliography{evospace-i,volunteer}
 %\bibliography{languages}

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
changes

--- 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 \begin{itemize}
-  \item { \tt Individual-Individual} The relationships between individuals in the
+  \item Individual-Individual. The relationships between individuals in the
   population are naturally expressed as a tree, or a graph when both ancestors 
@@
 
-  \item {\tt User-Individual}
+  \item User-Individual.
   Many kind of relationships between users and individuals,
@@
 
-  \item {\tt User-User}
+  \item User-User.
   Relationship between users can also be modeled 
@@
 
-\section{EvoSpace-Interactive}
+\subsection{EvoSpace-Interactive}
 \label{sec:evospace-i} 
@@
 
-\subsection{Fitness Assignment}
+\subsubsection{Fitness Assignment}
 \label{sec:col}
@@
 
-\subsection{Collaboration}
+\subsubsection{Collaboration}
 \label{sec:col}
@@
 
-\begin{figure*}
-  \captionsetup{justification=centering,margin=2cm}
+\begin{figure}[ht]
   \centering
-  \setlength\fboxsep{0pt}
-  \setlength\fboxrule{0.7pt}
-  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
-  \caption{User-Centerd Framework.}
+  \includegraphics[width=\textwidth,height=8cm,keepaspectratio]{img/framework.png}
+
+  \caption{Human-Centerd Framework.}
   \label{fig:uc_framework}       
-\end{figure*}
+\end{figure}
 
@@
 \begin{itemize}
-  \item { \tt Interactive Evolutionary Computational system.}  
+  \item Interactive Evolutionary Computational system. 
   This is the real world system that we are going to represent in the model, 
@@
 
-  \item {\tt IEC Data Model}
+  \item IEC Data Model.
   A data model is used to describe the IEC system prior to a physical 
@@
 
-  \item {\tt Evolutionary Algorithm} 
+  \item Evolutionary Algorithm. 
   The EA algorithm queries the graph model, and then updates those nodes that 
@@
 
-\subsection{IEC Graph Conceptual Model}
+\section{IEC Graph Conceptual Model}
 
-In this section the Graph Model is presented,
+A graph is composed of two fundamental entities: nodes and relationships,
+both having attributes. Nodes are abstractions of the objects in the real 
+world system, and relationships describe their interactions. Types describe
+the attributes describing an entity, First we describe the types of nodes and  relationships in the system.
 
-\subsection{Experiments}
+\subsection{Node Types}
+\begin{itemize}
+
+\item Users. This nodes represent the human users of the system and have 
+the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
+specifing the  creation date (time stamp) of the node.
+
+\item Individual. This type of node has the following attributes: 
+{\tt id}, {\tt chromosome}  and {\tt views} this is 
+the amount that users have seen or interacted with this individual and
+ {\tt created}.
+\item Collection. This type of nodes represent the collections belonging
+to users, their attributes are {\tt id}, 
+{\tt name}, {\tt created} and {\tt is_private} indicating if this collection
+contents are shared with others. 
+\end{itemize}
+
+\subsection{Relationships Type}
+%%
+%% Explain relations
+
+The {\tt likes} relation describes the interaction between a User and
+an Individual in which a rating value is assigned. In the current application
+a user is allowed to rate an Individual multiple times, so the date of the 
+interactions is also of interest.
+
+\begin{itemize} 
+\item id. 
+\item created. 
+\item rate.
+\end{itemize}
+
+The relation {\tt knows} connects two users that know each other. This relation
+is "unique" meaning that it can not be repeted between a pair of nodes.
+
+The relation {\tt parent} describes what Individual is the parent of a new
+Individual and is also unique.
+
+The relation {\tt has} describes an ownership relation between users and
+those collections they own. Also a collection {\tt has}  individuals stored
+in them.
+
+In figure \ref{fig:Has_edge} the edge {\tt has} attributes and the relationship between certain nodes $user$, $individual$ and $collection$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
+\caption{HAS edge Properties.}
+\label{fig:Has_edge}       
+\end{figure*}
+
+\section{Case Study}
 \label{sec:experiments}
 
-\subsection{Conclusions}
+\section{Conclusion}
 \label{sec:conclusions}

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Changes for the conflict

--- 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 \begin{itemize}
+
   \item {\bf Individual-Individual} The relationships between individuals in the
@@
   \item {\bf User-User}
+
   Relationship between users can also be modeled 
@@
 knowledge about the process and the population, for instance showing the
-most popular individual, or the the user with more participation \cite{picbreeder}.
+most popular individual, or the the user with more participation 
+\cite{picbreeder}.
 In order to do this, data about individuals %what? Maybe data
@@
 participation and also to minimize the amount of evaluations needed for the
-evolutionary process in given IEC application. We deliberately use the more general
+evolutionary process in given IEC application. We deliberately use the 
+more general
 term Human instead of User, because humans are not always playing the role of
@@
 
-\begin{figure*}[h!tbp]
-  \captionsetup{justification=centering,margin=2cm}
+\begin{figure}[ht]
   \centering
-  \setlength\fboxsep{0pt}
-  \setlength\fboxrule{0.7pt}
-  \fbox{\includegraphics[width=10cm,height=10cm,keepaspectratio]{img/framework.png}}
-  \caption{User-Centered Framework.} %Maybe a bit of more explanation,
-                                %showing the relationships and what
-                                %they mean.
+  \includegraphics[width=\textwidth,height=8cm,keepaspectratio]{img/framework.png}
+
+  \caption{Human-Centerd Framework.}
   \label{fig:uc_framework}       
-\end{figure*}
-%
+\end{figure}
+
 The human centered framework consists of three high level components depicted
@@
 \begin{itemize}
-  \item { \tt Interactive Evolutionary Computational system.}  
+  \item Interactive Evolutionary Computational system. 
   This is the real world system that we are going to represent in the model, 
-  it consist of human users and their interactions with one or more individuals from
+  it consist of human users and their interactions with one or more individuals
+  from
   the population. There are many ways in which humans could interact 
@@
 
-In this section the Graph Model is presented,
+A graph is composed of two fundamental entities: nodes and relationships,
+both having attributes. Nodes are abstractions of the objects in the real 
+world system, and relationships describe their interactions. Types describe
+the attributes describing an entity, First we describe the types of nodes and  relationships in the system.
 
-\section{Experiments}
+\subsection{Node Types}
+\begin{itemize}
+
+\item Users. This nodes represent the human users of the system and have 
+the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
+specifing the  creation date (time stamp) of the node.
+
+\item Individual. This type of node has the following attributes: 
+{\tt id}, {\tt chromosome}  and {\tt views} this is 
+the amount that users have seen or interacted with this individual and
+ {\tt created}.
+\item Collection. This type of nodes represent the collections belonging
+to users, their attributes are {\tt id}, 
+{\tt name}, {\tt created} and {\tt is_private} indicating if this collection
+contents are shared with others. 
+\end{itemize}
+
+\subsection{Relationships Type}
+%%
+%% Explain relations
+
+The {\tt likes} relation describes the interaction between a User and
+an Individual in which a rating value is assigned. In the current application
+a user is allowed to rate an Individual multiple times, so the date of the 
+interactions is also of interest.
+
+\begin{itemize} 
+\item id. 
+\item created. 
+\item rate.
+\end{itemize}
+
+The relation {\tt knows} connects two users that know each other. This relation
+is "unique" meaning that it can not be repeted between a pair of nodes.
+
+The relation {\tt parent} describes what Individual is the parent of a new
+Individual and is also unique.
+
+The relation {\tt has} describes an ownership relation between users and
+those collections they own. Also a collection {\tt has}  individuals stored
+in them.
+
+In figure \ref{fig:Has_edge} the edge {\tt has} attributes and the relationship 
+between certain nodes $user$, $individual$ and $collection$ are shown.
+
+\begin{figure*}
+\captionsetup{justification=centering,margin=2cm}
+\centering
+\setlength\fboxsep{0pt}
+\setlength\fboxrule{0.7pt}
+
+\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
+\caption{HAS edge Properties.}
+\label{fig:Has_edge}       
+\end{figure*}
+
+\section{Case Study}
 \label{sec:experiments}
 
-\subsection{Conclusion}
+\section{Conclusion}
 \label{sec:conclusions}

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Smaller figures

@@
 found in the system. A case study is presented as a proof of concept, 
-providing both conceptual and implementation details of a graph based model as it is 
+providing both conceptual and implementation details of a graph based model 
+as it is 
 applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
@@
   \item {\bf User-User}
-
   Relationship between users can also be modeled 
@@
   A graph model could eneable researchers to find other ways of identifyng leaders of 
-  opinion or measuring the similarity between user\'s preferences. 
+  opinion or measuring the similarity between user's preferences. 
   These measures can then be used by recommender algorithms selecting 
-  individuals according users\' preferences.
+  individuals according users' preferences.
 
@@
 
-\begin{figure*}[h!tbp]
-\captionsetup{justification=centering,margin=2cm}
+\begin{figure}[!t]
     \centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
+        \includegraphics[width=3.5in]{img/UI_ed01.png}
     \caption{User interface of an EvoSpace-Interactive application.}
     \label{fig:web}
-\end{figure*}
+\end{figure}
 
-\subsection{EvoSpace-Interactive}
+\section{EvoSpace-Interactive}
 \label{sec:evospace-i} 
@@
 
-\subsubsection{Collaboration}
+\subsection{Collaboration}
 \label{sec:col}
@@
 from their friend\'s collections by using the web interface depicted 
-in Figure \ref{fig:web}
-
+in Figure \ref{fig:web}.
 At the top left corner a list of Facebook friends is presented
@@
 details, the parents, genetic operators that created it, and genealogy information.
+
+\begin{figure*}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/framework.png}
+    \caption{IEC Human-centered framework.}
+    \label{fig:web}
+\end{figure*}
 
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\textwidth,height=8cm,keepaspectratio]{img/framework.png}
-
-  \caption{Human-Centerd Framework.}
-  \label{fig:uc_framework}       
-\end{figure}
-
 The human centered framework consists of three high level components depicted
@@
 \begin{itemize}
-  \item Interactive Evolutionary Computational system. 
+  \item {\bf Interactive Evolutionary Computational system} 
   This is the real world system that we are going to represent in the model, 
@@
 
-  \item IEC Data Model.
+  \item {\bf IEC Data Model}
   A data model is used to describe the IEC system prior to a physical 
@@
 
-  \item Evolutionary Algorithm. 
+  \item {\bf Evolutionary Algorithm} 
   The EA algorithm queries the graph model, and then updates those nodes that 
@@
 
-\subsection{Relationships Type}
+\subsection{Relationship Types}
 %%
@@
 is "unique" meaning that it can not be repeted between a pair of nodes.  
-
 The relation {\tt parent} describes what Individual is the parent of a new
 Individual and is also unique.  
-
 The relation {\tt has} describes an ownership relation between users and
@@
 
-In figure \ref{fig:Has_edge} the edge {\tt has} attributes and the relationship 
-between certain nodes $user$, $individual$ and $collection$ are shown.
+In figure \ref{fig:Has_edge} the {\tt has} relationship 
+between certain nodes $user$, $individual$ and $collection$ is shown.
 
-\begin{figure*}
-\captionsetup{justification=centering,margin=2cm}
+\begin{figure}[!t]
     \centering
-\setlength\fboxsep{0pt}
-\setlength\fboxrule{0.7pt}
-
-\fbox{\includegraphics[width=5cm,height=5cm,keepaspectratio]{img/edge_properties_has.png}}
-\caption{HAS edge Properties.}
-\label{fig:Has_edge}       
-\end{figure*}
+        \includegraphics[width=2.5in]{img/edge_properties_has.png}
+    \caption{ The {\tt has} relationship type example and attributes.}
+    \label{fig:evo}
+\end{figure}
 

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Basic description of case study
Gammification Techniques explained, to do: Set-up and Results.

@@
 
-\section{IEC Graph Conceptual Model}
+\section{IEC Graph Data Model}
 
@@
 
-\item Users. This nodes represent the human users of the system and have 
+\item {\bf Users} This nodes represent the human users of the system and have 
 the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
@@
 
-\item Individual. This type of node has the following attributes: 
+\item {\bf Individual} This type of node has the following attributes: 
 {\tt id}, {\tt chromosome}  and {\tt views} this is 
@@
  {\tt created}.
-\item Collection. This type of nodes represent the collections belonging
+
+\item {\bf Collection} This type of nodes represent the collections belonging
 to users, their attributes are {\tt id}, 
-{\tt name}, {\tt created} and {\tt is_private} indicating if this collection
-contents are shared with others. 
+{\tt name}, {\tt created} and {\tt is\_private} indicating if this collection
+contents can be are shared with others. 
 \end{itemize}
@@
 %% Explain relations
+\begin{itemize}
+\item {\bf Likes} This relation describes the interaction between a user and
+an individual in which a rating value is assigned this is represented by 
+the {\tt rate} attribute. In the current application a user is allowed to rate
+an individual multiple times, so the date of the interactions is also of
+interest.
 
-The {\tt likes} relation describes the interaction between a User and
-an Individual in which a rating value is assigned. In the current application
-a user is allowed to rate an Individual multiple times, so the date of the 
-interactions is also of interest.
+\item {\bf Knows} The relation connects two users that know each
+other. This relation is "unique" meaning that it can not be repeted between a pair of nodes.
 
-\begin{itemize} 
-\item id. 
-\item created. 
-\item rate.
+\item {\bf Parent} Describes what individual is the parent of a new individual and is also unique.
+
+\item {\bf Has} The relation describes an ownership relation between users and
+those collections they own. Also a collection {\tt has}  individuals stored
+in them.  In figure \ref{fig:has} the {\tt has} relationship 
+between certain nodes $user$, $individual$ and $collection$ is shown.
+
+\item {\bf Shared} When a user takes an individual from a friends collection, this interactions is described by a {\tt share} relationship, between the user and the individual.
+
+\item {\bf Saved} When a user takes an individual from the wall interface, this 
+interactions is described by a {\tt save} relationship. 
 \end{itemize}
-
-The relation {\tt knows} connects two users that know each other. This relation
-is "unique" meaning that it can not be repeted between a pair of nodes.  
-The relation {\tt parent} describes what Individual is the parent of a new
-Individual and is also unique.  
-The relation {\tt has} describes an ownership relation between users and
-those collections they own. Also a collection {\tt has}  individuals stored
-in them.
-
-In figure \ref{fig:Has_edge} the {\tt has} relationship 
-between certain nodes $user$, $individual$ and $collection$ is shown.
 
@@
     \caption{ The {\tt has} relationship type example and attributes.}
-    \label{fig:evo}
+    \label{fig:has}
 \end{figure}
+
+Is important to notice that this model is still partial view of the IEC system,
+and other models can could be used in practice when implementing the system. 
+For instance a relational database can be used for the implementation of a
+user profile and authentication module. Also a copy or cache of the population 
+could be kept in a key-value store in memmory for performance reasons. 
+Nevertheless, the model can give important insights of the behaviour of the 
+system.
 
@@
 \label{sec:experiments}
+
+We present in this section the results of a comparassion between two versions
+of the same IEC application. The goal of the experiment is to compare two 
+gamification techiniques to increase volunteer participation. 
+The main difference between both versions is how the fitness of individuals
+is calculated and the information presented to users. 
+\subsection{Gamification technique}
+\label{sec:gamification}
+The gammification technique
+consists in giving more importance to the preference of users with more
+participations. A feature found in video games is the concept of a 
+score and experience levels. Levels depend on the score and certain features
+of the game are only available to gamers when they reach a giving level.
+
+The score achieved by users depends on the actions he does in the system as
+each time a user does one of these actions their score is incremented by one.
+The actions considered in this application are: 
+\begin{itemize}
+\item Start a session.
+\item Rate an individual.
+\item Create a collection.
+\item Save an individual of the wall to a collection.
+\item Save an individual from a friend's collection.
+\item Explore collections of other friends.
+\end{itemize}
+Only the "Start a session" and "Explore collections of other friends". 
+Are not found in the graph model, 
+%% Note: I must ask Christian where they are stored 
+%%
+
+The weight of the preference of a users can depend on two scores:
+\begin{itemize}
+\item Experience. This value depends on the score and is a value 
+between 0 and 100.
+
+\item Participation. This value is simply the degree of the user node 
+in the graph. This measure also considers the number of users he knows,
+perhaps some of the users where invited to the system by him.   
+\end{itemize}
+
+The comparation consisted on three versions:
+\begin{itemize}
+\item Non Gammification. All users have the same weight.
+\item Non Graph Gammification. Only experience is considered
+\item Graph Gammification. Both experience and participation where considered.
+\end{itemize}
+When gammification was employed, all score values known where presented to users
+and a ranking of users by weight was shown in a window, a screen capture of
+this interface is shown in figure \ref{fig:top-users}.
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/user_known_2.png}
+    \caption{Sample of users, showing the {\tt knows} relationship.}
+    \label{fig:top-users}
+\end{figure}
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/leaderBoard.png}
+    \caption{User interface showing top users.}
+    \label{fig:top-users}
+\end{figure}
+
+\subsection{Set-up}
+\label{sec:setup}
+
+\subsection{Results}
+\label{sec:results}
+
+\begin{figure*}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/comparison.png}
+    \caption{Top ranked number individuals rated by user using a
+     logarithmic scale.}
+    \label{fig:top-ranked-participation}
+\end{figure*}
 

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
correcting duplicate labels
Abstract updated

@@
 a large number of individuals. 
-To tackle this issues, we propose the design of IEC systems using a user-centered 
+To tackle this issues, we propose the design of IEC systems using a human-centered 
 approach, with a framework consisting of a social network of 
-volunteers, interacting with a population consisting again of a network 
-of individuals. In this work we propose a graph model as a viable 
+volunteers, interacting with a population consisting also of a network 
+of individuals. In this work we propose a graph model as a practical 
 and efficient tool for mapping in a natural way the relationships 
@@
 providing both conceptual and implementation details of a graph based model 
-as it is 
-applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
-As a result, the graph model was then succesfully applied to measure the degree
-of relationship between users with the porpuse of quantifing 
-their participation.
+as it is  applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
+As a result, the graph model was successfully applied in a gamification technique 
+developed to increase user engagement.
 
@@
     \caption{IEC Human-centered framework.}
-    \label{fig:web}
+    \label{fig:uc_framework}
 \end{figure*}
@@
     \caption{Sample of users, showing the {\tt knows} relationship.}
-    \label{fig:top-users}
+    \label{fig:users-graph}
 \end{figure}

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Spell check

@@
 system.  Thus, the main loop of the EA needs the intervention of a
-human for the quality assessment of each proposed solution.
+human to perform the quality assessment of each proposed solution.
 IEAs have demonstrated their ability for effectively
@@
 engagement, and to tackle the issues mentioned above,  
-IEAs must follow a user centered design, giving extensive attention 
+IEAs must follow a human centered design, giving extensive attention 
 to volunteer users, not only because their
@@
 that a graph model is a viable representation for this purpose,  
-presenting a natural way of maping the relationships found in the system. 
+presenting a natural way of mapping the relationships found in the system. 
 Most IEAs have the following many-to-many relations:
@@
   Relationship between users can also be modeled 
-  as a social network, with well stablished semantics, algorithms and metrics (cite).
-  A graph model could eneable researchers to find other ways of identifyng leaders of 
+  as a social network, with well established semantics, algorithms and metrics (cite).
+  A graph model could enable researchers to find other ways of identifying leaders of 
   opinion or measuring the similarity between user's preferences. 
@@
 
-These relations are naturaly maped to graphs, while other actions and 
+These relations are naturally mapped to graphs, while other actions and 
 properties can also be modeled as a semantic network using
@@
 conceptual and implementation details of a graph data model applied in the 
-implementation of an IEA. The experiment was implmented using
+implementation of an IEA. The experiment was implemented using
 the EvoSpace-Interactive framework, and consisted of a web application
@@
 \label{sec:evospace-i} 
-As a case study, a IEA was implented using the 
+As a case study, a IEA was developed using the 
 EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
-A brief description of the aplication is presented next, focusing
+A brief description of the application is presented next, focusing
 on the data elements that where ported to the graph model.
@@
 users in software engineering sense, they could just be interacting 
-naturaly with their environment and with out knowing they are implicitly
+naturally with their environment and with out knowing they are implicitly
 evaluating a possible solution. In other cases the evaluation could be done
@@
   the population. There are many ways in which humans could interact 
-  with indivuals, this could be through a mobile device, a user interface 
+  with individuals, this could be through a mobile device, a user interface 
   or by interacting with real world objects generated as part of the algorithm 
   [cite evodraw o unplugged].
-  There is also the posibility that fitness or even part of the search 
-  to be done by devices lended by the human [cite nodio].
+  There is also the possibility that fitness or even part of the search 
+  to be done by devices lent by the human [cite nodio].
   The information gathered through the interaction is the primary concern
@@
   %%
-  A graph is proposed for modelling the social network
+  A graph is proposed for modeling the social network
   of users and their interaction with the population. This database 
@@
   are part of the population. The EA could also be done with the help of
-  humans, for instance in the XY project [cite] artists physicaly created
+  humans, for instance in the XY project [cite] artists physically created
   the new population.  
@@
 the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
-specifing the  creation date (time stamp) of the node.
+specifying the  creation date (time stamp) of the node.
 
@@
 \item {\bf Knows} The relation connects two users that know each
-other. This relation is "unique" meaning that it can not be repeted between a pair of nodes.
+other. This relation is "unique" meaning that it can not be repeated between a pair of nodes.
 
@@
 user profile and authentication module. Also a copy or cache of the population 
-could be kept in a key-value store in memmory for performance reasons. 
-Nevertheless, the model can give important insights of the behaviour of the 
+could be kept in a key-value store in memory for performance reasons. 
+Nevertheless, the model can give important insights of the behavior of the 
 system.
@@
 
-We present in this section the results of a comparassion between two versions
+We present in this section the results of a comparison between two versions
 of the same IEC application. The goal of the experiment is to compare two 
-gamification techiniques to increase volunteer participation. 
+gamification techniques to increase volunteer participation. 
 The main difference between both versions is how the fitness of individuals
@@
 \label{sec:gamification}
-The gammification technique
+The gamification technique
 consists in giving more importance to the preference of users with more
@@
 \begin{itemize}
-\item Non Gammification. All users have the same weight.
-\item Non Graph Gammification. Only experience is considered
-\item Graph Gammification. Both experience and participation where considered.
+\item Non Gamification. All users have the same weight.
+\item Non Graph Gamification. Only experience is considered
+\item Graph Gamification. Both experience and participation where considered.
 \end{itemize}
-When gammification was employed, all score values known where presented to users
+When gamification was employed, all score values known where presented to users
 and a ranking of users by weight was shown in a window, a screen capture of

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Spell check
Added references

@@
   can be expressed using common actions found in social networks;
-  for instance: share, like, rate or add to collection [cite Activity Streams 2.0
-  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
+  for instance: share, like, rate or add to collection \cite{Prodromou:16:AS}. In IEAs these
   relationships are used to assign a single or a collection of
-  fitness values to individuals [cite ES-I].
+  fitness values to individuals \cite{garcia2013evospace}.
 
@@
   Relationship between users can also be modeled 
-  as a social network, with well established semantics, algorithms and metrics (cite).
+  as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
   A graph model could enable researchers to find other ways of identifying leaders of 
@@
 properties can also be modeled as a semantic network using
-basic subject–-predicate-–object expressions [cite Semantic triples].
+basic subject–-predicate-–object expressions \cite{Prud'hommeaux:14:RT}.
 Another reason for using a graph model is the tools available for
 implementing and query the model, and the increased performance of
-native graph database systems [cite Christian].
+native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.
 
@@
 \label{sec:graph}
-The general goal of this research is to develop a human-centered [cite] framework
+The general goal of this research is to develop a human-centered \cite{gasson2003human} framework
 for interactive evolutionary computation (IEC) in order to increase
@@
   or by interacting with real world objects generated as part of the algorithm 
-  [cite evodraw o unplugged].
+  \cite{de2014artists,de2013unplugging}. 
   There is also the possibility that fitness or even part of the search 
-  to be done by devices lent by the human [cite nodio].
+  to be done by devices lent by the human \cite{merelo2016performance}.
   The information gathered through the interaction is the primary concern
@@
   are part of the population. The EA could also be done with the help of
-  humans, for instance in the XY project [cite] artists physically created
+  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
   the new population.  
@@
 %%
-
-The weight of the preference of a users can depend on two scores:
+Two scores are used to determine the weight of a user's preference:
 \begin{itemize}
@@
 
-The comparation consisted on three versions:
+Three versions were compared:
 \begin{itemize}

### 2017-01-20 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Old comments removed.
Related Work Planned

@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
 \title{A graph-based data model for interactive 
@@
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
 \author{\IEEEauthorblockN{Michael Shell}
@@
 
-% make the title area
 \maketitle
@@
 
-% what is this issue?
-%Maybe the abstract should start with the issues we want to tackle as
-%motivation. Motivation always goes first - JJ
 %and novelty. % These two last issues are not too clear, maybe they
              % could be explained in that motivation.
-%The proposed graph database could be used to model IEC systems that 
-%follow a user-centered approach. % Graph is the model, graph database
-                                % the implementation. Two different
-                                % things.
 
-% It must prove that the issues have been addressed - JJ
 \end{abstract}
@@
 % no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 
+% IEC 
+% Collaborative
+% Users - Volunteers - Engagement
+% Too much data
+% Relational MySQL, Key-Value, Mongo, Redis, Memcached
+% Not well suited for certain relations.
+% Graph DBs, Performance, Query Langs 
+% Are they any good? 
+% Titan, Neo4J
+
+\section{EvoSpace-Interactive}
+\label{sec:evospace-i}
 \begin{figure}[!t]
@@
 
-\section{EvoSpace-Interactive}
-\label{sec:evospace-i} 
 As a case study, a IEA was developed using the 
@@
 \cite{picbreeder}.
-In order to do this, data about individuals %what? Maybe data
-                                %about individuals? - JJ
+In order to do this, data about individuals 
 must be permanently stored, even
@@
 
-%I will keep it general for this section -Mario
-%The current model is tailored 
-%to the EvoSpace-I functionality and capabilities but it can be modified to 
-%handle other platforms, just by adding new clases of nodes.
-
 The human centered framework consists of three high level components depicted
@@
 
-%A hint on how
-% done -Mario
-
 \section{IEC Graph Data Model}
@@
 world system, and relationships describe their interactions. Types describe
-the attributes describing an entity, First we describe the types of nodes and  relationships in the system.
+the attributes describing an entity, First we describe the types of nodes and
+relationships in the system.
 
@@
 \subsection{Relationship Types}
-%%
-%% Explain relations
+
 \begin{itemize}
@@
 Are not found in the graph model, 
-%% Note: I must ask Christian where they are stored 
+%% Note: I must ask Christian where they're stored 
 %%
@@
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
 \section*{Acknowledgment}
@@
 \bibliography{evospace-i,volunteer}
-%\bibliography{languages}
+
 \end{document}
 
-% that's all folks
 \end{document}

### 2017-01-21 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Title ideas?

@@
 \begin{document}
-\title{A graph-based data model for interactive 
-evolutionary computation}
+% Title ideas here:
+
+\title{Exploiting the Graph: Increasing Engagement in Collaborative Interactive Evolution}
 
@@
 % Titan, Neo4J
+
+There is information regarding users and their collaborations, the interaction and subjective evaluation of individuals of the population.
+There is also information about the actual evolutionary process: relationships between individuals and evolutionary operators, used.
 
@@
 \label{sec:setup}
+% Anuncements
+% Table of GA Parameters
+% Time
 

### 2017-01-22 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
biblio from christian added
Table added

@@
 Section \ref{sec:interactive} presents related work on the topic 
-of Interactive Evolution data models.
+of Interactive Evolution with emphasis in data models.
 Then, Section \ref{sec:evospace-i} presents the computational platform on which 
@@
 \begin{itemize}
-\item Non Gamification. All users have the same weight.
-\item Non Graph Gamification. Only experience is considered
-\item Graph Gamification. Both experience and participation where considered.
+\item Base (B). All users have the same weight.
+\item Non Graph Gamification (G). Only experience is considered
+\item Graph Gamification (GG) . Both experience and participation are considered.
 \end{itemize}
@@
 \label{sec:setup}
+
+The three versions of the EvoDrawing web application were deployed to the
+Heroku cloud service, using the same deployment configuration, using the
+Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB plug-in for the GG experiment.
+Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.
+
+\begin{table}
+  \small
+  \caption{ Parameters for experiments.  }
+  \label{tab:params} 
+  \centering
+  \small
+  \begin{tabular}{l  c   }
+    \hline\noalign{\smallskip}
+     Parameter & Value \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    Initial Population Size   & 80 \\ \hline
+    Sample Size & 1 \\ \hline
+    Step Size & 8 Samples \\ \hline
+    Mutation &  \\ \hline
+    Selection & Tournament \\ \hline
+    Tournament Size &  6 \\ \hline
+  \end{tabular}
+\end{table}
+
 % Anuncements
@@
 \bibliographystyle{IEEEtran}
-\bibliography{evospace-i,volunteer}
+\bibliography{evospace-i,volunteer,biblio}
 

### 2017-01-23 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More tomorrow
Related Work

@@
 \label{sec:interactive}
-
 % IEC 
@@
 
-There is information regarding users and their collaborations, the interaction and subjective evaluation of individuals of the population.
-There is also information about the actual evolutionary process: relationships between individuals and evolutionary operators, used.
+The IEC the approach has been used in a wide variety of applications and problem domains
+and remains an active area of research. However, in this work, emphasis is given to IEAs 
+implemented as volunteer based collaborative systems; where volunteers interact and evaluate 
+an evolving population, thus guiding the search based on an aggregate their subjective 
+preferences. What follows is not intended as a comprehensive survey, only a review of 
+the most relevant contributions to the present work, with special attention put in
+data structures and data management requirements.
+
+From the earlier works in the IEC field the notion of user collaboration in a natural way
+has been proposed. A notable application was the Galapagos Project \cite{sims1997interactivity},
+an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
+twelve displays by selecting those they find most aesthetically interesting by standing on
+step sensors in front of each. An early example of web based collaboration is the work by 
+Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a distributed 
+EA using a global population stored on a central web server sending 
+portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
+and returned to the server those they liked more, to be distributed over the web.
+Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} use web-based IEAs 
+to evolve artistic artifacts using a generative encoding, compositional pattern producing networks.
+
+In all these systems, the evolutionary interactive process is captured by the system. The genetic lineage of each
+individual is presented where users can visualize the relationships between them and evolutionary operators used in
+their creation. Also there is information regarding users and their collaborations.  Common offerings show 
+interesting individuals order by categories, number of votes, best of the week among others.
+
+These features have the goal of encouraging participation but require the use of a database system 
+for their implementation. Picbreeder \cite{picbreeder} is implemented with a service oriented
+architecture using a relational database and the file system for genotypes and images.
 
@@
 \end{table}
+% Anuncements
 
-% Anuncements
-% Table of GA Parameters
+http://goo.gl/jLis4Q
+
 % Time
-
 \subsection{Results}

### 2017-01-23 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More related work

@@
 interesting individuals order by categories, number of votes, best of the week among others.
-
 These features have the goal of encouraging participation but require the use of a database system 
@@
 architecture using a relational database and the file system for genotypes and images.
+
+Collaborative IEC systems need to store highly connected data, as it is common in current applications
+like social networks, and the world wide web. In order to deal with large datasets of connected data,
+graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
+which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
+
+The present work builds on a previous proposals and extends the C-IEA approach.
+First, it proposes a graph data model for representing the interactive process of C-IEC systems.
+Second, it incorporates the graph model to the evolutionary algorithm.
+Finally, uses the same model to increase the collaboration of volunteers.
 

### 2017-01-23 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
URL table added

@@
 
+%%
+%% TO DO: Talk about the Model (in the MVC sense) of this system: PostgreSQL, Redis, 
+%%
+
  \section{IEC Human-centered framework}
@@
 Only the "Start a session" and "Explore collections of other friends". 
-Are not found in the graph model, 
-%% Note: I must ask Christian where they're stored 
-%%
+Are not found in the graph model, these are stored in the PostgreSQL relational
+database system.
+
 Two scores are used to determine the weight of a user's preference:
@@
 \end{table}
+
 % Anuncements
+In order to start each one of the experiments was promoted through social networks particularly on the Facebook
+and Twitter platforms. The link used in the call for participation  was shortened URL, in order to obtain additional
+analytic data of the visitors.  In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
+along with the short URL and the analytics link. In the same table a link to the GitHub application repository
+used to deploy to Heroku is also listed.
 
-http://goo.gl/jLis4Q
+\begin{table*}
+  \small
+  \caption{ URLs used for each deployment.}
+  \label{tab:urls} 
+  \centering
+  \small
+  \begin{tabular}{l l  l l l}
+    \hline\noalign{\smallskip}
+     Deployment & Heroku URL & Shortened URL & Analytics & Github Repository \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    B   & evodrawings01.herokuapp.com & goo.gl/jLis4Q & goo.gl/jLis4Q.info &  \\ \hline
+    G   & evodrawings02.herokuapp.com & goo.gl/jqjNy5 & goo.gl/jqjNy5.info &  \\ \hline
+    GG  & evodrawings03.herokuapp.com & goo.gl/J8TCe1 & goo.gl/J8TCe1.info &  \\ \hline
+    \end{tabular}
+\end{table*}
+
+Only data for the first week of deployment was considered for the experiments, and they where conducted 
+between January and May of 2016. Each deployment was tested first with a few beta testers, before launch
+in order to test design decisions.
 
@@
 \section*{Acknowledgment}
-
-The authors would like to thank...
+Space reserved for Acknowledgment text, Lorem ipsum dolor sit amet, consectetur adipiscing elit,
+sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis 
+nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
+in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
 

### 2017-01-23 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More

@@
 
-We present in this section the results of a comparison between two versions
+We present in this section the results of a comparison between three versions
 of the same IEC application. The goal of the experiment is to compare two 
-gamification techniques to increase volunteer participation. 
-The main difference between both versions is how the fitness of individuals
+gamification techniques to increase volunteer participation against a third 
+without any thechnique applied.
+
+The main difference between both versions employing gamification is how the fitness of individuals
 is calculated and the information presented to users.
-\subsection{Gamification technique}
+
+\subsection{Gamification techniques}
 \label{sec:gamification}
-The gamification technique
-consists in giving more importance to the preference of users with more
-participations. A feature found in video games is the concept of a 
-score and experience levels. Levels depend on the score and certain features
-of the game are only available to gamers when they reach a giving level.
+
+Gamification is defined by Deterding et al. as
+``the use of game design elements in non-game contexts'' \cite{deterding2011game}.
+The gamification element employed in this work is rewarding mechanism 
+\cite{dubois2013understanding}, in general they consisting of a reputation systems with score points, 
+levels and leader boards. Points are awarded to users in response of
+the accomplishment of certain activities that need to be encouraged. Levels depend
+on the score and certain features of the game are only available to gamers when 
+they reach a giving level.
+
+The rewarding mechanism as it is applied in these versions 
+consists in giving more importance to the preference of those users having a higher reputation
+as given by the score points and experience levels.
 
 The score achieved by users depends on the actions he does in the system as
-each time a user does one of these actions their score is incremented by one.
-The actions considered in this application are: 
+each time a user does one of these actions their score is incremented by one:
 \begin{itemize}
@@
 \end{itemize}
+
 Only the "Start a session" and "Explore collections of other friends". 
@@
 between 0 and 100. 
+% More details needed? Is done by a Fuzzy Inference System but I think
+% the description would take to much space.
 
@@
 The three versions of the EvoDrawing web application were deployed to the
-Heroku cloud service, using the same deployment configuration, using the
+Heroku cloud service with the same configuration: using the
 Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB plug-in for the GG experiment.
-Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.
+Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
+system used in the implementation was Neo4J, and it was allready available as a plug-in in Heroku.
 
@@
 % Anuncements
-In order to start each one of the experiments was promoted through social networks particularly on the Facebook
-and Twitter platforms. The link used in the call for participation  was shortened URL, in order to obtain additional
-analytic data of the visitors.  In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
-along with the short URL and the analytics link. In the same table a link to the GitHub application repository
-used to deploy to Heroku is also listed.
+In order to start each of the experiments was promoted through social networks
+particularly on the Facebook and Twitter platforms. The link used in the call for participation
+was shortened URL, in order to obtain additional analytic data of the visitors.
+In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
+along with the short URL and the analytics link. In the same table a link to the GitHub 
+application repository used to deploy to Heroku is also listed.
 
@@
 Only data for the first week of deployment was considered for the experiments, and they where conducted 
-between January and May of 2016. Each deployment was tested first with a few beta testers, before launch
-in order to test design decisions.
+between January and May of 2016.
 
@@
 \label{sec:results}
+Before launch, each deployment was first tried with a few beta testers. When applying the gammification
+technique for the first time a problem was found in this beta stage. Some users were chating by 
+giving a rating to an animation even before it was returned from the server, this was done by just
+constantly pressing the mouse button. This is a common problem found in systmes appling leader boards.
+By making the rewards for tasks achievements visible to other players or providing leader boards are ways of 
+encouraging players to compete \cite{hickman2010total}. Because the  problematic consequences of competition, 
+which can result in negative conduct, low cooperation and collaboration, or disadvantaging certain player
+demographics such as women \cite{kumar2013gamification}, best-practice gamification designs try to
+refrain from using this element. In order to measure the effect of the leader board in the system, the 
+problem of cheating was mitigated by disabling the buttons for a certain amount of time. But the problem 
+persists, competitive users will try to get in to the leader board by other means. 
+% 
+% Future work could compare the use of gamification with out leaderboards
+%
+
+Using a graph database provided certain advantages when monitoring the experiments, for instance 
+using the query language
 

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Changed image

@@
     \centering
-        \includegraphics[width=3.5in]{img/user_known_2.png}
-    \caption{Sample of users, showing the {\tt knows} relationship.}
+        \includegraphics[width=3.5in]{img/users_known_1.png}
+    \caption{Comparasion of two users showing the {\tt knows} relationship.}
     \label{fig:users-graph}
@@
 
-Using a graph database provided certain advantages when monitoring the experiments, for instance 
-using the query language
+Using the Noe4J graph database and the Cypher query language when monitoring the experiments provided 
+views on the connected data difficult to replicate in a relational database. An example view is presented 
+in figure \ref{fig:users-graph} showing comparasion between two users on the degree of the 
+{\tt has} relationship.
 

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Results

@@
 In order to start each of the experiments was promoted through social networks
-particularly on the Facebook and Twitter platforms. The link used in the call for participation
+particularly on the Facebook and Twitter platforms.
+Is important to note that participants needed to log-in to facebook and give permissions to the application
+in order to participate, there where no anonymous volunteers. 
+The link used in the call for participation
 was shortened URL, in order to obtain additional analytic data of the visitors.
@@
 \label{sec:results}
+
 Before launch, each deployment was first tried with a few beta testers. When applying the gammification
@@
 
+The results of each of the three experiments in terms of participation are detailed next.
+
+\subsubsection{ Base deployment}
+After a week of the anouncement the total number of volunteers for this experiment was 53 registered users. 
+The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
+experiment a total of 500 new individuals where generated. The total number of evaluated
+individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a blue line.
+
+\subsubsection{ Non Graph Gamification deployment}
+The total number of volunteers for this experiment was very close to those obtained in the previous
+experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
+with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
+with a marginal increase in participation. The increased participation also helped in the creation of
+more individuals for a total of 648. The total number of evaluated
+individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.
+
+\subsubsection{ Graph Gamification deployment}
+This experiment had the higher number of volunteers for for a total of 68 registered users. 
+In this experiment a total of 3594 new individuals where generated in the experiment. 
+The total number of evaluated individuals for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} in green.
+
+\subsubsection{ Comparasion Results}
+The results presented in figure \ref{fig:top-ranked-participation}, show that those users with
+higher level of participation in deployments B and G, had about the same behavior, but the difference
+came with users with medium level of participation, in this case participants where more engaged in deployment
+G. When comparing all the experiments the deployment GG had the higher number of participation in users
+of all levels.
+
 \begin{figure*}[!t]
     \centering
-        \includegraphics[width=3.5in]{img/comparison.png}
+        \includegraphics[width=5.5in]{img/comparison.png}
     \caption{Top ranked number individuals rated by user using a

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Social network figures added

@@
 Is important to note that participants needed to log-in to facebook and give permissions to the application
-in order to participate, there where no anonymous volunteers. 
+in order to participate, there where no anonymous volunteers. This permited the automatic access to
+each participant's social network permiting the construction a local network of friends participating.
+
 The link used in the call for participation
@@
 
-\subsubsection{ Comparasion Results}
 The results presented in figure \ref{fig:top-ranked-participation}, show that those users with
@@
 
+When comparing the networks of users, there is also a difference shown between experiments.
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/user_known_2.png}
+    \caption{}
+    \label{}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/user_known_3.png}
+    \caption{}
+    \label{}
+\end{figure}
 \begin{figure*}[!t]

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Related Work added
Closes #2

@@
 architecture using a relational database and the file system for genotypes and images.
+
+Other works employ the use of No SQL databases, the work of Seyama and Munetomo \cite{seyama2016development}
+employs in their implementation a Platform as-a Service (PaaS) and NoSQL database to facilitate collaboration 
+among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
+by using a collaborative filtering algorithm to show the information searched for and utilized by similar users.
 

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Captions and description for network of users

@@
 
-Using the Noe4J graph database and the Cypher query language when monitoring the experiments provided 
-views on the connected data difficult to replicate in a relational database. An example view is presented 
-in figure \ref{fig:users-graph} showing comparasion between two users on the degree of the 
-{\tt has} relationship.
-
 The results of each of the three experiments in terms of participation are detailed next.
@@
         \includegraphics[width=3.5in]{img/user_known_2.png}
-    \caption{}
-    \label{}
+    \caption{User network view of the Base deployment}
+    \label{fig:B-network}
 \end{figure}
@@
         \includegraphics[width=3.5in]{img/user_known_3.png}
-    \caption{}
-    \label{}
+    \caption{User network view of the Graph Gamification deployment}
+    \label{fig:GG-network}
 \end{figure}
@@
 \end{figure*}
+
+Using the Noe4J graph database and the Cypher query language when monitoring the experiments provided 
+views on the connected data difficult to replicate in a relational database. As an example, when comparing the
+deployments B and GG based on their network of users (shown in figures \ref{fig:B-network} and \ref{fig:GG-network}
+respectivly) additionaly to the differences in the number of nodes, certain patterns can also be observed.
 

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More related work

@@
 their creation. Also there is information regarding users and their collaborations.  Common offerings show 
-interesting individuals order by categories, number of votes, best of the week among others.
-These features have the goal of encouraging participation but require the use of a database system 
-for their implementation. Picbreeder \cite{picbreeder} is implemented with a service oriented
-architecture using a relational database and the file system for genotypes and images.
+interesting individuals order by categories, number of votes, best of the week among others. In a recent work by
+Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
+new designs of robot locomotiom. Collaboration is encoureged by gamifying the system, using the maximum distance indicator
+to inspire the user to try and “beat” previous designs.
 
-Other works employ the use of No SQL databases, the work of Seyama and Munetomo \cite{seyama2016development}
-employs in their implementation a Platform as-a Service (PaaS) and NoSQL database to facilitate collaboration 
+Implementing recommendation algorithms or adding gamification techniques have the goal of encouraging participation 
+but require the use of a database system for managing the additional data required. 
+For example Picbreeder \cite{picbreeder} is implemented with a service oriented
+architecture using a relational database and the file system for genotypes and images. Other works employ the 
+use of No SQL databases like the work of Seyama and Munetomo \cite{seyama2016development} where they
+employ a Platform as-a Service (PaaS) and NoSQL database to facilitate the collaboration 
 among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
-by using a collaborative filtering algorithm to show the information searched for and utilized by similar users.
+by using a collaborative filtering algorithm to show only the information utilized by similar users.
 
 Collaborative IEC systems need to store highly connected data, as it is common in current applications
-like social networks, and the world wide web. In order to deal with large datasets of connected data,
+like social networks. In order to deal with large datasets of connected data found in these systems,
 graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
 which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
-
-The present work builds on a previous proposals and extends the C-IEA approach.
-First, it proposes a graph data model for representing the interactive process of C-IEC systems.
-Second, it incorporates the graph model to the evolutionary algorithm.
-Finally, uses the same model to increase the collaboration of volunteers.
+In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
+Moreover, an proof-of-concept implemantation of IEC incorporating the graph model to the evolutionary algorithm is presented.
+%Finally, a comparasion of between three deployments of the system, with different increase the collaboration of volunteers.
 
@@
 
-\item {\bf Shared} When a user takes an individual from a friends collection, this interactions is described by a {\tt share} relationship, between the user and the individual.
+\item {\bf Shared} When a user takes an individual from a friends collection, this interactions 
+is described by a {\tt share} relationship, between the user and the individual.
 

### 2017-01-24 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Sentence corrected

@@
 Moreover, an proof-of-concept implemantation of IEC incorporating the graph model to the evolutionary algorithm is presented.
+
 %Finally, a comparasion of between three deployments of the system, with different increase the collaboration of volunteers.
@@
 
-Only the "Start a session" and "Explore collections of other friends". 
-Are not found in the graph model, these are stored in the PostgreSQL relational
+Only the ``Start a session'' and ``Explore collections of other friends'' actions 
+are not found in the graph model as these are stored in the PostgreSQL relational
 database system.

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
A few corrections

@@
 as web applications in order to give users easy access the application as they
-often depend on volunteers who visit the system to interactively help with the search. 
+often depend on volunteers who visit the system to interactively help with fitness assignement. 
 There are several drawbacks that limit user participation 
@@
 a large number of individuals. 
-To tackle this issues, we propose the design of IEC systems using a human-centered 
+To tackle these issues, we propose the design of IEC systems using a human-centered 
 approach, with a framework consisting of a social network of 
 volunteers, interacting with a population consisting also of a network 
-of individuals. In this work we propose a graph model as a practical 
-and efficient tool for mapping in a natural way the relationships 
-found in the system. A case study is presented as a proof of concept, 
+of individuals. In this work we suggest the use a graph model as a practical 
+and efficient tool for mapping the relationships 
+found in the system. A case study is presented as proof-of-concept, 
 providing both conceptual and implementation details of a graph based model 
-as it is  applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
-As a result, the graph model was successfully applied in a gamification technique 
+as it is  applied in the implementation of an IEC system. 
+As result show, the data model was successfully applied in a gamification technique 
 developed to increase user engagement.
@@
 Section \ref{sec:interactive} presents related work on the topic 
-of Interactive Evolution with emphasis in data models.
+of Interactive Evolution.
 Then, Section \ref{sec:evospace-i} presents the computational platform on which 
@@
 implemented as volunteer based collaborative systems; where volunteers interact and evaluate 
-an evolving population, thus guiding the search based on an aggregate their subjective 
+an evolving population, thus guiding the search based on an aggregate of their subjective 
 preferences. What follows is not intended as a comprehensive survey, only a review of 
@@
 twelve displays by selecting those they find most aesthetically interesting by standing on
-step sensors in front of each. An early example of web based collaboration is the work by 
-Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a distributed 
-EA using a global population stored on a central web server sending 
+step sensors in front of them. An early example of web based collaboration is the work by 
+Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a 
+distributed EA using a global population stored on a central web server sending 
 portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
@@
 
-In all these systems, the evolutionary interactive process is captured by the system. The genetic lineage of each
-individual is presented where users can visualize the relationships between them and evolutionary operators used in
-their creation. Also there is information regarding users and their collaborations.  Common offerings show 
-interesting individuals order by categories, number of votes, best of the week among others. In a recent work by
+In all these systems, the interactive process is captured by the system. The genetic lineage of each
+individual is presented in a graphical interface 
+where users can visualize the relationships between them, and evolutionary operators used in
+their creation; they also display information regarding users and their collaborations.  Common offerings show 
+interesting individuals ordered by categories, number of votes, and best of the week among others. In a recent work by
 Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
-new designs of robot locomotiom. Collaboration is encoureged by gamifying the system, using the maximum distance indicator
-to inspire the user to try and “beat” previous designs.
+new designs of robot locomotiom. Collaboration is encoureged by gamifying the system 
+using the maximum distance indicator to inspire the user to try and “beat” previous designs.
 
@@
 
-Collaborative IEC systems need to store highly connected data, as it is common in current applications
+Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common in current applications
 like social networks. In order to deal with large datasets of connected data found in these systems,
@@
 In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
-Moreover, an proof-of-concept implemantation of IEC incorporating the graph model to the evolutionary algorithm is presented.
+Moreover, an proof-of-concept implemantation of an IEA incorporating the graph model is presented.
 
@@
 Here, the user can interact with the system in two ways.
-He can assign a quality assessment to the individual using
-a five star rating measure.
-Additionally, a user can choose to add an image to one of their \emph{Collections}.
-A collection is a special directory to store individuals a user prefers and wishes
-to save. After the user finishes interacting the individual
-on the Wall, he can choose to retrieve a new one from population.
+First, he can assign a quality assessment to the individual using
+a five star rating measure or
+additionally, a user can choose to add an image to one of their \emph{Collections}.
+A collection is a special folder to store those individuals a user prefers and wishes
+to save. After the user finishes interacting with the individual
+on the Wall, he can choose to retrieve a new one from the population.
 At the left side, is the \emph{Collections} section.

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Neo4J

@@
 
-Using the Noe4J graph database and the Cypher query language when monitoring the experiments provided 
+Using the Neo4J graph database and the Cypher query language when monitoring the experiments provided 
 views on the connected data difficult to replicate in a relational database. As an example, when comparing the

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Few changes

@@
 Heroku cloud service with the same configuration: using the
-Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB plug-in for the GG experiment.
+Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB and RedisTOGO plug-ins.
 Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
-system used in the implementation was Neo4J, and it was allready available as a plug-in in Heroku.
+system used in the implementation was Neo4J as it was allready available as a plug-in in Heroku, is well documented
+and a scalable solution \cite{miller2013graph,holzschuher2013performance}.
 
@@
 % Anuncements
-In order to start each of the experiments was promoted through social networks
-particularly on the Facebook and Twitter platforms.
+At the start of each of the experiments, a call to participation was promoted through social networks
+particularly on Facebook and Twitter social networks.
 Is important to note that participants needed to log-in to facebook and give permissions to the application
-in order to participate, there where no anonymous volunteers. This permited the automatic access to
-each participant's social network permiting the construction a local network of friends participating.
+in order to participate, there were no anonymous volunteers. This permited the automatic access to
+each participant's social network and to store a local copy of those friends also participating.
 
 The link used in the call for participation
-was shortened URL, in order to obtain additional analytic data of the visitors.
+was shortened URL, in order to obtain additional data and analytic of the visitors.
 In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Eliminating comments on the top
Reviews abstract
Among some corrections, I have added the last sentence. Please check
that it's true and correct @mariosky.

@@
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[conference]{IEEEtran}
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
 
@@
 \usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
 
@@
 \fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
 
@@
 %
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
+\usepackage{url}
 
@@
 practice is to develop Interactive Evolutionary Computation (IEC) algorithms
-as web applications in order to give users easy access the application as they
+as web applications in order to give users easy access to the
+application as the algorithms 
 often depend on volunteers who visit the system to interactively help with fitness assignement. 
-There are several drawbacks that limit user participation 
+However, there are several drawbacks that limit user participation 
 such as human fatigue and boredom resulting from evaluating 
 a large number of individuals. 
-To tackle these issues, we propose the design of IEC systems using a human-centered 
+To tackle these issues, in this paper we propose the design of IEC systems using a human-centered 
 approach, with a framework consisting of a social network of 
 volunteers, interacting with a population consisting also of a network 
-of individuals. In this work we suggest the use a graph model as a practical 
+of individuals. In this work we suggest the use of a graph model as a practical 
 and efficient tool for mapping the relationships 
@@
 as it is  applied in the implementation of an IEC system. 
-As result show, the data model was successfully applied in a gamification technique 
-developed to increase user engagement.
+As the results show, the data model was successfully applied in a gamification technique 
+developed to increase user engagement, which means that this technique can successfully be used to
+decrease user fatigue and thus increase {\em computer power}
+delivered by the set of users.
 

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Changed those wrong where to were

--- 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 A brief description of the application is presented next, focusing
-on the data elements that where ported to the graph model.
+on the data elements that were ported to the graph model.
 
@@
 in the graph. This measure also considers the number of users he knows,
-perhaps some of the users where invited to the system by him.   
+perhaps some of the users were invited to the system by him.   
 \end{itemize}
@@
 \end{itemize}
-When gamification was employed, all score values known where presented to users
-and a ranking of users by weight was shown in a window, a screen capture of
-this interface is shown in figure \ref{fig:top-users}.
+When gamification was employed, all score values known were presented to users
+along with a ranking of users by their weight was shown in a window, as an example
+a screen capture of this interface is shown in figure \ref{fig:top-users}.
 
@@
 The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
-experiment a total of 500 new individuals where generated. The total number of evaluated
+experiment a total of 500 new individuals were generated. The total number of evaluated
 individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a blue line.
@@
 This experiment had the higher number of volunteers for for a total of 68 registered users. 
-In this experiment a total of 3594 new individuals where generated in the experiment. 
+In this experiment a total of 3594 new individuals were generated in the experiment. 
 The total number of evaluated individuals for each volunteer is presented in figure 
@@
 higher level of participation in deployments B and G, had about the same behavior, but the difference
-came with users with medium level of participation, in this case participants where more engaged in deployment
+came with users with medium level of participation, in this case participants were more engaged in deployment
 G. When comparing all the experiments the deployment GG had the higher number of participation in users

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Little explanation of the EvoDrawings app

@@
 %
-%\usepackage{url}
+\usepackage{url}
 % url.sty was written by Donald Arseneau. It provides better support for
@@
 knowledge about the process and the population, for instance showing the
-most popular individual, or the the user with more participation 
+most popular individual, or the the user with more participation.  
 \cite{picbreeder}.
@@
 \end{figure*}
+
+\subsection{EvoDrawings}
+\label{sec:evo_drawings}
+The particular application implemented for this work is called
+EvoDrawings. In this IEA individuals are drawing created by an animation of three types of 
+shapes: equilateral triangles, squares and cirlcles, varying in size and color. The movement of
+each shape is defined by a behaviour rule, for instance  ``move straight'', 
+``change direction in case of colission'',  etc., this animations are inspired in 
+the Process Compendium by Reas \cite{reas2004}, where visual forms 
+emerge by the relations of multiple forms and their behavior.
 
@@
 \begin{itemize}
-\item Base (B). All users have the same weight.
-\item Non Graph Gamification (G). Only experience is considered
-\item Graph Gamification (GG) . Both experience and participation are considered.
+\item Base (B). All users have the same weight. Still the graph is used internally to explore
+the social network and the evolution of the population.
+\item Non Graph Gamification (G). Only experience is considered for the reputation of users. 
+\item Graph Gamification (GG). Both experience and participation are considered.
 \end{itemize}

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Reviews stuff

--- 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 \section{Introduction}
-Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
+Interactive Evolutionary Algorithms (IEAs) are, in general, standard EAs whose
 fitness evaluations are performed by persons within an interactive 
-system.  Thus, the main loop of the EA needs the intervention of a
+system \cite{eiben2015interactive}.  Thus, the main evolutionary loop needs the intervention of a
 human to perform the quality assessment of each proposed solution.
-IEAs have demonstrated their ability for effectively
+These kind of IEAs have demonstrated their ability for effectively
 producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
-as well as other products in many other domains \cite{ie1}. 
-Several IECs have been developed as web based systems \cite{picbreeder},
+as well as other products in many other domains \cite{ie1}, and this
+success has led to the development of web based systems such as \cite{picbreeder},
 who depend on volunteers who visit the system to interactively help
-with the search, using both anonymous and registered users.
-These systems have several inherent drawbacks arising from the very nature of 
+with the search, using both anonymous and registered users. This
+volunteeer system lowers the requirements of anyone participating in
+the experiment thus increasing the {\em performance}, in terms of {\em
+  human brain cycles}, of the whole system.
+
+However, the necessary intervention of humans in these systems leads
+to them having several inherent drawbacks arising from the very nature of 
 the algorithms, namely, the human fatigue caused by the interaction, and
@@
 engagement, and to tackle the issues mentioned above,  
-IEAs must follow a human centered design, giving extensive attention 
+IEAs must follow a human centered design, % reference and explain what
+                                % you mean by that. It's an EC
+                                % conference - JJ
+giving extensive attention 
 to volunteer users, not only because their

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex

@@
 \begin{itemize}
-\item Base (B). All users have the same weight.
-\item Non Graph Gamification (G). Only experience is considered
-\item Graph Gamification (GG) . Both experience and participation are considered.
+\item Base (B). All users have the same weight. Still the graph is used internally to explore
+the social network and the evolution of the population.
+\item Non Graph Gamification (G). Only experience is considered for the reputation of users. 
+\item Graph Gamification (GG). Both experience and participation are considered.
 \end{itemize}
-When gamification was employed, all score values known where presented to users
-and a ranking of users by weight was shown in a window, a screen capture of
-this interface is shown in figure \ref{fig:top-users}.
+When gamification was employed, all score values known were presented to users
+along with a ranking of users by their weight was shown in a window, as an example
+a screen capture of this interface is shown in figure \ref{fig:top-users}.
 

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More changes here

--- 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 with other users, their previous evaluations and their own stream of activities
-and past experiences. Thus users in this proposal are defined as dynamic entities 
+and past experiences. Thus users in the method presented in this paper are defined as dynamic entities 
 interacting with an evolutionary computation having the fundamental purpose 
@@
 importance to users and their interactions as the population of 
-individuals have in a traditional EA, even to the extend as using this
-knowledge as an integral part of the evolutionary algorithm. We argue also
-that a graph model is a viable representation for this purpose,  
+individuals have in a traditional EA, even to the extent of using this
+knowledge as an integral part of the evolutionary algorithm. % how?
+                                % Explain a bit more
+We argue also
+that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
 presenting a natural way of mapping the relationships found in the system. 
-Most IEAs have the following many-to-many relations:
+Most IEAs have % have is not very precise and I'm a bit lost
+               % here. What do you mean? is this constrained to IEAs
+               % or EAs? At least the first is not exclusive, right? - JJ
+the following many-to-many relations:
 
@@
   population are naturally expressed as a tree, or a graph when both ancestors 
-  and descendants are considered. Other semantic relations can also be of interest
-  for instance similarity or composition.
+  and descendants are considered. Other semantic relations can also be
+  of interest, 
+  for instance similarity or composition. %This is interesting for
+                                %what reason? What do you mean by
+                                %composition? "to be composed of"? - JJ
 
-  \item {\bf User-Individual}
-  Many kind of relationships between users and individuals,
+  \item {\bf User-Individual} There are many kind of relationships
+    between users and individuals, and these can be
   can be expressed using common actions found in social networks;
@@
   relationships are used to assign a single or a collection of
-  fitness values to individuals \cite{garcia2013evospace}.
+  fitness values to individuals \cite{garcia2013evospace}.   % I am
+                                % afraid that individual is too
+                                % similar to user, and you might get
+                                % lost here. Maybe talk about
+                                % "genotypes" or "phenotypes"? - JJ
 
   \item {\bf User-User}
-  Relationship between users can also be modeled 
+  Relationship between users in an interactive evolutionary algorithm can also be modeled 
   as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
@@
 
-These relations are naturally mapped to graphs, while other actions and 
+These relations are naturally mapped to graphs, %They are bipartite
+                                %graphs, but I don't see your point...
+while other actions and 
 properties can also be modeled as a semantic network using
-basic subject–-predicate-–object expressions \cite{Prud'hommeaux:14:RT}.
-Another reason for using a graph model is the tools available for
+basic subject–-predicate-–object expressions
+\cite{Prud'hommeaux:14:RT}. % that's a graph, too.. - JJ
+Another reason for using a graph model % But for what exactly? A
+                                % bipartite graph to express which
+                                % relations? - JJ
+is the wealth of tools available for
 implementing and query the model, and the increased performance of
@@
 
-The IEC the approach has been used in a wide variety of applications and problem domains
+Up until now, IEC  has been used in a wide variety of applications and problem domains
 and remains an active area of research. However, in this work, emphasis is given to IEAs 
@@
 an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
-twelve displays by selecting those they find most aesthetically interesting by standing on
+twelve displays by selecting those they found most aesthetically interesting by standing on
 step sensors in front of them. An early example of web based collaboration is the work by 
-Langdon \cite{langdon:2004}, evolving fractal representations of virtual creatures. It proposed a 
+Langdon \cite{langdon:2004}, which evolved fractal representations of virtual creatures. It used a 
 distributed EA using a global population stored on a central web server sending 
@@
 
-In all these systems, the interactive process is captured by the system. The genetic lineage of each
+In all these systems, the interactive process is captured by the
+system and no special attention is lent to the process of capturing
+and engaging users themselves. In some cases, the genetic lineage of each
 individual is presented in a graphical interface 
 where users can visualize the relationships between them, and evolutionary operators used in
-their creation; they also display information regarding users and their collaborations.  Common offerings show 
-interesting individuals ordered by categories, number of votes, and best of the week among others. In a recent work by
+their creation; they also display information regarding users and
+their collaborations.  Common control panels  show, for instance,  
+interesting individuals ordered by categories, number of votes, and
+best of the week among others. In a recent work by 
 Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
@@
 but require the use of a database system for managing the additional data required. 
-For example Picbreeder \cite{picbreeder} is implemented with a service oriented
+For example, Picbreeder \cite{picbreeder} is implemented with a service oriented
 architecture using a relational database and the file system for genotypes and images. Other works employ the 

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
And this is it for the day
I'm going for dinner.

@@
 among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
-by using a collaborative filtering algorithm to show only the information utilized by similar users.
+by using a collaborative filtering algorithm to show only the
+information utilized by similar users. In any case, using gamification
+techniques imply that we must deal with interactive evolutionary
+system as sociotechnical constructs, where the social aspects are
+essential to understand its dynamics. In this sense, conclusions
+reached with other systems such as NodEO
+\cite{DBLP:conf/gecco/MereloCGCRV1} can also be applied to this
+system; and applying social network techniques such as graph analysis
+to their study will allow us to understand them more thoroughly.
 
-Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common in current applications
+Collaborative IEC (C-IEC) systems also need to store highly connected data, as it is common in current applications
 like social networks. In order to deal with large datasets of connected data found in these systems,
@@
 
-\section{EvoSpace-Interactive}
+\section{Experimental setup: the EvoSpace-Interactive framework}
 \label{sec:evospace-i}

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More on Leader Board

@@
 particularly on Facebook and Twitter social networks.
-Is important to note that participants needed to log-in to facebook and give permissions to the application
-in order to participate, there were no anonymous volunteers. This permited the automatic access to
-each participant's social network and to store a local copy of those friends also participating.
+It is important to note that volunteers needed to log-in in to facebook and give permissions to the 
+web application in order to participate, in thi case there were no anonymous volunteers. This permited the automatic access to
+each participant's social network and enabled the system to store a local copy of those friends participating.
+Another experiment is needed to see the effect this requirement has on participation.
 
@@
 
-Before launch, each deployment was first tried with a few beta testers. When applying the gammification
-technique for the first time a problem was found in this beta stage. Some users were chating by 
-giving a rating to an animation even before it was returned from the server, this was done by just
-constantly pressing the mouse button. This is a common problem found in systmes appling leader boards.
-By making the rewards for tasks achievements visible to other players or providing leader boards are ways of 
-encouraging players to compete \cite{hickman2010total}. Because the  problematic consequences of competition, 
-which can result in negative conduct, low cooperation and collaboration, or disadvantaging certain player
-demographics such as women \cite{kumar2013gamification}, best-practice gamification designs try to
-refrain from using this element. In order to measure the effect of the leader board in the system, the 
-problem of cheating was mitigated by disabling the buttons for a certain amount of time. But the problem 
-persists, competitive users will try to get in to the leader board by other means. 
-% 
+Before launch, each deployment was first tried with a few beta testers. When applying the leader board
+gammification technique for the first time a problem was found in this stage. The problem was that some 
+users were chating by giving a rating to an animation even before it was returned from the server, this was done by just
+constantly clicking the mouse button. This is a common problem found in systems using leader boards because
+by making the scores visible to other players they are encouraged to compete \cite{hickman2010total}. 
+According to \cite{kumar2013gamification} best-practice gamification designs try to
+refrain from using this element because of  problematic consequences of competition, which can result 
+in negative conduct, low cooperation and collaboration, or disadvantaging certain player
+demographics such as women. In order to still measure the effect of the leader board on the system, the 
+problem of cheating was mitigated by disabling the rate button for a certain amount of time giving the animation
+enough time to finish. But the problem persists, competitive users will try to get in to the leader board by other means.
+
+% Chating in an IEA is a serious but common risk in volunteer based systems.  
 % Future work could compare the use of gamification with out leaderboards

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More results

@@
 
-% Chating in an IEA is a serious but common risk in volunteer based systems.  
 % Future work could compare the use of gamification with out leaderboards
 %
-
 The results of each of the three experiments in terms of participation are detailed next.
@@
 
-When comparing the networks of users, there is also a difference shown between experiments.
-
 \begin{figure}[!t]
@@
 
-Using the Neo4J graph database and the Cypher query language when monitoring the experiments provided 
-views on the connected data difficult to replicate in a relational database. As an example, when comparing the
-deployments B and GG based on their network of users (shown in figures \ref{fig:B-network} and \ref{fig:GG-network}
-respectivly) additionaly to the differences in the number of nodes, certain patterns can also be observed.
+When monitoring the experiments, using the graph database and the Cypher query language 
+provided interesting views on the connected data difficult to replicate with a relational database. 
+As an example we can consider comparing the difference between the number of volunteers in deployments
+B and GG,(shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectivly) additional
+metrics and patterns are available when they are treated as a network. The same can be applied to 
+the population, or the interaction as a whole. For instance when creating a new individual 
+parents could be selected if there is a minimum distance between them in the graph.
 

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Reference

@@
 reached with other systems such as NodEO
-\cite{DBLP:conf/gecco/MereloCGCRV1} can also be applied to this
+\cite{merelo2014nodeo} can also be applied to this
 system; and applying social network techniques such as graph analysis

### 2017-01-25 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Results

@@
 metrics and patterns are available when they are treated as a network. The same can be applied to 
-the population, or the interaction as a whole. For instance when creating a new individual 
-parents could be selected if there is a minimum distance between them in the graph.
+the population, or the interaction as a whole. The graph could be used also by the evolutionary
+algorithm, for instance when creating a new individual 
+parents could be selected only if there is a minimum distance between them in the graph.
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Added images and reference

@@
 reached with other systems such as NodEO
-\cite{merelo2014nodeo} can also be applied to this
+\cite{DBLP:conf/gecco/MereloCGCRV16} can also be applied to this
 system; and applying social network techniques such as graph analysis

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More  details added

@@
                                 % conference - JJ
+                                % I explain more later in other section, with references, maybe move that to here? -Mario
+                                % If not I will add a few words later, such a large topic -Mario
+                                % even with a Journal: http://hcis-journal.springeropen.com/about
 giving extensive attention 
@@
                                 % Explain a bit more
+                                % This is the idea:
+Fitness assignement could depend on the actions of a group of users, and depend on how
+they tag, share, rate, or even delete the phenotype. The selection of parents could
+also depend on the previous activites: they have the same tag?, they were shared by
+similar users? a leader of opinion liked both? they are stored in the same collection?.
+That knowledge could also be applied in choosing which phenotypes present to a
+user, in order to give them those they will find more interesting or useful.
+
 We argue also
 that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
+% I was thinking to record all the activities and relationships
+
 presenting a natural way of mapping the relationships found in the system. 
@@
 the design of an exit sign in an airport, then the quality of this sign could be
-determined by the average the time it takes a sample of people to 
+determined by the average time it takes a sample of people to 
 find the exit. In this example there could be certain collaboration

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Fixing bibliography
Fixes biblio again and more comments

--- 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 sharing those individuals they like, or by taking individuals
-from their friend\'s collections by using the web interface depicted 
+from their friend's collections by using the web interface depicted 
 in Figure \ref{fig:web}.
@@
   There is also the possibility that fitness or even part of the search 
-  to be done by devices lent by the human \cite{merelo2016performance}.
+  to be done by devices lent by the human \cite{DBLP:conf/gecco/MereloCGCRV16}.
   The information gathered through the interaction is the primary concern
@@
 
-\item {\bf Shared} When a user takes an individual from a friends collection, this interactions 
+\item {\bf Shared} When a user takes an individual from a friend's collection, this interactions 
 is described by a {\tt share} relationship, between the user and the individual.
@@
 
+%This section on gamification should go to the intro and state of the
+%art, since it is one of the main points of the paper - JJ
 Gamification is defined by Deterding et al. as
@@
 The link used in the call for participation
-was shortened URL, in order to obtain additional data and analytic of the visitors.
+was shortened Google URL, that provides metadata and analytics for the
+users that click on it. 
 In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
@@
 application repository used to deploy to Heroku is also listed.    
+%Isn't this anonymous? - JJ
 
@@
         \includegraphics[width=5.5in]{img/comparison.png}
-    \caption{Top ranked number individuals rated by user using a
-     logarithmic scale.}
+    \caption{Top ranked number individuals rated by each user using a
+     logarithmic scale.} % Explain a little more and label axes. The
+                         % three lines have the same label,
+                         % "participation". Not too clear - JJ
     \label{fig:top-ranked-participation}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{evospace-i,volunteer,biblio}
+\bibliography{evospace-i,volunteer,biblio,geneura}
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Purpose of graph

@@
 We argue also
-that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
+%that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
 % I was thinking to record all the activities and relationships
+that a graph model is a viable representation for the above relationships and
+interactions. Graph based models are currently used in social networks to keep track
+of user interactions with media objects, places, and other users. Users of
+social networks (for instance the Facebook Graph) are accustomed to express these 
+complex relationships for instance: ``Jhon and Ann eating breakfast at Tony's''. 
+Other example is the W3C Activity Streams 2.0 specification used for representing activities 
+common in social web applications.
 
-presenting a natural way of mapping the relationships found in the system. 
+%presenting a natural way of mapping the relationships found in the system. 
 Most IEAs have % have is not very precise and I'm a bit lost

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Changing the title
Check it out, @mariosky
Adds reference to activity streams

@@
 
-\title{Exploiting the Graph: Increasing Engagement in Collaborative Interactive Evolution}
+\title{Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution}
+% Changed it to "Social". What do you think? - JJ
 
@@
 interacting with an evolutionary computation having the fundamental purpose 
-of evaluating individuals according to their preferences,
+of evaluating individuals accordi ng to their preferences,
 both explicitly or implicitly. In this work we propose giving the same
@@
 they tag, share, rate, or even delete the phenotype. The selection of parents could
-also depend on the previous activites: they have the same tag?, they were shared by
-similar users? a leader of opinion liked both? they are stored in the same collection?.
-That knowledge could also be applied in choosing which phenotypes present to a
+also depend on the previous actions, leveraging information such as the fac that  they have the same tag, or shared by
+similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
+That knowledge could also be applied in choosing which phenotypes whould be presented to a
 user, in order to give them those they will find more interesting or useful.
@@
 that a graph model is a viable representation for the above relationships and
-interactions. Graph based models are currently used in social networks to keep track
+interactions and can then be used to model and use, within the
+evolutionary algorithm, actions performed by every user. Graph based
+models are currently used in social networks to keep track  
 of user interactions with media objects, places, and other users. Users of
 social networks (for instance the Facebook Graph) are accustomed to express these 
-complex relationships for instance: ``Jhon and Ann eating breakfast at Tony's''. 
+complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
 Other example is the W3C Activity Streams 2.0 specification used for representing activities 
-common in social web applications.
+common in social web applications \cite{json:streams}..
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Eliminates addressed comments

@@
   \item {\bf Individual-Individual} The relationships between individuals in the
-  population are naturally expressed as a tree, or a graph when both ancestors 
+  population % do you mean different generations? Would be better to
+             % make this more precise - JJ
+are naturally expressed as a tree, or a graph when both ancestors 
   and descendants are considered. Other semantic relations can also be
@@
                                 %graphs, but I don't see your point...
+% Besides, _any_ relation can be mapped to a graph - JJ
 while other actions and 
@@
 native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.
-
-% Moving to related work - Mario
-%\cite{Performance for the masses and Christian's previous work},
-
- % Can it be quantified? Maybe a reference to NodIO? - JJ
-%Some authors have already tried to address this basic problem
-%\cite{Frade:2010:EvoGAMES}. %Non sequitur. Is it something we are
-                            %going to do here? - jj
-%% Moving this to Related Work Section -Mario
-% That section is not there yet. Forgot to paste? - JJ
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Splitting paragraphs that are terribly long

@@
 many of which are not interesting or are very similar to each other.
+
 IEAs also share some of the issues commonly found in volunteer based systems
@@
 interacting with an evolutionary computation having the fundamental purpose 
-of evaluating individuals accordi ng to their preferences,
-both explicitly or implicitly. In this work we propose giving the same
+of evaluating individuals according to their preferences,
+both explicitly or implicitly.
+
+In this work we propose giving the same
 importance to users and their interactions as the population of 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Some small corrections
But check @mariosky the last sentence to see if it reflects correctly
the intentions of the paper.

@@
 % http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
 
@@
 First, he can assign a quality assessment to the individual using
-a five star rating measure or
+a five star rating system or,
 additionally, a user can choose to add an image to one of their \emph{Collections}.
-A collection is a special folder to store those individuals a user prefers and wishes
+A collection is a special folder that stores those individuals a user likes and wishes
 to save. After the user finishes interacting with the individual
 on the Wall, he can choose to retrieve a new one from the population.
-At the left side, is the \emph{Collections} section.
+At the left hand side, the web page shows the \emph{Collections} section.
 The user can create several collections, to group and organize his favorite 
-artifacts. Moreover, a user can browse the content of each collection and from
-there share images through the social network.
-When a user browses over an individual a detail pane shows how many users have
+artifacts. Moreover, users can browse the content of each collection and from
+there share images through their social network.
+When a user hovers the cursor over an individual a pop-up pane shows how many users have
 liked the individual. The pane also includes a link to the individual's 
-details, the parents, genetic operators that created it, and genealogy information.
+details, the parents, genetic operators that created it, and genealogy
+information. This makes the assignment of fitness through the rating
+system a {\em social} activity, pursuing the objective of this work,
+which is to increase engagement of the users. % Adding this connecting
+                                % to paper's objetives - JJ
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Little extension and comments

@@
 by the natural interactions of a group of humans in a collaborative fashion
-or not. For instance the IEC algorithm could by trying to optimize 
+or not. For instance the IEC algorithm could be trying to optimize 
 the design of an exit sign in an airport, then the quality of this sign could be
@@
 find the exit. In this example there could be certain collaboration
-between users if they are not traveling alone, of course this is only one of
+between users if they are not traveling alone, but of course this is only one of
 the considerations designers could make.
@@
 \end{itemize}
+
+These components will be described in detail next.
+% Besides, add something about the innovation in this paper, for
+% instance, explaining that the data model is extended to a graph here
+% - JJ
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More suggestions

@@
 world system, and relationships describe their interactions. Types describe
-the attributes describing an entity, First we describe the types of nodes and
-relationships in the system.
+the attributes describing an entity.
+
+Next we describe the types of nodes and
+relationships in the system. % I think a bit of more explanation would
+                             % help why you have these types and
+                             % whether they are simple implementation
+                             % details or part of the method - JJ
 
@@
 
-\item {\bf Users} This nodes represent the human users of the system and have 
-the attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
+\item {\bf Users} This type of node represent the human users of the system and have 
+the following attributes: {\tt id} a unique identifier, {\tt name}  and {\tt created}
 specifying the  creation date (time stamp) of the node.
@@
 \item {\bf Individual} This type of node has the following attributes: 
-{\tt id}, {\tt chromosome}  and {\tt views} this is 
-the amount that users have seen or interacted with this individual and
- {\tt created}.
+{\tt id}, {\tt chromosome}  and {\tt views}, which is the number of
+times users have seen or interacted with this individual and, finally, 
+ {\tt created}, the time and date of creation as above..
 
-\item {\bf Collection} This type of nodes represent the collections belonging
+\item {\bf Collection} This type of node represents the collections belonging
 to users, their attributes are {\tt id}, 
-{\tt name}, {\tt created} and {\tt is\_private} indicating if this collection
-contents can be are shared with others. 
+{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
+belonging to this collection can be are shared with others. 
 \end{itemize}

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Proceeding to case study

@@
 
+% Add something at the end linking with the next paragraph
+
 \subsection{Relationship Types}
+
+These relationships are expressed as edges in the graph. We use
+several of them.
 
@@
 interest.
+%this is a bipartite graph them. - JJ
 
 \item {\bf Knows} The relation connects two users that know each
-other. This relation is "unique" meaning that it can not be repeated between a pair of nodes.
+other. This relation is "unique" meaning that it cannot be repeated
+between a pair of nodes.  % Not too clear what you mean by that - JJ
 
-\item {\bf Parent} Describes what individual is the parent of a new individual and is also unique.
+\item {\bf Parent} Describes what individual is the parent of a new
+  individual and is also unique.
 
@@
 \item {\bf Shared} When a user takes an individual from a friend's collection, this interactions 
-is described by a {\tt share} relationship, between the user and the individual.
+is described by a {\tt share} relationship between the user and the
+individual.  % Still think that user and individual is confusing - JJ
 
@@
 
-Is important to notice that this model is still partial view of the IEC system,
-and other models can could be used in practice when implementing the system. 
+Is important to notice that this model is still a partial view of the IEC system,
+and other models can could be used in practice when implementing the
+system. % It is also a bipartite graph, which could be converted to a
+% single-party graph with only one type of node.
+% I don't think "collections" should be nodes too. They are simply
+% attributes for an individual - JJ
 For instance a relational database can be used for the implementation of a
@@
 Nevertheless, the model can give important insights of the behavior of the 
-system.
+system.   % So what we have done here is this and that because... - JJ
 
@@
 gamification techniques to increase volunteer participation against a third 
-without any thechnique applied.
+without any technique applied.
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Check in

@@
 Only the ``Start a session'' and ``Explore collections of other friends'' actions 
-are not found in the graph model as these are stored in the PostgreSQL relational
-database system.
+are not reflected in the graph model since they are stored instead in the PostgreSQL relational
+database system. % because... - JJ
 
@@
 % the description would take to much space.
+% Better a small description than nothing. If you need space, some
+% sections above are maybe a bit too verbose - JJ
 
@@
 When gamification was employed, all score values known where presented to users
-and a ranking of users by weight was shown in a window, a screen capture of
+and a ranking of users by weight was shown in a window. A screen capture of
 this interface is shown in figure \ref{fig:top-users}.
@@
 Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
-system used in the implementation was Neo4J as it was allready available as a plug-in in Heroku, is well documented
+system used in the implementation was Neo4J as it was already available as a plug-in in Heroku, is well documented
 and a scalable solution \cite{miller2013graph,holzschuher2013performance}.
@@
 % Anuncements
-At the start of each of the experiments, a call to participation was promoted through social networks
-particularly on Facebook and Twitter social networks.
-It is important to note that volunteers needed to log-in in to facebook and give permissions to the 
-web application in order to participate, in thi case there were no anonymous volunteers. This permited the automatic access to
+At the start of every experiment, a call to participation was issued
+through social networks, 
+particularly on Facebook and Twitter.
+It is important to note that volunteers needed to log-in to Facebook and give permissions to the 
+web application in order to participate, in this case there were no anonymous volunteers. This enabled the automatic access to
 each participant's social network and enabled the system to store a local copy of those friends participating.
-Another experiment is needed to see the effect this requirement has on participation.
+However, another experiment would be needed to see the effect this
+requirement has on participation, since we are not dealing with it in
+this paper.
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Suggestions for conclusions
And everything is revised now. From now on, please make me at @
message to draw my attention to specific parts to work on.

@@
 experiment a total of 500 new individuals where generated. The total number of evaluated
-individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a blue line.
+individuals for each volunteer is presented in figure
+\ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ
 
@@
 As an example we can consider comparing the difference between the number of volunteers in deployments
-B and GG,(shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectivly) additional
-metrics and patterns are available when they are treated as a network. The same can be applied to 
+B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectivly) additional
+metrics and patterns are available when they are treated as a
+network. %Please read the whole sentence and maybe rewrite - JJ
+The same can be applied to 
 the population, or the interaction as a whole. The graph could be used also by the evolutionary
@@
 \label{sec:conclusions}
+
+% Using graphs improves engagement and also provides for a model for
+% visualizing different experiments better. It has to tie with the
+% initial objectives.
 

### 2017-01-26 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Changed Individual to Phenotype

@@
 % Changed it to "Social". What do you think? - JJ
+% Great! - M
 
@@
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
@@
 such as human fatigue and boredom resulting from evaluating 
-a large number of individuals. 
+a large number of phenotypes. 
 To tackle these issues, in this paper we propose the design of IEC systems using a human-centered 
@@
 volunteers, interacting with a population consisting also of a network 
-of individuals. In this work we suggest the use of a graph model as a practical 
+of phenotypes. In this work we suggest the use of a graph model as a practical 
 and efficient tool for mapping the relationships 
@@
 delivered by the set of users.
-
-%and novelty. % These two last issues are not too clear, maybe they
-             % could be explained in that motivation.
-
 \end{abstract}
@@
 the algorithms, namely, the human fatigue caused by the interaction, and
-the boredom arising when users evaluate a large number of individuals 
+the boredom arising when users evaluate a large number of phenotypes 
 many of which are not interesting or are very similar to each other.
@@
 interaction affects the system as a whole. Users interact not only with the graphical
-interface, and each individual\'s phenotype, there is also the interaction
+interface, and each phenotype, there is also the interaction
 with other users, their previous evaluations and their own stream of activities
@@
 interacting with an evolutionary computation having the fundamental purpose 
-of evaluating individuals according to their preferences,
+of evaluating phenotypes according to their preferences,
 both explicitly or implicitly.
@@
 importance to users and their interactions as the population of 
-individuals have in a traditional EA, even to the extent of using this
+phenotypes have in a traditional EA, even to the extent of using this
 knowledge as an integral part of the evolutionary algorithm. % how?
@@
 
-  \item {\bf Individual-Individual} The relationships between individuals in the
+  \item {\bf Phenotype-Phenotype} The relationships between phenotypes in the
   population % do you mean different generations? Would be better to
@@
 
-  \item {\bf User-Individual} There are many kind of relationships
-    between users and individuals, and these can be
+  \item {\bf User-Phenotype} There are many kind of relationships
+    between users and phenotypes, and these can be
   can be expressed using common actions found in social networks;
@@
   relationships are used to assign a single or a collection of
-  fitness values to individuals \cite{garcia2013evospace}.   % I am
+  fitness values to phenotypes \cite{garcia2013evospace}.   % I am
                                 % afraid that individual is too
@@
                                 % "genotypes" or "phenotypes"? - JJ
+                                % changed to phenotypes 
+                                % do you think an aditional explanation is needed?
 
@@
   These measures can then be used by recommender algorithms selecting 
-  individuals according users' preferences.
+  phenotypes according users' preferences.
 
@@
 distributed EA using a global population stored on a central web server sending 
-portions of the population to remote clients using JavaScript. Users evaluated individuals locally 
+portions of the population to remote clients using JavaScript. Users evaluated phenotypes locally 
 and returned to the server those they liked more, to be distributed over the web.
@@
 and engaging users themselves. In some cases, the genetic lineage of each
-individual is presented in a graphical interface 
+phenotypes is presented in a graphical interface 
 where users can visualize the relationships between them, and evolutionary operators used in
@@
 their collaborations.  Common control panels  show, for instance,  
-interesting individuals ordered by categories, number of votes, and
+interesting phenotypes ordered by categories, number of votes, and
 best of the week among others. In a recent work by 
@@
 where several registered users assign a quality assessment to a single
-individual and then an aggregated fitness value is calculated. The fitness
-assigned to each individual depends on the taste of each particular user, 
-resulting in a many-to-many relationship between users and individuals. 
-Many systems query this user-individual relation to extract relevant
+phenotypes and then an aggregated fitness value is calculated. The fitness
+assigned to each phenotypes depends on the taste of each particular user, 
+resulting in a many-to-many relationship between users and phenotypes. 
+Many systems query this user-phenotype relation to extract relevant
 knowledge about the process and the population, for instance showing the
-most popular individual, or the the user with more participation 
+most popular phenotype, or the the user with more participation 
 \cite{picbreeder}.
-In order to do this, data about individuals 
+In order to do this, data about phenotypes 
 must be permanently stored, even
@@
 users can collaborate with their Facebook friends, 
-sharing those individuals they like, or by taking individuals
+sharing those phenotypes they like, or by taking phenotypes
 from their friend's collections by using the web interface depicted 
@@
 to encourage users to interact with the system. In the central 
-\emph{ Wall } area, an individual sampled from the population that is
+\emph{ Wall } area, a phenotype sampled from the population that is
 being evolved via the evolutionary algorithm 
@@
 Here, the user can interact with the system in two ways.
-First, he can assign a quality assessment to the individual using
+First, he can assign a quality assessment to the phenotype using
 a five star rating system or,
 additionally, a user can choose to add an image to one of their \emph{Collections}.
-A collection is a special folder that stores those individuals a user likes and wishes
-to save. After the user finishes interacting with the individual
+A collection is a special folder that stores those phenotypes a user likes and wishes
+to save. After the user finishes interacting with the phenotype
 on the Wall, he can choose to retrieve a new one from the population.
@@
 there share images through their social network.
-When a user hovers the cursor over an individual a pop-up pane shows how many users have
-liked the individual. The pane also includes a link to the individual's 
+When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
+liked the phenotype. The pane also includes a link to the phenotype's 
 details, the parents, genetic operators that created it, and genealogy
@@
   This is the real world system that we are going to represent in the model, 
-  it consist of human users and their interactions with one or more individuals
+  it consist of human users and their interactions with one or more phenotypes
   from
   the population. There are many ways in which humans could interact 
-  with individuals, this could be through a mobile device, a user interface 
+  with phenotypes, this could be through a mobile device, a user interface 
   or by interacting with real world objects generated as part of the algorithm 
@@
 
-\item {\bf Individual} This type of node has the following attributes: 
+\item {\bf Phenotype} This type of node has the following attributes: 
 {\tt id}, {\tt chromosome}  and {\tt views}, which is the number of
-times users have seen or interacted with this individual and, finally, 
+times users have seen or interacted with this phenotype and, finally, 
  {\tt created}, the time and date of creation as above..
@@
 \item {\bf Likes} This relation describes the interaction between a user and
-an individual in which a rating value is assigned this is represented by 
+a phenotype in which a rating value is assigned this is represented by 
 the {\tt rate} attribute. In the current application a user is allowed to rate
-an individual multiple times, so the date of the interactions is also of
+a phenotype multiple times, so the date of the interactions is also of
 interest.
@@
 
-\item {\bf Parent} Describes what individual is the parent of a new
-  individual and is also unique.
+\item {\bf Parent} Describes what phenotype is the parent of a new
+  phenotype and is also unique.
 
 \item {\bf Has} The relation describes an ownership relation between users and
-those collections they own. Also a collection {\tt has}  individuals stored
+those collections they own. Also a collection {\tt has}  phenotypes stored
 in them.  In figure \ref{fig:has} the {\tt has} relationship 
-between certain nodes $user$, $individual$ and $collection$ is shown.
+between certain nodes $user$, $phenotype$ and $collection$ is shown.
 
-\item {\bf Shared} When a user takes an individual from a friend's collection, this interactions 
+\item {\bf Shared} When a user takes an phenotype from a friend's collection, this interactions 
 is described by a {\tt share} relationship between the user and the
-individual.  % Still think that user and individual is confusing - JJ
+phenotype.  % Still think that user and individual is confusing - JJ
+            % Agreed
 
-\item {\bf Saved} When a user takes an individual from the wall interface, this 
+\item {\bf Saved} When a user takes an phenotype from the wall interface, this 
 interactions is described by a {\tt save} relationship. 
@@
 
-The main difference between both versions employing gamification is how the fitness of individuals
+The main difference between both versions employing gamification is how the fitness of phenotypes
 is calculated and the information presented to users.
@@
 \item Start a session.
-\item Rate an individual.
+\item Rate a phenotype.
 \item Create a collection.
-\item Save an individual of the wall to a collection.
-\item Save an individual from a friend's collection.
+\item Save a phenotype of the wall to a collection.
+\item Save a phenotypes from a friend's collection.
 \item Explore collections of other friends.
@@
 The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
-experiment a total of 500 new individuals where generated. The total number of evaluated
-individuals for each volunteer is presented in figure
+experiment a total of 500 new phenotypes where generated. The total number of evaluated
+phenotypes for each volunteer is presented in figure
 \ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ
@@
 with a marginal increase in participation. The increased participation also helped in the creation of
-more individuals for a total of 648. The total number of evaluated
-individuals for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.
+more phenotypes for a total of 648. The total number of evaluated
+phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.
 
@@
 This experiment had the higher number of volunteers for for a total of 68 registered users. 
-In this experiment a total of 3594 new individuals where generated in the experiment. 
-The total number of evaluated individuals for each volunteer is presented in figure 
+In this experiment a total of 3594 new phenotypes where generated in the experiment. 
+The total number of evaluated phenotypes for each volunteer is presented in figure 
 \ref{fig:top-ranked-participation} in green.
@@
         \includegraphics[width=5.5in]{img/comparison.png}
-    \caption{Top ranked number individuals rated by each user using a
+    \caption{Top ranked number phenotypes rated by each user using a
      logarithmic scale.} % Explain a little more and label axes. The
@@
 the population, or the interaction as a whole. The graph could be used also by the evolutionary
-algorithm, for instance when creating a new individual 
+algorithm, for instance when creating a new phenotype 
 parents could be selected only if there is a minimum distance between them in the graph.

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Addressing comments

@@
                                 % Explain a bit more
-                                % This is the idea:
-Fitness assignement could depend on the actions of a group of users, and depend on how
+                                % This is the idea: -M
+Fitness assignment could depend on the actions of a group of users, and depend on how
 they tag, share, rate, or even delete the phenotype. The selection of parents could
@@
 %presenting a natural way of mapping the relationships found in the system. 
-Most IEAs have % have is not very precise and I'm a bit lost
+%Most IEAs have % have is not very precise and I'm a bit lost
                % here. What do you mean? is this constrained to IEAs
                % or EAs? At least the first is not exclusive, right? - JJ
-the following many-to-many relations:
+%the following many-to-many relations:
+When considering the type of relationships found in a genetic algorithm 
+we can think of the Parent-Child relationship between a pair of parent solutions and
+the child solution they produce. In IEAs there could be more entities involved, with 
+even more types of interaction:
 
@@
 
-  \item {\bf Phenotype-Phenotype} The relationships between phenotypes in the
-  population % do you mean different generations? Would be better to
+  \item {\bf Phenotype-Phenotype} The parent-child relationship between phenotypes in the
+  population, as mentioned before.  
+             % do you mean different generations? Would be better to
              % make this more precise - JJ
-are naturally expressed as a tree, or a graph when both ancestors 
-  and descendants are considered. Other semantic relations can also be
+  This relationships are naturally expressed as a tree.
+  %, or a graph when both ancestors 
+  %and descendants are considered. 
+  Other semantic relations can also be
   of interest, 
-  for instance similarity or composition. %This is interesting for
+  for instance the similarity relation between two phenotypes, could be used to measure
+  the population's diversity.   %This is interesting for
                                 %what reason? What do you mean by
                                 %composition? "to be composed of"? - JJ
+                                % Yes, I was thinking about collections as candidate solutions,
+                                % but maybe not yet.
 
@@
     between users and phenotypes, and these can be
-  can be expressed using common actions found in social networks;
-  for instance: share, like, rate or add to collection \cite{Prodromou:16:AS}. In IEAs these
+  expressed using common actions found in social networks;
+  as mentioned before.
+  %for instance: share, like, rate or add to collection 
+  %\cite{Prodromou:16:AS}. This was the Activity Stream referenced before. - Mario
+
+  In IEAs these
   relationships are used to assign a single or a collection of
@@
                                 % changed to phenotypes 
-                                % do you think an aditional explanation is needed?
+                                % do you think an additional explanation is needed? -Mario
 
@@
 \end{itemize}
-
-These relations are naturally mapped to graphs, %They are bipartite
+% Maybe this paragraph is somewhat redundant, 
+% I think your point in the comments below is:
+% so what if they are easily expressed as a network, thats not the main reason.
+% if this is so, agreed. I will remove it- 
+%These relations are naturally mapped to graphs, %They are bipartite
                                 %graphs, but I don't see your point...
 % Besides, _any_ relation can be mapped to a graph - JJ
-while other actions and 
-properties can also be modeled as a semantic network using
-basic subject–-predicate-–object expressions
-\cite{Prud'hommeaux:14:RT}. % that's a graph, too.. - JJ
-Another reason for using a graph model % But for what exactly? A
+%while other actions and 
+%properties can also be modeled as a semantic network using
+%basic subject‚Äì-predicate-‚Äìobject expressions
+%\%cite{Prud'hommeaux:14:RT}. % that's a graph, too.. - JJ
+%Another reason for using a graph model % But for what exactly? A
                                 % bipartite graph to express which
                                 % relations? - JJ
-is the wealth of tools available for
-implementing and query the model, and the increased performance of
-native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.
+%is the wealth of tools available for
+%implementing and query the model, and the increased performance of
+%native graph database systems \cite{holzschuher2013performance,holzschuher2016querying}.
 
@@
 Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
-new designs of robot locomotiom. Collaboration is encoureged by gamifying the system 
-using the maximum distance indicator to inspire the user to try and “beat” previous designs.
+new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
+using the maximum distance indicator to inspire the user to try and Ä'úbeat'Äù previous designs.
 
@@
 employ a Platform as-a Service (PaaS) and NoSQL database to facilitate the collaboration 
-among a large number of users for the interactive modelling of 3D glasess.  This system reduces the user fatigue
+among a large number of users for the interactive modeling of 3D glasses.  This system reduces the user fatigue
 by using a collaborative filtering algorithm to show only the
@@
 techniques imply that we must deal with interactive evolutionary
-system as sociotechnical constructs, where the social aspects are
+system as socio-technical constructs, where the social aspects are
 essential to understand its dynamics. In this sense, conclusions
@@
 
-%Finally, a comparasion of between three deployments of the system, with different increase the collaboration of volunteers.
-
 \section{Experimental setup: the EvoSpace-Interactive framework}
@@
 As a case study, a IEA was developed using the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. %citation
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
 A brief description of the application is presented next, focusing
@@
                                 % to paper's objetives - JJ
+                                % Thanks yes -Mario
 
@@
 
-%%
-%% TO DO: Talk about the Model (in the MVC sense) of this system: PostgreSQL, Redis, 
-%%
-
  \section{IEC Human-centered framework}
-\label{sec:graph}
+\label{sec:HCF}
 The general goal of this research is to develop a human-centered \cite{gasson2003human} framework
@@
 participation and also to minimize the amount of evaluations needed for the
-evolutionary process in given IEC application. We deliberately use the 
+evolutionary process in a given IEC application. We deliberately use the 
 more general
@@
   A data model is used to describe the IEC system prior to a physical 
-  implementation. 
-  %% This block is maybe not needed but in our ES-I implementation
-  %% we use Relational (PosgreSQL), Objects (JSON-Redis) and Graphs (Neo4J)
-  Depending on the domain several techniques can be used.
+  implementation.  Depending on the domain several techniques can be used.
   Parts of the system could be better described by an Entity relationship 
-  modeling to be used in a relational database or a Class Digram for a 
+  modeling to be used in a relational database or a class digram for a 
   key-value store.
-  %%
-  %%
-  A graph is proposed for modeling the social network
-  of users and their interaction with the population. This database 
-  is the back-end of the system, is used by the EA and the user interface.
+
+  A graph is proposed for modeling the social network of users 
+  and their interaction with the population. When implemented a graph database 
+  will be the back-end of the system. This  backend will then be
+  used by the EA and the user interface, in a Model-View-Controller architecture.
 
@@
 
-These components will be described in detail next.
+%These components will be described in detail next.
 % Besides, add something about the innovation in this paper, for
@@
 % - JJ
+
+%Ok, I hope is not to much. Please edit at will - Mario
+
+Besides, this framework is a proposal to view IEC as human centered computing [cite],
+in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive
+limitations, culture, creativity and other human beings aspects are an integral part of the system. 
+This view requires a multidisciplinary approach, as well as multiple knowledge representation formalisms.
+As a first step toward, the traditional data model is extended with a graph, in order to have 
+yet another view to exploit. Following a description of the data model is presented.
 
@@
 
-Next we describe the types of nodes and
-relationships in the system. % I think a bit of more explanation would
+As it was mentioned before, an IEA must be modeled considering the 
+entities involved in the interaction, if we consider the EvoSpace-Interactive framework
+described in section  \ref{sec:evospace-i} we could find the following types of
+nodes and relationships:
+
+%Next we describe the types of nodes and
+%relationships in the system. % I think a bit of more explanation would
                              % help why you have these types and
@@
                              % details or part of the method - JJ
+                             % Explaining above, the graph will depend on the elements of the interaction - Mario
 
@@
 {\tt name}, {\tt created} and {\tt is\_private} indicating if contents
-belonging to this collection can be are shared with others. 
+belonging to this collection can be shared with others. Collections can contain
+phenotypes and a single phenotype could be shared by many collections. % This is why they are not attributes - Mario 
 \end{itemize}
@@
 % Add something at the end linking with the next paragraph
+% Done - Mario
+The interaction between these entities are represented by the following relationships 
+between nodes:
 
@@
 %this is a bipartite graph them. - JJ
+% Do you mean like a figure?  - Mario
 
 \item {\bf Knows} The relation connects two users that know each
-other. This relation is "unique" meaning that it cannot be repeated
-between a pair of nodes.  % Not too clear what you mean by that - JJ
+other. 
+% This relation is "unique" meaning that it cannot be repeated
+% between a pair of nodes.  % Not too clear what you mean by that - JJ
+% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
+ % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
+ % There is even cases where knows relationship only goes in one direction, this could be because
+ % a user does not publishes their friends or one only follows another but are not friends of each other.  
+% Deleting the restriction, I think is not important and can be seen as a graph.
 
 \item {\bf Parent} Describes what phenotype is the parent of a new
-  phenotype and is also unique.
+  phenotype.
 
@@
 phenotype.  % Still think that user and individual is confusing - JJ
-            % Agreed
+            % Agreed corrected -Mario
+            % I was thinking about calling them Solutions or Candidate Solutions,
 
@@
 % single-party graph with only one type of node. 
+% I don't get this comment, yes some relations are bipartite, but If we see the whole
+% graph I think is a multi-graph.   - Mario 
+% I will check the  single-party graph conversion you suggest, I don't know the concept. - Mario
+
 % I don't think "collections" should be nodes too. They are simply
 % attributes for an individual - JJ
+% They are nodes, I added the reason above. - Mario
 For instance a relational database can be used for the implementation of a

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Spell checking

@@
 with the search, using both anonymous and registered users. This
-volunteeer system lowers the requirements of anyone participating in
+volunteer system lowers the requirements of anyone participating in
 the experiment thus increasing the {\em performance}, in terms of {\em
@@
 In this work the use of a graph data model for representing the interactive process of C-IEC systems is proposed.
-Moreover, an proof-of-concept implemantation of an IEA incorporating the graph model is presented.
+Moreover, an proof-of-concept implementation of an IEA incorporating the graph model is presented.
 
@@
 Before launch, each deployment was first tried with a few beta testers. When applying the leader board
-gammification technique for the first time a problem was found in this stage. The problem was that some 
-users were chating by giving a rating to an animation even before it was returned from the server, this was done by just
+gamification technique for the first time a problem was found in this stage. The problem was that some 
+users were cheating by giving a rating to an animation even before it was returned from the server, this was done by just
 constantly clicking the mouse button. This is a common problem found in systems using leader boards because
@@
 
-% Future work could compare the use of gamification with out leaderboards
-%
 The results of each of the three experiments in terms of participation are detailed next.
@@
 \subsubsection{ Base deployment}
-After a week of the anouncement the total number of volunteers for this experiment was 53 registered users. 
+After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
 The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
@@
 As an example we can consider comparing the difference between the number of volunteers in deployments
-B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectivly) additional
+B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) additional
 metrics and patterns are available when they are treated as a

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Starting Results

@@
 % They are nodes, I added the reason above. - Mario
-For instance a relational database can be used for the implementation of a
-user profile and authentication module. Also a copy or cache of the population 
-could be kept in a key-value store in memory for performance reasons. 
-Nevertheless, the model can give important insights of the behavior of the 
-system.   % So what we have done here is this and that because... - JJ
-
+For instance a relational database was used for the implementation of a 
+basic user profile and authentication module. Also the active population 
+was kept in a key-value store in memory for performance reasons.  
+Nevertheless, having a multi-model implementation can give important 
+insights of the behavior of the system. This architecture was validated
+by a case study, requiring the interaction of all of the components.        
+ % So what we have done here is this and that because... - JJ
+ % Changed the paragraph. - Mario 
+% refinement
 \section{Case Study}
@@
 are not reflected in the graph model since they are stored instead in the PostgreSQL relational
-database system. % because... - JJ
+database system. This is because user sessions and authentication, as well as basic
+web requests, are handled directly by the django web framework \cite{garcia2013evospace}.  
+     % because... - JJ
+     % done - Mario
 
@@
 \item Experience. This value depends on the score and is a value 
-between 0 and 100. 
+between 0 and 100. It is assumed for this case, that once a user
+reaches a 100 actions, it has enough experience on using the application.   
 % More details needed? Is done by a Fuzzy Inference System but I think
@@
 % sections above are maybe a bit too verbose - JJ
+% I checked and this is not a fuzzy value. Explained any way. - Mario
 
 \item Participation. This value is simply the degree of the user node 
-in the graph. This measure also considers the number of users he knows,
-perhaps some of the users where invited to the system by him.   
+in the graph. This measure considers the number of users he knows,
+collections and phenotypes rated.    
 \end{itemize}
@@
 %Isn't this anonymous? - JJ
+% like in double blind? I think so, that is why no link is given in the table. - Mario
 
@@
 
-% Time
 \subsection{Results}

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Adds another utility package
And some corrections here and there.

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 the main proposal of this work is developed. The proposed graph model is 
-presented in Section \ref{sec:graph}.
+presented in Section \ref{sec:HCF}.
 The experimental work and results are presented in Section \ref{sec:experiments}.
@@
 % Done - Mario
-The interaction between these entities are represented by the following relationships 
-between nodes:
+The interaction between these entities and how they are represented by relationships 
+between nodes are presented in the next subsection.
 

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Deleting some addressed issues
And conclusion is still missing, so I will wait until it is ready to
work on it. Ping me later when you do it. Maybe start working on the
next paper, too, right?
Comments addressed, going to Figures and Conclusion.
@jj most comments addressed and a few question, please have a look.

@@
 system a {\em social} activity, pursuing the objective of this work,
-which is to increase engagement of the users. % Adding this connecting
-                                % to paper's objetives - JJ
-                                % Thanks yes -Mario
+which is to increase engagement of the users.
 
@@
 belonging to this collection can be shared with others. Collections can contain
-phenotypes and a single phenotype could be shared by many collections. % This is why they are not attributes - Mario 
+phenotypes and a single phenotype could be shared by many
+collections. % This is why they are not attributes - Mario
+% You can simply have an attribute that is a list of the collections
+% it belongs to - JJ
 \end{itemize}
 
-% Add something at the end linking with the next paragraph
-% Done - Mario
 The interaction between these entities and how they are represented by relationships 
@@
 % Do you mean like a figure?  - Mario
+% No, it's a kind of graph - JJ
 
@@
  % a user does not publishes their friends or one only follows another but are not friends of each other.  
-% Deleting the restriction, I think is not important and can be seen as a graph.
+% Deleting the restriction, I think is not important and can be seen
+% as a graph.
+% It's a directed, bipartite graph then - JJ
 
@@
             % Agreed corrected -Mario
-            % I was thinking about calling them Solutions or Candidate Solutions,
+            % I was thinking about calling them Solutions or Candidate
+            % Solutions,
+% Solutions is OK, but also phenotype - JJ
 
@@
 phenotypes for each volunteer is presented in figure
-\ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ
+\ref{fig:top-ranked-participation} in blue. Users are ranked by
+the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+the number of phenotypes evaluated using a logarithmic scale.   % Ranked by... - JJ
 
@@
 more phenotypes for a total of 648. The total number of evaluated
-phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.
+phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.
 
@@
         \includegraphics[width=5.5in]{img/comparison.png}
-    \caption{Top ranked number phenotypes rated by each user using a
-     logarithmic scale.} % Explain a little more and label axes. The
+    \caption{Users are ranked by
+		the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+		the number of phenotypes evaluated using a logarithmic scale. } % Explain a little more and label axes. The
                          % three lines have the same label,
                          % "participation". Not too clear - JJ
+                         % TO DO -Mario
     \label{fig:top-ranked-participation}
@@
 
-When monitoring the experiments, using the graph database and the Cypher query language 
-provided interesting views on the connected data difficult to replicate with a relational database. 
-As an example we can consider comparing the difference between the number of volunteers in deployments
-B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) additional
-metrics and patterns are available when they are treated as a
-network. %Please read the whole sentence and maybe rewrite - JJ
+When monitoring the experiments by querying the graph database,  the Cypher query language 
+provided interesting views on the data that it would have been difficult to replicate with a relational database.
+For instance, to compare the number of volunteers in deployments
+B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively)  a simple query
+provided a view of the social network of volunteers. 
+%Please read the whole sentence and maybe rewrite - JJ
+% Done - Mario
+
 The same can be applied to 
@@
 
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
 \bibliographystyle{IEEEtran}

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Response to comments
Answering some questions
And extending some explanations. Please @mariosky check them out if
you can.

@@
                                 % If not I will add a few words later, such a large topic -Mario
-                                % even with a Journal: http://hcis-journal.springeropen.com/about
+                                % even with a Journal:
+                                % http://hcis-journal.springeropen.com/about
+% a reference here would do no harm - JJ
 giving extensive attention 
@@
 phenotypes have in a traditional EA, even to the extent of using this
-knowledge as an integral part of the evolutionary algorithm. % how?
-                                % Explain a bit more
-                                % This is the idea: -M
-Fitness assignment could depend on the actions of a group of users, and depend on how
+knowledge as an integral part of the evolutionary algorithm.
+Instead of letting the user assign fitness directly via a rating
+or selection system, fitness assignment could depend on the actions of
+a group of users, possibly connected in a social network, and depend on how 
 they tag, share, rate, or even delete the phenotype. The selection of parents could
@@
 We argue also
-%that a graph model is a viable representation for this purpose,  % define better the purpose. Not clear. "giving the same importance to users and their interactions"? - JJ
-% I was thinking to record all the activities and relationships
 that a graph model is a viable representation for the above relationships and
 interactions and can then be used to model and use, within the
-evolutionary algorithm, actions performed by every user. Graph based
+evolutionary algorithm, actions performed by every user, which, as
+mentioned above, will be used to assign fitness to solutions. Graph based
 models are currently used in social networks to keep track  
@@
 Other example is the W3C Activity Streams 2.0 specification used for representing activities 
-common in social web applications \cite{json:streams}..
+common in social web applications \cite{json:streams}.
 
-%presenting a natural way of mapping the relationships found in the system. 
-%Most IEAs have % have is not very precise and I'm a bit lost
-               % here. What do you mean? is this constrained to IEAs
-               % or EAs? At least the first is not exclusive, right? - JJ
-%the following many-to-many relations:
 When considering the type of relationships found in a genetic algorithm 
@@
                                 % changed to phenotypes 
-                                % do you think an additional explanation is needed? -Mario
+                                % do you think an additional
+                                % explanation is needed? -Mario
+  % Maybe explain that phenotypes are the expression of solutions
+  % stored in the population, but do that above - JJ
 
@@
 implementation of an IEA. The experiment was implemented using
-the EvoSpace-Interactive framework, and consisted of a web application
+the EvoSpace-Interactive framework \cite{garcia2013evospace}, and consisted of a web application
 for evolving artistic drawings. The graph database is then used to
@@
 % it belongs to - JJ
+% Right, but:
+%  1) The relationship could have additional atributes, (when it was stored, order, etc.)
+%  2) Then it could be difficult to query with Cypher (I think) - Mario
 \end{itemize}
@@
 % No, it's a kind of graph - JJ
+% I read graph them :)  but is:  graph then , now I get it.
 
@@
 % It's a directed, bipartite graph then - JJ
+% Ok, I will mention the types of graphs. -Mario
 

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Adds citation for human-centered computing

@@
 in figure \ref{fig:uc_framework}:
-
 \begin{itemize}
@@
 
-Besides, this framework is a proposal to view IEC as human centered computing [cite],
+Besides, the methodology presented in this paper for interactive
+evolutionary computation is a proposal to frame it as human centered computing \cite{sebe2010human},
 in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Adding future lines of work
Which link this to other lines of research. You can add more, of
course.

--- 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
@@
 
-Besides, this framework is a proposal to view IEC as human centered computing [cite],
+Besides, the methodology presented in this paper for interactive
+evolutionary computation is a proposal to frame it as human centered computing \cite{sebe2010human},
 in which the context, environment, interfaces, preferences, accesibility, human relations, cognitive
-limitations, culture, creativity and other human beings aspects are an integral part of the system. 
-This view requires a multidisciplinary approach, as well as multiple knowledge representation formalisms.
-As a first step toward, the traditional data model is extended with a graph, in order to have 
-yet another view to exploit. Following a description of the data model is presented.
+limitations, culture, creativity and other human aspects are an integral part of the system. 
+This view requires a multidisciplinary approach, as well as multiple
+knowledge representation formalisms. 
+As a first step forward, the traditional data model is extended with a graph, in order to have 
+yet another view to exploit. In the following section a description of
+the data model is presented.
 
@@
 
-A graph is composed of two fundamental entities: nodes and relationships,
+A graph is composed of two fundamental entities: nodes and edges,
 both having attributes. Nodes are abstractions of the objects in the real 
 world system, and relationships describe their interactions. Types describe
-the attributes describing an entity.
+the attributes describing an entity, and edges express relationships
+between nodes.
 
@@
 %  1) The relationship could have additional atributes, (when it was stored, order, etc.)
-%  2) Then it could be difficult to query with Cypher (I think) - Mario
+%  2) Then it could be difficult to query with Cypher (I think) -
+%  Mario
+% Then it's more of an implementation detail, but anyway, point taken
+% - JJ
 \end{itemize}
@@
 interest.
-%this is a bipartite graph them. - JJ
-% Do you mean like a figure?  - Mario
-% No, it's a kind of graph - JJ
-% I read graph them :)  but is:  graph then , now I get it.
 
@@
 other.
-% This relation is "unique" meaning that it cannot be repeated
-% between a pair of nodes.  % Not too clear what you mean by that - JJ
-% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
- % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
- % There is even cases where knows relationship only goes in one direction, this could be because
- % a user does not publishes their friends or one only follows another but are not friends of each other.  
-% Deleting the restriction, I think is not important and can be seen
-% as a graph.
-% It's a directed, bipartite graph then - JJ
-% Ok, I will mention the types of graphs. -Mario
 
@@
 % graph I think is a multi-graph.   - Mario 
-% I will check the  single-party graph conversion you suggest, I don't know the concept. - Mario
+% I will check the  single-party graph conversion you suggest, I don't
+% know the concept. - Mario
+% Does not matter much... it converts a bipartite graph to another
+% graph with a single kind of nodes, phenotypes or users, for instance
+% - JJ
 
@@
 by a case study, requiring the interaction of all of the components.
- % So what we have done here is this and that because... - JJ
- % Changed the paragraph. - Mario 
-% refinement
+
 \section{Case Study}
@@
 database system. This is because user sessions and authentication, as well as basic
-web requests, are handled directly by the django web framework \cite{garcia2013evospace}.  
-     % because... - JJ
-     % done - Mario
+web requests, are handled directly by the Django web framework \cite{garcia2013evospace}.
 
@@
 \begin{itemize}
-\item Experience. This value depends on the score and is a value 
+\item {\em Experience}: This value depends on the score and is a value 
 between 0 and 100. It is assumed for this case, that once a user
 reaches a 100 actions, it has enough experience on using the application.
-% More details needed? Is done by a Fuzzy Inference System but I think
-% the description would take to much space.
-% Better a small description than nothing. If you need space, some
-% sections above are maybe a bit too verbose - JJ
-% I checked and this is not a fuzzy value. Explained any way. - Mario
 
-\item Participation. This value is simply the degree of the user node 
+\item {\em Participation}: This value is simply the degree of the user node 
 in the graph. This measure considers the number of users he knows,
@@
 \subsubsection{ Non Graph Gamification deployment}
-The total number of volunteers for this experiment was very close to those obtained in the previous
+The total number of volunteers for this experiment was very close to
+the one obtained in the previous
 experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
@@
 When monitoring the experiments by querying the graph database,  the Cypher query language 
-provided interesting views on the data that it would have been difficult to replicate with a relational database.
+provided interesting views on the data that it would have been
+difficult to replicate with a relational database. 
 For instance, to compare the number of volunteers in deployments
@@
 provided a view of the social network of volunteers.
-%Please read the whole sentence and maybe rewrite - JJ
-% Done - Mario
 
@@
 % initial objectives.
+
+One of the interesting future lines of work would be to look a bit
+more closely at the behavior of users rating the system. These initial
+experiments hint at a possible power law, which might indicate that
+the IEC system could be self-organizing, a process that would allow it
+to reach a critical state, as has been found in software repositories,
+for instance \cite{Merelo2016:repomining}. The dynamics of this kind
+of system are fundamentally different, and our future research will
+include exploring these aspects of the system.
 

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Conflicts resolved
Reference added

@@
 engagement, and to tackle the issues mentioned above,  
-IEAs must follow a human centered design, % reference and explain what
+IEAs must follow a human centered design \cite{greenhouse2012human}, % reference and explain what
                                 % you mean by that. It's an EC

### 2017-01-31 2017-GECCO-Modeling-User-Interactions-in-IEC/sample-sigconf.tex
Title

@@
 \begin{document}
-\title{SIG Proceedings Paper in LaTeX Format}
-\titlenote{Produces the permission block, and
-  copyright information}
-\subtitle{Extended Abstract}
-\subtitlenote{The full version of the author's guide is available as
-  \texttt{acmart.pdf} document}
+\title{Modeling Human Interactions in Collaborative Interactive 
+ Evolutionary Computation}
+%\titlenote{Produces the permission block, and
+%  copyright information}
+\subtitle{Anonymized Version}
+%\subtitlenote{The full version of the author's guide is available as
+%  \texttt{acmart.pdf} document}
 
@@
 
-\author{Sean Fogarty}
-\affiliation{%
-  \institution{NASA Ames Research Center}
-  \city{Moffett Field}
-  \state{California} 
-  \postcode{94035}}
-\email{fogartys@amesres.org}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas} 
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P.~Kumquat}
-\affiliation{\institution{The Kumquat Consortium}}
-\email{jpkumquat@consortium.net}
-
-% The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{B. Trovato et. al.}
-
 \begin{abstract}
@@
 <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+<concept_id>10003120.10003130.10003233</concept_id>
+<concept_desc>Human-centered computing~Collaborative and social computing systems and tools</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+<concept_id>10010147.10010178.10010205.10010206</concept_id>
+<concept_desc>Computing methodologies~Heuristic function construction</concept_desc>
 <concept_significance>300</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10010147.10010178.10010205.10010207</concept_id>
+<concept_desc>Computing methodologies~Discrete space search</concept_desc>
+<concept_significance>300</concept_significance>
 </concept>
 <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10010147.10010919.10010172</concept_id>
+<concept_desc>Computing methodologies~Distributed algorithms</concept_desc>
+<concept_significance>300</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
+\ccsdesc[500]{Human-centered computing~Collaborative and social computing systems and tools}
+\ccsdesc[300]{Computing methodologies~Heuristic function construction}
+\ccsdesc[300]{Computing methodologies~Discrete space search}
+\ccsdesc[300]{Computing methodologies~Distributed algorithms}
 % We no longer use \terms command

### 2017-01-31 2017-GECCO-Modeling-User-Interactions-in-IEC/sample-sigconf.tex
bib folder created

@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for
-ACM SIG Proceedings\footnote{This is an abstract footnote}. 
+The necessary intervention of humans in interactive evolutionary
+computational systems has inherent drawbacks arising from the very nature of 
+the algorithms, namely, the human fatigue caused by the interaction, and
+the boredom arising when users evaluate a large number of artifacts 
+many of which are not interesting or are very similar to each other.
+To tackle these issues, in this paper we
+propose a human-centered framework to model 
+the complex interactions on these system.
+Three case studies are presented where the model is applied in the 
+development of evolutionary interactive systems, with different techniques:
+volunteer-collaborative, unplugged and sensor-based. 
+Both conceptual and implementation details are provided where the   
+technique is applied to measure and increase user engagement and participation.
+Our experiments show that the model can be successfully applied in a
+gamification technique developed to increase user engagement, which
+implies that this technique can successfully be used to 
+decrease user fatigue and thus increase {\em performance} of the
+interactive system.
 \end{abstract}
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sigproc}
+\bibliography{sigproc,../bib/biblio}
 

### 2017-01-31 2017-GECCO-Modeling-User-Interactions-in-IEC/samplebody-conf.tex
bib folder created

\section{Introduction}

The \textit{proceedings} are the records of a conference\footnote{This
  is a footnote}.  ACM seeks
to give these conference by-products a uniform, high-quality
appearance.  To do this, ACM has some rigid requirements for the
format of the proceedings documents: there is a specified format
(balanced double columns), a specified set of fonts (Arial or
Helvetica and Times Roman) in certain specified sizes, a specified
live area, centered on the page, specified size of margins, specified
column width and gutter size.

\section{The Body of The Paper}
\cite{dawkins1986blind}
Typically, the body of a paper is organized into a hierarchical
structure, with numbered or unnumbered headings for sections,
subsections, sub-subsections, and even smaller sections.  The command
\texttt{{\char'134}section} that precedes this paragraph is part of
such a hierarchy.\footnote{This is a footnote.} \LaTeX\ handles the
numbering and placement of these headings for you, when you use the
appropriate heading commands around the titles of the headings.  If
you want a sub-subsection or smaller part to be unnumbered in your
output, simply append an asterisk to the command name.  Examples of
both numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in the \textbf{document}
environment, you can indicate the start of a new paragraph with a
blank line in your input file; that is why this sentence forms a
separate paragraph.

\subsection{Type Changes and {\itshape Special} Characters}

We have already seen several typeface changes in this sample.  You can
indicate italicized words or phrases in your text with the command
\texttt{{\char'134}textit}; emboldening with the command
\texttt{{\char'134}textbf} and typewriter-style (for instance, for
computer code) with \texttt{{\char'134}texttt}.  But remember, you do
not have to indicate typestyle changes when such changes are part of
the \textit{structural} elements of your article; for instance, the
heading of this subsection will be in a sans serif\footnote{Another
  footnote, here.  Let's make this a rather short one to see how it
  looks.} typeface, but that is handled by the document class file.
Take care with the use of\footnote{A third, and last, footnote.}  the
curly braces in typeface changes; they mark the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or non-English
characters you need anywhere in your document; you can find a complete
list of what is available in the \textit{\LaTeX\ User's Guide}
\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin\,\ldots{\char'134}end}
construction or with the short form \texttt{\$\,\ldots\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX\@; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles~\cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings~\cite{clark:pct} or maybe
books \cite{Lamport:LaTeX, salas:calculus} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file~\cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}~\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
are found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed
output of this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more desirable.
Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

It is strongly recommended to use the package booktabs~\cite{Fear05}
and follow its main principles of typography with respect to tables:
\begin{enumerate}
\item Never, ever use vertical rules.
\item Never use double rules.
\end{enumerate}
It is also a good idea not to overuse horizontal rules.

\subsection{Figures}

Like tables, figures cannot be split across pages; the best placement
for them is typically the top or the bottom of the page nearest their
initial cite.  To ensure this proper ``floating'' placement of
figures, use the environment \textbf{figure} to enclose the figure and
its caption.

This sample document contains examples of \texttt{.eps} files to be
displayable with \LaTeX.  If you work with pdf\LaTeX, use files in the
\texttt{.pdf} format.  Note that most modern \TeX\ systems will convert
\texttt{.eps} to \texttt{.pdf} for you on the fly.  More details on
each of these are found in the \textit{Author's Guide}.

\begin{figure}
\includegraphics{fly}
\caption{A sample black and white graphic.}
\end{figure}

\begin{figure}
\includegraphics[height=1in, width=1in]{fly}
\caption{A sample black and white graphic
that has been resized with the \texttt{includegraphics} command.}
\end{figure}

As was the case with tables, you may want a figure that spans two
columns.  To do this, and still to ensure proper ``floating''
placement of tables, use the environment \textbf{figure*} to enclose
the figure and its caption.  And don't forget to end the environment
with \textbf{figure*}, not \textbf{figure}!

\begin{figure*}
\includegraphics{flies}
\caption{A sample black and white graphic
that needs to span two columns of text.}
\end{figure*}

\begin{figure}
\includegraphics[height=1in, width=1in]{rosette}
\caption{A sample black and white graphic that has
been resized with the \texttt{includegraphics} command.}
\end{figure}

\subsection{Theorem-like Constructs}

Other common constructs that may occur in your article are the forms
for logical constructs like theorems, axioms, corollaries and proofs.
ACM uses two types of these constructs:  theorem-like and
definition-like.

Here is a theorem:
\begin{theorem}
  Let $f$ be continuous on $[a,b]$.  If $G$ is
  an antiderivative for $f$ on $[a,b]$, then
  \begin{displaymath}
    \int^b_af(t)\,dt = G(b) - G(a).
  \end{displaymath}
\end{theorem}

Here is a definition:
\begin{definition}
  If $z$ is irrational, then by $e^z$ we mean the
  unique number that has
  logarithm $z$:
  \begin{displaymath}
    \log e^z = z.
  \end{displaymath}
\end{definition}

The pre-defined theorem-like constructs are \textbf{theorem},
\textbf{conjecture}, \textbf{proposition}, \textbf{lemma} and
\textbf{corollary}.  The pre-defined de\-fi\-ni\-ti\-on-like constructs are
\textbf{example} and \textbf{definition}.  You can add your own
constructs using the \textsl{amsthm} interface~\cite{Amsthm15}.  The
styles used in the \verb|\theoremstyle| command are \textbf{acmplain}
and \textbf{acmdefinition}.

Another construct is \textbf{proof}, for example,

\begin{proof}
  Suppose on the contrary there exists a real number $L$ such that
  \begin{displaymath}
    \lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
  \end{displaymath}
  Then
  \begin{displaymath}
    l=\lim_{x\rightarrow c} f(x)
    = \lim_{x\rightarrow c}
    \left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
    = \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
    \frac{f(x)}{g(x)} = 0\cdot L = 0,
  \end{displaymath}
  which contradicts our assumption that $l\neq 0$.
\end{proof}

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e., the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{References}
Generated by bibtex from your \texttt{.bib} file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the \texttt{.bbl} file.  Insert that \texttt{.bbl}
file into the \texttt{.tex} source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}

Of course, reading the source code is always useful.  The file
\path{acmart.pdf} contains both the user guide and the commented
code.

\begin{acks}
  The authors would like to thank Dr. Yuhua Li for providing the
  matlab code of  the \textit{BEPS} method.

  The authors would also like to thank the anonymous referees for
  their valuable comments and helpful suggestions. The work is
  supported by the \grantsponsor{GS501100001809}{National Natural
    Science Foundation of
    China}{http://dx.doi.org/10.13039/501100001809} under Grant
  No.:~\grantnum{GS501100001809}{61273304}
  and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
    Scientsts' Support Program}.

\end{acks}

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Layout

--- 2017-01-31 2017-GECCO-Modeling-User-Interactions-in-IEC/samplebody-conf.tex
@@
 \section{Introduction}
+% C-IEAs
+  % Problems
+  % Volunteers
+% Human Based Computation 
+% Human Centered
+  % User -> Human
+  % Social Network
+  % Devices IoT
+  % Activities
+  % Engagement
+% C-IEA Complex Interactions
+% Why an HC Framework (Objectives)  
+% Presentation
 
-The \textit{proceedings} are the records of a conference\footnote{This
-  is a footnote}.  ACM seeks
-to give these conference by-products a uniform, high-quality
-appearance.  To do this, ACM has some rigid requirements for the
-format of the proceedings documents: there is a specified format
-(balanced double columns), a specified set of fonts (Arial or
-Helvetica and Times Roman) in certain specified sizes, a specified
-live area, centered on the page, specified size of margins, specified
-column width and gutter size.
+\section{Related Work}
+% C-IEAS
+% Volunteer Based
+% HBC
+% Engagement Techniques
+% Models
 
-\section{The Body of The Paper}
-\cite{dawkins1986blind}
-Typically, the body of a paper is organized into a hierarchical
-structure, with numbered or unnumbered headings for sections,
-subsections, sub-subsections, and even smaller sections.  The command
-\texttt{{\char'134}section} that precedes this paragraph is part of
-such a hierarchy.\footnote{This is a footnote.} \LaTeX\ handles the
-numbering and placement of these headings for you, when you use the
-appropriate heading commands around the titles of the headings.  If
-you want a sub-subsection or smaller part to be unnumbered in your
-output, simply append an asterisk to the command name.  Examples of
-both numbered and unnumbered headings will appear throughout the
-balance of this sample document.
+\section{HC  IEC Framework}
+  \subsection{Real World}
+  \subsection{Data Model}
+  \subsection{Evolutionary Algorithm}
 
-Because the entire article is contained in the \textbf{document}
-environment, you can indicate the start of a new paragraph with a
-blank line in your input file; that is why this sentence forms a
-separate paragraph.
+\section{EvoSpace-Interactive}
+  \subsection{Interfaces}
+  \subsection{Data Models}
+    \subsubsection{EvoSpace}
+    \subsubsection{Graph}
+    \subsubsection{PostgreSQL}
+  \subsection{Evolution}
 
-\subsection{Type Changes and {\itshape Special} Characters}
+\section{Case Studies}
+\subsection{EvoDraw}
+\subsection{EvoDraw-Kinect}
+\subsection{XYZ-Unplugged}
 
-We have already seen several typeface changes in this sample.  You can
-indicate italicized words or phrases in your text with the command
-\texttt{{\char'134}textit}; emboldening with the command
-\texttt{{\char'134}textbf} and typewriter-style (for instance, for
-computer code) with \texttt{{\char'134}texttt}.  But remember, you do
-not have to indicate typestyle changes when such changes are part of
-the \textit{structural} elements of your article; for instance, the
-heading of this subsection will be in a sans serif\footnote{Another
-  footnote, here.  Let's make this a rather short one to see how it
-  looks.} typeface, but that is handled by the document class file.
-Take care with the use of\footnote{A third, and last, footnote.}  the
-curly braces in typeface changes; they mark the beginning and end of
-the text that is to be in the different typeface.
-
-You can use whatever symbols, accented characters, or non-English
-characters you need anywhere in your document; you can find a complete
-list of what is available in the \textit{\LaTeX\ User's Guide}
-\cite{Lamport:LaTeX}.
-
-\subsection{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
-
-\subsubsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{{\char'134}begin\,\ldots{\char'134}end}
-construction or with the short form \texttt{\$\,\ldots\$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation:
-\begin{math}
-  \lim_{n\rightarrow \infty}x=0
-\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsubsection{Display Equations}
-A numbered display equation---one set off by vertical space from the
-text and centered horizontally---is produced by the \textbf{equation}
-environment. An unnumbered display equation is produced by the
-\textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX\@; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-\begin{equation}
-  \lim_{n\rightarrow \infty}x=0
-\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}
-  \sum_{i=0}^{\infty} x + 1
-\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}
-  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
-\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\subsection{Citations}
-Citations to articles~\cite{bowman:reasoning,
-clark:pct, braams:babel, herlihy:methodology},
-conference proceedings~\cite{clark:pct} or maybe
-books \cite{Lamport:LaTeX, salas:calculus} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file~\cite{Lamport:LaTeX}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
-
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}~\cite{Lamport:LaTeX}.
-
-This article shows only the plainest form
-of the citation command, using \texttt{{\char'134}cite}.
-
-\subsection{Tables}
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-are found in the \textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table~\ref{tab:freq} is included in the input file; compare the
-placement of the table here with the table in the printed
-output of this document.
-
-\begin{table}
-  \caption{Frequency of Special Characters}
-  \label{tab:freq}
-  \begin{tabular}{ccl}
-    \toprule
-    Non-English or Math&Frequency&Comments\\
-    \midrule
-    \O & 1 in 1,000& For Swedish names\\
-    $\pi$ & 1 in 5& Common in math\\
-    \$ & 4 in 5 & Used in business\\
-    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-  \bottomrule
-\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of the page's
-live area, use the environment \textbf{table*} to enclose the table's
-contents and the table caption.  As with a single-column table, this
-wide table will ``float'' to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table~\ref{tab:commands} is included in the input file; again, it is
-instructive to compare the placement of the table here with the table
-in the printed output of this document.
-
-\begin{table*}
-  \caption{Some Typical Commands}
-  \label{tab:commands}
-  \begin{tabular}{ccl}
-    \toprule
-    Command &A Number & Comments\\
-    \midrule
-    \texttt{{\char'134}author} & 100& Author \\
-    \texttt{{\char'134}table}& 300 & For tables\\
-    \texttt{{\char'134}table*}& 400& For wider tables\\
-    \bottomrule
-  \end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-It is strongly recommended to use the package booktabs~\cite{Fear05}
-and follow its main principles of typography with respect to tables:
-\begin{enumerate}
-\item Never, ever use vertical rules.
-\item Never use double rules.
-\end{enumerate}
-It is also a good idea not to overuse horizontal rules.
-
-\subsection{Figures}
-
-Like tables, figures cannot be split across pages; the best placement
-for them is typically the top or the bottom of the page nearest their
-initial cite.  To ensure this proper ``floating'' placement of
-figures, use the environment \textbf{figure} to enclose the figure and
-its caption.
-
-This sample document contains examples of \texttt{.eps} files to be
-displayable with \LaTeX.  If you work with pdf\LaTeX, use files in the
-\texttt{.pdf} format.  Note that most modern \TeX\ systems will convert
-\texttt{.eps} to \texttt{.pdf} for you on the fly.  More details on
-each of these are found in the \textit{Author's Guide}.
-
-\begin{figure}
-\includegraphics{fly}
-\caption{A sample black and white graphic.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[height=1in, width=1in]{fly}
-\caption{A sample black and white graphic
-that has been resized with the \texttt{includegraphics} command.}
-\end{figure}
-
-As was the case with tables, you may want a figure that spans two
-columns.  To do this, and still to ensure proper ``floating''
-placement of tables, use the environment \textbf{figure*} to enclose
-the figure and its caption.  And don't forget to end the environment
-with \textbf{figure*}, not \textbf{figure}!
-
-\begin{figure*}
-\includegraphics{flies}
-\caption{A sample black and white graphic
-that needs to span two columns of text.}
-\end{figure*}
-
-\begin{figure}
-\includegraphics[height=1in, width=1in]{rosette}
-\caption{A sample black and white graphic that has
-been resized with the \texttt{includegraphics} command.}
-\end{figure}
-
-\subsection{Theorem-like Constructs}
-
-Other common constructs that may occur in your article are the forms
-for logical constructs like theorems, axioms, corollaries and proofs.
-ACM uses two types of these constructs:  theorem-like and
-definition-like.
-
-Here is a theorem:
-\begin{theorem}
-  Let $f$ be continuous on $[a,b]$.  If $G$ is
-  an antiderivative for $f$ on $[a,b]$, then
-  \begin{displaymath}
-    \int^b_af(t)\,dt = G(b) - G(a).
-  \end{displaymath}
-\end{theorem}
-
-Here is a definition:
-\begin{definition}
-  If $z$ is irrational, then by $e^z$ we mean the
-  unique number that has
-  logarithm $z$:
-  \begin{displaymath}
-    \log e^z = z.
-  \end{displaymath}
-\end{definition}
-
-The pre-defined theorem-like constructs are \textbf{theorem},
-\textbf{conjecture}, \textbf{proposition}, \textbf{lemma} and
-\textbf{corollary}.  The pre-defined de\-fi\-ni\-ti\-on-like constructs are
-\textbf{example} and \textbf{definition}.  You can add your own
-constructs using the \textsl{amsthm} interface~\cite{Amsthm15}.  The
-styles used in the \verb|\theoremstyle| command are \textbf{acmplain}
-and \textbf{acmdefinition}.
-
-Another construct is \textbf{proof}, for example,
-
-\begin{proof}
-  Suppose on the contrary there exists a real number $L$ such that
-  \begin{displaymath}
-    \lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-  \end{displaymath}
-  Then
-  \begin{displaymath}
-    l=\lim_{x\rightarrow c} f(x)
-    = \lim_{x\rightarrow c}
-    \left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-    = \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-    \frac{f(x)}{g(x)} = 0\cdot L = 0,
-  \end{displaymath}
-  which contradicts our assumption that $l\neq 0$.
-\end{proof}
-
+\section{Discussion}
 \section{Conclusions}
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
-%\end{document}  % This is where a 'short' article might terminate
-
-\appendix
-%Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e., the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{References}
-Generated by bibtex from your \texttt{.bib} file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the \texttt{.bbl} file.  Insert that \texttt{.bbl}
-file into the \texttt{.tex} source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-
-Of course, reading the source code is always useful.  The file
-\path{acmart.pdf} contains both the user guide and the commented
-code.
 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Intro

@@
 \section{Introduction}
+% Human Based Computation
 % C-IEAs
+% Volunteers 
 % Problems
-  % Volunteers
-% Human Based Computation 
 % Human Centered
@@
 % Presentation 
+Interactive evolutionary computation (IEC) systems are, in general, methods
+of evolutionary computation whose fitness evaluations are performed by humans 
+within an interactive system \cite{eiben2015interactive}.  Human evaluation 
+is needed because the fitness function is not known and the result
+of optimization should fit certain human need. Usually human preference is
+needed to evaluate subjective characteristics like visual appeal or 
+attractiveness \cite{biomorphs}, but in other cases human behavior is 
+considered, for instance to optimize teamwork \cite{kosorukoff2002evolutionary}
+or creativity cite[cook]. When human interaction is responsible of other 
+operations of the evolutionary process some authors classify these IEC methods 
+as human-based evolutionary computation cite[kosorukoff] or as human-based computation cite[].
+IEC systems have demonstrated their ability for effectively
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
+as well as other artifacts in many other domains \cite{ie1}. 
+In order to reach more human users, Interactive Evolutionary Algorithms (IEAs) are
+some times developed as web applications,
+who depend on volunteers who visit the system to interactively help
+with the search, using both anonymous and registered users. Some systems 
+employ a collaborative technique, where several users assign a quality assessment 
+to a single solution and then an aggregated fitness value is calculated 
+\cite{picbreeder,seyama2016development,wagy2014collective}.
+However, the necessary intervention of humans in these systems leads
+to them having several inherent problems arising from the very nature of 
+the algorithms, namely, human evaluations are slow and expensive, there is a
+human fatigue caused by the interaction \cite{ie1}, and
+the boredom arising when users evaluate a large number of phenotypes 
+many of which are not interesting or are very similar to each other. 
+The number of evaluations that IEC can receive is limited by the number of users
+collaborating and the amount of interaction of each. Having a volunteer system 
+can lower the requirements of anyone participating in
+the experiment thus increasing the {\em performance}, in terms of {\em
+human brain cycles}, of the whole system. Using a volunteer based system  
+brings other issues \cite{sarmenta2001volunteer,web:BOINC}, 
+such as the volunteer\'s lack of accountability,
+and the need to build trust between participants and application providers. 
+For the provider there is also the difficulty of establishing 
+the amount of time and resources
+a volunteer is willing to spend on the system, and how they decide if they
+participate or not \cite{JJ:2016}.
+
+In order to increase volunteer participation and to tackle some of the issues mentioned above,  
+IEAs must developed following a human centered design \cite{greenhouse2012human},
+giving extensive attention to volunteers, not only because their
+explicit evaluation is essential, but also because the context of the 
+interaction affects the system as a whole. In collaborative systems, users interact 
+with each other, and are aware of the actions. Users also consider their previous evaluations
+and past experiences.
+
+%Please help with this section, 
+Evolutionary computations algorithms are
+
+In this work we propose giving the same
+importance to users and their interactions as the population of 
+individuals have in a traditional EA, even to the extent of using this
+knowledge as an integral part of the evolutionary algorithm.
+Instead of letting the user assign fitness directly via a rating
+or selection system, fitness assignment could depend on the actions of
+a group of users, possibly connected in a social network, and depend on how 
+they tag, share, rate, or even delete the phenotype. The selection of parents could
+also depend on the previous actions, leveraging information such as the fact that  they have the same tag, or shared by
+similar users, or whether a leader of opinion liked both or if they are stored in the same collection.
+That knowledge could also be applied in choosing which phenotypes would be presented to a
+user, in order to give them those they will find more interesting or useful.
 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
paper-name
Intro

--- 2017-01-31 2017-GECCO-Modeling-User-Interactions-in-IEC/sample-sigconf.tex
@@
 <concept_id>10010147.10010178.10010205.10010207</concept_id>
-<concept_desc>Computing methodologies~Discrete space search</concept_desc>
+<concept_desc>Computing methodologies~Bio Inspired Algorithms</concept_desc>
 <concept_significance>300</concept_significance>
@@
 \ccsdesc[300]{Computing methodologies~Heuristic function construction}
-\ccsdesc[300]{Computing methodologies~Discrete space search}
 \ccsdesc[300]{Computing methodologies~Distributed algorithms}
@@
 
-\input{samplebody-conf}
+\input{paper-body}
 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Intro draft
@JJ Please have a look, I tried to highlight the differences with the
CEC paper. Focus is on the model and other applications.

--- 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 \section{Introduction}
+% Main Ideas:
+
+% IEAs
+% Human Based Computation
 % C-IEAs
@@
 % Volunteers 
-% Human Based Computation 
 % Human Centered
+% C-IEA Complex Interactions
   % User -> Human
@@
   % Engagement
-% C-IEA Complex Interactions
+
 % Why an HC Framework (Objectives)  
 % Presentation
+
+% IEAs
+Interactive evolutionary computation (IEC) systems are, in general, methods
+of evolutionary computation whose fitness evaluations are performed by humans 
+within an interactive system \cite{eiben2015interactive}.  Human evaluation 
+is needed because the fitness function is not known and the result
+of optimization should fit certain human need. Usually human preference is
+needed to evaluate subjective characteristics like visual appeal or 
+attractiveness \cite{biomorphs}, but in other cases human behavior is 
+considered, for instance to optimize teamwork \cite{kosorukoff2002evolutionary}
+or creativity cite[cook].
+% Human Based Computation
+When human interaction is responsible of other 
+operations of the evolutionary process some authors classify these IEC methods 
+as human-based evolutionary computation cite[kosorukoff] or as human-based computation cite[].
+IEC systems have demonstrated their ability for effectively
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
+as well as other artifacts in many other domains \cite{ie1}. 
+% C-IEAs
+In order to reach more human users, Interactive Evolutionary Algorithms (IEAs) are
+some times developed as web applications,
+who depend on volunteers who visit the system to interactively help
+with the search, using both anonymous and registered users. Some systems 
+employ a collaborative technique, where several users assign a quality assessment 
+to a single solution and then an aggregated fitness value is calculated 
+\cite{picbreeder,seyama2016development,wagy2014collective}.
+% Problems
+However, the necessary intervention of humans in these systems leads
+to them having several inherent problems arising from the very nature of 
+the algorithms, namely, human evaluations are slow and expensive, there is a
+human fatigue caused by the interaction \cite{ie1}, and
+the boredom arising when users evaluate a large number of phenotypes 
+many of which are not interesting or are very similar to each other. 
+The number of evaluations that IEC can receive is limited by the number of users
+collaborating and the amount of interaction of each. Having a volunteer system 
+can lower the requirements of anyone participating in
+the experiment thus increasing the {\em performance}, in terms of {\em
+human brain cycles}, of the whole system.
+% Volunteers 
+Using a volunteer based system  
+brings other issues \cite{sarmenta2001volunteer,web:BOINC}, 
+such as the volunteer\'s lack of accountability,
+and the need to build trust between participants and application providers. 
+For the provider there is also the difficulty of establishing 
+the amount of time and resources
+a volunteer is willing to spend on the system, and how they decide if they
+participate or not \cite{JJ:2016}. 
+% Human Centered
+In order to increase volunteer participation and to tackle some of the issues mentioned above,  
+we proposed a framework following a human centered design \cite{greenhouse2012human},
+giving extensive attention to volunteers, not only because their
+explicit evaluation is essential, but also because the context of the 
+interaction affects the system as a whole. In collaborative systems, users interact 
+with each other, and are aware of the actions ref[]. Users also consider their previous evaluations
+and past experiences ref[].
+%Please help with this paragraph, is evidence needed? 
+In evolutionary computation algorithms are data driven, as it is important to keep data about the 
+population of each generation, in order to generate the next or even change certain parameters
+or analyze the evolutionary process. 
+% Human Centered Importance
+The framework models the context of interaction along with the population and
+this knowledge is available to be an integral part of the evolutionary computation.
+%This could be generalized?
+For example, in a C-IEC application, fitness assignment depend on the actions of a social network of users. 
+A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. The selection of parents could
+also depend also on the previous actions, leveraging information such as the fact that they have the same tag, or are shared by
+similar users. That knowledge could also be applied in choosing which phenotypes present to a certain
+users, in order to give them those they will find more interesting or useful. 
+% IoT
+The context of interaction, can also include data retrieved from sensors and other IoT devices.
+% Engagement
+Data available from the interaction is also used to increase the performance of the system by applying 
+gamification techniques. Gamification is defined by Deterding et al. as
+``the use of game design elements in non-game contexts'' \cite{deterding2011game}.
+The gamification element employed in a case study is a rewarding mechanism  
+\cite{dubois2013understanding}, in general they consist of a reputation system with score points, 
+levels and leader boards. Points are awarded to users in response of 
+the accomplishment of certain activities that need to be encouraged. Levels depend
+on the score and certain features of the game are only available to gamers when 
+they reach a giving level.
+
+Three case studies are presented as a proof of concept, providing both
+conceptual and implementation details of the framework. 
+The experiments were implemented using the EvoSpace-Interactive framework \cite{garcia2013evospace}, 
+and where employed in different contexts of interaction. 
+Additional details of each case study application is presented elsewhere refs[][][]. 
+Giving this paper the objective of presenting the underlying conceptual framework.
+
+The remainder of this paper is organized as follows.
+Section \ref{sec:interactive} presents related work on the topic 
+of Interactive Evolution.
+Then, Section \ref{sec:e} presents the Model of Human Interactions in Collaborative Interactive 
+ Evolutionary Computation which 
+the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:HCF}.
+The case studies  and results are presented in Section \ref{sec:experiments}.
+Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
 
@@
 
-\section{Discussion}
 \section{Conclusions}

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Slight revision with some comments

@@
 % IEAs
-Interactive evolutionary computation (IEC) systems are, in general, methods
-of evolutionary computation whose fitness evaluations are performed by humans 
-within an interactive system \cite{eiben2015interactive}.  Human evaluation 
-is needed because the fitness function is not known and the result
-of optimization should fit certain human need. Usually human preference is
-needed to evaluate subjective characteristics like visual appeal or 
-attractiveness \cite{biomorphs}, but in other cases human behavior is 
+Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
+whose fitness evaluations are performed by humans through an
+interactive system \cite{eiben2015interactive}. They are usually
+applied in problems where the fitness function is not known  or simply
+does not exist, and the result
+of optimization should fit certain human and possibly aesthetic
+need. That is why their use case includes the evolution of objects
+with subjective characteristics such visual appeal or 
+attractiveness \cite{biomorphs} as well as others where human behavior is 
 considered, for instance to optimize teamwork \cite{kosorukoff2002evolutionary}
@@
 % Volunteers 
-Using a volunteer based system  
-brings other issues \cite{sarmenta2001volunteer,web:BOINC}, 
+Using a volunteer based system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, 
 such as the volunteer\'s lack of accountability,
@@
 %Please help with this paragraph, is evidence needed?
-In evolutionary computation algorithms are data driven, as it is important to keep data about the 
+% It's the tuning of the algorithm which is data driven, not the
+% algorithm itself, right? - JJ
+In evolutionary computation algorithms are data driven since it is important to keep data about the 
 population of each generation, in order to generate the next or even change certain parameters
@@
 %This could be generalized?
-For example, in a C-IEC application, fitness assignment depend on the actions of a social network of users. 
+For example, in a C-IEC application, fitness assignment depend on the
+actions of a social network of users. 
 A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. The selection of parents could
@@
 
-The remainder of this paper is organized as follows.
+The rest of this paper is organized as follows.
 Section \ref{sec:interactive} presents related work on the topic 
@@
 \begin{acks}
-  The authors would like to thank Dr. Yuhua Li for providing the
-  matlab code of  the \textit{BEPS} method.
+  % The authors would like to thank Dr. Yuhua Li for providing the
+  % matlab code of  the \textit{BEPS} method.
 
-  The authors would also like to thank the anonymous referees for
-  their valuable comments and helpful suggestions. The work is
-  supported by the \grantsponsor{GS501100001809}{National Natural
-    Science Foundation of
-    China}{http://dx.doi.org/10.13039/501100001809} under Grant
-  No.:~\grantnum{GS501100001809}{61273304}
-  and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
-    Scientsts' Support Program}.
+  % The authors would also like to thank the anonymous referees for
+  % their valuable comments and helpful suggestions. The work is
+  % supported by the \grantsponsor{GS501100001809}{National Natural
+  %   Science Foundation of
+  %   China}{http://dx.doi.org/10.13039/501100001809} under Grant
+  % No.:~\grantnum{GS501100001809}{61273304}
+  % and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
+  %   Scientsts' Support Program}.
 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Splitting in paragraphs
And doing a bit of more revision

@@
 % Human Based Computation
-When human interaction is responsible of other 
+These IEC systems are an interesting venue of research, since they have demonstrated their ability for effectively 
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
+as well as other artifacts in many other domains \cite{ie1}. In these cases when human interaction is responsible of other 
 operations of the evolutionary process some authors classify these IEC methods 
 as human-based evolutionary computation cite[kosorukoff] or as human-based computation cite[].
-IEC systems have demonstrated their ability for effectively
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-as well as other artifacts in many other domains \cite{ie1}. 
+
 % C-IEAs
-In order to reach more human users, Interactive Evolutionary Algorithms (IEAs) are
-some times developed as web applications,
-who depend on volunteers who visit the system to interactively help
+One of the issues with these systems is that their performance effectively depends on the number of users
+they are able to include; and in order to reach more human users,
+Interactive Evolutionary Algorithms (IEAs) are 
+some times developed as web applications depending on visitors to the web to interactively help
 with the search, using both anonymous and registered users. Some systems 
@@
 % Volunteers
+
 Using a volunteer based system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, 
 such as the volunteer\'s lack of accountability,
-and the need to build trust between participants and application providers. 
-For the provider there is also the difficulty of establishing 
+and the need to build trust between participants and application
+providers. Issues for the provider of the system or organizer  are also the difficulty of establishing 
 the amount of time and resources
@@
 participate or not \cite{JJ:2016}.
+
 % Human Centered
@@
 and past experiences ref[].
+
 %Please help with this paragraph, is evidence needed?
@@
 % algorithm itself, right? - JJ
+% And evidence for what? - jj
 In evolutionary computation algorithms are data driven since it is important to keep data about the 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Minor changes
Defining framework

@@
 % algorithm itself, right? - JJ
+%  Tuning yes - Mario
 % And evidence for what? - jj
+%  That most EAs use only population data - Mario
 In evolutionary computation algorithms are data driven since it is important to keep data about the 
@@
 actions of a social network of users. 
-A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. The selection of parents could
-also depend also on the previous actions, leveraging information such as the fact that they have the same tag, or are shared by
+A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. 
+The selection of parents could also depend also on the previous actions, leveraging information 
+such as the fact that they have the same tag, or are shared by
 similar users. That knowledge could also be applied in choosing which phenotypes present to a certain
@@
 \section{HC  IEC Framework}
+The general goal of this research is to develop a human-centered \cite{gasson2003human} 
+software framework for interactive evolutionary computation (IEC). 
+The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}.
+We deliberately use the more general term Human instead of User, because humans are not always playing the role of
+users in software engineering sense, they could just be interacting 
+naturally with their environment and with out knowing they are implicitly
+evaluating a possible solution.
+
+% Comment: 
+% Maybe in this case we should use the term software framework or architecture. 
+% is a reusable design together with an implementation
+
+The framework
+
+in order to increase
+participation and also to minimize the amount of evaluations needed for the
+evolutionary process in a given IEC application.
+
   \subsection{Real World}

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
More framework description

--- 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 \section{HC  IEC Framework}
+% Comment: 
+% Maybe in this case we should use the term software framework or architecture. 
+The general goal of this research is to develop a human-centered \cite{gasson2003human} 
+software framework for interactive evolutionary computation (IEC). 
+The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
+providing generalized components to developers of IEC systems.
+In the description we deliberately use the more general term Human instead of User, because humans are not always playing the role of users in software engineering sense, they could just be interacting naturally with their environment and with out knowing they are implicitly
+evaluating a possible solution. The framework includes components that can be refined to increase
+participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
+
   \subsection{Real World}

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Opinionated framework
More
Users

@@
 The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
-providing generalized components to developers of IEC systems.
-In the description we deliberately use the more general term Human instead of User, because humans are not always playing the role of users in software engineering sense, they could just be interacting naturally with their environment and with out knowing they are implicitly
+providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
+term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
+could just be interacting naturally with their environment and with out knowing they are implicitly
 evaluating a possible solution. The framework includes components that can be refined to increase
 participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
+Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
+design. Before entering to details the main design decisions of the framework are explained next:
+
+\begin{itemize}
+\item {\bf Users are Human} 
+  The user will have certain goals and has the intention to perform certain tasks.
+  The framework follows the approach of human centered computing \cite{sebe2010human},
+  in which the context, environment, interfaces, preferences, accessibility, human relations, cognitive
+  limitations, culture, creativity and other human aspects are an integral part of the system.
+
+\item {\bf Users are not alone}
+  Relationship between users in an interactive evolutionary algorithm can also be modeled 
+  as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
+  A graph model could enable researchers to find other ways of identifying leaders of 
+  opinion or measuring the similarity between user's preferences. 
+  These measures can then be used by recommender algorithms selecting 
+  phenotypes according users' preferences.
+
+\item {\bf Context of use matters}
+ The product will  be used within a certain range of technical, physical and social or
+ organizational environments \cite{maguire2001context} that may affect its use.
+ We can refer to these environments as the \textit{context of use} (see
+ Figure \ref{fig:logicalmodel}), this concept has been formally defined
+ by ISO standard 9241-11 \cite{international1998iso} as \textit{``users,
+ tasks and equipment (hardware, software and materials), and the
+ physical and social environments in which a product is used"}. \\ 
+ In daily life we often find products that are difficult to use or
+ understand. This type of difficulties are \textit{usability problems}
+ that arise from  diverse issues that have not been addressed in the
+ design of the product's \textit{user-artifact interaction}.
+
+\item {\bf Interaction is a stream of actions}
+
+\item {\bf Users are volunteers}
+
+\item {\bf Solutions could be: real objects, processes, digital artifacts}
+
+\end{itemize}
 

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Frame work

--- 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
-providing generalized components to developers of IEC systems.
-In the description we deliberately use the more general term Human instead of User, because humans are not always playing the role of users in software engineering sense, they could just be interacting naturally with their environment and with out knowing they are implicitly
+providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
+term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
+could just be interacting naturally with their environment and with out knowing they are implicitly
 evaluating a possible solution. The framework includes components that can be refined to increase
 participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
+Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
+design. Before entering to details the main design decisions of the framework are explained next:
 
-  \subsection{Real World}
-  \subsection{Data Model}
+\begin{itemize}
+\item {\bf Users are human} 
+  The framework follows the approach of human centered computing \cite{sebe2010human},
+  in which the context, environment, interfaces, preferences, accessibility, human relations, cognitive
+  limitations, culture, creativity and other human aspects are an integral part of the system.
+  Humans are the computing resorces of the system, having unique characteristics as those identified by 
+  Sun \& Dance \cite{Sun2013}  ``(1) humans can solve computer hard problems; (2) humans are very good at exception handling, (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are vulnerable to
+  psychological manipulation, (6) humans are prone to errors, especially for reflective tasks.''
+
+\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
+  some times they try to game the system. Providers must actively promote and design the interactive
+  system to engage volunteers \cite{oh2015clicking}.
+
+\item {\bf Users are not alone}
+  Relationship between users in an interactive evolutionary algorithm can be modeled
+  as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
+  A graph model could enable researchers to find other ways of identifying leaders of 
+  opinion or measuring the similarity between user's preferences. 
+  These measures can then be used by recommender algorithms selecting 
+  phenotypes according users' preferences.
+
+\item {\bf Context of use matters}
+  \textit{Context} In computer science Fischer \cite{fischer2012context}
+  defines context as \textit{``the interaction between humans and
+  computers in socio-technical systems that takes place in a certain
+  context referring to the physical and social situation in which
+  computational devices and environments are embedded"}. 
+  Fischer also identifies the important aspects to consider when the context is used: how it is
+  the contextual data obtained, how the context is represented and what
+  goals and purposes the context has when is used in a particular
+  application.   An IEC system will  be used within a certain range of technical, physical and social or
+  organizational environments \cite{maguire2001context} that may affect its use.
+  The concept has been formally defined
+  by ISO standard 9241-11 \cite{international1998iso} as \textit{``users,
+  tasks and equipment (hardware, software and materials), and the
+  physical and social environments in which a product is used"}.
+
+\item {\bf Interaction is a stream of actions}
+  Real time processing of users' actions could be needed for certain applications when data is 
+  captured by sensors, or directly captured as user input.
+  Graph based models are currently used in social networks to keep track  
+  of user interactions with media objects, places, and other users. Users of 
+  social networks (for instance the Facebook Graph) are accustomed to express these 
+  complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
+  Other example is the W3C Activity Streams 2.0 specification used for representing activities 
+  common in social web applications \cite{json:streams}. An important component of the 
+  framework is the stream of user-activities.
+\end{itemize}
+
+The human centered framework consists of three high level components depicted
+in figure \ref{fig:uc_framework}:
+
+\begin{itemize}
+  \item {\bf Interactive Evolutionary Computational system} 
+  This is the real world system that we are going to represent in the model, 
+  it consist of human users and their interactions with one or more phenotypes
+  from
+  the population. There are many ways in which humans could interact 
+  with phenotypes, this could be through a mobile device, a user interface 
+  or by interacting with real world objects generated as part of the algorithm 
+  \cite{de2014artists,de2013unplugging}. 
+  There is also the possibility that fitness or even part of the search 
+  to be done by devices lent by the human \cite{DBLP:conf/gecco/MereloCGCRV16}.
+  The information gathered through the interaction is the primary concern
+  as it will guide the search.
+
+  \item {\bf IEC Data Model}
+  A data model is used to describe the IEC system prior to a physical 
+  implementation.  Depending on the domain several techniques can be used.
+  Parts of the system could be better described by an Entity relationship 
+  modeling to be used in a relational database or a class digram for a 
+  key-value store.
+
+  A graph is proposed for modeling the social network of users 
+  and their interaction with the population. When implemented a graph database 
+  will be the back-end of the system. This  backend will then be
+  used by the EA and the user interface, in a Model-View-Controller architecture.
+
+  \item {\bf Evolutionary Algorithm} 
+  The EA algorithm queries the graph model, and then updates those nodes that 
+  are part of the population. The EA could also be done with the help of
+  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
+  the new population.  
+\end{itemize}
+
   \subsection{Evolutionary Algorithm}

### 2017-02-01 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Framework fig

@@
 
-  \subsection{Evolutionary Algorithm}
+\section{EvoSpace-Interactive}
+\label{sec:evospace-i}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/UI_ed01.png}
+    \caption{User interface of an EvoSpace-Interactive application.}
+    \label{fig:web}
+\end{figure}
 
-\section{EvoSpace-Interactive}
+As a case study, a IEA was developed using the 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
+A brief description of the application is presented next, focusing
+on the data elements that where ported to the graph model.
+
+\subsection{Fitness Assignment}
+\label{sec:assignment}
+The ES-I platform employs a collaborative technique,
+where several registered users assign a quality assessment to a single
+phenotypes and then an aggregated fitness value is calculated. The fitness
+assigned to each phenotypes depends on the taste of each particular user, 
+resulting in a many-to-many relationship between users and phenotypes. 
+Many systems query this user-phenotype relation to extract relevant
+knowledge about the process and the population, for instance showing the
+most popular phenotype, or the the user with more participation 
+\cite{picbreeder}.
+In order to do this, data about phenotypes 
+must be permanently stored, even
+if they are no longer in the active population.
+
+\subsection{Collaboration}
+\label{sec:col}
+After entering the web application by using their Facebook account,
+users can collaborate with their Facebook friends, 
+sharing those phenotypes they like, or by taking phenotypes
+from their friend's collections by using the web interface depicted 
+in Figure \ref{fig:web}.
+At the top left corner a list of Facebook friends is presented
+to encourage users to interact with the system. In the central 
+\emph{ Wall } area, a phenotype sampled from the population that is
+being evolved via the evolutionary algorithm 
+is shown to the user.
+Here, the user can interact with the system in two ways.
+First, he can assign a quality assessment to the phenotype using
+a five star rating system or,
+additionally, a user can choose to add an image to one of their \emph{Collections}.
+A collection is a special folder that stores those phenotypes a user likes and wishes
+to save. After the user finishes interacting with the phenotype
+on the Wall, he can choose to retrieve a new one from the population.
+At the left hand side, the web page shows the \emph{Collections} section.
+The user can create several collections, to group and organize his favorite 
+artifacts. Moreover, users can browse the content of each collection and from
+there share images through their social network.
+When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
+liked the phenotype. The pane also includes a link to the phenotype's 
+details, the parents, genetic operators that created it, and genealogy
+information. This makes the assignment of fitness through the rating
+system a {\em social} activity, pursuing the objective of this work,
+which is to increase engagement of the users.
+
+\begin{figure*}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/framework.png}
+    \caption{IEC Human-centered framework.}
+    \label{fig:uc_framework}
+\end{figure*}
+
   \subsection{Interfaces}

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
EvoSpace-I

@@
 
-\section{HC  IEC Framework}
+\begin{figure*}[!t]
+    \centering
+        \includegraphics[width=4.5in]{img/framework.eps}
+    \caption{IEC Human-centered framework.}
+    \label{fig:uc_framework}
+\end{figure*}
+
+\section{Human-Centered IEC Framework}
 % Comment: 
@@
   Real time processing of users' actions could be needed for certain applications when data is 
-  captured by sensors, or directly captured as user input.
-  Graph based models are currently used in social networks to keep track  
-  of user interactions with media objects, places, and other users. Users of 
+  captured by sensors, or directly captured as user input. For example, social networks keep track of 
+  users' interactions with other users, media objects and places. Users of 
   social networks (for instance the Facebook Graph) are accustomed to express these 
@@
   Other example is the W3C Activity Streams 2.0 specification used for representing activities 
-  common in social web applications \cite{json:streams}. An important component of the 
-  framework is the stream of user-activities.
+  common in social web applications \cite{json:streams}. 
 \end{itemize}
 
-The human centered framework consists of three high level components depicted
+The human centered IEC architecture consists of three high level components depicted
 in figure \ref{fig:uc_framework}:
@@
   \item {\bf Interactive Evolutionary Computational system} 
-  This is the real world system that we are going to represent in the model, 
+  This is the real world system that we are going to represent in the data model, 
   it consist of human users and their interactions with one or more phenotypes
-  from
-  the population. There are many ways in which humans could interact 
-  with phenotypes, this could be through a mobile device, a user interface 
-  or by interacting with real world objects generated as part of the algorithm 
+  from the population. There are many ways in which humans could interact 
+  with phenotypes. The interaction consists of a set of actions and 
+  takes place in a certain context. For example, through a mobile device  
+  or by interacting with real world objects 
   \cite{de2014artists,de2013unplugging}. 
   There is also the possibility that fitness or even part of the search 
-  to be done by devices lent by the human \cite{DBLP:conf/gecco/MereloCGCRV16}.
+  is done by devices lent by humans \cite{DBLP:conf/gecco/MereloCGCRV16}.
   The information gathered through the interaction is the primary concern
@@
   modeling to be used in a relational database or a class digram for a 
-  key-value store.
-
-  A graph is proposed for modeling the social network of users 
+  key-value store. A graph is proposed for modeling the social network of users 
   and their interaction with the population. When implemented a graph database 
-  will be the back-end of the system. This  backend will then be
-  used by the EA and the user interface, in a Model-View-Controller architecture.
+  will be the back-end of the system.
 
   \item {\bf Evolutionary Algorithm} 
-  The EA algorithm queries the graph model, and then updates those nodes that 
-  are part of the population. The EA could also be done with the help of
-  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
-  the new population.  
+  The EA algorithm interacts with the data model. The EA could also be done with the help of
+  humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
+  the solutions of the new population.  
 \end{itemize}
 
-\section{EvoSpace-Interactive}
+\section{Data Model Implementation}
+The core of the framework is the data model because both the IEC system and the EA will 
+depend on the application. In this section implementation details of the Data Model are presented. 
+  \subsection{EvoSpace}
+  \subsection{Graph}
+  \subsection{PostgreSQL}
+
+  \subsection{Interfaces}
+
+  \subsection{Evolution}
+
+\section{Data Model Implementation}
+
+\section{Case Studies}
 \label{sec:evospace-i}
@@
     \centering
-        \includegraphics[width=3.5in]{img/UI_ed01.png}
-    \caption{User interface of an EvoSpace-Interactive application.}
+        \includegraphics[width=3.5in]{img/interface.png}
+    \caption{User interface of the EvoDrawings application.}
     \label{fig:web}
@@
 
+\subsection{EvoDraw}
 As a case study, a IEA was developed using the 
@@
 which is to increase engagement of the users.
+\subsubsection{Graph Model} 
+\subsubsection{Human Interaction Model}
 
-\begin{figure*}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/framework.png}
-    \caption{IEC Human-centered framework.}
-    \label{fig:uc_framework}
-\end{figure*}
-
-  \subsection{Interfaces}
-  \subsection{Data Models}
-    \subsubsection{EvoSpace}
-    \subsubsection{Graph}
-    \subsubsection{PostgreSQL}
-  \subsection{Evolution}
-
-\section{Case Studies}
-\subsection{EvoDraw}
 \subsection{EvoDraw-Kinect}

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Starting Case Study
@JJ please have a look, I think only the EvoDrawings case study fits.
Done for today.

@@
 % Maybe in this case we should use the term software framework or architecture. 
-The general goal of this research is to develop a human-centered \cite{gasson2003human} 
+The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
 software framework for interactive evolutionary computation (IEC). 
-The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
-providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
+A framework is defined as a reusable architectural design together
+with an implementation \cite{campbell1991choices}, in this case 
+providing generalized components to developers of IEC systems. In the
+description %below? If that's te chase, we _will_ - JJ
+ we deliberately use the more general 
 term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
-could just be interacting naturally with their environment and with out knowing they are implicitly
-evaluating a possible solution. The framework includes components that can be refined to increase
+could just be interacting naturally with their environment and
+unwittingly and at the same time they are implicitly
+evaluating a possible solution.
+
+%Which framework? Are we talking about a particular framework or that
+%is the general rule for frameworks? - JJ
+The framework includes components that can be refined to increase
 participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
 Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
-design. Before entering to details the main design decisions of the framework are explained next:
+design. Before diving into details the main design decisions of the
+framework are explained next:  % constraints? Rules? You don't decide
+                               % that users are humans, they simply
+                               % are - JJ
 
@@
   some times they try to game the system. Providers must actively promote and design the interactive
-  system to engage volunteers \cite{oh2015clicking}.
+  system to engage volunteers \cite{oh2015clicking}. % Define
+                                % providers somewhere. Experiment
+                                % owners? - JJ
 
@@
   context referring to the physical and social situation in which
-  computational devices and environments are embedded"}. 
+  computational devices and environments are embedded''}. 
   Fischer also identifies the important aspects to consider when the context is used: how it is
@@
 
+% Say something about how this is important for the framework and how
+% you incorporate it - JJ
 The human centered IEC architecture consists of three high level components depicted
@@
 depend on the application. In this section implementation details of the Data Model are presented.
-  \subsection{EvoSpace}
+Three database models are employed to store different elements of the system:
+
+  \subsection{EvoSpace-Redis}
+EvoSpace is a population store \cite{Evospace}  for the development of 
+evolutionary algorithms that are intended to run on a cloud computing model. 
+The population is decoupled from any particular evolutionary algorithm. 
+Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
+processed and replaced using a specified set of methods \cite{GValdez2015}. The population
+is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
+database management system, because it provides a hash based implementation of sets and 
+queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
+candidate solutions are retieved from the server, evaluated, and then sent back. 
+The same operations are used to Evolve the population. In EvoSpace individuals replaced 
+in the population are stored indefinitly, to permit users to store permalinks to them.
+Implementation details are presented in \cite{garcia2013evospace}.
+
 \subsection{Graph}
+  In order to deal with large datasets of connected data found in these systems,
+  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
+  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
+  A graph is proposed for modeling the social network of users, their interaction with 
+  candidate solutions, and the relationships between them in the population.
+  The graph database system used in the implementation is Neo4J as it is
+  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
+  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
+  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
+  are presented.
+
+  \begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/gui-neo.png}
+    \caption{Example query in Cypher.}
+    \label{fig:cypher}
+  \end{figure}
+
 \subsection{PostgreSQL}
+  The PostgreSQL relational database system, is also employed because user sessions and authentication,
+  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.
 
 \subsection{Interfaces}
+%They are Sensor and Web,
 
@@
 
-\section{Data Model Implementation}
-
 \section{Case Studies}
 \label{sec:evospace-i}
+
 \begin{figure}[!t]
@@
 \label{sec:assignment}
-The ES-I platform employs a collaborative technique,
-where several registered users assign a quality assessment to a single
-phenotypes and then an aggregated fitness value is calculated. The fitness
-assigned to each phenotypes depends on the taste of each particular user, 
-resulting in a many-to-many relationship between users and phenotypes. 
-Many systems query this user-phenotype relation to extract relevant
+The ES-I % You should use always the same acronym, not start using it
+         % here - JJ
+ platform employs a collaborative methodology for performing fitness
+ assignment. Several registered users assign a quality assessment to a single
+phenotype and then an aggregated fitness value is calculated,
+depending on the tastes of each particular user,  % how do you define
+                                % "taste"? - JJ
+resulting in a many-to-many relationship between users and
+phenotypes. %scored by them? - JJ 
+Many systems % formerly published methods? Other researchers? - JJ
+query this user-phenotype relation to extract relevant
 knowledge about the process and the population, for instance showing the
@@
 \cite{picbreeder}.
-In order to do this, data about phenotypes 
+In order to do this, data about phenotypes % metadata better? - JJ
 must be permanently stored, even
@@
 \label{sec:col}
+Users need to authenticate themselves to the system using their
+Facebook account. In that sense, some users might not be interested
+either because they do not want to give that that or simply because
+they do not use that particular social network. Even as we might lose
+some users that way, the additional information we obtain for scoring
+phenotypes more than compensates for that. % Added this, check it
+                                % out. 
 After entering the web application by using their Facebook account,
@@
 Here, the user can interact with the system in two ways.
-First, he can assign a quality assessment to the phenotype using
+First, he can assign a quality assessment % Why don't you use other
+                                % thing instead of quality? We are not
+                                % talking about quality, but about
+                                % tastes or similar - JJ
+ to the phenotype using
 a five star rating system or,
@@
 on the Wall, he can choose to retrieve a new one from the population.
-At the left hand side, the web page shows the \emph{Collections} section.
+At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
 The user can create several collections, to group and organize his favorite 
@@
 \subsubsection{Graph Model}
-\subsubsection{Human Interaction Model}
+
+\subsubsection{Gammification}
+
+\subsubsection{Setup}
+\subsubsection{Results}
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_1.png}
+    \caption{Experiment 1. Weibull Fit}
+    \label{fig:}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_2.png}
+    \caption{Experiment 2. Weibull Fit}
+    \label{fig:}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_3.png}
+    \caption{Experiment 3. Weibull Fit}
+    \label{fig:}
+\end{figure}
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Some comments and corrections

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 % Maybe in this case we should use the term software framework or architecture. 
-The general goal of this research is to develop a human-centered \cite{gasson2003human} 
+The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
 software framework for interactive evolutionary computation (IEC). 
-The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
-providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
+A framework is defined as a reusable architectural design together
+with an implementation \cite{campbell1991choices}, in this case 
+providing generalized components to developers of IEC systems. In the
+description %below? If that's te chase, we _will_ - JJ
+ we deliberately use the more general 
 term Human instead of User, because humans are not always playing the role of users in software engineering sense, they

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Revision of definitions and introduction

@@
 term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
-could just be interacting naturally with their environment and with out knowing they are implicitly
-evaluating a possible solution. The framework includes components that can be refined to increase
+could just be interacting naturally with their environment and
+unwittingly and at the same time they are implicitly
+evaluating a possible solution.
+
+%Which framework? Are we talking about a particular framework or that
+%is the general rule for frameworks? - JJ
+The framework includes components that can be refined to increase
 participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
 Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
-design. Before entering to details the main design decisions of the framework are explained next:
+design. Before diving into details the main design decisions of the
+framework are explained next:  % constraints? Rules? You don't decide
+                               % that users are humans, they simply
+                               % are - JJ
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Revised available data

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 depend on the application. In this section implementation details of the Data Model are presented. 
-Three database models are employed to store different elements of the system:
-
-  \subsection{EvoSpace-Redis}
-EvoSpace is a population store \cite{Evospace}  for the development of 
-evolutionary algorithms that are intended to run on a cloud computing model. 
-The population is decoupled from any particular evolutionary algorithm. 
-Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
-processed and replaced using a specified set of methods \cite{GValdez2015}. The population
-is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
-database management system, because it provides a hash based implementation of sets and 
-queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
-candidate solutions are retieved from the server, evaluated, and then sent back. 
-The same operations are used to Evolve the population. In EvoSpace individuals replaced 
-in the population are stored indefinitly, to permit users to store permalinks to them.
-Implementation details are presented in \cite{garcia2013evospace}.
-
+  \subsection{EvoSpace}
   \subsection{Graph}
-  In order to deal with large datasets of connected data found in these systems,
-  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
-  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
-  A graph is proposed for modeling the social network of users, their interaction with 
-  candidate solutions, and the relationships between them in the population.
-  The graph database system used in the implementation is Neo4J as it is
-  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
-  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
-  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
-  are presented.
-
-  \begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/gui-neo.png}
-    \caption{Example query in Cypher.}
-    \label{fig:cypher}
-  \end{figure}
-
   \subsection{PostgreSQL}
-  The PostgreSQL relational database system, is also employed because user sessions and authentication,
-  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.
 
   \subsection{Interfaces}
-%They are Sensor and Web,
 
@@
 
+\section{Data Model Implementation}
+
 \section{Case Studies}
 \label{sec:evospace-i}
-
 \begin{figure}[!t]
@@
 \subsubsection{Graph Model} 
-
-\subsubsection{Gammification}
-
-\subsubsection{Setup}
-\subsubsection{Results}
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/weibull_1.png}
-    \caption{Experiment 1. Weibull Fit}
-    \label{fig:}
-\end{figure}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/weibull_2.png}
-    \caption{Experiment 2. Weibull Fit}
-    \label{fig:}
-\end{figure}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/weibull_3.png}
-    \caption{Experiment 3. Weibull Fit}
-    \label{fig:}
-\end{figure}
+\subsubsection{Human Interaction Model}
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Uncommitted from yesterday

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 
-\begin{figure*}[!t]
-    \centering
-        \includegraphics[width=4.5in]{img/framework.eps}
-    \caption{IEC Human-centered framework.}
-    \label{fig:uc_framework}
-\end{figure*}
-
-\section{Human-Centered IEC Framework}
+\section{HC  IEC Framework}
 % Comment: 
 % Maybe in this case we should use the term software framework or architecture. 
-The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
+The general goal of this research is to develop a human-centered \cite{gasson2003human} 
 software framework for interactive evolutionary computation (IEC). 
-A framework is defined as a reusable architectural design together
-with an implementation \cite{campbell1991choices}, in this case 
-providing generalized components to developers of IEC systems. In the
-description %below? If that's te chase, we _will_ - JJ
- we deliberately use the more general 
+The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
+providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
 term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
@@
   Real time processing of users' actions could be needed for certain applications when data is 
-  captured by sensors, or directly captured as user input. For example, social networks keep track of 
-  users' interactions with other users, media objects and places. Users of 
+  captured by sensors, or directly captured as user input.
+  Graph based models are currently used in social networks to keep track  
+  of user interactions with media objects, places, and other users. Users of 
   social networks (for instance the Facebook Graph) are accustomed to express these 
@@
   Other example is the W3C Activity Streams 2.0 specification used for representing activities 
-  common in social web applications \cite{json:streams}. 
+  common in social web applications \cite{json:streams}. An important component of the 
+  framework is the stream of user-activities.
 \end{itemize}
 
-The human centered IEC architecture consists of three high level components depicted
+The human centered framework consists of three high level components depicted
 in figure \ref{fig:uc_framework}:
@@
   \item {\bf Interactive Evolutionary Computational system} 
-  This is the real world system that we are going to represent in the data model, 
+  This is the real world system that we are going to represent in the model, 
   it consist of human users and their interactions with one or more phenotypes
-  from the population. There are many ways in which humans could interact 
-  with phenotypes. The interaction consists of a set of actions and 
-  takes place in a certain context. For example, through a mobile device  
-  or by interacting with real world objects 
+  from
+  the population. There are many ways in which humans could interact 
+  with phenotypes, this could be through a mobile device, a user interface 
+  or by interacting with real world objects generated as part of the algorithm 
   \cite{de2014artists,de2013unplugging}. 
   There is also the possibility that fitness or even part of the search 
-  is done by devices lent by humans \cite{DBLP:conf/gecco/MereloCGCRV16}.
+  to be done by devices lent by the human \cite{DBLP:conf/gecco/MereloCGCRV16}.
   The information gathered through the interaction is the primary concern
@@
   modeling to be used in a relational database or a class digram for a 
-  key-value store. A graph is proposed for modeling the social network of users 
+  key-value store.
+
+  A graph is proposed for modeling the social network of users 
   and their interaction with the population. When implemented a graph database 
-  will be the back-end of the system.
+  will be the back-end of the system. This  backend will then be
+  used by the EA and the user interface, in a Model-View-Controller architecture.
 
   \item {\bf Evolutionary Algorithm} 
-  The EA algorithm interacts with the data model. The EA could also be done with the help of
-  humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
-  the solutions of the new population.  
+  The EA algorithm queries the graph model, and then updates those nodes that 
+  are part of the population. The EA could also be done with the help of
+  humans, for instance in the XY project \cite{de2013unplugging} artists physically created
+  the new population.  
 \end{itemize}
 
-\section{Data Model Implementation}
-The core of the framework is the data model because both the IEC system and the EA will 
-depend on the application. In this section implementation details of the Data Model are presented. 
-  \subsection{EvoSpace}
-  \subsection{Graph}
-  \subsection{PostgreSQL}
-
-  \subsection{Interfaces}
-
-  \subsection{Evolution}
-
-\section{Data Model Implementation}
-
-\section{Case Studies}
+\section{EvoSpace-Interactive}
 \label{sec:evospace-i}
@@
     \centering
-        \includegraphics[width=3.5in]{img/interface.png}
-    \caption{User interface of the EvoDrawings application.}
+        \includegraphics[width=3.5in]{img/UI_ed01.png}
+    \caption{User interface of an EvoSpace-Interactive application.}
     \label{fig:web}
@@
 
-\subsection{EvoDraw}
 As a case study, a IEA was developed using the 
@@
 which is to increase engagement of the users.
-\subsubsection{Graph Model} 
-\subsubsection{Human Interaction Model}
 
+\begin{figure*}[!t]
+    \centering
+        \includegraphics[width=3.5in]{framework.eps}
+    \caption{IEC Human-centered framework.}
+    \label{fig:uc_framework}
+\end{figure*}
+
+  \subsection{Interfaces}
+  \subsection{Data Models}
+    \subsubsection{EvoSpace}
+    \subsubsection{Graph}
+    \subsubsection{PostgreSQL}
+  \subsection{Evolution}
+
+\section{Case Studies}
+\subsection{EvoDraw}
 \subsection{EvoDraw-Kinect}

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 The rest of this paper is organized as follows.
-Section \ref{sec:interactive} presents related work on the topic 
+Section \ref{sec:related} presents related work on the topic 
 of Interactive Evolution.
-Then, Section \ref{sec:e} presents the Model of Human Interactions in Collaborative Interactive 
+Then, Section \ref{sec:framework} presents the framework of Human Interactions in Collaborative Interactive 
  Evolutionary Computation which 
-the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:HCF}.
+the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:implementation}.
 The case studies  and results are presented in Section \ref{sec:experiments}.
@@
 \section{Related Work}
+\label{sec:related}
 % C-IEAS
@@
 \section{Human-Centered IEC Framework}
+\label{sec:framework}
 % Comment: 
@@
 \section{Data Model Implementation}
+\label{sec:implementation}
+
 The core of the framework is the data model because both the IEC system and the EA will 
 depend on the application. In this section implementation details of the Data Model are presented.
-  \subsection{EvoSpace}
+Three database models are employed to store different elements of the system:
+
+  \subsection{EvoSpace-Redis}
+EvoSpace is a population store \cite{Evospace}  for the development of 
+evolutionary algorithms that are intended to run on a cloud computing model. 
+The population is decoupled from any particular evolutionary algorithm. 
+Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
+processed and replaced using a specified set of methods \cite{GValdez2015}. The population
+is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
+database management system, because it provides a hash based implementation of sets and 
+queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
+candidate solutions are retieved from the server, evaluated, and then sent back. 
+The same operations are used to Evolve the population. In EvoSpace individuals replaced 
+in the population are stored indefinitly, to permit users to store permalinks to them.
+Implementation details are presented in \cite{garcia2013evospace}.
+
 \subsection{Graph}
+  In order to deal with large datasets of connected data found in these systems,
+  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
+  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
+  A graph is proposed for modeling the social network of users, their interaction with 
+  candidate solutions, and the relationships between them in the population.
+  The graph database system used in the implementation is Neo4J as it is
+  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
+  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
+  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
+  are presented.
+
+  \begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/gui-neo.png}
+    \caption{Example query in Cypher.}
+    \label{fig:cypher}
+  \end{figure}
+
 \subsection{PostgreSQL}
+  The PostgreSQL relational database system, is also employed because user sessions and authentication,
+  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.
 
 \subsection{Interfaces}
+%They are Sensor and Web,
 
@@
 
-\section{Data Model Implementation}
+\section{Case Studies}
+\label{sec:experiments}
 
-\section{Case Studies}
-\label{sec:evospace-i}
 \begin{figure}[!t]
@@
 \subsubsection{Graph Model}
-\subsubsection{Human Interaction Model}
+
+\subsubsection{Gammification}
+
+\subsubsection{Setup}
+\subsubsection{Results}
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_1.png}
+    \caption{Experiment 1. Weibull Fit}
+    \label{fig:w1}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_2.png}
+    \caption{Experiment 2. Weibull Fit}
+    \label{fig:w2}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/weibull_3.png}
+    \caption{Experiment 3. Weibull Fit}
+    \label{fig:w3}
+\end{figure}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex

@@
 providing generalized components to developers of IEC systems. In the
-description %below? If that's te chase, we _will_ - JJ
- we deliberately use the more general 
+description below %below? If that's te chase, we _will_ - JJ
+% Not only in the description, more like a paradigm shitf. But ok for now - Mario
+ we will deliberately use the more general 
 term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
@@
 %is the general rule for frameworks? - JJ
-The framework includes components that can be refined to increase
-participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application.
-Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
+%Is general bu I was talking about this one - Mario
+The proposed framework includes components that can be refined to increase
+participation and also to minimize the amount of evaluations needed for the evolutionary 
+process in a given IEC application. Software frameworks often have a vison or a way 
+of thinking \cite{carneiro2010introducing} guiding their
 design. Before diving into details the main design decisions of the
@@
                                % are - JJ 
+                               % Ok, is more like we want to acknowledge that, 
+                               % To Do - Mario
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Just a few corrections here and there. Charts look cool

@@
 As a case study, a IEA was developed using the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}.
 A brief description of the application is presented next, focusing
@@
 Facebook account. In that sense, some users might not be interested
-either because they do not want to give that that or simply because
+either because they do not want to give that information or simply because
 they do not use that particular social network. Even as we might lose
@@
 
-\subsubsection{Gammification}
+\subsubsection{Gamification}
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
primer añadido a related work

@@
 % C-IEAS
+Given the fatigue fitness evaluations produce in human users when applying IEAs, some authors have tried to mitigate the problem by allowing the algorithm to collaborate with the user, so that sometimes users perform the evaluation, but also specific measures are included into the algorithm to perform the evaluation of some features automatically.
+For instance, Reis et al.
+added some terrain measures (such as accessibility and edge length) to a standard IEC in \cite{DBLP:journals/soco/FradeVC12} .
+This way the algorithm was capable of providing terrains that would otherwise have need users evaluation for these specific features.
+
 % Volunteer Based

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Addressing JJ comments

@@
 conceptual and implementation details of the framework. 
-The experiments were implemented using the EvoSpace-Interactive framework \cite{garcia2013evospace}, 
+The experiments were implemented using the EvoSpace-Interactive (ES-I) framework \cite{garcia2013evospace}, 
 and where employed in different contexts of interaction. 
@@
  Evolutionary Computation which 
-the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:implementation}.
+the main proposal of this work. An implementation using the ES-I framework is detailed in Section \ref{sec:implementation}.
 The case studies  and results are presented in Section \ref{sec:experiments}.
@@
     \caption{IEC Human-centered framework.}
-    \label{fig:uc_framework}
+    \label{fig:hc_framework}
 \end{figure*}
@@
 with an implementation \cite{campbell1991choices}, in this case 
-providing generalized components to developers of IEC systems. In the
-description below %below? If that's te chase, we _will_ - JJ
+providing generalized components to developers of IEC systems. 
+% Just by saying is Human Centered this is not needed, it was from the User-Centered days - Mario
+%In the
+%description below %below? If that's te chase, we _will_ - JJ
 % Not only in the description, more like a paradigm shitf. But ok for now - Mario
- we will deliberately use the more general 
-term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
-could just be interacting naturally with their environment and
-unwittingly and at the same time they are implicitly
-evaluating a possible solution.
+% we will deliberately use the more general 
+%term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
+%could just be interacting naturally with their environment and
+%unwittingly and at the same time they are implicitly
+%evaluating a possible solution.
 
@@
 of thinking \cite{carneiro2010introducing} guiding their
-design. Before diving into details the main design decisions of the
+design. Before diving into details the main design considerations of the
 framework are explained next:  % constraints? Rules? You don't decide
@@
                                % Ok, is more like we want to acknowledge that, 
-                               % To Do - Mario
+                               % considerations? - Mario
 
@@
 \item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
-  some times they try to game the system. Providers must actively promote and design the interactive
+  some times they try to game the system. Project owners must actively promote and design the interactive
   system to engage volunteers \cite{oh2015clicking}. % Define
@@
                                 % owners? - JJ
+                                % Using project, like BOINC  https://boinc.berkeley.edu/trac/wiki/VolunteerComputing - M
 
@@
 % you incorporate it - JJ
-The human centered IEC architecture consists of three high level components depicted
-in figure \ref{fig:uc_framework}:
+% Done - Mario
+The above considerations have guided the design of the framework, and througout they have been
+treated as application constraints. In order to satisfy theses requirements the
+human centered IEC architecture consists of three high level components depicted
+in figure \ref{fig:hc_framework}:
 
@@
 As a case study, a IEA was developed using the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}.
+ES-I platform \cite{garcia2013evospace}.
 A brief description of the application is presented next, focusing
@@
          % here - JJ 
+         % Done -Mario
  platform employs a collaborative methodology for performing fitness
@@
 phenotype and then an aggregated fitness value is calculated,
-depending on the tastes of each particular user,  % how do you define
+depending on the rating (from 1 to 5 stars) of each particular user,  % how do you define
                                 % "taste"? - JJ
+                                % Using actions
 resulting in a many-to-many relationship between users and
-phenotypes. %scored by them? - JJ 
-Many systems % formerly published methods? Other researchers? - JJ
+phenotypes. 
+ %scored by them? - JJ 
+Other components or users of the systems can % formerly published methods? Other researchers? - JJ
+% Corrected -Mario
 query this user-phenotype relation to extract relevant
@@
 \cite{picbreeder}.
-In order to do this, data about phenotypes % metadata better? - JJ
+In order to do this, metadata about phenotypes % metadata better? - JJ Yes -Mario
 must be permanently stored, even
@@
                                 % out.  
+                                % Ok! - Mario (and knowing how much more is future work)
 After entering the web application by using their Facebook account,
@@
 Here, the user can interact with the system in two ways.
-First, he can assign a quality assessment % Why don't you use other
+First, he can assign a rating % Why don't you use other
                                 % thing instead of quality? We are not
@@
                                 % tastes or similar - JJ
+                                % is rating Ok? 
  to the phenotype using

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Changing quotes to its own environment

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 Data available from the interaction is also used to increase the performance of the system by applying 
-gamification techniques. Gamification is defined by Deterding et al. as
-``the use of game design elements in non-game contexts'' \cite{deterding2011game}.
-The gamification element employed in a case study is a rewarding mechanism  
-\cite{dubois2013understanding}, in general they consist of a reputation system with score points, 
+gamification techniques. Gamification is defined by Deterding et al. \cite{deterding2011game} as
+\begin{quote}
+  the use of game design elements in non-game contexts.
+\end{quote}  
+The gamification element employed in this case study has been a rewarding mechanism  
+\cite{dubois2013understanding}. In general rewards  consist of a reputation system with score points, 
 levels and leader boards. Points are awarded to users in response of 
@@
 on the score and certain features of the game are only available to gamers when 
-they reach a giving level.
+they reach a giving level. % Say what we are doing in this case - JJ
 
@@
   Humans are the computing resorces of the system, having unique characteristics as those identified by 
-  Sun \& Dance \cite{Sun2013}  ``(1) humans can solve computer hard problems; (2) humans are very good at exception handling, (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are vulnerable to
-  psychological manipulation, (6) humans are prone to errors, especially for reflective tasks.''
+  Sun \& Dance \cite{Sun2013}
+\begin{quote}
+  (1) humans can solve computer hard problems; (2) humans are very good at exception handling, (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are vulnerable to 
+  psychological manipulation, (6) humans are prone to errors,
+  especially for reflective tasks.
+\end{quote}
 
@@
   \textit{Context} In computer science Fischer \cite{fischer2012context}
-  defines context as \textit{``the interaction between humans and
+  defines context as
+  \begin{quote}
+  the interaction between humans and
   computers in socio-technical systems that takes place in a certain
   context referring to the physical and social situation in which
-  computational devices and environments are embedded''}. 
+  computational devices and environments are embedded.
+\end{quote}   
   Fischer also identifies the important aspects to consider when the context is used: how it is
@@
   The concept has been formally defined
-  by ISO standard 9241-11 \cite{international1998iso} as \textit{``users,
+  by ISO standard 9241-11 \cite{international1998iso} as\begin{quote}
+  users,
   tasks and equipment (hardware, software and materials), and the
-  physical and social environments in which a product is used"}.
+  physical and social environments in which a product is used.
+\end{quote}
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Erasing addressed comments and some fixes

@@
 \subsection{PostgreSQL}
-  The PostgreSQL relational database system, is also employed because user sessions and authentication,
-  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.
+  The PostgreSQL relational database system is also employed because user sessions and authentication,
+  as well as dynamic web pages are handled directly by the Django web
+  framework \cite{garcia2013evospace}. % these implementations details
+                                % could go if needed.
 
@@
 \label{sec:assignment}
-The ES-I % You should use always the same acronym, not start using it
-         % here - JJ 
-         % Done -Mario
- platform employs a collaborative methodology for performing fitness
+% changed to this, because I understand it does not always work that
+% way - JJ
+The ES-I platform has been programmed to employ a collaborative methodology for performing fitness
  assignment. Several registered users assign a quality assessment to a single
 phenotype and then an aggregated fitness value is calculated,
-depending on the rating (from 1 to 5 stars) of each particular user,  % how do you define
+depending on the rating (from 1 to 5 stars) by each particular user,  % how do you define
                                 % "taste"? - JJ
@@
  %scored by them? - JJ 
-Other components or users of the systems can % formerly published methods? Other researchers? - JJ
-% Corrected -Mario
+Other components or users of the systems can 
 query this user-phenotype relation to extract relevant
@@
 \cite{picbreeder}.
-In order to do this, metadata about phenotypes % metadata better? - JJ Yes -Mario
+In order to do this, metadata about phenotypes 
 must be permanently stored, even
@@
                                 % is rating Ok?
+% Rating is probably OK. 
  to the phenotype using
@@
   % Scientsts' Support Program}.
+  Reserved\\
+  Space\\
+  for\\
+  Acks
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Complete Image
Interfaces

@@
 % C-IEAS
-Given the fatigue fitness evaluations produce in human users when applying IEAs, some authors have tried to mitigate the problem by allowing the algorithm to collaborate with the user, so that sometimes users perform the evaluation, but also specific measures are included into the algorithm to perform the evaluation of some features automatically.
-For instance, Reis et al.
-added some terrain measures (such as accessibility and edge length) to a standard IEC in \cite{DBLP:journals/soco/FradeVC12} .
-This way the algorithm was capable of providing terrains that would otherwise have need users evaluation for these specific features.
+Given the fatigue fitness evaluations produce in human users when applying IEAs, some authors have tried to mitigate
+the problem by allowing the algorithm to collaborate with the user, so that sometimes users perform the evaluation, 
+but also specific measures are included into the algorithm to perform the evaluation of some features automatically.  
+For instance, Reis et al. added some terrain measures (such as accessibility and edge length) to a standard 
+IEC in \cite{DBLP:journals/soco/FradeVC12} .  This way the algorithm was capable of providing terrains that would
+otherwise have need users evaluation for these specific features.
 
@@
     \centering
-        \includegraphics[width=4.5in]{img/framework.eps}
+        \includegraphics[width=5in]{img/framework.eps}
     \caption{IEC Human-centered framework.}
@@
 \subsection{Interfaces}
-%They are Sensor and Web,
+The appropiate human interface will depend on the application domain, the current framework implemantation  
+employs a web based application. Developers can use different templates, to present one or more phenotypes 
+at a time, and the type of rating system: like or one to five stars. Another implementations do not need a
+graphical interface at all, for example the EvoDraw instalation or the XYZ Project \cite{de2014artists}.
 

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Results

@@
 Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
-whose fitness evaluations are performed by humans through an
-interactive system \cite{eiben2015interactive}. They are usually
-applied in problems where the fitness function is not known  or simply
-does not exist, and the result
-of optimization should fit certain human and possibly aesthetic
-need. That is why their use case includes the evolution of objects
-with subjective characteristics such visual appeal or 
+whose fitness evaluations are performed by humans through an interactive system \cite{eiben2015interactive}.
+They are usually applied in problems where the fitness function is not known  or simply
+does not exist, and the result of optimization should fit certain human need. That is why their use 
+case includes the evolution of objects with subjective characteristics such visual appeal or 
 attractiveness \cite{biomorphs} as well as others where human behavior is 
@@
 producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-as well as other artifacts in many other domains \cite{ie1}. In these cases when human interaction is responsible of other 
+as well as other artifacts in many other domains \cite{ie1}. In these cases when 
+human interaction is responsible of other 
 operations of the evolutionary process some authors classify these IEC methods 
 as human-based evolutionary computation cite[kosorukoff] or as human-based computation cite[].
-
+% Problems
+However, the necessary intervention of humans, brings certain challenges to designers of IEC methods  
+namely, human evaluations are slow and expensive, there is a
+human fatigue caused by the interaction \cite{ie1}, and
+there is also the boredom arising when users evaluate a large number of phenotypes 
+many of which are not interesting or are very similar to each other.
 % C-IEAs
-One of the issues with these systems is that their performance effectively depends on the number of users
+Moreover, the performance of theses systems effectively depends on the number of users
 they are able to include; and in order to reach more human users,
@@
 \cite{picbreeder,seyama2016development,wagy2014collective}.
-% Problems
-However, the necessary intervention of humans in these systems leads
-to them having several inherent problems arising from the very nature of 
-the algorithms, namely, human evaluations are slow and expensive, there is a
-human fatigue caused by the interaction \cite{ie1}, and
-the boredom arising when users evaluate a large number of phenotypes 
-many of which are not interesting or are very similar to each other. 
-The number of evaluations that IEC can receive is limited by the number of users
-collaborating and the amount of interaction of each. Having a volunteer system 
-can lower the requirements of anyone participating in
+%The number of evaluations that IEC can receive is limited by the number of users
+%collaborating and the amount of interaction of each. 
+Having a volunteer system  can lower the requirements of anyone participating in
 the experiment thus increasing the {\em performance}, in terms of {\em
@@
 % Volunteers 
-
 Using a volunteer based system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, 
@@
 In order to increase volunteer participation and to tackle some of the issues mentioned above,  
-we proposed a framework following a human centered design \cite{greenhouse2012human},
+we proposed a software framework following a human centered design \cite{greenhouse2012human},
 giving extensive attention to volunteers, not only because their
 explicit evaluation is essential, but also because the context of the 
-interaction affects the system as a whole. In collaborative systems, users interact 
-with each other, and are aware of the actions ref[]. Users also consider their previous evaluations
-and past experiences ref[].
-
-%Please help with this paragraph, is evidence needed?
-% It's the tuning of the algorithm which is data driven, not the
-% algorithm itself, right? - JJ
-%  Tuning yes - Mario
-% And evidence for what? - jj
-%  That most EAs use only population data - Mario
-In evolutionary computation algorithms are data driven since it is important to keep data about the 
-population of each generation, in order to generate the next or even change certain parameters
-or analyze the evolutionary process. 
+interaction affects the system as a whole.
+%Evolutionary computation algorithms are data driven since it is important to keep data about the 
+%population of each generation, in order to generate the next or even change certain parameters
+%or analyze the evolutionary process. 
 % Human Centered Importance
-The framework models the context of interaction along with the population and
-this knowledge is available to be an integral part of the evolutionary computation.
+%The framework models the context of interaction along with the population and
+%this knowledge is available to be an integral part of the evolutionary computation.
 %This could be generalized?
 For example, in a C-IEC application, fitness assignment depend on the
-actions of a social network of users. 
-A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. 
-The selection of parents could also depend also on the previous actions, leveraging information 
-such as the fact that they have the same tag, or are shared by
-similar users. That knowledge could also be applied in choosing which phenotypes present to a certain
-users, in order to give them those they will find more interesting or useful. 
+actions of a social network of users.  Actions are triggered when 
+they tag, share, rate, store or even delete a phenotype. 
+The selection of parents could depend on the previous actions, leveraging information 
+such as the fact that they have the same tag, or where shared by
+similar users.
 % IoT
-The context of interaction, can also include data retrieved from sensors and other IoT devices.
+%The context of interaction, can also include data retrieved from sensors and other IoT devices.
 % Engagement
@@
 conceptual and implementation details of the framework.
-The experiments were implemented using the EvoSpace-Interactive (ES-I) framework \cite{garcia2013evospace}, 
-and where employed in different contexts of interaction. 
-Additional details of each case study application is presented elsewhere refs[][][]. 
-Giving this paper the objective of presenting the underlying conceptual framework.
 
@@
  Evolutionary Computation which 
-the main proposal of this work. An implementation using the ES-I framework is detailed in Section \ref{sec:implementation}.
+the main proposal of this work. An implementation using the ES-I framework 
+is detailed in Section \ref{sec:implementation}.
 The case studies  and results are presented in Section \ref{sec:experiments}.
@@
 \label{sec:related}
+
+An early example of web based collaboration is the work by 
+Langdon \cite{langdon:2004}, which evolved fractal representations of virtual creatures. 
+Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} use web-based IEAs 
+to evolve artistic artifacts using a generative encoding, compositional pattern producing networks.
+Another example was the Galapagos Project \cite{sims1997interactivity},
+an exhibit in the Tokyo Multimedia Museum (1997--2000) were visitors interacted with images presented in 
+twelve displays by selecting those they found most aesthetically interesting by standing on
+step sensors in front of them. 
 % C-IEAS
@@
 IEC in \cite{DBLP:journals/soco/FradeVC12} .  This way the algorithm was capable of providing terrains that would
-otherwise have need users evaluation for these specific features.
+otherwise have need users evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
+also propose the reduction of the user fatigue by using a collaborative filtering algorithm to show only the
+information utilized by similar users as they collaborate with a large number of users for 
+the interactive modeling of 3D glasses.
 
+Some Collaborative IEC (C-IEC) systems promote user engagement by presenting interesting information to 
+users, for instance the genetic lineage of each phenotype or the most popular or 
+best rated solucion\cite{picbreeder,forms}.In a recent work by 
+Wagy \& Bongard \cite{wagy2014collective} user interaction is needed for evaluating the fitness and developing
+new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
+using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
+In any case, using gamification techniques imply that we must deal with interactive evolutionary
+system as socio-technical constructs, where the social aspects are
+essential to understand its dynamics. In this sense, conclusions
+reached with other systems such as NodEO \cite{DBLP:conf/gecco/MereloCGCRV16} can also be applied to this
+system; and applying social network techniques such as graph analysis
+to their study will allow us to understand them more thoroughly.
+
+Collaborative IEC (C-IEC) systems also need to store highly connected data, as it is common in current applications
+like social networks. In order to deal with large datasets of connected data found in these systems,
+graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
+which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
+The framework and data model for representing the interactive process of C-IEC systems is proposed next
 % Volunteer Based
@@
 
-\subsection{Graph}
+\subsection{Graph-Neo4J}
   In order to deal with large datasets of connected data found in these systems,
@@
 
-\subsection{PostgreSQL}
+\subsection{Relational Objects-PostgreSQL}
   The PostgreSQL relational database system is also employed because user sessions and authentication,
@@
 \subsection{Evolution}
+The evolutionary algorithm is decoupled from the data model, the algorithm can be implemented as an
+EvoWorker \cite{garcia2013evospace}, use an external library, or even it can be done by humans.
+An additional tool was implmented for a human based evolution , where volunteers select the parents of
+the next generation and upload the new individuals \cite{de2014artists}.
 
-\section{Case Studies}
-\label{sec:experiments}
+\section{Case Study: EvoDraw}
+As a case study, a IEA was developed using the 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
+A brief description of the application is presented next, focusing
+on the data elements that where ported to the graph model.
 
@@
 
-\subsection{EvoDraw}
-As a case study, a IEA was developed using the 
-ES-I platform \cite{garcia2013evospace}.
-A brief description of the application is presented next, focusing
-on the data elements that where ported to the graph model.
+\label{sec:experiments}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
+    \caption{User interface of the XYZ application.}
+    \label{fig:xyz}
+\end{figure}
 
-\subsection{Fitness Assignment}
+\subsubsection{Fitness Assignment}
 \label{sec:assignment}
@@
 some users that way, the additional information we obtain for scoring
-phenotypes more than compensates for that. % Added this, check it
-                                % out.  
-                                % Ok! - Mario (and knowing how much more is future work)
+phenotypes more than compensates for that. 
 After entering the web application by using their Facebook account,
@@
 Here, the user can interact with the system in two ways.
-First, he can assign a rating % Why don't you use other
-                                % thing instead of quality? We are not
-                                % talking about quality, but about
-                                % tastes or similar - JJ
-                                % is rating Ok?
-% Rating is probably OK. 
- to the phenotype using
+First, he can assign a rating to the phenotype using
 a five star rating system or,
@@
 which is to increase engagement of the users.
-\subsubsection{Graph Model}
 
-\subsubsection{Gamification}
+\subsection{Graph Model}
 
-\subsubsection{Setup}
+The graph model for the EvoDraw applications has the following types of
+nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
+a collection of drawings belonging to users. One collection can contain many phenotypes 
+or be empty and a single phenotype could be shared by many collections. The interaction 
+between these entities are represented by the following edges:
+
+\begin{itemize}
+\item {\bf Likes} This relation describes the interaction between a user and
+a phenotype in which a rating value is assigned this is represented by 
+the {\tt rate} attribute. In the current application a user is allowed to rate
+a phenotype multiple times, so the date of the interactions is also of
+interest.
+
+\item {\bf Knows} The relation connects two users that know each
+other in the Facebook social graph.
+
+\item {\bf Parent} Describes what phenotype is the parent of a new
+  phenotype.
+
+\item {\bf Has} The relation describes an ownership relation between users and
+those collections they own. Also a collection {\tt has}  phenotypes stored
+in them.
+\end{itemize}
+
+\subsection{Gamification}
+The rewarding mechanism as it is applied in EvoDraw 
+consists in giving more importance to the preference of those users having a higher reputation
+as given by the score points and experience levels.
+
+The score achieved by users depends on the actions he does in the system as
+each time a user does one of these actions their score is incremented by one:
+\begin{itemize}
+\item Start a session.
+\item Rate a phenotype.
+\item Create a collection.
+\item Save a phenotype of the wall to a collection.
+\item Save a phenotypes from a friend's collection.
+\item Explore collections of other friends.
+\end{itemize}
+
+Only the ``Start a session'' and ``Explore collections of other friends'' actions 
+are not reflected in the graph model since they are stored instead in the PostgreSQL relational
+database system. This is because user sessions and authentication, as well as basic
+web requests, are handled directly by the Django web framework \cite{garcia2013evospace}.
+
+Two scores are used to determine the weight of a user's preference:
+\begin{itemize}
+\item {\em Experience}: This value depends on the score and is a value 
+between 0 and 100. It is assumed for this case, that once a user
+reaches a 100 actions, it has enough experience on using the application.
+
+\item {\em Participation}: This value is simply the degree of the user node 
+in the graph. This measure considers the number of users he knows,
+collections and phenotypes rated.    
+\end{itemize}
+
+\section{Experimental Setup}
+\label{sec:experiments}
+We present in this section the results of a comparison between three versions
+of the same IEC application. The goal of the experiment is to compare two 
+gamification techniques to increase volunteer participation against a third 
+without any technique applied. The main difference between both versions employing 
+gamification is how the fitness of phenotypes
+is calculated and the information presented to users.
+
+Three versions were compared:
+\begin{itemize}
+\item Base (B). All users have the same weight.
+\item Non Graph Gamification (G). Only experience is considered
+\item Graph Gamification (GG) . Both experience and participation are considered.
+\end{itemize}
+When gamification was employed, all score values known where presented to users
+and a ranking of users by weight was shown in a window.
+
+The three versions of the EvoDrawing web application were deployed to the
+Heroku cloud service with the same configuration: using the
+Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB and RedisTOGO plug-ins.
+Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
+system used in the implementation was Neo4J as it was already available as a plug-in in Heroku, is well documented
+and a scalable solution \cite{miller2013graph,holzschuher2013performance}.
+
+\begin{table}
+  \small
+  \caption{ Parameters for experiments.  }
+  \label{tab:params} 
+  \centering
+  \small
+  \begin{tabular}{l  c   }
+    \hline\noalign{\smallskip}
+     Parameter & Value \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    Initial Population Size   & 80 \\ \hline
+    Sample Size & 1 \\ \hline
+    Step Size & 8 Samples \\ \hline
+    Mutation &  \\ \hline
+    Selection & Tournament \\ \hline
+    Tournament Size &  6 \\ \hline
+  \end{tabular}
+\end{table}
+
+% Anuncements
+At the start of every experiment, a call to participation was issued
+through social networks, 
+particularly on Facebook and Twitter.
+It is important to note that volunteers needed to log-in to Facebook and give permissions to the 
+web application in order to participate, in this case there were no anonymous volunteers. This enabled the automatic access to
+each participant's social network and enabled the system to store a local copy of those friends participating.
+However, another experiment would be needed to see the effect this
+requirement has on participation, since we are not dealing with it in
+this paper.
+
+The link used in the call for participation
+was shortened Google URL, that provides metadata and analytics for the
+users that click on it. 
+In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
+along with the short URL and the analytics link. In the same table a link to the GitHub 
+application repository used to deploy to Heroku is also listed.    
+%Isn't this anonymous? - JJ
+% like in double blind? I think so, that is why no link is given in the table. - Mario
+
+\begin{table}
+  \small
+  \caption{ Deployment data. URLs were {\tt
+      evodrawings0(1,2,3).herokuapp.com}. For analytics URLs, add {\tt
+    .info}.}
+  \label{tab:urls} 
+  \centering
+  \small
+  \begin{tabular}{l l l }
+    \hline\noalign{\smallskip}
+     Deployment &  Shortened URL &  Github Repository \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    B   & goo.gl/jLis4Q &   \\ \hline
+    G   & goo.gl/jqjNy5 &  \\ \hline
+    GG  & goo.gl/J8TCe1 &   \\ \hline
+    \end{tabular}
+\end{table}
+
+Only data for the first week of deployment was considered for the experiments, and they where conducted 
+between January and May of 2016.
+
 \subsubsection{Results} 
+Before launch, each deployment was first tried with a few beta testers. When applying the leader board
+gamification technique for the first time a problem was found in this stage. The problem was that some 
+users were cheating by giving a rating to an animation even before it was returned from the server, this was done by just
+constantly clicking the mouse button. This is a common problem found in systems using leader boards because
+by making the scores visible to other players they are encouraged to compete \cite{hickman2010total}. 
+According to \cite{kumar2013gamification} best-practice gamification designs try to
+refrain from using this element because of  problematic consequences of competition, which can result 
+in negative conduct, low cooperation and collaboration, or disadvantaging certain player
+demographics such as women. In order to still measure the effect of the leader board on the system, the 
+problem of cheating was mitigated by disabling the rate button for a certain amount of time giving the animation
+enough time to finish. But the problem persists, competitive users will try to get in to the leader board by other means.
+
+The results of each of the three experiments in terms of participation are detailed next.
+
+\subsection{ Base deployment}
+After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
+The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
+experiment a total of 500 new phenotypes where generated. The total number of evaluated
+phenotypes for each volunteer is presented in figure
+\ref{fig:top-ranked-participation} in blue. Users are ranked by
+the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+the number of phenotypes evaluated using a logarithmic scale.   % Ranked by... - JJ
+% Done - Mario
+
+\subsection{ Non Graph Gamification deployment}
+The total number of volunteers for this experiment was very close to
+the one obtained in the previous
+experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
+with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
+with a marginal increase in participation. The increased participation also helped in the creation of
+more phenotypes for a total of 648. The total number of evaluated
+phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.
+
+\subsection{ Graph Gamification deployment}
+This experiment had the higher number of volunteers for for a total of 68 registered users. 
+In this experiment a total of 3594 new phenotypes where generated in the experiment. 
+The total number of evaluated phenotypes for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} in green.
+
+The results presented in figure \ref{fig:top-ranked-participation}, show that those users with
+higher level of participation in deployments B and G, had about the same behavior, but the difference
+came with users with medium level of participation, in this case participants where more engaged in deployment
+G. When comparing all the experiments the deployment GG had the higher number of participation in users
+of all levels.
+
+\section{Results}
+Before launch, each deployment was first tried with a few beta testers. When applying the leader board
+gamification technique for the first time a problem was found in this stage. The problem was that some 
+users were cheating by giving a rating to an animation even before it was returned from the server, this was done by just
+constantly clicking the mouse button. This is a common problem found in systems using leader boards because
+by making the scores visible to other players they are encouraged to compete \cite{hickman2010total}. 
+According to \cite{kumar2013gamification} best-practice gamification designs try to
+refrain from using this element because of  problematic consequences of competition, which can result 
+in negative conduct, low cooperation and collaboration, or disadvantaging certain player
+demographics such as women. In order to still measure the effect of the leader board on the system, the 
+problem of cheating was mitigated by disabling the rate button for a certain amount of time giving the animation
+enough time to finish. But the problem persists, competitive users will try to get in to the leader board by other means.
+
+The results of each of the three experiments in terms of participation are detailed next.
+
+\subsubsection{ Base deployment}
+After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
+The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
+experiment a total of 500 new phenotypes where generated. The total number of evaluated
+phenotypes for each volunteer is presented in figure
+\ref{fig:top-ranked-participation} in blue. Users are ranked by
+the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+the number of phenotypes evaluated using a logarithmic scale.   % Ranked by... - JJ
+% Done - Mario
+
+\subsubsection{ Non Graph Gamification deployment}
+The total number of volunteers for this experiment was very close to
+the one obtained in the previous
+experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
+with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
+with a marginal increase in participation. The increased participation also helped in the creation of
+more phenotypes for a total of 648. The total number of evaluated
+phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.
+
+\subsubsection{ Graph Gamification deployment}
+This experiment had the higher number of volunteers for for a total of 68 registered users. 
+In this experiment a total of 3594 new phenotypes where generated in the experiment. 
+The total number of evaluated phenotypes for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} in green.
+
+The results presented in figure \ref{fig:top-ranked-participation}, show that those users with
+higher level of participation in deployments B and G, had about the same behavior, but the difference
+came with users with medium level of participation, in this case participants where more engaged in deployment
+G. When comparing all the experiments the deployment GG had the higher number of participation in users
+of all levels.
 
@@
 
-\subsection{EvoDraw-Kinect}
-\subsection{XYZ-Unplugged}
+%\subsection{EvoDraw-Kinect}
+%\subsection{XYZ-Unplugged}
 
@@
 \label{sec:conclusions}
+
+Collaborative and volunteer based IEC systems involve the dynamic 
+interaction of many entities and artifacts. Employing a multi-model approach
+can benefit researchers to understand and visualize this kind of systems better.
+In this paper a graph model was proposed as a complement to current
+models, initially to improve the visualization of different
+experiments, but in this paper we prove that this model  enables
+the implementation of a gamification technique to improve engagement in
+a case study. Providing a common arena where users are aware of the
+activities of other users in the social vicinity, as well as using
+this graph to assign fitness allows in one hand for a greater
+engagement by the users and from the algorithmic point of view a more
+complex way of assigning fitness and push the generation of content
+forward.
+
+One of the interesting future lines of work would be to look a bit
+more closely at the behavior of users rating the system. These initial
+experiments hint at a possible power law, which might indicate that
+the IEC system could be self-organizing, a process that would allow it
+to reach a critical state, as has been found in software repositories,
+for instance \cite{Merelo2016:repomining}. The dynamics of this kind
+of system are fundamentally different, and our future research will
+include exploring these aspects of the system. Another line of work would be to study the negative effects of using 
+gamification techniques to improve engagement, like cheating or competition.
+Finally, the refinement of the proposed Human-Centered framework will need
+more case studies and further multi-disciplinary research.
 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Results EvoDraw

--- 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
-whose fitness evaluations are performed by humans through an
-interactive system \cite{eiben2015interactive}. They are usually
-applied in problems where the fitness function is not known  or simply
-does not exist, and the result
-of optimization should fit certain human and possibly aesthetic
-need. That is why their use case includes the evolution of objects
-with subjective characteristics such visual appeal or 
-attractiveness \cite{biomorphs} as well as others where human behavior is 
-considered, for instance to optimize teamwork \cite{kosorukoff2002evolutionary}
-or creativity cite[cook].
+whose fitness evaluations are performed by humans through an interactive 
+system \cite{eiben2015interactive}.
+They are usually applied in problems where the fitness function is not known or simply
+does not exist, and the result of optimization should fit a certain human need. 
+That is why their use case includes the evolution of objects with subjective characteristics
+such visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
+considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
+or creativity \cite{yu2011cooks}.
 % Human Based Computation
-These IEC systems are an interesting venue of research, since they have demonstrated their ability for effectively 
+These IEC systems are an interesting venue of research, since they have demonstrated 
+their ability for effectively 
 producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-as well as other artifacts in many other domains \cite{ie1}. In these cases when human interaction is responsible of other 
-operations of the evolutionary process some authors classify these IEC methods 
-as human-based evolutionary computation cite[kosorukoff] or as human-based computation cite[].
-
+as well as other artifacts in many other domains \cite{ie1}. In those cases when 
+human interaction is responsible of other 
+operations of the evolutionary process, some authors classify these IEC methods 
+as human-based evolutionary computation \cite{kosorukoff2001human} 
+or as human-based computation \cite{quinn2011human}.
+% Problems
+However, the necessary intervention of humans, brings certain challenges 
+to designers of IEC methods namely, human evaluations are slow and expensive, there is a
+human fatigue caused by the interaction \cite{ie1}, and
+there is also the boredom arising when users evaluate a large number of phenotypes 
+many of which are not interesting or are very similar to each other.
 % C-IEAs
-One of the issues with these systems is that their performance effectively depends on the number of users
+Moreover, the performance of theses systems effectively depend on the number of users
 they are able to include; and in order to reach more human users,
@@
 with the search, using both anonymous and registered users. Some systems 
-employ a collaborative technique, where several users assign a quality assessment 
-to a single solution and then an aggregated fitness value is calculated 
+employ a collaborative technique, where several users participate in 
+the evaluation, this method is called Collaborative-IEC (C-IEC)
 \cite{picbreeder,seyama2016development,wagy2014collective}.
-% Problems
-However, the necessary intervention of humans in these systems leads
-to them having several inherent problems arising from the very nature of 
-the algorithms, namely, human evaluations are slow and expensive, there is a
-human fatigue caused by the interaction \cite{ie1}, and
-the boredom arising when users evaluate a large number of phenotypes 
-many of which are not interesting or are very similar to each other. 
-The number of evaluations that IEC can receive is limited by the number of users
-collaborating and the amount of interaction of each. Having a volunteer system 
-can lower the requirements of anyone participating in
-the experiment thus increasing the {\em performance}, in terms of {\em
-human brain cycles}, of the whole system.
 % Volunteers 
-
-Using a volunteer based system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, 
-such as the volunteer\'s lack of accountability,
-and the need to build trust between participants and application
-providers. Issues for the provider of the system or organizer  are also the difficulty of establishing 
-the amount of time and resources
-a volunteer is willing to spend on the system, and how they decide if they
-participate or not \cite{JJ:2016}.
+Having a volunteer system can lower the requirements of anyone participating in
+the experiment thus increasing the {\em performance}, in terms of 
+{\em human brain cycles}, of the whole system. But using a volunteer based 
+system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
+volunteer\'s lack of accountability, and the need to build trust between participants and project
+owners. Other issues for project owners are also the difficulty of establishing 
+the amount of time and resources a volunteer is willing to spend on the system, 
+and how they decide if they participate or not \cite{JJ:2016}.
 
@@
 In order to increase volunteer participation and to tackle some of the issues mentioned above,  
-we proposed a framework following a human centered design \cite{greenhouse2012human},
+we proposed a software framework following a human centered design \cite{greenhouse2012human},
 giving extensive attention to volunteers, not only because their
 explicit evaluation is essential, but also because the context of the 
-interaction affects the system as a whole. In collaborative systems, users interact 
-with each other, and are aware of the actions ref[]. Users also consider their previous evaluations
-and past experiences ref[].
+interaction affects the system as a whole.
 
-%Please help with this paragraph, is evidence needed?
-% It's the tuning of the algorithm which is data driven, not the
-% algorithm itself, right? - JJ
-%  Tuning yes - Mario
-% And evidence for what? - jj
-%  That most EAs use only population data - Mario
-In evolutionary computation algorithms are data driven since it is important to keep data about the 
-population of each generation, in order to generate the next or even change certain parameters
-or analyze the evolutionary process. 
-% Human Centered Importance
-The framework models the context of interaction along with the population and
-this knowledge is available to be an integral part of the evolutionary computation.
-%This could be generalized?
-For example, in a C-IEC application, fitness assignment depend on the
-actions of a social network of users. 
-A stream of actions is triggered when they tag, share, rate, store or even delete a phenotype. 
-The selection of parents could also depend also on the previous actions, leveraging information 
-such as the fact that they have the same tag, or are shared by
-similar users. That knowledge could also be applied in choosing which phenotypes present to a certain
-users, in order to give them those they will find more interesting or useful. 
-% IoT
-The context of interaction, can also include data retrieved from sensors and other IoT devices.
-% Engagement
-Data available from the interaction is also used to increase the performance of the system by applying 
-gamification techniques. Gamification is defined by Deterding et al. \cite{deterding2011game} as
+For example, in a C-IEC application, fitness assignment depends on the
+actions of a social network of users.  These actions are triggered when 
+they tag, share, rate, store or even delete a phenotype. 
+Then, the selection of parents could depend on the previous actions, leveraging information 
+such as the fact that they both have the same tag, or where shared by
+similar users.
+
+Data available from the interaction is also used to increase the engegment of 
+users in the system by applying  gamification techniques. Gamification is defined by 
+Deterding et al. \cite{deterding2011game} as
 \begin{quote}
@@
 The gamification element employed in this case study has been a rewarding mechanism  
-\cite{dubois2013understanding}. In general rewards  consist of a reputation system with score points, 
-levels and leader boards. Points are awarded to users in response of 
+\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
+with score points, levels and leader boards. Points are awarded to users in response of 
 the accomplishment of certain activities that need to be encouraged. Levels depend
@@
 they reach a giving level. % Say what we are doing in this case - JJ
-
-Three case studies are presented as a proof of concept, providing both
-conceptual and implementation details of the framework. 
-The experiments were implemented using the EvoSpace-Interactive (ES-I) framework \cite{garcia2013evospace}, 
-and where employed in different contexts of interaction. 
-Additional details of each case study application is presented elsewhere refs[][][]. 
-Giving this paper the objective of presenting the underlying conceptual framework.
+A case study consisting in a web application for evolving artistic drawings
+is presented, providing both conceptual and implementation details of the framework. 
+In order to test the framework three versions of the application are compared to
+measure the degree of relationship between users and their participation,
+when appling a gamification technique.
 
@@
 Section \ref{sec:related} presents related work on the topic 
-of Interactive Evolution.
-Then, Section \ref{sec:framework} presents the framework of Human Interactions in Collaborative Interactive 
- Evolutionary Computation which 
-the main proposal of this work. An implementation using the ES-I framework is detailed in Section \ref{sec:implementation}.
-The case studies  and results are presented in Section \ref{sec:experiments}.
-Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
+of Interactive Evolution. Then, Section \ref{sec:framework} presents the framework of Human
+Interactions in Collaborative Interactive Evolutionary Computation (C-IEC) which 
+the main proposal of this work. An implementation using the ES-I framework 
+is detailed in Section \ref{sec:implementation}.
+The case study and results are presented in Sections \ref{sec:experiments} and  
+\ref{sec:results}. Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
 
@@
 \label{sec:related}
+
+An early example of web based collaboration is the work by 
+Langdon \cite{langdon:2004}, which evolved fractal representations of virtual creatures. 
+Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} 
+use web-based IEAs to evolve artistic artifacts using a generative encoding.
+Another example is the Galapagos Project \cite{sims1997interactivity},
+an exhibit in the Tokyo Multimedia Museum (1997--2000) 
+were visitors interacted with images presented in 
+twelve displays by selecting those they found most aesthetically interesting by standing on
+step sensors in front of them, presenting an example of a non web C-IEC system.
+
+Some C-IEC systems promote user engagement by presenting interesting information to 
+users, for instance the genetic lineage of each phenotype or the most popular or 
+best rated solucion\cite{picbreeder,forms}.In a recent work by 
+Wagy \& Bongard \cite{wagy2014collective} user interaction 
+is needed for evaluating the fitness and developing
+new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
+using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
+In any case, using gamification techniques imply that we must deal with interactive evolutionary
+system as socio-technical constructs, where the social aspects are
+essential to understand its dynamics. In this sense, conclusions
+reached with other systems such as NodEO \cite{DBLP:conf/gecco/MereloCGCRV16}
+can also be applied to this system; and applying social 
+network techniques such as graph analysis
+to their study will allow us to understand them more thoroughly. 
 % C-IEAS
-Given the fatigue fitness evaluations produce in human users when applying IEAs, some authors have tried to mitigate
-the problem by allowing the algorithm to collaborate with the user, so that sometimes users perform the evaluation, 
-but also specific measures are included into the algorithm to perform the evaluation of some features automatically.  
-For instance, Reis et al. added some terrain measures (such as accessibility and edge length) to a standard 
-IEC in \cite{DBLP:journals/soco/FradeVC12} .  This way the algorithm was capable of providing terrains that would
-otherwise have need users evaluation for these specific features.
-
+Given the human fatigue limitation when applying IEAs, some authors 
+have tried to mitigate the problem by allowing the algorithm to 
+collaborate with the user, so that sometimes 
+users perform the evaluation,  but also specific measures are included 
+into the algorithm to perform the
+evaluation of some features automatically. For instance, 
+Reis et al. added some terrain measures (such as accessibility and edge length) 
+to a standard  IEC in \cite{DBLP:journals/soco/FradeVC12}. 
+This way the algorithm was capable of providing  terrains that would otherwise have need
+users evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
+also propose the reduction of the user fatigue by using 
+a collaborative filtering algorithm to show only the information utilized by similar users as 
+they collaborate with a large number of users for the interactive modeling of 3D glasses. 
+The framework and data model for 
+representing the interactive process of C-IEC systems is proposed next
 % Volunteer Based
@@
 \label{sec:framework}
-% Comment: 
-% Maybe in this case we should use the term software framework or architecture. 
 The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
@@
 providing generalized components to developers of IEC systems.
-% Just by saying is Human Centered this is not needed, it was from the User-Centered days - Mario
-%In the
-%description below %below? If that's te chase, we _will_ - JJ
-% Not only in the description, more like a paradigm shitf. But ok for now - Mario
-% we will deliberately use the more general 
-%term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
-%could just be interacting naturally with their environment and
-%unwittingly and at the same time they are implicitly
-%evaluating a possible solution.
 
-%Which framework? Are we talking about a particular framework or that
-%is the general rule for frameworks? - JJ
-%Is general bu I was talking about this one - Mario
 The proposed framework includes components that can be refined to increase
@@
 design. Before diving into details the main design considerations of the
-framework are explained next:  % constraints? Rules? You don't decide
-                               % that users are humans, they simply
-                               % are - JJ 
-                               % Ok, is more like we want to acknowledge that, 
-                               % considerations? - Mario
+framework are explained next:
 
@@
   The framework follows the approach of human centered computing \cite{sebe2010human},
-  in which the context, environment, interfaces, preferences, accessibility, human relations, cognitive
-  limitations, culture, creativity and other human aspects are an integral part of the system.
-  Humans are the computing resorces of the system, having unique characteristics as those identified by 
-  Sun \& Dance \cite{Sun2013}
+  in which the context, environment, interfaces, preferences, accessibility, human relations,
+  cognitive limitations, culture, creativity and other human aspects are an integral part 
+  of the system. Humans are the computing resorces of the system, having unique characteristics
+  as those identified by Sun \& Dance \cite{Sun2013}
 \begin{quote}
-  (1) humans can solve computer hard problems; (2) humans are very good at exception handling, (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are vulnerable to 
-  psychological manipulation, (6) humans are prone to errors,
+  (1) humans can solve computer hard problems; (2) humans are very good at exception handling,
+  (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are
+  vulnerable to psychological manipulation, (6) humans are prone to errors,
   especially for reflective tasks.
@@
 
-\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
-  some times they try to game the system. Project owners must actively promote and design the interactive
-  system to engage volunteers \cite{oh2015clicking}. % Define
-                                % providers somewhere. Experiment
-                                % owners? - JJ
-                                % Using project, like BOINC  https://boinc.berkeley.edu/trac/wiki/VolunteerComputing - M
-
+\item {\bf Users are volunteers} Users donate their computing resources, so they are 
+unaccountable, some times they try to game the system. Project owners must actively promote and
+design the interactive system to engage volunteers \cite{oh2015clicking}. % Define
 \item {\bf Users are not alone}
   Relationship between users in an interactive evolutionary algorithm can be modeled
-  as a social network, with well established semantics, algorithms and metrics \cite{ahuja1993network}.
+  as a social network, with well established semantics, algorithms and metrics 
+  \cite{ahuja1993network}.
   A graph model could enable researchers to find other ways of identifying leaders of 
@@
   goals and purposes the context has when is used in a particular
-  application.   An IEC system will  be used within a certain range of technical, physical and social or
+  application.   An IEC system will  be used within a certain range 
+  of technical, physical and social or
   organizational environments \cite{maguire2001context} that may affect its use.
@@
 
-% Say something about how this is important for the framework and how
-% you incorporate it - JJ
-% Done - Mario
 The above considerations have guided the design of the framework, and througout they have been
@@
 
-\subsection{Graph}
-  In order to deal with large datasets of connected data found in these systems,
-  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
-  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
+\subsection{Graph-Neo4J}
+ Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
+ in current applications like social networks. In order to deal with large datasets of connected 
+ data found in these systems, graph databases \cite{angles2012comparison} have been proposed 
+ as an alternative to relational databases which have performance limitations when dealing with 
+ highly connected data \cite{holzschuher2013performance}.
  A graph is proposed for modeling the social network of users, their interaction with 
@@
  The graph database system used in the implementation is Neo4J as it is
-  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
-  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
-  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
-  are presented.
+ a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well 
+ supported and documented in PaaS infrastructures like Heroku. The Cypher query 
+ language it used to retrieve views from the graph.
+ An exaplmple query is shown in figure \ref{fig:cypher} where the relations 
+ between users and solutions are presented.
 
@@
 
-\subsection{PostgreSQL}
-  The PostgreSQL relational database system is also employed because user sessions and authentication,
+\subsection{Relational Objects-PostgreSQL}
+  The PostgreSQL relational database system is also employed because user 
+  sessions and authentication,
   as well as dynamic web pages are handled directly by the Django web
@@
 \subsection{Interfaces}
-The appropiate human interface will depend on the application domain, the current framework implemantation  
-employs a web based application. Developers can use different templates, to present one or more phenotypes 
-at a time, and the type of rating system: like or one to five stars. Another implementations do not need a
-graphical interface at all, for example the EvoDraw instalation or the XYZ Project \cite{de2014artists}.
+The appropiate human interface will depend on the application domain, 
+the current framework implemantation employs a web based application. Developers 
+can use different templates, to present one or more phenotypes 
+at a time, and the type of rating system: like or one to five stars. 
+Another implementations do not need a graphical interface at all, for example the
+EvoDraw instalation or the XYZ Project \cite{de2014artists}.
 
 \subsection{Evolution}
+The evolutionary algorithm is decoupled from the data model, the algorithm 
+could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
+or even assigned to humans. An additional tool was implmented for a human based evolution, 
+where volunteers select the parents of the next generation and upload the 
+new individuals \cite{de2014artists}.
 
-\section{Case Studies}
-\label{sec:experiments}
+\section{Case Study: EvoDraw}
+As a case study, a IEA was developed using the 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
+A brief description of the application is presented next, focusing
+on the data elements that where ported to the graph model.
 
@@
 
-\subsection{EvoDraw}
-As a case study, a IEA was developed using the 
-ES-I platform \cite{garcia2013evospace}.
-A brief description of the application is presented next, focusing
-on the data elements that where ported to the graph model.
+\label{sec:experiments}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
+    \caption{User interface of the XYZ application.}
+    \label{fig:xyz}
+\end{figure}
 
@@
 \label{sec:assignment}
-% changed to this, because I understand it does not always work that
-% way - JJ
-The ES-I platform has been programmed to employ a collaborative methodology for performing fitness
- assignment. Several registered users assign a quality assessment to a single
+
+The ES-I platform has been programmed to employ a 
+collaborative methodology for performing fitness assignment. 
+Several registered users assign a quality assessment to a single
 phenotype and then an aggregated fitness value is calculated,
-depending on the rating (from 1 to 5 stars) by each particular user,  % how do you define
-                                % "taste"? - JJ
-                                % Using actions
+depending on the rating (from 1 to 5 stars) by each particular user,  
 resulting in a many-to-many relationship between users and
-phenotypes. 
- %scored by them? - JJ 
-Other components or users of the systems can 
+phenotypes. Other components or users of the systems can 
 query this user-phenotype relation to extract relevant
@@
 artifacts. Moreover, users can browse the content of each collection and from
-there share images through their social network.
-When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
-liked the phenotype. The pane also includes a link to the phenotype's 
-details, the parents, genetic operators that created it, and genealogy
-information. This makes the assignment of fitness through the rating
-system a {\em social} activity, pursuing the objective of this work,
-which is to increase engagement of the users.
-\subsubsection{Graph Model}
+there share images through their social network. This makes the assignment 
+of fitness through the rating system a {\em social} activity, 
+pursuing the objective of this work, which is to increase engagement of the users.
 
-\subsubsection{Gamification}
+\subsection{Graph Model} 
+The graph model for the EvoDraw applications has the following types of
+nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
+a collection of drawings belonging to users. One collection can contain many phenotypes 
+or be empty. A single phenotype could be shared by many collections. The interaction 
+between these entities are represented by the following edges:
 
-\subsubsection{Setup}
-\subsubsection{Results}
+\begin{itemize}
+\item {\bf Likes} This relation describes the interaction between a user and
+a phenotype in which a rating value is assigned.
+
+\item {\bf Knows} The relation connects two users that know each
+other in the Facebook social graph.
+
+\item {\bf Parent} Describes what phenotype is the parent of a new
+  phenotype.
+
+\item {\bf Has} The relation describes an ownership relation between users and
+those collections they own. 
+\end{itemize}
+
+\subsection{Gamification}
+The rewarding mechanism as it is applied in EvoDraw gives more importance 
+to the preference of those users with higher reputation
+as given by their score points and experience levels.  
+Each time a user does on of these actions their score is incremented by one:
+\begin{itemize}
+\item Start a session.
+\item Rate a phenotype.
+\item Create a collection.
+\item Save a phenotype of the wall to a collection.
+\item Save a phenotypes from a friend's collection.
+\item Explore collections of other friends.
+\end{itemize}
+
+Two variables are used to determine the weight of a user's preference:
+\begin{itemize}
+\item {\em Experience}: This variable depends on the score and is a value 
+between 0 and 100. It is assumed for this case, that once a user
+reaches a 100 actions, it has enough experience on using the application.
+
+\item {\em Participation}: This variable is simply the degree of the user node 
+in the graph.    
+\end{itemize}
+
+\subsection{Experimental Setup}
+\label{sec:experiments}
+
+Three versions of EvoDraw were compared:
+\begin{itemize}
+\item Base (B). All users have the same weight.
+\item Non Graph Gamification (G). Only experience is considered
+\item Graph Gamification (GG) . Both experience and participation are considered.
+\end{itemize}
+When gamification was employed, all score values known where presented to users
+and a ranking of users by weight was shown in a window.
+The three versions of the EvoDrawing web application were deployed to the
+Heroku cloud service with the same configuration: Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB and RedisTOGO plug-ins.
+
+Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.
+
+\begin{table}
+  \small
+  \caption{ Parameters for experiments.  }
+  \label{tab:params} 
+  \centering
+  \small
+  \begin{tabular}{l  c   }
+    \hline\noalign{\smallskip}
+     Parameter & Value \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    Initial Population Size   & 80 \\ \hline
+    Sample Size & 1 \\ \hline
+    Step Size & 8 Samples \\ \hline
+    Mutation &  \\ \hline
+    Selection & Tournament \\ \hline
+    Tournament Size &  6 \\ \hline
+  \end{tabular}
+\end{table}
+
+% Anuncements
+At the start of every experiment, a call to participation was issued
+through social networks.  The link used in the call for participation
+was shortened Google URL, that provides metadata and analytics for the
+users that click on it.  In table \ref{tab:urls} the URL for each 
+deployment in the Heroku platform is shown,
+along with the short URL and the analytics link. In the same table a link to the GitHub 
+application repository used to deploy to Heroku is also listed.
+
+Only data for the first week of deployment was considered 
+for the experiments, and they where conducted 
+between January and May of 2016.
+
+\subsection{Results} 
+Before launch, each deployment was first tried with a few beta testers. 
+When applying the leader board gamification technique for the first time a 
+problem was found in this stage. The problem was that some users were cheating by giving a
+rating to an animation even before it was returned from the server, this was done by just
+constantly clicking the mouse button. This is a common problem found in systems using leader
+boards because by making the scores visible to other players they are encouraged 
+to compete \cite{hickman2010total}. The results of each of the three experiments in 
+terms of participation are detailed next.
+
+\begin{table}
+  \small
+  \caption{ After a week of the announcement the total number of volunteers, 
+  nodes and edges in the graph and analytics URLs}
+  \label{tab:urls} 
+  \centering
+  \small
+  \begin{tabular}{l l l l l}
+    \hline\noalign{\smallskip}
+     Deployment &  Users &  Nodes &  Edges & URL \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    B   & 53 &  595   & 2220  & goo.gl/jLis4Q.info \\ \hline
+    G   & 54 &  648   & 2596  & goo.gl/jqjNy5.info \\ \hline
+    GG  & 68 &        & 3594  & goo.gl/J8TCe1.info \\ \hline
+    \end{tabular}
+\end{table}
+
+After a week of the announcement the total number of volunteers,nodes and edges 
+in the graph, for for each experiment is shown in table \ref{tab:urls}.
+Moreover, the total number of evaluated phenotypes for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} in blue. Users are ranked by the 
+number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+the number of phenotypes evaluated using a logarithmic scale. 
+Results  show that those users with higher level of participation in deployments B and G, had
+about the same behavior, but the difference came with users with medium level of participation,
+in this case participants where more engaged in deployment G. When comparing all the experiments
+the deployment GG had the higher number of participation in users of all levels.
 
@@
     \centering
-        \includegraphics[width=3.5in]{img/weibull_1.png}
-    \caption{Experiment 1. Weibull Fit}
-    \label{fig:w1}
-\end{figure}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/weibull_2.png}
-    \caption{Experiment 2. Weibull Fit}
-    \label{fig:w2}
-\end{figure}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/weibull_3.png}
-    \caption{Experiment 3. Weibull Fit}
-    \label{fig:w3}
+        \includegraphics[width=4.5in]{img/comparison.png}
+    \caption{Users ranked by the number of phenotypes they rated vs. the number of phenotypes evaluated using a logarithmic scale. }
+    \label{fig:top-ranked-participation}
 \end{figure}
 
-\subsection{EvoDraw-Kinect}
-\subsection{XYZ-Unplugged}
+\section{Case Study: EvoDraw-Kinect}
+\section{Case Study: XYZ-Unplugged}
 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Adds biblio files
But still many errors...
Some more changes
Corrects many addresses and things in bib
Results EvoDraw

@@
 propose a human-centered framework to model 
-the complex interactions on these system.
+the complex interactions  \\on these system.
 Three case studies are presented where the model is applied in the 
@@
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{Interactive evolutionary computation, Human Centered Computing}
 
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sigproc,../bib/biblio}
+\bibliography{sigproc,../bib/biblio,../bib/evospace-i,../bib/volunteer,../bib/geneura}
 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
EvoDraw Installation

@@
     \centering
-        \includegraphics[width=5in]{img/framework.eps}
+        \includegraphics[width=5.5in]{img/framework.eps}
     \caption{IEC Human-centered framework.}
@@
   organizational environments \cite{maguire2001context} that may affect its use.
-  The concept has been formally defined
-  by ISO standard 9241-11 \cite{international1998iso} as\begin{quote}
-  users,
-  tasks and equipment (hardware, software and materials), and the
-  physical and social environments in which a product is used.
-\end{quote}
 
@@
   Real time processing of users' actions could be needed for certain applications when data is 
-  captured by sensors, or directly captured as user input. For example, social networks keep track of 
-  users' interactions with other users, media objects and places. Users of 
+  captured by sensors, or directly captured as user input. For example, social networks keep
+  track of users' interactions with other users, media objects and places. Users of 
   social networks (for instance the Facebook Graph) are accustomed to express these 
@@
     \label{fig:web}
-\end{figure}
-
-\label{sec:experiments}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
-    \caption{User interface of the XYZ application.}
-    \label{fig:xyz}
 \end{figure}
@@
     \centering
-        \includegraphics[width=4.5in]{img/comparison.png}
+        \includegraphics[width=3.3in]{img/comparison.png}
     \caption{Users ranked by the number of phenotypes they rated vs. the number of phenotypes evaluated using a logarithmic scale. }
@@
 \section{Case Study: EvoDraw-Kinect}
+The EvoDraw installation is an IEA in which viewers interact with animations by simply
+looking at them. Drawings emerge from animations scripted in the Processing 
+language, and each is configured by a list of parameters represented by their
+chromosome. To start the evolutionary cycle, first a population of animations is 
+randomly generated and stored in a container server called EvoSpace. 
+Once the population is active,  client computers remove animations from the evospace 
+server to be presented in a monitor one at a time. After a certain amount of time, 
+animations are put back to the server along with information generated from the interaction.
+To continue, another is again removed and presented to viewers. An important feature 
+is that once an animation is removed from the server, no other client can remove it, 
+until the client that remove it puts the animation back, this means that any
+animation is been shown only on one client at a given time. This cycle continues until there
+is enough interactions to create a new generation of animations.
+
+In the previous version of this system \citep{garcia2013evospace} 
+viewers directly assigned a rating to each animation via a web page; in this work, viewers
+only need to see the animations on different monitors, but this time they can look at them 
+or not, just like the paintings inside a gallery. In order to assign a fitness value to 
+each animation and thus  enabling the EA the selection of the more interesting animations, 
+a natural user interface device (Kinect sensor for Xbox One)
+is used for facial tracking. The sensor can track up to 6 persons at once and  
+it provides a basic set of information for every face: where the face is, where it is looking, 
+basic expressive information, and if the face has glasses, which are used to compute the
+attention received by each simulation. Each animation is 
+presented several times in a single generation and the best are always passed to the next, 
+with its fitness recomputed. To generate the new population GA operators
+are applied to the population. One or more client computers are needed in order to run the 
+installation, each needing a Kinect sensor to keep track of viewers.
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/kinect.png}
+    \caption{EvoDraw Instalation on Alife XV.}
+    \label{fig:kinect}
+\end{figure}
+
 \section{Case Study: XYZ-Unplugged}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
+    \caption{User interface of the XYZ application.}
+    \label{fig:xyz}
+\end{figure}
 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
EvoDraw Installation

@@
 \author{Ben Trovato}
-\authornote{Dr.~Trovato insisted his name be first.}
+%\authornote{Dr.~Trovato insisted his name be first.}
 \orcid{1234-5678-9012}
@@
 \author{G.K.M. Tobin}
-\authornote{The secretary disavows any knowledge of this author's actions.}
+%\authornote{The secretary disavows any knowledge of this author's actions.}
 \affiliation{%
@@
 \author{Lars Th{\o}rv{\"a}ld}
-\authornote{This author is the
-  one who did all the really hard work.}
+%\authornote{This author is the
+%  one who did all the really hard work.}
 \affiliation{%
@@
 the algorithms, namely, the human fatigue caused by the interaction, and
-the boredom arising when users evaluate a large number of artifacts 
-many of which are not interesting or are very similar to each other.
+the boredom arising when users evaluate a large number of artifacts.
 To tackle these issues, in this paper we
-propose a human-centered framework to model 
-the complex interactions  \\on these system.
+propose a human-centered framework to modelthe complex interactions on these systems.
 Three case studies are presented where the model is applied in the 
@@
 </concept>
-<concept>
-<concept_id>10010147.10010919.10010172</concept_id>
-<concept_desc>Computing methodologies~Distributed algorithms</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
 </ccs2012>
@@
 \ccsdesc[300]{Computing methodologies~Heuristic function construction}
-\ccsdesc[300]{Computing methodologies~Distributed algorithms}
 % We no longer use \terms command

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Texto añadido para XYZ

@@
   \item {\bf Evolutionary Algorithm} 
-  The EA algorithm interacts with the data model. The EA could also be done with the help of
+  The EA algorithm interacts with the data model. The EA could also be run with the help of
   humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
-  the solutions of the new population.  
+  the solutions of the new population and applied every genetic operator.
 \end{itemize}
@@
 \section{Case Study: XYZ-Unplugged}
+
+XYZ is a collective artwork generated by means of the Unplugged version of the Evolutionary Algorithm.
+While in the previous cases software tools are in charge of running the EA and users collaborate evaluating fitness quality, the software tool in XYZ is only in charge of storing digital versions of the works produced by the team of artists, who are really in charge of applying every step of the evolutionary algorithm.
+
+Thus, the main difference between the standard IEC and the Unplugged EA, is that in the latest, every genetic operation as well as the general evolutionary loop is in charge of human beings.
+Thus the team of artists are in charge of:  (i) setting up the initial population by selecting well known works from the history of arts;  (ii) each of the artists selects every generation two works from the previous one, and produce a new painting by applying any kind of crossover and mutation he may consider (iii) the process repeats for the number of generations initially decided, so that at the end, a collective artwork has been produced including as many paintings per generation, as artists are participating in the experiment.
+
+Artists resort to i-EvoSpace as the tool allowing them to store the paintings, deciding their preferred ones from the previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
+The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to the next generation, etc.
+
 \begin{figure}[!t]

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Weibul fit
@JJ please check, Paco there is a missing reference.

@@
     G   & 54 &  648   & 2596  & goo.gl/jqjNy5.info \\ \hline
-    GG  & 68 &        & 3594  & goo.gl/J8TCe1.info \\ \hline
+    GG  & 68 &  932   & 3594  & goo.gl/J8TCe1.info \\ \hline
     \end{tabular}
@@
 
-After a week of the announcement the total number of volunteers,nodes and edges 
-in the graph, for for each experiment is shown in table \ref{tab:urls}.
-Moreover, the total number of evaluated phenotypes for each volunteer is presented in figure 
-\ref{fig:top-ranked-participation} in blue. Users are ranked by the 
+Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
+in the graph after each experiment. Moreover, the total number of evaluated 
+phenotypes for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} where users are ranked by the 
 number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
-the number of phenotypes evaluated using a logarithmic scale. 
-Results  show that those users with higher level of participation in deployments B and G, had
-about the same behavior, but the difference came with users with medium level of participation,
-in this case participants where more engaged in deployment G. When comparing all the experiments
-the deployment GG had the higher number of participation in users of all levels.
+the number of phenotypes evaluated using a logarithmic scale. Results show that when 
+considering deployments B and G, the difference came with users with a medium level of participation.
+When comparing all the experiments the deployment GG had the higher number of participation in users of all levels.
 
@@
 
+Moreover, modeling the performance of an C-IEC system
+involves understanding its dynamics. Previous works on browser-based
+volunteer computing have used metrics basic such as: 
+the number of users or the time spent by every one in the
+computation \cite{DBLP:journals/gpem/LaredoBGVAGF14, 2016arXiv160101607M}. 
+While works on other platforms such as SETI@home
+\cite{javadi2009mining} have found that the Weibull, log-normal, and
+Gamma distributions where viable modeles of the  
+the availability of computing resources in several clusters.
+This is in concordance with the results obtained in \cite{agajaj}, and also
+browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
+The shape of those distributions is a skewed bell
+with more resources in the {\em low} areas than in the high areas:
+there are many users that give a small amount of cycles, while there
+are just a few that give many cycles.
+
+In orther to assess if C-IEC follow the same pattern, even when
+using gamification techniques, participation was fittted to a 
+Weibull distribution, results in figures \ref{fig:w1}--\ref{fig:w3}  show that 
+the same pattern appears in this case study.
+
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3in]{img/weibull_1.png}
+    \caption{Experiment 1. Weibull Fit}
+    \label{fig:w1}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3in]{img/weibull_2.png}
+    \caption{Experiment 2. Weibull Fit}
+    \label{fig:w2}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3in]{img/weibull_3.png}
+    \caption{Experiment 3. Weibull Fit}
+    \label{fig:w3}
+\end{figure}
+
 \section{Case Study: EvoDraw-Kinect}
-The EvoDraw installation is an IEA in which viewers interact with animations by simply
-looking at them. Drawings emerge from animations scripted in the Processing 
-language, and each is configured by a list of parameters represented by their
-chromosome. To start the evolutionary cycle, first a population of animations is 
-randomly generated and stored in a container server called EvoSpace. 
-Once the population is active,  client computers remove animations from the evospace 
-server to be presented in a monitor one at a time. After a certain amount of time, 
-animations are put back to the server along with information generated from the interaction.
-To continue, another is again removed and presented to viewers. An important feature 
-is that once an animation is removed from the server, no other client can remove it, 
+
+A variation of the previous application is The EvoDraw installation , using the same
+framework but with a different form of interaction. Humans interact with the animations 
+by simply looking at them. Again the population is stored in the EvoSpace container and
+client computers remove animations from the EvoSpace to be presented in a display one at a time. 
+After a certain amount of time,  animations are put back to the server along with information 
+generated from the interaction. To continue, another is again removed and presented to viewers. 
+An important feature is that once an animation is removed from the server, no other client can remove it, 
 until the client that remove it puts the animation back, this means that any
@@
 
-In the previous version of this system \citep{garcia2013evospace} 
-viewers directly assigned a rating to each animation via a web page; in this work, viewers
-only need to see the animations on different monitors, but this time they can look at them 
-or not, just like the paintings inside a gallery. In order to assign a fitness value to 
-each animation and thus  enabling the EA the selection of the more interesting animations, 
-a natural user interface device (Kinect sensor for Xbox One)
-is used for facial tracking. The sensor can track up to 6 persons at once and  
-it provides a basic set of information for every face: where the face is, where it is looking, 
-basic expressive information, and if the face has glasses, which are used to compute the
-attention received by each simulation. Each animation is 
+In order to assign a fitness value to each animation and thus enabling the EA the selection of the more interesting animations, a natural user interface device (Kinect sensor for Xbox One)
+is used for facial tracking. Each animation is 
 presented several times in a single generation and the best are always passed to the next, 
-with its fitness recomputed. To generate the new population GA operators
-are applied to the population. One or more client computers are needed in order to run the 
-installation, each needing a Kinect sensor to keep track of viewers.
+with its fitness recomputed. To generate the new population the same operators as the previous case study
+are applied to the population.
 
@@
         \includegraphics[width=3.5in]{img/kinect.png}
-    \caption{EvoDraw Instalation on Alife XV.}
+    \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}
     \label{fig:kinect}
 \end{figure}
+
+In this instalation the components of the framework used where the EvoSpace and the Evolutionary Algorithm.
+Information from users was limited to a stream of readings from the sensor.
 
@@
 XYZ is a collective artwork generated by means of the Unplugged version of the Evolutionary Algorithm.
-While in the previous cases software tools are in charge of running the EA and users collaborate evaluating fitness quality, the software tool in XYZ is only in charge of storing digital versions of the works produced by the team of artists, who are really in charge of applying every step of the evolutionary algorithm.
+While in the previous cases software tools are in charge of running the EA and users collaborate evaluating
+fitness quality, the sframework in XYZ is only in charge of storing digital versions of the works produced 
+by the team of artists, who are really in charge of applying every step of the evolutionary algorithm. In 
+this case only the relations are stored, and a basic user profile.
 
-Thus, the main difference between the standard IEC and the Unplugged EA, is that in the latest, every genetic operation as well as the general evolutionary loop is in charge of human beings.
-Thus the team of artists are in charge of:  (i) setting up the initial population by selecting well known works from the history of arts;  (ii) each of the artists selects every generation two works from the previous one, and produce a new painting by applying any kind of crossover and mutation he may consider (iii) the process repeats for the number of generations initially decided, so that at the end, a collective artwork has been produced including as many paintings per generation, as artists are participating in the experiment.
+Thus, the main difference between the standard IEC and the Unplugged EA, is that in the latest,
+every genetic operation as well as the general evolutionary loop is in charge of human beings. 
+Thus the team of artists are in charge of:  (i) setting up the initial population by selecting
+well known works from the history of arts;  (ii) each of the artists selects every generation two works 
+from the previous one, and produce a new painting by applying any kind of crossover and mutation he may consider 
+(iii) the process repeats for the number of generations initially decided, so that at the end, a collective 
+artwork has been produced including as many paintings per generation, as artists are participating in the experiment.
 
-Artists resort to i-EvoSpace as the tool allowing them to store the paintings, deciding their preferred ones from the previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
-The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to the next generation, etc.
+Artists resort to framework as the tool allowing them to store the paintings, deciding their preferred ones from the
+previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
+The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to 
+the next generation, etc.
 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Back to 8 pages, complete with three case studies and distribution fit

@@
 
-  \subsection{EvoSpace-Redis}
+\begin{itemize}
+\item {\bf EvoSpace-Redis} 
 EvoSpace is a population store \cite{Evospace}  for the development of 
@@
 
-\subsection{Graph-Neo4J}
+\item {\bf Graph-Neo4J} 
  Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
@@
 
-\subsection{Relational Objects-PostgreSQL}
+\item {\bf Relational Objects-PostgreSQL}
   The PostgreSQL relational database system is also employed because user 
@@
 
-\subsection{Interfaces}
+\item {\bf Interfaces}
 The appropiate human interface will depend on the application domain, 
@@
 
-\subsection{Evolution}
+\item {\bf Evolution}    
 The evolutionary algorithm is decoupled from the data model, the algorithm 
@@
 new individuals \cite{de2014artists}.  
+\end{itemize}
 
@@
 The three versions of the EvoDrawing web application were deployed to the
-Heroku cloud service with the same configuration: Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB and RedisTOGO plug-ins.
+Heroku cloud service with the same configuration: Heroku Free option with a 512 MB of RAM, 
+adding only the GrapheneDB and RedisTOGO plug-ins.
 
@@
   \small
-  \begin{tabular}{l  c   }
+  \begin{tabular}{l  c}
     \hline\noalign{\smallskip}
@@
 using gamification techniques, participation was fittted to a 
-Weibull distribution, results in figures \ref{fig:w1}--\ref{fig:w3}  show that 
-the same pattern appears in this case study.
+Weibull distribution, results in figure \ref{fig:w2} and figure \ref{fig:w3}
+show that the same pattern appears in this case study.
 
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3in]{img/weibull_1.png}
-    \caption{Experiment 1. Weibull Fit}
-    \label{fig:w1}
-\end{figure}
 \begin{figure}[!t]
@@
 
+Collaborative and volunteer based IEC systems involve the dynamic 
+interaction of many entities and artifacts. Employing human centered approach 
+can benefit researchers to understand and visualize this kind of systems better.
+In this paper a human-centered software framework was proposed, and was validated
+through the implementaion and refinement of three C-IEC applications.
+This framework enabled the implementation of a gamification technique 
+to improve engagement in a case study. Providing a common arena where users are aware of the
+activities of other users in the social vicinity. Following a hybrid data model approach
+gives developers and project owners options when innovating with other means
+of interaction or new applications.
+
+One of the interesting future lines of work would be to look a bit
+more closely at the behavior of users as they are rating artifacts 
+in the web system. These initial experiments hint at a possible power law, which might indicate that
+the IEC system could be self-organizing, a process that would allow it
+to reach a critical state, as has been found in software repositories,
+for instance \cite{Merelo2016:repomining}. 
+The dynamics of this kind of system are fundamentally different, and our future research will
+include exploring these aspects of the system.
+
+Another line of work would be to study the negative effects of using 
+gamification techniques to improve engagement, like cheating or competition.
+Finally, the refinement of the proposed Human-Centered framework will need
+more case studies and further multi-disciplinary research.
+
 \begin{acks}

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Typo arreglado
otro typo arreglado

@@
 to a standard  IEC in \cite{DBLP:journals/soco/FradeVC12}. 
-This way the algorithm was capable of providing  terrains that would otherwise have need
+This way the algorithm was capable of providing  terrains that would otherwise have needed
 users evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
@@
 queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
-candidate solutions are retieved from the server, evaluated, and then sent back. 
+candidate solutions are retrieved from the server, evaluated, and then sent back. 
 The same operations are used to Evolve the population. In EvoSpace individuals replaced 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
exaplmple arreglado
implmented arreglado

@@
  language it used to retrieve views from the graph.
- An exaplmple query is shown in figure \ref{fig:cypher} where the relations 
+ An example query is shown in figure \ref{fig:cypher} where the relations 
  between users and solutions are presented.
@@
 could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
-or even assigned to humans. An additional tool was implmented for a human based evolution, 
+or even assigned to humans. An additional tool was implemented for a human based evolution, 
 where volunteers select the parents of the next generation and upload the 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
the the arreglado

@@
 knowledge about the process and the population, for instance showing the
-most popular phenotype, or the the user with more participation 
+most popular phenotype, or the user with more participation 
 \cite{picbreeder}.

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
figure arreglado
Addressing Leo´s comments

@@
 They are usually applied in problems where the fitness function is not known or simply
-does not exist, and the result of optimization should fit a certain human need. 
-That is why their use case includes the evolution of objects with subjective characteristics
+does not exist, and the result of optimization should fit a certain human need or desire. 
+That is why their use-case includes the evolution of objects with subjective characteristics
 such visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
@@
 human interaction is responsible of other 
-operations of the evolutionary process, some authors classify these IEC methods 
+aspects of the evolutionary process, some authors classify these IEC methods 
 as human-based evolutionary computation \cite{kosorukoff2001human} 
@@
 However, the necessary intervention of humans, brings certain challenges 
-to designers of IEC methods namely, human evaluations are slow and expensive, there is a
-human fatigue caused by the interaction \cite{ie1}, and
-there is also the boredom arising when users evaluate a large number of phenotypes 
+to designers of IEC methods namely, human evaluations are slow and expensive, there is
+human fatigue caused by the interaction \cite{ie1}, and also boredom arise
+when users evaluate a large number of phenotypes, 
 many of which are not interesting or are very similar to each other.
 % C-IEAs
-Moreover, the performance of theses systems effectively depend on the number of users
-they are able to include; and in order to reach more human users,
-Interactive Evolutionary Algorithms (IEAs) are 
-some times developed as web applications depending on visitors to the web to interactively help
+Moreover, the performance of these systems effectively depends on the number of users
+they are able to include; to reach more users,
+IEC systems are some times developed as web applications depending on web visitors to help
 with the search, using both anonymous and registered users. Some systems 
@@
 on the Wall, he can choose to retrieve a new one from the population.
-At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
+At the left hand side of Figure \ref{fig:web}, the web page shows the \emph{Collections} section.
 The user can create several collections, to group and organize his favorite 

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Texto añadido para XYZ
Addressing Leo´s comments

@@
 To tackle these issues, in this paper we
-propose a human-centered framework to modelthe complex interactions on these systems.
+propose a human-centered framework to model complex interactions on these systems.
 Three case studies are presented where the model is applied in the 
@@
 implies that this technique can successfully be used to 
-decrease user fatigue and thus increase {\em performance} of the
+decrease user fatigue and boredom, and thus increase {\em performance} of the
 interactive system.

### 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
going home

@@
 % Volunteers 
-Having a volunteer system can lower the requirements of anyone participating in
-the experiment thus increasing the {\em performance}, in terms of 
-{\em human brain cycles}, of the whole system. But using a volunteer based 
+Having a volunteer system can lower the requirements for anyone to participate in
+the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
 system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
@@
 
-For example, in a C-IEC application, fitness assignment depends on the
+For example, in a C-IEC application fitness assignment depends on the
 actions of a social network of users.  These actions are triggered when 
-they tag, share, rate, store or even delete a phenotype. 
-Then, the selection of parents could depend on the previous actions, leveraging information 
-such as the fact that they both have the same tag, or where shared by
+they tag, share, rate, store or delete a phenotype. 
+Then, the selection of parents could depend on the previous actions, leveraging information, 
+such as the fact that they both have the same tag, or were shared by
 similar users.
@@
 they reach a giving level. % Say what we are doing in this case - JJ
-A case study consisting in a web application for evolving artistic drawings
-is presented, providing both conceptual and implementation details of the framework. 
-In order to test the framework three versions of the application are compared to
-measure the degree of relationship between users and their participation,
-when appling a gamification technique.
+%%%
+%%% Change if case studies change
+%In this work, a case study consisting in a web application for evolving artistic drawings
+%is presented, providing both conceptual and implementation details of the framework. 
+%In order to test the framework three versions of the application are compared to
+%measure the degree of relationship between users and their participation,
+%when appling a gamification technique. 
+In this work, three case studies are presented. First, the development of a web based C-IEC application 
+for evolving artistic drawings is described, to give the reader details about 
+the utilization of the proposed framework. Then, a rewarding mechanism is 
+implemented in the same application, to study the changes in participation,
+when appling a gamification technique. Later two other case studies are briefly presented
+to show how the framework is applied in other domains.
 
@@
 Section \ref{sec:related} presents related work on the topic 
-of Interactive Evolution. Then, Section \ref{sec:framework} presents the framework of Human
-Interactions in Collaborative Interactive Evolutionary Computation (C-IEC) which 
-the main proposal of this work. An implementation using the ES-I framework 
+of Interactive Evolution. Then, Section \ref{sec:framework} presents the human centered framework of
+for C-IEC appliactions which the main proposal of this work. Next three case studies are presented in
+sections
+
+An implementation using the ES-I framework 
 is detailed in Section \ref{sec:implementation}.

### 2017-02-04 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
EvoDrawings,EvoDraw, organization

@@
 In this work, three case studies are presented. First, the development of a web based C-IEC application 
-for evolving artistic drawings is described, to give the reader details about 
+for evolving artistic drawings is described to give the reader details about 
 the utilization of the proposed framework. Then, a rewarding mechanism is 
-implemented in the same application, to study the changes in participation,
-when appling a gamification technique. Later two other case studies are briefly presented
-to show how the framework is applied in other domains.
+implemented in the same application to study the changes in participation
+when applying a gamification technique. Finally, two other case studies are briefly presented
+to show how the framework is applied to other domains.
 
@@
 Section \ref{sec:related} presents related work on the topic 
-of Interactive Evolution. Then, Section \ref{sec:framework} presents the human centered framework of
-for C-IEC appliactions which the main proposal of this work. Next three case studies are presented in
-sections
-
-An implementation using the ES-I framework 
-is detailed in Section \ref{sec:implementation}.
-The case study and results are presented in Sections \ref{sec:experiments} and  
-\ref{sec:results}. Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.
+of interactive evolution. Then, Section \ref{sec:framework} presents the human centered framework of
+for C-IEC appliactions which the main proposal of this work. The implementation of the data model
+is detailed in Section \ref{sec:implementation}. Next three case studies are presented: EvoDrawings in
+Section \ref{sec:evodraw}, EvoDraw Installation in Section\ref{sec:evodraw}, 
+and XYZ in Section\ref{sec:evodraw}. Finally, a concluding remarks are provided 
+in Section \ref{sec:conclusions}.
 
@@
 Another implementations do not need a graphical interface at all, for example the
-EvoDraw instalation or the XYZ Project \cite{de2014artists}.
+EvoDrawings instalation or the XYZ Project \cite{de2014artists}.
 
@@
 
-\section{Case Study: EvoDraw}
+\section{Case Study: EvoDrawings}
+\label{sec:evodraw}
 As a case study, a IEA was developed using the 
@@
 \subsection{Fitness Assignment}
-\label{sec:assignment}
-
 The ES-I platform has been programmed to employ a 
@@
 \subsection{Graph Model} 
-The graph model for the EvoDraw applications has the following types of
+The graph model for the EvoDrawings application has the following types of
 nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
@@
 \subsection{Gamification}
-The rewarding mechanism as it is applied in EvoDraw gives more importance 
+The rewarding mechanism as it is applied in EvoDrawings gives more importance 
 to the preference of those users with higher reputation
@@
 \subsection{Experimental Setup}
-\label{sec:experiments}
 
-Three versions of EvoDraw were compared:
+Three versions of EvoDrawings were compared:
 \begin{itemize}
@@
 and a ranking of users by weight was shown in a window.
-The three versions of the EvoDrawing web application were deployed to the
-Heroku cloud service with the same configuration: Heroku Free option with a 512 MB of RAM, 
-adding only the GrapheneDB and RedisTOGO plug-ins.
-
 Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.
@@
 \section{Case Study: EvoDraw-Kinect}
-
+\label{sec:kinect}
 A variation of the previous application is The EvoDraw installation , using the same
@@
 \section{Case Study: XYZ-Unplugged}
+\label{sec:xyz}
 

### 2017-02-04 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
More corrections

@@
 users, for instance the genetic lineage of each phenotype or the most popular or 
-best rated solucion\cite{picbreeder,forms}.In a recent work by 
+best rated solucion \cite{picbreeder,forms}. In a recent work by 
 Wagy \& Bongard \cite{wagy2014collective} user interaction 
-is needed for evaluating the fitness and developing
+is needed for evaluating fitness and developing
 new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
 using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
-In any case, using gamification techniques imply that we must deal with interactive evolutionary
-system as socio-technical constructs, where the social aspects are
+In any case, using gamification techniques imply that we must deal with IEC
+systems as socio-technical constructs, where the social aspects are
 essential to understand its dynamics. In this sense, conclusions
 reached with other systems such as NodEO \cite{DBLP:conf/gecco/MereloCGCRV16}
-can also be applied to this system; and applying social 
+can also be applied to these systems; and applying social 
 network techniques such as graph analysis
@@
 to a standard  IEC in \cite{DBLP:journals/soco/FradeVC12}. 
-This way the algorithm was capable of providing  terrains that would otherwise have needed
-users evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
-also propose the reduction of the user fatigue by using 
+This way the algorithm was capable of providing terrains that would otherwise have needed the
+users' evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
+also propose the reduction of user fatigue by using 
 a collaborative filtering algorithm to show only the information utilized by similar users as 
@@
 The framework and data model for 
-representing the interactive process of C-IEC systems is proposed next
+representing the interactive process of our C-IEC systems is proposed next.
 % Volunteer Based
@@
 \label{sec:framework}
-The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
-software framework for interactive evolutionary computation (IEC). 
+The general goal of this research is to develop a human-centered \cite{gasson2003human} 
+software framework for C-IEC. 
 A framework is defined as a reusable architectural design together
@@
 participation and also to minimize the amount of evaluations needed for the evolutionary 
-process in a given IEC application. Software frameworks often have a vison or a way 
-of thinking \cite{carneiro2010introducing} guiding their
+process in a given IEC application. Software frameworks often have 
+a vison \cite{carneiro2010introducing} guiding their
 design. Before diving into details the main design considerations of the
@@
 \begin{itemize}
-\item {\bf Users are human} 
+\item {\bf Users are human.} 
   The framework follows the approach of human centered computing \cite{sebe2010human},
@@
   of the system. Humans are the computing resorces of the system, having unique characteristics
-  as those identified by Sun \& Dance \cite{Sun2013}
+  as those identified by Sun \& Dance \cite{Sun2013}:
 \begin{quote}
@@
 
-\item {\bf Users are volunteers} Users donate their computing resources, so they are 
-unaccountable, some times they try to game the system. Project owners must actively promote and
+\item {\bf Users are volunteers.} Users donate their computing resources, so they are 
+unaccountable, sometimes they try to game the system. Project owners must actively promote and
 design the interactive system to engage volunteers \cite{oh2015clicking}. % Define
@@
 
-\item {\bf Context of use matters}
-  \textit{Context} In computer science Fischer \cite{fischer2012context}
+\item {\bf Context of use matters.}
+  Fischer \cite{fischer2012context}
   defines context as
@@
 \end{quote}   
-  Fischer also identifies the important aspects to consider when the context is used: how it is
-  the contextual data obtained, how the context is represented and what
-  goals and purposes the context has when is used in a particular
+  Fischer also identifies the important aspects to consider when the context is used: how is
+  contextual data obtained, how is context represented and what
+  goals and purposes the context has when it is used in a particular
   application. An IEC system will  be used within a certain range 
@@
 
-\item {\bf Interaction is a stream of actions}
+\item {\bf Interaction is a stream of actions.}
   Real time processing of users' actions could be needed for certain applications when data is 
@@
 
-\begin{itemize}
-  \item {\bf Interactive Evolutionary Computational system} 
+\begin{enumerate}
+  \item {\bf Interactive System.} 
   This is the real world system that we are going to represent in the data model, 
@@
 
-  \item {\bf IEC Data Model}
+  \item {\bf IEC Data Model.}
   A data model is used to describe the IEC system prior to a physical 
   implementation.  Depending on the domain several techniques can be used.
-  Parts of the system could be better described by an Entity relationship 
-  modeling to be used in a relational database or a class digram for a 
+  Parts of the system could be better described by an entity-relationship 
+  modeling, to be used in a relational database or a class digram for a 
   key-value store. A graph is proposed for modeling the social network of users 
-  and their interaction with the population. When implemented a graph database 
+  and their interaction with the population. When implemented, a graph database 
   will be the back-end of the system.
 
-  \item {\bf Evolutionary Algorithm} 
-  The EA algorithm interacts with the data model. The EA could also be run with the help of
+  \item {\bf Evolutionary Algorithm.} 
+  The Evolutionary Algorithm (EA) algorithm interacts with the data model. The EA could also be run with the help of
   humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
   the solutions of the new population and applied every genetic operator.
-\end{itemize}
+\end{enumerate}
 
@@
 \begin{itemize}
-\item {\bf EvoSpace-Redis} 
+\item {\bf EvoSpace-Redis.} 
 EvoSpace is a population store \cite{Evospace}  for the development of 
@@
 is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
-database management system, because it provides a hash based implementation of sets and 
-queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
-candidate solutions are retrieved from the server, evaluated, and then sent back. 
-The same operations are used to Evolve the population. In EvoSpace individuals replaced 
+database management system because it provides a hash based implementation of sets and 
+queues which are natural data structures for the EvoSpace model. Basicaly, a sample of 
+candidate solutions are retrieved from the server, evaluated and then sent back. 
+The same operations are used to evolve the population. In EvoSpace individuals replaced 
 in the population are stored indefinitly, to permit users to store permalinks to them.
@@
 
-\item {\bf Graph-Neo4J} 
+\item {\bf Graph-Neo4J.} 
  Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
@@
  language it used to retrieve views from the graph.
- An example query is shown in figure \ref{fig:cypher} where the relations 
+ An example query is shown in Figure \ref{fig:cypher} where the relations 
  between users and solutions are presented.
@@
 
-\item {\bf Relational Objects-PostgreSQL}
+\item {\bf Relational Objects-PostgreSQL.}
   The PostgreSQL relational database system is also employed because user 
@@
 
-\item {\bf Interfaces}
+\item {\bf Interfaces.}
 The appropiate human interface will depend on the application domain, 
@@
 can use different templates, to present one or more phenotypes 
-at a time, and the type of rating system: like or one to five stars. 
-Another implementations do not need a graphical interface at all, for example the
+at a time, and the type of rating system: like based or a rating from one to five (stars). 
+Other implementations do not need a graphical interface at all, for example the
 EvoDrawings instalation or the XYZ Project \cite{de2014artists}.
 
-\item {\bf Evolution}    
-The evolutionary algorithm is decoupled from the data model, the algorithm 
+\item {\bf Evolution.}    
+The EA is decoupled from the data model, the algorithm 
 could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
@@
 \label{sec:evodraw}
-As a case study, a IEA was developed using the 
+As a case study, a C-IEC application was developed using the 
 EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
@@
 query this user-phenotype relation to extract relevant
-knowledge about the process and the population, for instance showing the
+knowledge about the process and the population. For instance, showing the
 most popular phenotype, or the user with more participation 
@@
 either because they do not want to give that information or simply because
-they do not use that particular social network. Even as we might lose
+they do not use that social network. Even if we might lose
 some users that way, the additional information we obtain for scoring
@@
 Here, the user can interact with the system in two ways.
-First, he can assign a rating to the phenotype using
-a five star rating system or,
-additionally, a user can choose to add an image to one of their \emph{Collections}.
+First, he can assign a rating to the phenotype or choose to add an image to one of their \emph{Collections}.
 A collection is a special folder that stores those phenotypes a user likes and wishes

### 2017-02-04 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Corrections

@@
 of fitness through the rating system a {\em social} activity, 
-pursuing the objective of this work, which is to increase engagement of the users.
+pursuing the objective of this work, which is to increase user engagement.
 
@@
 Each time a user does on of these actions their score is incremented by one:
-\begin{itemize}
-\item Start a session.
-\item Rate a phenotype.
-\item Create a collection.
-\item Save a phenotype of the wall to a collection.
-\item Save a phenotypes from a friend's collection.
-\item Explore collections of other friends.
-\end{itemize}
+start a session, rate a phenotype, create a collection, save a phenotype of 
+the wall to a collection, save a phenotypes from a friend's collection, and
+explore collections of other friends.
 
@@
 \begin{itemize}
-\item {\em Experience}: This variable depends on the score and is a value 
+\item {\bf Experience.}: This variable depends on the score and is a value 
 between 0 and 100. It is assumed for this case, that once a user
@@
 
-\item {\em Participation}: This variable is simply the degree of the user node 
-in the graph.    
+\item {\bf Participation.}: This variable is simply the degree of the user node 
+in the graph (number of edges).    
 \end{itemize}
@@
 \begin{itemize}
-\item Base (B). All users have the same weight.
-\item Non Graph Gamification (G). Only experience is considered
-\item Graph Gamification (GG) . Both experience and participation are considered.
+\item Base (B): All users have the same weight.
+\item Non Graph Gamification (G): Only experience is considered
+\item Graph Gamification (GG): Both experience and participation are considered.
 \end{itemize}
 When gamification was employed, all score values known where presented to users
-and a ranking of users by weight was shown in a window.
-Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.
+and a ranking of users by weight was shown in a window. Table \ref{tab:params} shows 
+the parameters used for the evolutionary algorithm.
 
@@
 was shortened Google URL, that provides metadata and analytics for the
-users that click on it.  In table \ref{tab:urls} the URL for each 
+users that click on it.  In Table \ref{tab:urls} the URL for each 
 deployment in the Heroku platform is shown,
 along with the short URL and the analytics link. In the same table a link to the GitHub 
-application repository used to deploy to Heroku is also listed.
-
-Only data for the first week of deployment was considered 
-for the experiments, and they where conducted 
+application repository used to deploy to Heroku is also listed.Only data 
+for the first week of deployment was considered for the experiments, and they where conducted 
 between January and May of 2016.
@@
 boards because by making the scores visible to other players they are encouraged 
-to compete \cite{hickman2010total}. The results of each of the three experiments in 
-terms of participation are detailed next.
+to compete \cite{hickman2010total}. The final version disabled the button until 
+the drawing animation was over. The results of each of the three experiments in 
+terms of participation are dtailed next.
 
@@
 \ref{fig:top-ranked-participation} where users are ranked by the 
-number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+number of phenotypes they rated. In the \emph{x} axis is the rank and in the \emph{y} axis 
 the number of phenotypes evaluated using a logarithmic scale. Results show that when 
@@
 involves understanding its dynamics. Previous works on browser-based
-volunteer computing have used metrics basic such as: 
+volunteer computing have used basic metrics such as: 
 the number of users or the time spent by every one in the
@@
 The shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas:
+with more resources in the {\em low} areas than in the high areas; i.e., 
 there are many users that give a small amount of cycles, while there
@@
 
-In orther to assess if C-IEC follow the same pattern, even when
+In order to assess if C-IEC follows the same pattern, even when
 using gamification techniques, participation was fittted to a 
-Weibull distribution, results in figure \ref{fig:w2} and figure \ref{fig:w3}
+Weibull distribution. Results in Figure \ref{fig:w2} and Figure \ref{fig:w3}
 show that the same pattern appears in this case study.
@@
     \centering
-        \includegraphics[width=3in]{img/weibull_2.png}
-    \caption{Experiment 2. Weibull Fit}
+        \includegraphics[width=4in]{img/weibull_2.png}
+    \caption{Experiment G. Weibull Fit}
     \label{fig:w2}
@@
     \centering
-        \includegraphics[width=3in]{img/weibull_3.png}
-    \caption{Experiment 3. Weibull Fit}
+        \includegraphics[width=4in]{img/weibull_3.png}
+    \caption{Experiment GG. Weibull Fit}
     \label{fig:w3}
@@
 \label{sec:kinect}
-A variation of the previous application is The EvoDraw installation , using the same
+A variation of the previous application is The EvoDraw installation, using the same
 framework but with a different form of interaction. Humans interact with the animations 
@@
 generated from the interaction. To continue, another is again removed and presented to viewers. 
-An important feature is that once an animation is removed from the server, no other client can remove it, 
-until the client that remove it puts the animation back, this means that any
-animation is been shown only on one client at a given time. This cycle continues until there
-is enough interactions to create a new generation of animations.
+This cycle continues until there is enough interactions to create a new generation of animations.
 
-In order to assign a fitness value to each animation and thus enabling the EA the selection of the more interesting animations, a natural user interface device (Kinect sensor for Xbox One)
-is used for facial tracking. Each animation is 
+In order to assign a fitness value to each animation and thus enabling 
+the EA the selection of the most interesting animations, a natural user interface 
+device (Kinect sensor for Xbox One) is used for facial tracking. Each animation is 
 presented several times in a single generation and the best are always passed to the next, 
@@
     \centering
-        \includegraphics[width=3.5in]{img/kinect.png}
+        \includegraphics[width=4in]{img/kinect.png}
     \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}
@@
 
-In this instalation the components of the framework used where the EvoSpace and the Evolutionary Algorithm.
+In this instalation the components of the framework used where EvoSpace and the EA.
 Information from users was limited to a stream of readings from the sensor.
@@
 
-XYZ is a collective artwork generated by means of the Unplugged version of the Evolutionary Algorithm.
-While in the previous cases software tools are in charge of running the EA and users collaborate evaluating
-fitness quality, the sframework in XYZ is only in charge of storing digital versions of the works produced 
-by the team of artists, who are really in charge of applying every step of the evolutionary algorithm. In 
-this case only the relations are stored, and a basic user profile.
+XYZ is a collective artwork generated by means of the Unplugged version of the EA \cite{de2014artists}.
+While in the previous cases interactive tools are in charge of running the EA and users collaborate evaluating
+fitness quality, the proposed framework in XYZ is mainly in charge of storing digital versions of the works produced 
+by the team of artists, who are really in charge of applying every step of the EA. Thus the team of artists 
+are in charge of:  (i) setting up the initial population by selecting in the web interface
+well known works from the history of arts;  (ii) each of the artists selects, every generation, two works 
+from the previous one (see Figure \ref{fig:xyz}), and produce a new painting by applying 
+any kind of crossover and mutation he or she may consider  (iii) the process repeats for the number 
+of generations initially decided, so that at the end, a collective  artwork has been produced 
+including as many paintings per generation, as artists are participating in the experiment. In this application
+the final product is the motivation.
 
-Thus, the main difference between the standard IEC and the Unplugged EA, is that in the latest,
-every genetic operation as well as the general evolutionary loop is in charge of human beings. 
-Thus the team of artists are in charge of:  (i) setting up the initial population by selecting
-well known works from the history of arts;  (ii) each of the artists selects every generation two works 
-from the previous one, and produce a new painting by applying any kind of crossover and mutation he may consider 
-(iii) the process repeats for the number of generations initially decided, so that at the end, a collective 
-artwork has been produced including as many paintings per generation, as artists are participating in the experiment.
-
-Artists resort to framework as the tool allowing them to store the paintings, deciding their preferred ones from the
+Artists resort to the framework as the tool allowing them to store the paintings, deciding their preferred ones from the
 previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
@@
         \includegraphics[width=3.5in]{img/interfaceXYZ.png}
-    \caption{User interface of the XYZ application.}
+    \caption{User interface of the XYZ application, used by artists to select the best pair of the current generation.}
     \label{fig:xyz}
@@
 
-Collaborative and volunteer based IEC systems involve the dynamic 
-interaction of many entities and artifacts. Employing human centered approach 
+Volunteer-based and C-IEC systems involve the dynamic 
+interaction of many entities and artifacts. Employing a human-centered approach 
 can benefit researchers to understand and visualize this kind of systems better.
@@
 \begin{acks}
-  % The authors would like to thank Dr. Yuhua Li for providing the
-  % matlab code of  the \textit{BEPS} method.
-
-  % The authors would also like to thank the anonymous referees for
-  % their valuable comments and helpful suggestions. The work is
-  % supported by the \grantsponsor{GS501100001809}{National Natural
-  %   Science Foundation of
-  %   China}{http://dx.doi.org/10.13039/501100001809} under Grant
-  % No.:~\grantnum{GS501100001809}{61273304}
-  % and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
-  % Scientsts' Support Program}.
   Reserved\\

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
More corrections
@JJ have a look to last changes when you have time, I will come back
Sunday night, so we have all Monday.
Revising the introduction.
Mainly first paragraphs.

@@
 They are usually applied in problems where the fitness function is not known or simply
-does not exist, and the result of optimization should fit a certain human need or desire. 
-That is why their use-case includes the evolution of objects with subjective characteristics
-such visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
+does not exist, and the result of optimization should fit a certain human need or desire such as an aesthetic ideal.  
+That is why their use cases includes the evolution of objects with subjective characteristics
+such as visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
 considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
-or creativity \cite{yu2011cooks}.
-% Human Based Computation
-These IEC systems are an interesting venue of research, since they have demonstrated 
-their ability for effectively 
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-as well as other artifacts in many other domains \cite{ie1}. In those cases when 
+or creativity \cite{yu2011cooks}. In the cases when 
 human interaction is responsible of other 
@@
 or as human-based computation \cite{quinn2011human}.
+
+% Human Based Computation
+IEC systems are an interesting venue of research, since they have demonstrated 
+their ability for effectively 
+producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
+as well as other types of artifacts in many other domains \cite{ie1}. 
 % Problems
-However, the necessary intervention of humans, brings certain challenges 
-to designers of IEC methods namely, human evaluations are slow and expensive, there is
-human fatigue caused by the interaction \cite{ie1}, and also boredom arise
+However, the necessary intervention of humans brings certain challenges 
+to designers of IEC methods; namely, human evaluations are scarce, slow and expensive, there is
+human fatigue caused by the interaction \cite{ie1}, and also boredom arises
 when users evaluate a large number of phenotypes, 
@@
 % Volunteers 
-Having a volunteer system can lower the requirements for anyone to participate in
-the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
+Including an IEC in a volunteer system can lower the requirements for
+participants in the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
 system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
 volunteer\'s lack of accountability, and the need to build trust between participants and project
-owners. Other issues for project owners are also the difficulty of establishing 
+owners. Other issues for project owners are also the difficulty of predicting 
 the amount of time and resources a volunteer is willing to spend on the system, 
@@
     \centering
-        \includegraphics[width=5.5in]{img/framework.eps}
+        \includegraphics[width=6in]{img/framework.eps}
     \caption{IEC Human-centered framework.}
@@
     \centering
-        \includegraphics[width=4in]{img/weibull_2.png}
+        \includegraphics[width=3.5in]{img/weibull_2.png}
     \caption{Experiment G. Weibull Fit}
@@
     \centering
-        \includegraphics[width=4in]{img/weibull_3.png}
+        \includegraphics[width=3.5in]{img/weibull_3.png}
     \caption{Experiment GG. Weibull Fit}
@@
     \centering
-        \includegraphics[width=4in]{img/kinect.png}
+        \includegraphics[width=3.3in]{img/kinect.png}
     \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Revised intro

@@
 Data available from the interaction is also used to increase the engegment of 
-users in the system by applying  gamification techniques. Gamification is defined by 
+users in the system by applying  gamification techniques. Gamification
+is a technique defined by 
 Deterding et al. \cite{deterding2011game} as
@@
 when applying a gamification technique. Finally, two other case studies are briefly presented
-to show how the framework is applied to other domains.
+to show how the framework can applied to other domains. %The
+                                %objectives of these lasts are... - JJ
 
@@
 for C-IEC appliactions which the main proposal of this work. The implementation of the data model
-is detailed in Section \ref{sec:implementation}. Next three case studies are presented: EvoDrawings in
+is detailed in Section \ref{sec:implementation}. % Why not all case
+                                % studies in the same section? - JJ
+Next three case studies are presented: EvoDrawings in
 Section \ref{sec:evodraw}, EvoDraw Installation in Section\ref{sec:evodraw}, 
-and XYZ in Section\ref{sec:evodraw}. Finally, a concluding remarks are provided 
+and XYZ in Section\ref{sec:evodraw}. Finally, concluding remarks are provided 
 in Section \ref{sec:conclusions}.

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Fixed some sentences
Also typos. @mariosky remember to spell-check before submitting

@@
 
-An early example of web based collaboration is the work by 
-Langdon \cite{langdon:2004}, which evolved fractal representations of virtual creatures. 
-Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} 
-use web-based IEAs to evolve artistic artifacts using a generative encoding.
-Another example is the Galapagos Project \cite{sims1997interactivity},
+An early example of IEC is the Galapagos Project \cite{sims1997interactivity},
 an exhibit in the Tokyo Multimedia Museum (1997--2000) 
@@
 twelve displays by selecting those they found most aesthetically interesting by standing on
-step sensors in front of them, presenting an example of a non web C-IEC system.
+step sensors in front of them, presenting an example of a non web
+C-IEC system. Web based systems were introduced later, with  
+Langdon's system \cite{langdon:2004} which evolved fractal representations of virtual creatures. 
+Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} 
+use web-based IEAs to evolve artistic artifacts using a generative encoding.
 
@@
 users, for instance the genetic lineage of each phenotype or the most popular or 
-best rated solucion \cite{picbreeder,forms}. In a recent work by 
+best rated solucion \cite{picbreeder,forms}. En example is the recent work by 
 Wagy \& Bongard \cite{wagy2014collective} user interaction 
@@
 new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
-using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
-In any case, using gamification techniques imply that we must deal with IEC
+using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
+In any case, using gamification techniques imply dealing with IEC
 systems as socio-technical constructs, where the social aspects are
@@
 to their study will allow us to understand them more thoroughly.
+
 % C-IEAS
@@
 % Models
-
+% Is that left for later? No space, I'm afraid - JJ
 \begin{figure*}[!t]
@@
 \label{sec:framework}
+% You have to clearly express the objectives of the paper from tine
+% introduction. If the objective is to reduce fatigue via social
+% engagement, you have to frame the results that way. 
 The general goal of this research is to develop a human-centered \cite{gasson2003human} 
@@
 process in a given IEC application. Software frameworks often have 
-a vison \cite{carneiro2010introducing} guiding their
+a vision \cite{carneiro2010introducing} guiding their
 design. Before diving into details the main design considerations of the
@@
   from the population. There are many ways in which humans could interact 
-  with phenotypes. The interaction consists of a set of actions and 
-  takes place in a certain context. For example, through a mobile device  
+  with these phenotypes. The interaction consists of a set of actions and 
+  takes place in a certain context; for example, through a mobile device  
   or by interacting with real world objects 
@@
   is done by devices lent by humans \cite{DBLP:conf/gecco/MereloCGCRV16}.
-  The information gathered through the interaction is the primary concern
+  The information gathered through the interaction is the primary focus
   as it will guide the search.
@@
  candidate solutions, and the relationships between them in the population.
- The graph database system used in the implementation is Neo4J as it is
- a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well 
+ The graph database system used in the implementation is Neo4J, which is
+ a scalable solution \cite{miller2013graph,holzschuher2013performance}, well 
  supported and documented in PaaS infrastructures like Heroku. The Cypher query 
@@
 The appropiate human interface will depend on the application domain, 
-the current framework implemantation employs a web based application. Developers 
+the current framework implementation employs a web based application. Developers 
 can use different templates, to present one or more phenotypes 

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Some sentences corrected

@@
 can use different templates, to present one or more phenotypes 
-at a time, and the type of rating system: like based or a rating from one to five (stars). 
+at a time, and the type of rating system:  based on ``likes'' or in a rating from one to five (stars). 
 Other implementations do not need a graphical interface at all, for example the
@@
 some users that way, the additional information we obtain for scoring
-phenotypes more than compensates for that. 
-After entering the web application by using their Facebook account,
+phenotypes more than balances that. 
+After logging in,
 users can collaborate with their Facebook friends, 
-sharing those phenotypes they like, or by taking phenotypes
+sharing the phenotypes they like, or by taking phenotypes
 from their friend's collections by using the web interface depicted 
@@
 \end{itemize}
+% Once again, these implementation details could go - JJ
 
@@
 \begin{itemize}
-\item {\bf Experience.}: This variable depends on the score and is a value 
+\item {\bf Experience}: This variable depends on the score and is a value 
 between 0 and 100. It is assumed for this case, that once a user
-reaches a 100 actions, it has enough experience on using the application.
+reaches 100 actions, it has enough experience on using the
+application. % What happens before that? - JJ
 
-\item {\bf Participation.}: This variable is simply the degree of the user node 
+\item {\bf Participation}: This variable is simply the degree of the user node 
 in the graph (number of edges).    
@@
 through social networks.  The link used in the call for participation
-was shortened Google URL, that provides metadata and analytics for the
-users that click on it.  In Table \ref{tab:urls} the URL for each 
+was a shortened Google URL, that provides metadata and analytics for the
+users that click on it. %Will we comment on that? - JJ
+  In Table \ref{tab:urls} the URL for each 
 deployment in the Heroku platform is shown,
 along with the short URL and the analytics link. In the same table a link to the GitHub 
-application repository used to deploy to Heroku is also listed.Only data 
+application repository used to deploy to Heroku is also listed. Only data 
 for the first week of deployment was considered for the experiments, and they where conducted 
@@
 \subsection{Results} 
-Before launch, each deployment was first tried with a few beta testers. 
+Before release, each deployment was first tried with a few beta testers. 
 When applying the leader board gamification technique for the first time a 
-problem was found in this stage. The problem was that some users were cheating by giving a
+problem was found: some users were cheating by giving a
 rating to an animation even before it was returned from the server, this was done by just
@@
 the drawing animation was over. The results of each of the three experiments in 
-terms of participation are dtailed next.
+terms of participation are detailed next.
 
@@
   \small
-  \caption{ After a week of the announcement the total number of volunteers, 
+  \caption{After a week of the announcement the total number of volunteers, 
   nodes and edges in the graph and analytics URLs}

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Finished revision.
Now we have 9 pages, but the problem is that once we add addresses and publishers to bibliography is going to be even more. We could eliminate the "unplugged" section of the paper that has nothing to do with the rest.

@@
 considering deployments B and G, the difference came with users with a medium level of participation.
-When comparing all the experiments the deployment GG had the higher number of participation in users of all levels.
+When comparing all the experiments the deployment GG had the higher
+number of participation in users of all levels, besides attracting a
+bigger number of users.
 
@@
 involves understanding its dynamics. Previous works on browser-based
-volunteer computing have used basic metrics such as: 
+volunteer computing have used basic metrics such as
 the number of users or the time spent by every one in the
@@
 Gamma distributions where viable modeles of the  
-the availability of computing resources in several clusters.
-This is in concordance with the results obtained in \cite{agajaj}, and also
+the availability of computing resources in several clusters, which  is in concordance with the results obtained in \cite{agajaj}, and also
 browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
@@
 there are many users that give a small amount of cycles, while there
-are just a few that give many cycles.
-
-In order to assess if C-IEC follows the same pattern, even when
-using gamification techniques, participation was fittted to a 
-Weibull distribution. Results in Figure \ref{fig:w2} and Figure \ref{fig:w3}
-show that the same pattern appears in this case study.
-
+are just a few that give many cycles. In order to assess gamification techniques alongside C-IEC follows the
+same pattern, participation was fittted to a 
+Weibull distribution,
+ and shown in Figures \ref{fig:w2} and \ref{fig:w3}, confirming  this model for user interaction, although with different fitted
+values in each case. 
+%
 \begin{figure}[!t]
@@
 \end{figure}
+%
 \begin{figure}[!t]
@@
 In this instalation the components of the framework used where EvoSpace and the EA.
-Information from users was limited to a stream of readings from the sensor.
+Information from users was limited to a stream of readings from the
+sensor.
+% This has nothing to do with the rest of the paper. You are not
+% modeling anything. It's just another application of the
+% framework. Could be mentioned in the state of the art, that's
+% all. We should devote more space to the study of user interactions.
 
@@
 \end{figure}
+% Once again, no model, no relation to the rest of the paper. Either
+% you talk about a model of interactions with humans and their
+% similarity with the first case study, or you eliminate it. How do
+% you tackle boredom? How do you increase engagement? Those are the
+% paper's topics, you can't just drop it here because it's interactive
+% - JJ
 
@@
 This framework enabled the implementation of a gamification technique 
-to improve engagement in a case study. Providing a common arena where users are aware of the
+to improve engagement in a case study which provides a common arena where users are aware of the
 activities of other users in the social vicinity. Following a hybrid data model approach
@@
 of interaction or new applications.
+
+% More conclusions and some way of relating the three case studies
+% would be something that is missing and should definitely go here - JJ
 
@@
 
-Another line of work would be to study the negative effects of using 
-gamification techniques to improve engagement, like cheating or competition.
+Another line of work would be to study the possible negative effects of using  
+gamification techniques to improve engagement, like cheating or
+literally {\em gaming} the system to defeat competition. We already
+found some hints of this behavior at the beginning of the release of
+this system, but more subtle effect could be taking place. 
 Finally, the refinement of the proposed Human-Centered framework will need
@@
   Space\\
-  for\\
-  Acks
+  for acks\\
 

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Spell check

@@
 They are usually applied in problems where the fitness function is not known or simply
-does not exist, and the result of optimization should fit a certain human need or desire such as an aesthetic ideal.  
+does not exist, and the result of optimization should fit a certain human need or
+desire such as an aesthetic ideal.  
 That is why their use cases includes the evolution of objects with subjective characteristics
@@
 
-Data available from the interaction is also used to increase the engegment of 
+Data available from the interaction is also used to increase the engagement of 
 users in the system by applying  gamification techniques. Gamification
@@
 of interactive evolution. Then, Section \ref{sec:framework} presents the human centered framework of
-for C-IEC appliactions which the main proposal of this work. The implementation of the data model
+for C-IEC applications which the main proposal of this work. The implementation of the data model
 is detailed in Section \ref{sec:implementation}. % Why not all case
@@
 users, for instance the genetic lineage of each phenotype or the most popular or 
-best rated solucion \cite{picbreeder,forms}. En example is the recent work by 
-Wagy \& Bongard \cite{wagy2014collective} user interaction 
+best rated solution \cite{picbreeder,forms}. An example is the recent work by 
+Wagy \& Bongard \cite{wagy2014collective} where user interaction 
 is needed for evaluating fitness and developing
@@
 into the algorithm to perform the
-evaluation of some features automatically. For instance, 
-Reis et al. added some terrain measures (such as accessibility and edge length) 
+evaluation of some features automatically. For instance,Reis et al. added
+some terrain measures (such as accessibility and edge length) 
 to a standard  IEC in \cite{DBLP:journals/soco/FradeVC12}. 
@@
   cognitive limitations, culture, creativity and other human aspects are an integral part 
-  of the system. Humans are the computing resorces of the system, having unique characteristics
+  of the system. Humans are the computing resources of the system, having unique characteristics
   as those identified by Sun \& Dance \cite{Sun2013}:
@@
 
-The above considerations have guided the design of the framework, and througout they have been
+The above considerations have guided the design of the framework, and throughout they have been
 treated as application constraints. In order to satisfy theses requirements the
@@
 database management system because it provides a hash based implementation of sets and 
-queues which are natural data structures for the EvoSpace model. Basicaly, a sample of 
+queues which are natural data structures for the EvoSpace model. Basically, a sample of 
 candidate solutions are retrieved from the server, evaluated and then sent back. 
 The same operations are used to evolve the population. In EvoSpace individuals replaced 
-in the population are stored indefinitly, to permit users to store permalinks to them.
+in the population are stored indefinitely, to permit users to store permalinks to them.
 Implementation details are presented in \cite{garcia2013evospace}.
@@
 \item {\bf Interfaces.}
-The appropiate human interface will depend on the application domain, 
+The appropriate human interface will depend on the application domain, 
 the current framework implementation employs a web based application. Developers 
@@
 Other implementations do not need a graphical interface at all, for example the
-EvoDrawings instalation or the XYZ Project \cite{de2014artists}.
+EvoDrawings installation or the XYZ Project \cite{de2014artists}.
 
@@
 \cite{picbreeder}.
-In order to do this, metadata about phenotypes 
+In order to do this, meta-data about phenotypes 
 must be permanently stored, even
@@
 through social networks.  The link used in the call for participation
-was a shortened Google URL, that provides metadata and analytics for the
+was a shortened Google URL, that provides meta-data and analytics for the
 users that click on it. %Will we comment on that? - JJ
@@
         \includegraphics[width=3.3in]{img/comparison.png}
-    \caption{Users ranked by the number of phenotypes they rated vs. the number of phenotypes evaluated using a logarithmic scale. }
+    \caption{Users ranked by the number of phenotypes they 
+    rated vs. the number of phenotypes evaluated using a logarithmic scale. }
     \label{fig:top-ranked-participation}
@@
 \cite{javadi2009mining} have found that the Weibull, log-normal, and
-Gamma distributions where viable modeles of the  
-the availability of computing resources in several clusters, which  is in concordance with the results obtained in \cite{agajaj}, and also
+Gamma distributions where viable models of the  
+the availability of computing resources in several clusters, which  is 
+in concordance with the results obtained in \cite{agajaj}, and also
 browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
@@
 are just a few that give many cycles. In order to assess gamification techniques alongside C-IEC follows the
-same pattern, participation was fittted to a 
+same pattern, participation was flitted to a 
 Weibull distribution,
- and shown in Figures \ref{fig:w2} and \ref{fig:w3}, confirming  this model for user interaction, although with different fitted
+and shown in Figures \ref{fig:w2} and \ref{fig:w3}, 
+confirming  this model for user interaction, although with different fitted
 values in each case. 
@@
         \includegraphics[width=3.3in]{img/kinect.png}
-    \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}
+    \caption{EvoDraw Installation on ------ -- --- ----.}
     \label{fig:kinect}
@@
 
-In this instalation the components of the framework used where EvoSpace and the EA.
+In this installation the components of the framework used where EvoSpace and the EA.
 Information from users was limited to a stream of readings from the
@@
 
-Artists resort to the framework as the tool allowing them to store the paintings, deciding their preferred ones from the
-previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
-The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to 
+Artists resort to the framework as the tool allowing them 
+to store the paintings, deciding their preferred ones from the
+previous generation, as well as describing the reasons for their 
+selections, mutation and crossover operations applied.
+The tool is also in charge of generating statistics, 
+computing \textit{elitism}, thus adding the most selected work to 
 the next generation, etc.
@@
 In this paper a human-centered software framework was proposed, and was validated
-through the implementaion and refinement of three C-IEC applications.
+through the implementation and refinement of three C-IEC applications.
 This framework enabled the implementation of a gamification technique 

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Fixed bibliography. Phew

@@
 aspects of the evolutionary process, some authors classify these IEC methods 
-as human-based evolutionary computation \cite{kosorukoff2001human} 
+as human-based evolutionary computation \cite{972056} 
 or as human-based computation \cite{quinn2011human}.

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Abstract and new Image

@@
 Weibull distribution,
-and shown in Figures \ref{fig:w2} and \ref{fig:w3}, 
+and shown in Figures \ref{fig:w1} and \ref{fig:w3}, 
 confirming  this model for user interaction, although with different fitted
@@
     \centering
-        \includegraphics[width=3.5in]{img/weibull_2.png}
-    \caption{Experiment G. Weibull Fit}
-    \label{fig:w2}
+        \includegraphics[width=3.5in]{img/weibull_1.png}
+    \caption{Deployment B. Weibull Fit.}
+    \label{fig:w1}
 \end{figure}
@@
     \centering
+        \includegraphics[width=3.5in]{img/weibull_2.png}
+    \caption{Deployment G. Weibull Fit.}
+    \label{fig:w2}
+\end{figure}
+
+\begin{figure}[!t]
+    \centering
         \includegraphics[width=3.5in]{img/weibull_3.png}
-    \caption{Experiment GG. Weibull Fit}
+    \caption{Deployment GG. Weibull Fit.}
     \label{fig:w3}
@@
 
-\section{Case Study: EvoDraw-Kinect}
-\label{sec:kinect}
-A variation of the previous application is The EvoDraw installation, using the same
-framework but with a different form of interaction. Humans interact with the animations 
-by simply looking at them. Again the population is stored in the EvoSpace container and
-client computers remove animations from the EvoSpace to be presented in a display one at a time. 
-After a certain amount of time,  animations are put back to the server along with information 
-generated from the interaction. To continue, another is again removed and presented to viewers. 
-This cycle continues until there is enough interactions to create a new generation of animations.
-
-In order to assign a fitness value to each animation and thus enabling 
-the EA the selection of the most interesting animations, a natural user interface 
-device (Kinect sensor for Xbox One) is used for facial tracking. Each animation is 
-presented several times in a single generation and the best are always passed to the next, 
-with its fitness recomputed. To generate the new population the same operators as the previous case study
-are applied to the population.
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.3in]{img/kinect.png}
-    \caption{EvoDraw Installation on ------ -- --- ----.}
-    \label{fig:kinect}
-\end{figure}
-
-In this installation the components of the framework used where EvoSpace and the EA.
-Information from users was limited to a stream of readings from the
-sensor.
 % This has nothing to do with the rest of the paper. You are not
@@
 
-\section{Case Study: XYZ-Unplugged}
-\label{sec:xyz}
-
-XYZ is a collective artwork generated by means of the Unplugged version of the EA \cite{de2014artists}.
-While in the previous cases interactive tools are in charge of running the EA and users collaborate evaluating
-fitness quality, the proposed framework in XYZ is mainly in charge of storing digital versions of the works produced 
-by the team of artists, who are really in charge of applying every step of the EA. Thus the team of artists 
-are in charge of:  (i) setting up the initial population by selecting in the web interface
-well known works from the history of arts;  (ii) each of the artists selects, every generation, two works 
-from the previous one (see Figure \ref{fig:xyz}), and produce a new painting by applying 
-any kind of crossover and mutation he or she may consider  (iii) the process repeats for the number 
-of generations initially decided, so that at the end, a collective  artwork has been produced 
-including as many paintings per generation, as artists are participating in the experiment. In this application
-the final product is the motivation.
-
-Artists resort to the framework as the tool allowing them 
-to store the paintings, deciding their preferred ones from the
-previous generation, as well as describing the reasons for their 
-selections, mutation and crossover operations applied.
-The tool is also in charge of generating statistics, 
-computing \textit{elitism}, thus adding the most selected work to 
-the next generation, etc.
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
-    \caption{User interface of the XYZ application, used by artists to select the best pair of the current generation.}
-    \label{fig:xyz}
-\end{figure}
 % Once again, no model, no relation to the rest of the paper. Either

### 2017-02-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Abstract and new Image

--- 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
@@
 To tackle these issues, in this paper we
-propose a human-centered framework to modelthe complex interactions on these systems.
-Three case studies are presented where the model is applied in the 
-development of evolutionary interactive systems, with different techniques:
-volunteer-collaborative, unplugged and sensor-based. 
-Both conceptual and implementation details are provided where the   
-technique is applied to measure and increase user engagement and participation.
-Our experiments show that the model can be successfully applied in a
-gamification technique developed to increase user engagement, which
+propose a human-centered framework to model the complex interactions 
+found on these systems.
+A case study is presented, providing both conceptual and implementation 
+details where the model is applied to measure and increase user engagement and participation
+in a collaborative, volunteer based, evolutionary interactive system. 
+Our experiment shows that the model can be successfully applied for developing a
+gamification technique to increase user engagement, which
 implies that this technique can successfully be used to 
-decrease user fatigue and thus increase {\em performance} of the
+decrease user fatigue and boredom, and thus increase the {\em performance} of the
 interactive system.

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Only one case study

@@
 on the score and certain features of the game are only available to gamers when 
-they reach a giving level. % Say what we are doing in this case - JJ
-%%%
-%%% Change if case studies change
-%In this work, a case study consisting in a web application for evolving artistic drawings
-%is presented, providing both conceptual and implementation details of the framework. 
-%In order to test the framework three versions of the application are compared to
-%measure the degree of relationship between users and their participation,
-%when appling a gamification technique. 
-In this work, three case studies are presented. First, the development of a web based C-IEC application 
+they reach a giving level. 
+In this work, a case study is presented. The development of a web based C-IEC application 
 for evolving artistic drawings is described to give the reader details about 
@@
 implemented in the same application to study the changes in participation
-when applying a gamification technique. Finally, two other case studies are briefly presented
-to show how the framework can applied to other domains. %The
-                                %objectives of these lasts are... - JJ
+when applying a gamification technique. Finally, an experiment is conducted
+comparing three versions of the application, with, and with out, gamification
+in order to compare the amount of user participation.
 
@@
 of interactive evolution. Then, Section \ref{sec:framework} presents the human centered framework of
-for C-IEC applications which the main proposal of this work. The implementation of the data model
-is detailed in Section \ref{sec:implementation}. % Why not all case
-                                % studies in the same section? - JJ
-Next three case studies are presented: EvoDrawings in
-Section \ref{sec:evodraw}, EvoDraw Installation in Section\ref{sec:evodraw}, 
-and XYZ in Section\ref{sec:evodraw}. Finally, concluding remarks are provided 
+for C-IEC applications which the main proposal of this work. Implementation details of the data model
+are presented in Section \ref{sec:implementation}. 
+Next the EvoDrawings application case study is presented in
+Section \ref{sec:evodraw}. The experimental set-up is described in Section \ref{sec:experiment},
+an the results are discussed in Section \ref{sec:results}. Finally, concluding remarks are provided 
 in Section \ref{sec:conclusions}.
@@
 
-\section{Human-Centered IEC Framework}
+\section{Human-Centered C-IEC Framework}
 \label{sec:framework}
@@
 The general goal of this research is to develop a human-centered \cite{gasson2003human} 
-software framework for C-IEC. 
-A framework is defined as a reusable architectural design together
+software framework that can be used to increase volunteer participation 
+in C-IEC systems. A framework is defined as a reusable architectural design together
 with an implementation \cite{campbell1991choices}, in this case 
-providing generalized components to developers of IEC systems.
+providing generalized components to developers and researchers of C-IEC systems.
 
@@
   \item {\bf Evolutionary Algorithm.} 
-  The Evolutionary Algorithm (EA) algorithm interacts with the data model. The EA could also be run with the help of
+  The Evolutionary Algorithm (EA) algorithm interacts 
+  with the data model. The EA could also be run with the help of
   humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
@@
 \item {\bf Experience}: This variable depends on the score and is a value 
-between 0 and 100. It is assumed for this case, that once a user
-reaches 100 actions, it has enough experience on using the
+between 0 and 100. A new user starts at zero, and the experience increases until
+it reaches 100 actions. It is assumed for this case, that once a user
+reaches this value, it has enough experience on using the
 application. % What happens before that? - JJ  
-
+% explained - Mario
 \item {\bf Participation}: This variable is simply the degree of the user node 
@@
 \subsection{Experimental Setup}
+\label{sec:experiment}
 
@@
 was a shortened Google URL, that provides meta-data and analytics for the
-users that click on it. %Will we comment on that? - JJ
+users that click on it. 
+ %Will we comment on that? - JJ
+ % To do.
   In Table \ref{tab:urls} the URL for each 
@@
 
-\subsection{Results} 
+\section{Results}
+\label{sec:results} 
 Before release, each deployment was first tried with a few beta testers. 

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Anonimized figures

@@
 desire such as an aesthetic ideal.  
-That is why their use cases includes the evolution of objects with subjective characteristics
+That is why their use cases include the evolution of objects with subjective characteristics
 such as visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
@@
 \end{quote}  
-The gamification element employed in this case study has been a rewarding mechanism  
+The gamification element employed in this work is a rewarding mechanism  
 \cite{dubois2013understanding}. In general rewards  consist of a reputation system 
@@
 they reach a giving level.
-In this work, a case study is presented. The development of a web based C-IEC application 
+
+The development of a web based C-IEC application 
 for evolving artistic drawings is described to give the reader details about 
@@
 when applying a gamification technique. Finally, an experiment is conducted
-comparing three versions of the application, with, and with out, gamification
-in order to compare the amount of user participation.
+comparing three versions of the application: with out gamification and
+two gamification techniques.
 
@@
 Section \ref{sec:related} presents related work on the topic 
-of interactive evolution. Then, Section \ref{sec:framework} presents the human centered framework of
+of collaborative interactive evolution. Then, Section \ref{sec:framework} 
+presents the human centered framework
 for C-IEC applications which the main proposal of this work. Implementation details of the data model
@@
 
-An early example of IEC is the Galapagos Project \cite{sims1997interactivity},
+An early example of C-IEC is the Galapagos Project \cite{sims1997interactivity},
 an exhibit in the Tokyo Multimedia Museum (1997--2000) 
@@
 twelve displays by selecting those they found most aesthetically interesting by standing on
-step sensors in front of them, presenting an example of a non web
-C-IEC system. Web based systems were introduced later, with  
+step sensors in front of them. Web based systems were introduced later, with  
 Langdon's system \cite{langdon:2004} which evolved fractal representations of virtual creatures. 
@@
 % engagement, you have to frame the results that way.
+% Ok, added - Mario 
 The general goal of this research is to develop a human-centered \cite{gasson2003human} 
@@
 providing generalized components to developers and researchers of C-IEC systems. 
-
 The proposed framework includes components that can be refined to increase
@@
 \item Base (B): All users have the same weight.
-\item Non Graph Gamification (G): Only experience is considered
+\item Non Graph Gamification (G): Only experience is considered.
 \item Graph Gamification (GG): Both experience and participation are considered.
@@
 % - JJ
-
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/user_known_2.png}
+    \caption{User network view of the Base deployment}
+    \label{fig:B-network}
+\end{figure}
+\begin{figure}[!t]
+    \centering
+        \includegraphics[width=3.5in]{img/user_known_3.png}
+    \caption{User network view of the Graph Gamification deployment}
+    \label{fig:GG-network}
+\end{figure}
 \section{Conclusions}

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
First uploaded version
@JJ  have a look if possible

@@
 are just a few that give many cycles. In order to assess gamification techniques alongside C-IEC follows the
-same pattern, participation was flitted to a 
+same pattern, participation was fitted to a 
 Weibull distribution,
@@
 confirming  this model for user interaction, although with different fitted
-values in each case. 
+values in each case. A graph query was used to compare the number of volunteers in deployments
+B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively).
+
 %
@@
     \centering
-        \includegraphics[width=3.5in]{img/user_known_2.png}
+        \includegraphics[width=3.2in]{img/user_known_2.png}
     \caption{User network view of the Base deployment}
@@
     \centering
-        \includegraphics[width=3.5in]{img/user_known_3.png}
+        \includegraphics[width=3.2in]{img/user_known_3.png}
     \caption{User network view of the Graph Gamification deployment}
@@
 can benefit researchers to understand and visualize this kind of systems better.
+
 In this paper a human-centered software framework was proposed, and was validated
-through the implementation and refinement of three C-IEC applications.
+through the implementation and refinement of a C-IEC application.
 This framework enabled the implementation of a gamification technique 
 to improve engagement in a case study which provides a common arena where users are aware of the
-activities of other users in the social vicinity. Following a hybrid data model approach
-gives developers and project owners options when innovating with other means
-of interaction or new applications.
+activities of other users in the social vicinity.
 
+In concordance with the results obtained in other 
+browser-based volunteer systems, after 
+applying the gamification techniques, user participation follows the
+same pattern and was fitted to a Weibull distribution.
+
+It was shown how the estimation of fitness of candidate solutions
+can take advantage of a graph model representation. Also, the graph
+offers another view of the system that it would have been
+difficult to replicate with a relational database. 
 % More conclusions and some way of relating the three case studies
 % would be something that is missing and should definitely go here - JJ
+% No longer needed as they where removed - Mario
 

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Version submitted 1

--- 2017-02-03 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
@@
 propose a human-centered framework to model complex interactions on these systems.
-Three case studies are presented where the model is applied in the 
-development of evolutionary interactive systems, with different techniques:
-volunteer-collaborative, unplugged and sensor-based. 
+A case study is presented where the model is applied in the 
+development of a collaborative evolutionary interactive system. 
 Both conceptual and implementation details are provided where the   
-technique is applied to measure and increase user engagement and participation.
+technique is used to measure and increase user engagement and participation.
 Our experiments show that the model can be successfully applied in a
@@
 implies that this technique can successfully be used to 
-decrease user fatigue and boredom, and thus increase {\em performance} of the
+decrease user fatigue and boredom, and thus increase the {\em performance} of the
 interactive system.

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Intro revised

@@
 desire such as an aesthetic ideal.  
-That is why their use cases include the evolution of objects with subjective characteristics
-such as visual appeal or attractiveness \cite{biomorphs} as well as others where human behavior is 
+That is why their use cases include the evolution of objects with subjective characteristics,
+such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
 considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
@@
 human interaction is responsible of other 
-aspects of the evolutionary process, some authors classify these IEC methods 
+aspects of the evolutionary process, these IEC methods are classified by some authors  
 as human-based evolutionary computation \cite{972056} 
@@
 they are able to include; to reach more users,
-IEC systems are some times developed as web applications depending on web visitors to help
+IEC systems are some times developed as web applications depending on visitors to help
 with the search, using both anonymous and registered users. Some systems 
@@
 % Volunteers 
-Including an IEC in a volunteer system can lower the requirements for
+Including an C-IEC in a volunteer system can lower the requirements for
 participants in the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
 system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
-volunteer\'s lack of accountability, and the need to build trust between participants and project
-owners. Other issues for project owners are also the difficulty of predicting 
+volunteer's lack of accountability, and the need to build trust between participants and project
+owners. Other issues of interest for project owners are also the difficulty of predicting 
 the amount of time and resources a volunteer is willing to spend on the system, 
@@
 when applying a gamification technique. Finally, an experiment is conducted
-comparing three versions of the application: with out gamification and
-two gamification techniques.
+comparing three versions of the application: one with out gamification and
+two employing a different gamification technique.
 
@@
 
-\subsection{Experimental Setup}
+\section{Experimental Setup}
 \label{sec:experiment}

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Cutting it down to 8 pages

@@
     \centering
-        \includegraphics[width=3.5in]{img/weibull_1.png}
+        \includegraphics[width=3.2in]{img/weibull_1.png}
     \caption{Deployment B. Weibull Fit.}
@@
     \centering
-        \includegraphics[width=3.5in]{img/weibull_2.png}
+        \includegraphics[width=3.2in]{img/weibull_2.png}
     \caption{Deployment G. Weibull Fit.}
@@
     \centering
-        \includegraphics[width=3.5in]{img/weibull_3.png}
+        \includegraphics[width=3.2in]{img/weibull_3.png}
     \caption{Deployment GG. Weibull Fit.}
@@
   \end{figure}
+
 \section{Conclusions}
@@
 interaction of many entities and artifacts. Employing a human-centered approach 
-can benefit researchers to understand and visualize this kind of systems better.
+will allow researchers to understand and visualize this kind of
+systems better. % But you would have to provide some proof of this - JJ
 
@@
 to improve engagement in a case study which provides a common arena where users are aware of the
-activities of other users in the social vicinity.
+activities of other users in the social neighborhood.
 
-In concordance with the results obtained in other 
-browser-based volunteer systems, after 
-applying the gamification techniques, user participation follows the
-same pattern and was fitted to a Weibull distribution.
+After applying the gamification techniques user participation follows the
+same pattern and was fitted to a Weibull distribution, which matches
+results obtained in other browser-based volunteer systems.
+% We would have to check the actual coefficients. - JJ
 
@@
   Reserved\\
-  Space\\
-  for acks\\
+  Space for acks.
 

### 2017-02-07 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Version submitted 2

@@
 into the algorithm to perform the
-evaluation of some features automatically. For instance,Reis et al. added
+evaluation of some features automatically. For instance, Reis et al. added
 some terrain measures (such as accessibility and edge length) 
@@
 they collaborate with a large number of users for the interactive modeling of 3D glasses.
-The framework and data model for 
-representing the interactive process of our C-IEC systems is proposed next.
+
+Promoting user engagement and reducing user fatigue requires the management of the complex 
+interactions found in these systems. Our approach is proposed next.
 % Volunteer Based
@@
     \centering
-        \includegraphics[width=6in]{img/framework.eps}
+        \includegraphics[width=7in]{img/framework.eps}
     \caption{IEC Human-centered framework.}
@@
 \item {\bf Users are volunteers.} Users donate their computing resources, so they are 
-unaccountable, sometimes they try to game the system. Project owners must actively promote and
+unaccountable and sometimes they try to game the system. Project owners must actively promote and
 design the interactive system to engage volunteers \cite{oh2015clicking}. % Define
 \item {\bf Users are not alone}
-  Relationship between users in an interactive evolutionary algorithm can be modeled
+  Relationships between users in an interactive evolutionary algorithm can be modeled
   as a social network, with well established semantics, algorithms and metrics 
@@
   These measures can then be used by recommender algorithms selecting 
-  phenotypes according users' preferences.
+  phenotypes according to their preferences.
 
@@
   Real time processing of users' actions could be needed for certain applications when data is 
-  captured by sensors, or directly captured as user input. For example, social networks keep
-  track of users' interactions with other users, media objects and places. Users of 
+  captured by sensors, or directly captured as user input. For example, social networks encourage
+  users to publish their interactions with other users, media objects and places. Users of 
   social networks (for instance the Facebook Graph) are accustomed to express these 
   complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
-  Other example is the W3C Activity Streams 2.0 specification used for representing activities 
-  common in social web applications \cite{json:streams}. 
+  As this is gaining accptence, there are iniciatives like the W3C Activity Streams 2.0 
+  specification, used for representing common activities in social 
+  web applications \cite{json:streams}. 
 \end{itemize}
@@
 treated as application constraints. In order to satisfy theses requirements the
-human centered IEC architecture consists of three high level components depicted
+human-centered C-IEC architecture consists of three high level components depicted
 in figure \ref{fig:hc_framework}:
@@
 
-  \item {\bf IEC Data Model.}
+  \item {\bf Data Model.}
   A data model is used to describe the IEC system prior to a physical 
-  implementation.  Depending on the domain several techniques can be used.
-  Parts of the system could be better described by an entity-relationship 
-  modeling, to be used in a relational database or a class digram for a 
-  key-value store. A graph is proposed for modeling the social network of users 
-  and their interaction with the population. When implemented, a graph database 
-  will be the back-end of the system.
+  implementation.  Depending on the domain several models can be used, these
+  options and their physical implememtations are explained in Section \ref{sec:implementation}.
+
+  \item {\bf Human Interface.}
+  The appropriate human interface will depend on the application domain, 
+  the current framework implementation employs a web based application. Developers 
+  can use different templates, to present one or more phenotypes 
+  at a time, and the type of rating system:  based on ``likes'' or in a rating from one to five (stars). 
+  Other implementations do not need a graphical interface at all, using sensors and actuators.
 
   \item {\bf Evolutionary Algorithm.}
-  The Evolutionary Algorithm (EA) algorithm interacts 
-  with the data model. The EA could also be run with the help of
-  humans, for instance in the XY project \cite{de2013unplugging} artists actually painted
-  the solutions of the new population and applied every genetic operator.
+  The Evolutionary Algorithm (EA) algorithm interacts with the data model.     
+  The EA is decoupled from the data model, the algorithm could be implemented internally, 
+  use an external library, or even be human based evolution, 
+  where volunteers select the parents of the next generation and upload the 
+  new individuals as in the XYZ project \cite{de2014artists}.
+
+  \item {\bf Analytics.} 
+  Researchers and deveopers will need to query the data model in order to analyzing the 
+  system's behaviour.
+
 \end{enumerate}
@@
 
-The core of the framework is the data model because both the IEC system and the EA will 
-depend on the application. In this section implementation details of the Data Model are presented.
-Three database models are employed to store different elements of the system:
+In this section, the core component of the framework is presented.
+Three database models are employed to store different elements of the system, next 
+these are described including the technology selected for their physical implementation:
 
@@
 
-\item {\bf Interfaces.}
-The appropriate human interface will depend on the application domain, 
-the current framework implementation employs a web based application. Developers 
-can use different templates, to present one or more phenotypes 
-at a time, and the type of rating system:  based on ``likes'' or in a rating from one to five (stars). 
-Other implementations do not need a graphical interface at all, for example the
-EvoDrawings installation or the XYZ Project \cite{de2014artists}.
-
-\item {\bf Evolution.}    
-The EA is decoupled from the data model, the algorithm 
-could be implemented as an EvoWorker \cite{garcia2013evospace}, use an external library, 
-or even assigned to humans. An additional tool was implemented for a human based evolution, 
-where volunteers select the parents of the next generation and upload the 
-new individuals \cite{de2014artists}.  
 \end{itemize}
@@
 \label{sec:evodraw}
-As a case study, a C-IEC application was developed using the 
+As a case study, a C-IEC application was developed by extending the 
 EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
-A brief description of the application is presented next, focusing
-on the data elements that where ported to the graph model.
+A brief description of the application is presented next.
 
@@
         \includegraphics[width=3.5in]{img/interface.png}
-    \caption{User interface of the EvoDrawings application.}
+    \caption{User interface of the EvoDrawings application. (ANONIMIZED IN THIS VERSION)}
     \label{fig:web}
 \end{figure}
-
-\subsection{Fitness Assignment}
-The ES-I platform has been programmed to employ a 
-collaborative methodology for performing fitness assignment. 
-Several registered users assign a quality assessment to a single
-phenotype and then an aggregated fitness value is calculated,
-depending on the rating (from 1 to 5 stars) by each particular user,  
-resulting in a many-to-many relationship between users and
-phenotypes. Other components or users of the systems can 
-query this user-phenotype relation to extract relevant
-knowledge about the process and the population. For instance, showing the
-most popular phenotype, or the user with more participation 
-\cite{picbreeder}.
-In order to do this, meta-data about phenotypes 
-must be permanently stored, even
-if they are no longer in the active population.
 
@@
 The graph model for the EvoDrawings application has the following types of
-nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
-a collection of drawings belonging to users. One collection can contain many phenotypes 
+nodes: {\bf User}, {\bf Phenotype} and {\bf Collection}. The Collection node represents
+a collection of drawings belonging to a user. One collection can contain many phenotypes 
 or be empty. A single phenotype could be shared by many collections. The interaction 
@@
 along with the short URL and the analytics link. In the same table a link to the GitHub 
-application repository used to deploy to Heroku is also listed. Only data 
+application repository used to deploy to Heroku is also listed (ANONIMIZED IN THIS VERSION). Only data 
 for the first week of deployment was considered for the experiments, and they where conducted 
@@
 boards because by making the scores visible to other players they are encouraged 
-to compete \cite{hickman2010total}. The final version disabled the button until 
+to compete \cite{hickman2010total}. The version used in experiments disabled the button until 
 the drawing animation was over. The results of each of the three experiments in 
@@
 When comparing all the experiments the deployment GG had the higher
-number of participation in users of all levels, besides attracting a
-bigger number of users.
+number of participation, besides attracting also the higher number of users.
 
@@
 Weibull distribution,
-and shown in Figures \ref{fig:w1} and \ref{fig:w3}, 
-confirming  this model for user interaction, although with different fitted
+and shown in Figures \ref{fig:w1} to \ref{fig:w3}, 
+confirming this a model for user interaction, although with different fitted
 values in each case. A graph query was used to compare the number of volunteers in deployments
-B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively).
+B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) showing 
+additional patterns on both social networks.
 
@@
     \centering
-        \includegraphics[width=3.2in]{img/weibull_1.png}
+        \includegraphics[width=3.4in]{img/weibull_1.png}
     \caption{Deployment B. Weibull Fit.}
@@
     \centering
-        \includegraphics[width=3.2in]{img/weibull_2.png}
+        \includegraphics[width=3.4in]{img/weibull_2.png}
     \caption{Deployment G. Weibull Fit.}
@@
     \centering
-        \includegraphics[width=3.2in]{img/weibull_3.png}
+        \includegraphics[width=3.4in]{img/weibull_3.png}
     \caption{Deployment GG. Weibull Fit.}
@@
         \includegraphics[width=3.2in]{img/user_known_2.png}
-    \caption{User network view of the Base deployment}
+    \caption{User network view of the Base deployment (B). (ANONIMIZED IN THIS VERSION) }
     \label{fig:B-network}
@@
         \includegraphics[width=3.2in]{img/user_known_3.png}
-    \caption{User network view of the Graph Gamification deployment}
+    \caption{User network view of the Graph Gamification deployment(GG) (ANONIMIZED IN THIS VERSION) .}
     \label{fig:GG-network}
@@
 
-After applying the gamification techniques user participation follows the
-same pattern and was fitted to a Weibull distribution, which matches
-results obtained in other browser-based volunteer systems.
-% We would have to check the actual coefficients. - JJ
+In concordance with the results obtained in other 
+browser-based volunteer systems, after 
+applying the gamification techniques, user participation follows the
+same pattern and it was fitted to a Weibull distribution.
 

### 2017-02-08 2017-EUROPAR/Gamification/paper.tex
abstract and latex files for LNCS

\documentclass{llncs}
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\sloppy

\title{Increasing User Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System}
\titlerunning{Increasing Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System}

\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}

\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
\and
Universidad de Granada, Granada, Spain
\email{mario@tectijuana.edu.mx}\\
\email{jmerelo@geneura.ugr.es}}

\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }

\maketitle

\begin{abstract}

Volunteer computing is a distributed computing system, in which people
provide their own computing resources or storage to contribute to a common effort.
By runnning a script in a web page, collaboration is straightforward, but also ephemeral.
Resources depend on the amount of time a user lends, whicn means that 
the user has to be kept engaged to obtain as many computing cycles as
possible.

In this paper, we analyze a volunteer-based evolutionary computing system called
NodIO with the objective of discovering rules that encourage volunteer
participation, making it more or less efficient. We present the results of
an experiment where a gammification technique is applied by adding a leaderboard 
showing the top registered contributors. In NodIO volunteers can
participate without the need to create an account, so the question was
if the need to register would hava a negative impact on user participation. 
The experiment results show that even if only a small percentege of users created an account,
those participating in the competition provided around 90.

\keywords{Distributed Evolutionary Algorithms, Volunteer Computing}
\end{abstract}

\end{document}

### 2017-02-09 2017-EUROPAR/Gamification/paper.tex
Revised abstract

@@
 
-Volunteer computing is a distributed computing system, in which people
+One way of creating distributed computing system is to use volunteers who
 provide their own computing resources or storage to contribute to a common effort.
@@
 the user has to be kept engaged to obtain as many computing cycles as
-possible.
-
-In this paper, we analyze a volunteer-based evolutionary computing system called
+possible. In this paper, we analyze a volunteer-based evolutionary computing system called
 NodIO with the objective of discovering rules that encourage volunteer
-participation, making it more or less efficient. We present the results of
+participation thus increasing the overall computing power. We present the results of
 an experiment where a gammification technique is applied by adding a leaderboard 
-showing the top registered contributors. In NodIO volunteers can
+showing the top scores achieved by registered contributors. In the NodIO system volunteers can
 participate without the need to create an account, so the question was
@@
 The experiment results show that even if only a small percentege of users created an account,
-those participating in the competition provided around 90.
+those participating in the competition provided around 90\%.
 

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Figures uploaded with dummy captions
@JJ Please have a look

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\usepackage{subfigure}
 \begin{document}
@@
 
+\section{Experiments}
+\label{sec:experiments}
+
+\subsection{Results}
+\label{sec:results}
+
+\begin{figure*}[t]
+    \centering
+    \subfigure  [6 workers]
+    {
+        \includegraphics[width=4in]{img/puts_user.png}
+    }
+    \subfigure  [12 workers]
+    {
+        \includegraphics[width=4in]{img/puts_ip.png}
+    }
+
+    \caption{100 experiments with random parameters for the 128 Bit Griewank 
+    single-objective optimization test function. Experiments are ranked by 
+    the mean time to solution of 5 runs, with (a) 6 workers, and (b) 12 workers.}
+    \label{fig:griewank}
+\end{figure*}
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=4in]{img/puts_box.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/puts_by_time.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/workers_best_user.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/workers_put_ip.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+
 \end{document}

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Making changes

--- 2017-02-09 2017-EUROPAR/Gamification/paper.tex
@@
 participate without the need to create an account, so the question was
-if the need to register would hava a negative impact on user participation. 
+if the need to register would have a negative impact on user participation. 
 The experiment results show that even if only a small percentege of users created an account,

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Changes to original NodIO
Adds biblio

@@
 \title{Increasing User Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System}
-\titlerunning{Increasing Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System}
+\titlerunning{Increasing User Engagement in Volunteer Evolutionary Computing}
 
@@
 \end{abstract}
+
+\section{Introduction}
+
+The World Wide Web provides not only a platform for content
+distribution, but also, thanks to the maturity and reliability of the
+HTTP protocol, an increasingly reliable and high-performance
+operating system for running distributed applications. Besides the
+protocol itself, there are two factors that contribute to this fact:
+the JavaScript virtual machine every browser runs \cite{paulson2005building}
+and the simplified standard interface for interacting with servers
+exemplified by the REST application interface convention \cite{masse2011rest}.
+Thus creating a distributed computing experiment is just a matter of
+making a JavaScript application interchange information with a server,
+by using REST. From the point of view of the programmer, this involves
+relatively common skills and no special libraries, since the
+interface is built in the browser, and a simple application that
+responds to those requests on the server side; both involve just a few
+dozens lines of code additionally to whatever business logic the
+application has. But, more importantly and from the point of view of
+the user, that application can be run by simply visiting a web page.
+
+Using this approach for creating distributed experiments is called {\em
+  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
+\cite{sarmenta2001volunteer} and it dates back, in different shapes
+and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
+here, however, is to use it as a resource for evolutionary
+computation, as our group has done for a long time \cite{jj-ppsn98}.
+
+In this line of research that uses volunteer computing for
+evolutionary algorithms, there are several pending issues. The first
+and maybe most important is approaching volunteer computing as a
+socio-technical system \cite{vespignani2009predicting} which integrates
+user decisions and behavioral patterns in the system model; this
+includes trying to optimize the number of users in a particular
+experiment. The second line of research, although related to the
+first, is more focused on the evolutionary algorithm itself and how
+different design decisions will affect its performance.
+% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
+We have approached the first issue in our previous work \cite{jj-ppsn98},
+% I have justified to focus, in this case, on the second
+but in this paper our focus will be in the second aspect: we will try to
+design a decentralized system that, at the same time, is able to use
+all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
+changing client and the server and measuring its
+impact on the overall performance: time and evaluations needed to find
+the solution. Eventually, we want to find a system that, whatever the
+number of users available to perform the experiment, is able to
+maximize their contribution to the evolutionary algorithm, at the same
+time that the evolutionary algorithm itself makes the most of those
+contributions and is able to find
+the solution to the problem in a minimum time, with the least
+number of contributions.
+
+The rest of the paper is organized as follows: Next we will briefly
+present the state of the art in opportunistic distributed evolutionary
+computation (EC). Section \ref{sec:description} will describe the
+framework and problem used in the experiments, which are publicly
+available under a free license. We will present the results of the
+different steps in the incremental design in Section
+\ref{sec:experiments}, to finally wrap up with the conclusions.
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
 
@@
 
+\bibliographystyle{splncs03}
+\bibliography{volunteer,GA-general,geneura,javascript,ror-js}
 \end{document}

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Added gamification section,
TO DO: Add Lucero’s Related Work

--- 2017-02-16 2017-EUROPAR/Gamification/paper.tex
@@
 \end{abstract}
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+
+\section{Related Work}
+\label{sec:experiments}
+
+\section{Gamification.}
+\label{sec:experiments}
+A definition given by Huotari  \cite{huotari2012defining} is ``Gamification is
+the process by which gaming concepts are brought to the real world tasks associated with
+real people''. Gamification uses game design elements out of the domain of games 
+with the objective of enhancing the user's experience, engagement, productivity, 
+learning, among others. Deterding et al. proposes the following definition:
+ ``Gamification” is the use of game design elements in non-game contexts'' 
+ \cite{deterding2011gamification}.
+
+\subsection{Techniques.}
+Gamification techniques in a volunteer context seeks to persuade 
+users to use their natural desire to compete, learn and socialize in 
+given non-game context application \cite{deterding2011game, hamari2014does}.  
+Some works give a form of reward to users, these include 
+points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework}, 
+the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
+By Making the rewards for  tasks achievements visible to other players or 
+providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
+Competitio can also have problematic consequences, which can result in
+negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
+such as women \cite{kumar2013gamification}. Another techniques to gamification 
+is to make existing tasks feel more like games \cite{deterding2010just}. 
+Some techniques used in this approach include adding meaningful choice, 
+onboarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
 
@@
 
+\bibliographystyle{abbrv}
+\bibliography{../../bib/biblio,../../bib/evospace-i,../../bib/volunteer,../../bib/geneura}
 \end{document}

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Trap problem explained

@@
 \usepackage{subfigure}
+\usepackage{amsmath}
+
 \begin{document}
@@
 \end{abstract}
+
+In the browser, each page visiting the experiment loads an HTML Worker
+that runs a local island of an evolutionary algorithm to solve a
+multimodal problem called {\em l-trap}, which has been used extensively 
+as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
+This function counts the number of bits in a sequence $l$ and assigns
+the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
+bits. As seen in Figure \ref{fig:trap} the fitness falls into a {\em trap} 
+as the number of bits is increased, decreasing linearly until a change in slope 
+is reached at point $z$, adding deceptive component for evolutionary algorithms. 
+To increase the difficulty trap functions can be concatenated.
+In our case we have used $40$ concatenated traps. The trap function  defined as:   
+\[ f(u)= 
+    \begin{cases} 
+      \frac{a}{z}(z-u) & \text{if } u\leq z\\
+      \frac{b}{l-z} (u-z)& \text{otherwise} 
+   \end{cases}
+\]
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=3in]{img/trap.png}
+    \caption{Trap function.
+    }
+    \label{fig:trap}
+\end{figure*}
+
+Each local GA had the following parameters, the initial population was randomly generated 
+with sizes between 128 and 256 individuals, the period to send individuals to the server
+was set at 100 local generations, the parameters for the trap function are $l$ = 4 and $b$ = 2.
 

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Gasification

@@
 
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:comp-onemax}
+\end{figure*}
+
+\section{Related Work}
+\label{sec:experiments}
+In volunteer distributed computing the BOINC projects applies a gamification technique 
+by having a web page to present the ``Top 100 multi-project BOINC participants'' where
+the name of the volunteer, the number of projects, GFLOPS, country and team are presented 
+(\url{https://boinc.berkeley.edu/chart_list.php}).
+
+\section{Gamification Technique}
+\label{sec:experiments}
+A definition given by Huotari  \cite{huotari2012defining} is ``Gamification is
+the process by which gaming concepts are brought to the real world tasks associated with
+real people''. Gamification uses game design elements out of the domain of games 
+with the objective of enhancing the user's experience, engagement, productivity, 
+learning, among others. Deterding et al. proposes the following definition:
+ ``Gamification” is the use of game design elements in non-game contexts'' 
+ \cite{deterding2011gamification}.
+
+Gamification techniques in a volunteer context seeks to persuade 
+users to use their natural desire to compete, learn and socialize in 
+given non-game context application \cite{deterding2011game, hamari2014does}.  
+Some works give a form of reward to users, these include 
+points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework}, 
+the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
+By Making the rewards for  tasks achievements visible to other players or 
+providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
+Competitio can also have problematic consequences, which can result in
+negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
+such as women \cite{kumar2013gamification}. Another techniques to gamification 
+is to make existing tasks feel more like games \cite{deterding2010just}. 
+Some techniques used in this approach include adding meaningful choice, 
+onboarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
+
+\section{Experiment}
+\label{sec:experiments}
 In the browser, each page visiting the experiment loads an HTML Worker
@@
 was set at 100 local generations, the parameters for the trap function are $l$ = 4 and $b$ = 2.
-
-\begin{figure*}[t]
-    \centering
-        \includegraphics[width=5in]{img/login.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
-    }
-    \label{fig:comp-onemax}
-\end{figure*}
-
-\section{Related Work}
-\label{sec:experiments}
-
-\section{Gamification.}
-\label{sec:experiments}
-A definition given by Huotari  \cite{huotari2012defining} is ``Gamification is
-the process by which gaming concepts are brought to the real world tasks associated with
-real people''. Gamification uses game design elements out of the domain of games 
-with the objective of enhancing the user's experience, engagement, productivity, 
-learning, among others. Deterding et al. proposes the following definition:
- ``Gamification” is the use of game design elements in non-game contexts'' 
- \cite{deterding2011gamification}.
-
-\subsection{Techniques.}
-Gamification techniques in a volunteer context seeks to persuade 
-users to use their natural desire to compete, learn and socialize in 
-given non-game context application \cite{deterding2011game, hamari2014does}.  
-Some works give a form of reward to users, these include 
-points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework}, 
-the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
-By Making the rewards for  tasks achievements visible to other players or 
-providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
-Competitio can also have problematic consequences, which can result in
-negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
-such as women \cite{kumar2013gamification}. Another techniques to gamification 
-is to make existing tasks feel more like games \cite{deterding2010just}. 
-Some techniques used in this approach include adding meaningful choice, 
-onboarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
-
-\section{Experiments}
-\label{sec:experiments}
 

### 2017-02-16 2017-EUROPAR/Gamification/paper.tex
Starting experiments

@@
 
-\begin{figure*}[t]
-    \centering
-        \includegraphics[width=5in]{img/login.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
-    }
-    \label{fig:comp-onemax}
-\end{figure*}
-
 \section{Related Work}
@@
 \label{sec:experiments}
+
+In order to test the kind of impact using a gamification technique has on 
+user engagement in an ephemeral evolutionary system, a gamification 
+technique was applied to the {\sf NodIO} volunteer computing framework 
+described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
+In this experiment users have the option to register to then login to 
+the web application, using the graphical interface shown in Figure \ref{fig:login}.
+After login registered users are welcomed by their name, and have
+the oportunity to be listed in a leader-board.
+
+\subsection{Setup}
+
+\subsection{Problem}
 In the browser, each page visiting the experiment loads an HTML Worker
@@
 was set at 100 local generations, the parameters for the trap function are $l$ = 4 and $b$ = 2.
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:login}
+\end{figure*}
+
+The Gamification techinique applied to the NodIO framework
 

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Experimental setup

--- 2017-02-16 2017-EUROPAR/Gamification/paper.tex
@@
 users to use their natural desire to compete, learn and socialize in 
-given non-game context application \cite{deterding2011game, hamari2014does}.  
+given non-game context application \cite{deterding2011game,hamari2014does}.  
 Some works give a form of reward to users, these include 
@@
 providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
-Competitio can also have problematic consequences, which can result in
+Competition can also have problematic consequences, which can result in
 negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
@@
 Some techniques used in this approach include adding meaningful choice, 
-onboarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
+on-boarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
 
@@
 \label{sec:experiments}
+
+In this experiment a gamification technique is applied to 
+the {\sf NodIO} volunteer computing framework, using the 
+particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
+The objective of the experiment is to test the kind of impact 
+applying a gamification technique has on user engagement in this
+particular application. In order to apply a rewarding system
+user authentication had to be developed first. In earlier designs
+this functionality was not desired because it was seen as a barrier 
+for participation. After all the advantage of a browser based volunteer system
+is precisely the minimum amount of user intervention needed to start.
+The last thing a user wants to see is yet another registration form.
+Then, the first design decision for this version is that registration is optional
+and not promoted by the interface. There is only a button with a ``Login/Register'' label,
+using the graphical interface shown in Figure \ref{fig:login}. 
+The registration form is very basic, having only the user-name and 
+password (with out confirmation) as the required fields. Email is optional
+and is not verified. The problem of users that forgot their password
+happened, and it was resolved by sending an email to the administrators
+or some users just created another account.
+
+The gamification technique employed in this work is based on a rewarding mechanism  
+\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
+with score points, levels and leader boards. Points are awarded to users in response of 
+the accomplishment of certain activities that need to be encouraged. In these case
+a point is awarded for each PUT sent to the server. Levels are a long
+term achievement, in this case the level depends on the score:
+\[ \text{level}(\text{score};a,b)= 
+    \begin{cases} 
+      0,                                    &  \text{score}\leq a\\    
+      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
+      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
+      1,                                    & \text{score}\geq b 
+   \end{cases}
+\]
+The function returns a normalized value that is multiplied by the maximum level 
+in this case 100. The variables $a$=100 and $b$=7360 give the slope of the function as
+seen in Figure~\ref{fig:s}. The function was set to give users a rapid increase of 
+levels at the beginning.
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=3in]{img/s.png}
+    \caption{$S$ function used to assign the user's level.
+    }
+    \label{fig:s}
+\end{figure*}
+
+After login, registered users are welcomed by their name, and they can see
+their score and level. All users can see in a modal dialog the leader board.
+While the modal is opened, the scores are refreshed every second. All other 
+functionality is available to all users and is the same as the previous 
+{\sf NodIO} version, showing the state of the current algorithm. 
+In the front page a link to the open source code of the experiment was also 
+available \url{https://github.com/lucero21/login-master}
+
+\subsection{Problem}
 In the browser, each page visiting the experiment loads an HTML Worker
 that runs a local island of an evolutionary algorithm to solve a
-multimodal problem called {\em l-trap}, which has been used extensively 
+multi-modal problem called {\em l-trap}, which has been used extensively 
 as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
@@
 with sizes between 128 and 256 individuals, the period to send individuals to the server
-was set at 100 local generations, the parameters for the trap function are $l$ = 4 and $b$ = 2.
+was set at 100 local generations, the parameters for the trap function are $l$ = 4,
+$a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.
+
+In order to begin the experiment a call to participation was issued on November 23th, 
+2016 through the authors social networks:
+``Asking again for your help, we are conducting a computational experiment
+that requires computer power. Can we borrow some of your CPU? just visit the web page (link)
+and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
+The experiment will run until November 27th, Thanks!''. In the message, the registration
+to the application is presented, and also a deadline of the experiment. When the
+experiment was over, users visiting the page were presented with a thank you message 
+and a static leader board.
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
+    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
+    homogeneous configuration on the left side, and Heterogeneous configuration on the
+    right side of each.
+    }
+    \label{fig:login}
+\end{figure*}
 

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
On with Results

@@
 \label{sec:results}
+At the end of the experiment the resulting log file contained 933,513 contributions both 
+registered users and anonymous volunteers. Registered users where responsible of around
+90\% of the contributions. The JSON data recorded for each contribution is  presented 
+in Table~\ref{tab:record}.  
+\begin{table}
+  \small
+  \caption{ Logfile record structure}
+  \label{tab:record} 
+  \centering
+  \small
+  \begin{tabular}{l  l}
+    \hline\noalign{\smallskip}
+     Attribute & Value \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
+    fitness & fitness as an integer.  \\ \hline
+    IP & IP address of the participant (later anonymized) .\\ \hline
+    user & user-name or the string `anonymous'.  \\ \hline
+    worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
+    level &  level of user or the string `info' \\ \hline
+    message & Only PUT messages were recorded in this experiment. \\ \hline
+    timestamp & A time-stamp based on the Unix epoch\\ \hline
+  \end{tabular}
+\end{table}
 

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
More

@@
 on-boarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
+
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=5in]{img/login.png}
+    \caption{ User interface: Landing page, and Login and registration dialogs.  
+    }
+    \label{fig:login}
+\end{figure*}
 
@@
 
-\begin{figure*}[t]
-    \centering
-        \includegraphics[width=5in]{img/login.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
-    }
-    \label{fig:login}
-\end{figure*}
-
 \subsection{Results}
@@
   \small
-  \caption{ Logfile record structure}
+  \caption{ Log file record structure}
   \label{tab:record} 
@@
 \end{table}
+Contributions are expressed as the number of HTTP PUT requests sent to the server, 
+in Figure~\ref{fig:puts} a comparison between registered vs anonymous users is 
+presented.
 
@@
     \centering
-    \subfigure  [6 workers]
+    \subfigure  [Registered Users]
     {
@@
     }
-    \subfigure  [12 workers]
+    \subfigure  []
     {
@@
 
-    \caption{100 experiments with random parameters for the 128 Bit Griewank 
-    single-objective optimization test function. Experiments are ranked by 
-    the mean time to solution of 5 runs, with (a) 6 workers, and (b) 12 workers.}
-    \label{fig:griewank}
+    \caption{
+        Registered vs Anonymous users.  Users are ranked by
+        the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+        number or PUTs in a logarithmic scale.    }
+    \label{fig:puts}
 \end{figure*}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Discussing Figs

@@
     \centering
-        \includegraphics[width=3in]{img/s.png}
+        \includegraphics[width=2.5in]{img/s.png}
     \caption{$S$ function used to assign the user's level.
@@
     \centering
-        \includegraphics[width=3in]{img/trap.png}
+        \includegraphics[width=2.5in]{img/trap.png}
     \caption{Trap function.
@@
 \end{figure*}
+\begin{table}
+  \small
+  \caption{ Log file record structure}
+  \label{tab:record} 
+  \centering
+  \small
+  \begin{tabular}{l  l}
+    \hline\noalign{\smallskip}
+    Attribute & Value \\
+    \noalign{\smallskip}\hline\noalign{\smallskip}
+    chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
+    fitness & fitness as an integer.  \\ \hline
+    IP & IP address of the participant (later anonymized) .\\ \hline
+    user & user-name or the string `anonymous'.  \\ \hline
+    worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
+    level &  level of user or the string `info' \\ \hline
+    message & Only PUT messages were recorded in this experiment. \\ \hline
+    timestamp & A time-stamp based on the Unix epoch\\ \hline
+  \end{tabular}
+\end{table}
 
@@
 in Table~\ref{tab:record}.
-\begin{table}
-  \small
-  \caption{ Log file record structure}
-  \label{tab:record} 
-  \centering
-  \small
-  \begin{tabular}{l  l}
-    \hline\noalign{\smallskip}
-    Attribute & Value \\
-    \noalign{\smallskip}\hline\noalign{\smallskip}
-    chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
-    fitness & fitness as an integer.  \\ \hline
-    IP & IP address of the participant (later anonymized) .\\ \hline
-    user & user-name or the string `anonymous'.  \\ \hline
-    worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
-    level &  level of user or the string `info' \\ \hline
-    message & Only PUT messages were recorded in this experiment. \\ \hline
-    timestamp & A time-stamp based on the Unix epoch\\ \hline
-  \end{tabular}
-\end{table}
+
 Contributions are expressed as the number of HTTP PUT requests sent to the server, 
 in Figure~\ref{fig:puts} a comparison between registered vs anonymous users is 
-presented.
+presented. There where only 18 registered users, and the exact number of anonymous
+is not known as we only recorded the IP of the request, also a registered users 
+could sometimes be anonymous too, this could be because the first time they visit
+the application they automatically begin to work as anonymous. Nevertheless,
+there were 91 distinct IPs that had `anonymous' as the user. In the plot of
+anonymous users, some of the last in the rank could in fact be registered users, 
+momentarily participating as anonymous. Even if this is not considered, only 
+about 16\% of users attending the call decided to participate as named users. 
+There is a notable difference in the amount of participation between the two
+groups, and also in the slope of the plot.
 
@@
     \centering
-    \subfigure  [Registered Users]
+    \subfigure  [Registered users]
     {
@@
     }
-    \subfigure  []
+    \subfigure  [Anonymous users]
     {
@@
         \includegraphics[width=4in]{img/puts_box.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
+    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
+     In the \emph{y} axis are number or PUTs in a logarithmic scale.
     }
-    \label{fig:comp-onemax}
+    \label{fig:box}
 \end{figure*}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
More results

@@
 and a static leader board. 
-
-\subsection{Results}
-\label{sec:results}
-At the end of the experiment the resulting log file contained 933,513 contributions both 
-registered users and anonymous volunteers. Registered users where responsible of around
-90\% of the contributions. The JSON data recorded for each contribution is  presented 
-in Table~\ref{tab:record}.
-
-Contributions are expressed as the number of HTTP PUT requests sent to the server, 
-in Figure~\ref{fig:puts} a comparison between registered vs anonymous users is 
-presented. There where only 18 registered users, and the exact number of anonymous
-is not known as we only recorded the IP of the request, also a registered users 
-could sometimes be anonymous too, this could be because the first time they visit
-the application they automatically begin to work as anonymous. Nevertheless,
-there were 91 distinct IPs that had `anonymous' as the user. In the plot of
-anonymous users, some of the last in the rank could in fact be registered users, 
-momentarily participating as anonymous. Even if this is not considered, only 
-about 16\% of users attending the call decided to participate as named users. 
-There is a notable difference in the amount of participation between the two
-groups, and also in the slope of the plot.
-
 \begin{figure*}[t]
@@
     \centering
+        \includegraphics[width=5in]{img/workers_put_ip.png}
+    \caption{ Participation of registered users. Each circle represents a user relating  
+        in the \emph{x} axis the number of workers used and in the \emph{y} axis 
+        number or PUTs in a logarithmic scale. The area of the circle is the number 
+        of unique IPs used by the user.
+    }
+    \label{fig:worker-put-ips}
+\end{figure*}
+
+\begin{figure*}[t]
+    \centering
         \includegraphics[width=5in]{img/puts_by_time.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
+    \caption{Number of PUTs by User or IP in 15 minute slots.
     }
-    \label{fig:comp-onemax}
+    \label{fig:puts-time}
 \end{figure*}
@@
         \includegraphics[width=5in]{img/workers_best_user.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
+    \caption{ Number of workers used by the top ranked user in 15 minute slots.
     }
-    \label{fig:comp-onemax}
+    \label{fig:top-user}
 \end{figure*}
-\begin{figure*}[t]
-    \centering
-        \includegraphics[width=5in]{img/workers_put_ip.png}
-    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
-    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
-    homogeneous configuration on the left side, and Heterogeneous configuration on the
-    right side of each.
-    }
-    \label{fig:comp-onemax}
-\end{figure*}
+
+\subsection{Results}
+\label{sec:results}
+At the end of the experiment the resulting log file contained 933,513 contributions both 
+registered users and anonymous volunteers. Registered users where responsible of around
+90\% of the contributions. The JSON data recorded for each contribution is  presented 
+in Table~\ref{tab:record}.
+
+Contributions are expressed as the number of HTTP PUT requests sent to the server, 
+in Figure~\ref{fig:puts} a comparison between registered vs anonymous users is 
+presented. There where only 18 registered users, and the exact number of anonymous
+is not known as we only recorded the IP of the request, also a registered users 
+could sometimes be anonymous too, this could be because the first time they visit
+the application they automatically begin to work as anonymous. Nevertheless,
+there were 91 distinct IPs that had `anonymous' as the user. In the plot of
+anonymous users, some of the last in the rank could in fact be registered users, 
+momentarily participating as anonymous. Even if this is not considered, only 
+about 16\% of users attending the call decided to participate as named users. 
+There is a notable difference in the amount of participation between the two
+groups, and also in the slope of the plot.
+
+The box plot of the amount of participation between the two groups, highlights
+the difference and also the number of outliers in the groups. Two registered 
+outliers where very important participants maybe competing between them. But
+there where more outliers between anonymous IPs, and some of them could be related
+to the same users. The lower quartile in in the users box-plot has more
+spread, indicating a participation similar to the median of anonymous IPs.
+
+Figure~\ref{fig:worker-put-ips} gives an interesting view on the amount of
+resources shared by registered users. The area of the circle is proportional
+to the number of unique IPs used by each user, the user with more participation
+used a total of 66 different IPs, this number could be related with the number
+of devices used during the participation. There are some users that using less
+devices participated more, this could mean a more powerful device, faster 
+Internet connection or simply more time spent participating.
+
+User engagement is related to the amount of time a user spends in the 
+application. Figure~\ref{puts-time} shows to important aspects of
+engagement, the overall time spent and the amount of resources shared
+in that time. Registered users where more engaged during the experiment.
+They contribute more resources during longer periods of time.
+
+The participation of the top ranked user is presented in 15 seconds time slots in
+Figure~\ref{fig:top-user}. In the beginning this user was using more than 30
+workers, this means that more than 30 tabs of the page where open at the
+same time, perhaps using several computing devices. Some users reported that 
+they wanted to test the limits of their own systems, checking the percentage of
+CPU they where using.
 

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Adding boilerplate

--- 2017-02-17 2017-EUROPAR/Gamification/paper.tex
@@
 \section{Related Work}
-\label{sec:experiments}
+\label{sec:soa}
 In volunteer distributed computing the BOINC projects applies a gamification technique 
@@
     }
-    \label{fig:comp-onemax}
+    \label{fig:om:puts}
   \end{figure*}
+  %
 \begin{figure*}[t]
@@
     }
-    \label{fig:comp-onemax}
+    \label{fig:om:workers}
 \end{figure*}
@@
     }
-    \label{fig:comp-onemax}
+    \label{fig:om:wp}
 \end{figure*}
 
-\bibliographystyle{abbrv}
+\section*{Acknowledgment}
+
+This work has been supported in part by: de Ministerio espa\~{n}ol de
+Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
+(UGR-EPHEMECH).
+
+\bibliographystyle{splncs03}
 \bibliography{../../bib/biblio,../../bib/evospace-i,../../bib/volunteer,../../bib/geneura}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Sections
@JJ Please check, Im running out of energy. LOW BAT

--- 2017-02-17 2017-EUROPAR/Gamification/paper.tex
@@
 \section{Related Work}
-\label{sec:experiments}
+\label{sec:relatedwork}
 In volunteer distributed computing the BOINC projects applies a gamification technique 
@@
 \section{Gamification Technique}
-\label{sec:experiments}
+\label{sec:gamification}
 A definition given by Huotari  \cite{huotari2012defining} is ``Gamification is
@@
 User engagement is related to the amount of time a user spends in the 
-application. Figure~\ref{puts-time} shows to important aspects of
+application. Figure~\ref{fig:puts-time} shows to important aspects of
 engagement, the overall time spent and the amount of resources shared

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex

@@
 \section{Related Work}
-\label{sec:relatedwork}
+\label{sec:soa}
 In volunteer distributed computing the BOINC projects applies a gamification technique 
@@
 
-\bibliographystyle{abbrv}
+\section*{Acknowledgment}
+
+This work has been supported in part by: de Ministerio espa\~{n}ol de
+Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
+(UGR-EPHEMECH).
+
+\bibliographystyle{splncs03}
+
 \bibliography{../../bib/biblio,../../bib/evospace-i,../../bib/volunteer,../../bib/geneura}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Fixing biblio

@@
 
+\section{Introduction}
+
+The World Wide Web provides not only a platform for content
+distribution, but also, thanks to the maturity and reliability of the
+HTTP protocol, an increasingly reliable and high-performance
+operating system for running distributed applications. Besides the
+protocol itself, there are two factors that contribute to this fact:
+the JavaScript virtual machine every browser runs \cite{paulson2005building}
+and the simplified standard interface for interacting with servers
+exemplified by the REST application interface convention \cite{masse2011rest}.
+Thus creating a distributed computing experiment is just a matter of
+making a JavaScript application interchange information with a server,
+by using REST. From the point of view of the programmer, this involves
+relatively common skills and no special libraries, since the
+interface is built in the browser, and a simple application that
+responds to those requests on the server side; both involve just a few
+dozens lines of code additionally to whatever business logic the
+application has. But, more importantly and from the point of view of
+the user, that application can be run by simply visiting a web page.
+
+Using this approach for creating distributed experiments is called {\em
+  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
+\cite{sarmenta2001volunteer} and it dates back, in different shapes
+and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
+here, however, is to use it as a resource for evolutionary
+computation, as our group has done for a long time \cite{jj-ppsn98}.
+
+In this line of research that uses volunteer computing for
+evolutionary algorithms, there are several pending issues. The first
+and maybe most important is approaching volunteer computing as a
+socio-technical system \cite{vespignani2009predicting} which integrates
+user decisions and behavioral patterns in the system model; this
+includes trying to optimize the number of users in a particular
+experiment. The second line of research, although related to the
+first, is more focused on the evolutionary algorithm itself and how
+different design decisions will affect its performance.
+% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
+We have approached the first issue in our previous work \cite{jj-ppsn98},
+% I have justified to focus, in this case, on the second
+but in this paper our focus will be in the second aspect: we will try to
+design a decentralized system that, at the same time, is able to use
+all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
+changing client and the server and measuring its
+impact on the overall performance: time and evaluations needed to find
+the solution. Eventually, we want to find a system that, whatever the
+number of users available to perform the experiment, is able to
+maximize their contribution to the evolutionary algorithm, at the same
+time that the evolutionary algorithm itself makes the most of those
+contributions and is able to find
+the solution to the problem in a minimum time, with the least
+number of contributions.
+
+The rest of the paper is organized as follows: Next we will briefly
+present the state of the art in opportunistic distributed evolutionary
+computation (EC). Section \ref{sec:description} will describe the
+framework and problem used in the experiments, which are publicly
+available under a free license. We will present the results of the
+different steps in the incremental design in Section
+\ref{sec:experiments}, to finally wrap up with the conclusions.
+
 \section{Related Work}
@@
 on-boarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
+
+>>>>>>> e0b4fdb1a1917df769b649897a4a91e53129b0a3
 
@@
 
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 
-This work has been supported in part by: de Ministerio espa\~{n}ol de
+This work has been supported in part by: Ministerio espa\~{n}ol de
 Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
@@
 \bibliography{../../bib/biblio,../../bib/evospace-i,../../bib/volunteer,../../bib/geneura}
+
 \end{document}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Incorporating old state of the art

@@
 present the state of the art in opportunistic distributed evolutionary
-computation (EC). Section \ref{sec:description} will describe the
+computation (EC). Section \ref{sec:gamification} will describe the
 framework and problem used in the experiments, which are publicly
@@
 \label{sec:soa}
+
+Volunteer computing involves users deciding to run a program that acts
+as a client or as a peer in an experiment
+and, as such, has been deployed in many different ways from the
+beginning of the Internet, starting with the SETI@home framework for
+processing extraterrestrial signals \cite{david-seti:home}, or a 
+high-throughput queuing system such as HTCondor \cite{HTCondor}.
+However,
+the dual facts of the introduction of JavaScript as a universal language for the
+browser and the browser itself as both an ubiquitous web and Internet client has
+made this combination the most popular for volunteer computing
+frameworks such as the one we are using here, and whose first version
+was described in \cite{DBLP:conf/gecco/GuervosG15}. Systems based on the
+JavaScript/browser combination
+emphasize the ephemerality, ease of use, and universality, while
+systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
+require higher availability of volunteer resources, achieving this
+with downloadable clients which in some cases, like HTCondor, need
+administration access to the resources.
+
+Several authors have already described systems using JavaScript  either
+for unwitting
+\cite{unwitting-ec,boldrin2007distributed,apolonia2012enhancing} or volunteer 
+\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
+evolutionary computation and it has been used ever since by several
+authors, including more recent efforts
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}. In fact, this last paper \cite{pan2015gray} performs
+an analysis of what it calls {\em Gray computing} doing an analysis of
+feasibility, cost-effectiveness, change in the user experience and
+architectural optimization needed, concluding that the computing power
+available can be vast and it that it can be cost-effective to use it.
+
+Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness 
+evaluations across a heterogeneous pool of cloud compute nodes and volunteer
+peer computers running a browser. In SPACE peers establish a bi-directional 
+communication with the server using the \texttt{Socket.io} JavaScript library.  
+As an experiment they provide an \texttt{ASM.js} compiled version of the RoboGen 
+software platform, showing that JavaScript can be used in a broad 
+spectrum of applications.
+
+Recent works have been using crowdsourcing in order to train robots 
+for human interaction, a representative work of this trend is that of  
+Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose 
+training simulated robots for the grounding of language symbols. They do this by using human 
+observers that issue arbitrary commands to these robots via the web, 
+providing positive or negative reinforcement in response to the
+robot\'s resulting actions. For interaction, they use the Twitch 
+video streaming platform where users can issue commands through the
+platform's integrated chat service. Subjects were incentivized to 
+interact with the system by GUI features that provided participants
+with a sense of involvement with the simulation. This kind of systems
+could add additional dynamics, because users can collaborate or 
+even compete with each other when issuing their commands.
+
+Most
+of these papers focused on testing the system by itself as proofs of
+concepts, more than on its performance or their actual influence on
+the result of an evolutionary algorithm. In fact, some of them do not
+even work in as a distributed computing system
+\cite{EvoStar2014:jsEO}, describing the expressive capabilities of
+JavaScript for evolutionary computation.
+
+The proofs of concept described above do not go further than trying to
+find out  how many users join the effort and how many the system is
+able to support. In fact, systems described in
+\cite{gecco07:workshop:dcor} had serious scaling issues; some of them
+also tried to find out how much time was needed to find the solution
+or, alternatively, how many users would be needed to be competitive
+with single-user single-computer implementations of the same
+algorithm. However, the latest systems try to emphasize the seamless
+integration of peer and volunteer computing systems \cite{leclerc2016seamless}.
+
+This latest system offloads evaluation to browsers so the actual
+algorithm runs in some other place; in general, most researchers do
+not try to assess the performance of the algorithm 
+itself in the precise conditions they find in a volunteer computing
+system, although papers such as the one by Laredo et
+al. \cite{churn08:ijhpsa}, using models, try to find out the resulting
+performance when the users are changing continuously according to a
+Weibull distribution. In the case of Klein and Spector \cite{unwitting-ec}, the algorithm
+is actually run on the server, using the browser mainly for fitness
+evaluation, and thus no actual contribution to the evaluation of the
+evolutionary algorithm itself is made.
+
+There is another factor that should be taken into account: since the
+user has control of the browser, there is a limited amount of
+interaction with it, namely, the fact that by reloading the webpage they
+can apply a kind of hypermutation, killing the current population and
+generating new individuals some of which will make their way to the
+common pool via migration. In that sense, volunteer computing is also
+a way of {\em human computation} \cite{quinn2011human}, a concept that
+has also been applied to evolutionary algorithms \cite{972056}, in
+this case extensively and with all operators. It is quite difficult to
+find out how many times this happens and what is their effect on the
+overall algorithm, but this only reaffirms the fact that there is a
+long list of issues with volunteer evolutionary computation, and that
+the volunteer him or herself is at the center of many of them.
+
+In this paper we will make contributions towards the design of
+efficient volunteer evolutionary algorithms by studying the effect of
+several design decisions on performance. In the next Section we will
+present the general framework and the initial setup.
+
 In volunteer distributed computing the BOINC projects applies a gamification technique 
@@
 
->>>>>>> e0b4fdb1a1917df769b649897a4a91e53129b0a3
-
-\begin{figure*}[t]
+\begin{figure*}[htbp]
     \centering
@@
 levels at the beginning.
-\begin{figure*}[t]
+\begin{figure*}[htbp]
     \centering
@@
 \]
-\begin{figure*}[t]
+\begin{figure*}[htbp]
     \centering
@@
 \end{figure*}
-\begin{table}
+%
+\begin{table}[htbp]
   \small
@@
 and a static leader board. 
-\begin{figure*}[t]
+%
+\begin{figure*}[htbp]
     \centering
@@
 \end{figure*}
-
-\begin{figure*}[t]
+%
+\begin{figure*}[htbp]
     \centering
@@
 \end{figure*}
-\begin{figure*}[t]
+\begin{figure*}[htbp]
     \centering

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Rewriting intro

@@
 
-The World Wide Web provides not only a platform for content
-distribution, but also, thanks to the maturity and reliability of the
-HTTP protocol, an increasingly reliable and high-performance
+The World Wide Web is an increasingly reliable and high-performance
 operating system for running distributed applications. Besides the
-protocol itself, there are two factors that contribute to this fact:
-the JavaScript virtual machine every browser runs \cite{paulson2005building}
-and the simplified standard interface for interacting with servers
-exemplified by the REST application interface convention \cite{masse2011rest}.
-Thus creating a distributed computing experiment is just a matter of
-making a JavaScript application interchange information with a server,
-by using REST. From the point of view of the programmer, this involves
-relatively common skills and no special libraries, since the
-interface is built in the browser, and a simple application that
-responds to those requests on the server side; both involve just a few
-dozens lines of code additionally to whatever business logic the
-application has. But, more importantly and from the point of view of
-the user, that application can be run by simply visiting a web page.
-
-Using this approach for creating distributed experiments is called {\em
+maturity of HTTP, the underlying protocol itself, there are other
+factors: the JavaScript virtual machine embedded in every browser
+and the REST interface, a {\em de facto} standard interface for
+communication. A distributed system based on the browser can be
+created easily using these mechanisms, which have lately been enhanced
+by WebSockets for lower latency; you can gather users by just
+announcing the URL. This approach for creating distributed experiments is called {\em
   volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
 \cite{sarmenta2001volunteer} and it dates back, in different shapes
-and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
-here, however, is to use it as a resource for evolutionary
-computation, as our group has done for a long time \cite{jj-ppsn98}.
+and underlying mechanisms, to the origin of the web
+\cite{david-seti:home}.
+
+We are mainly interested in evolutionary algorithms \cite{jj-ppsn98},
+which are population-based search and optimization methods inspired by
+the theory of Evolution and its molecular basis. The fact that they
+are population-based makes easy its distribution in different clients
+and since their asynchronous nature has no impact on their performance
+and might even enhance it, makes them ideal candidates for volunteer
+computing setups such as the one presented in this paper.
 
 In this line of research that uses volunteer computing for
-evolutionary algorithms, there are several pending issues. The first
-and maybe most important is approaching volunteer computing as a
+evolutionary algorithms, there are several pending issues, of which
+the most important is approaching volunteer computing as a 
 socio-technical system \cite{vespignani2009predicting} which integrates
@@
 includes trying to optimize the number of users in a particular
-experiment. The second line of research, although related to the
-first, is more focused on the evolutionary algorithm itself and how
-different design decisions will affect its performance.
-% Is it ok that we say that the first line of research is the most important... but then we focus on the second?
-We have approached the first issue in our previous work \cite{jj-ppsn98},
-% I have justified to focus, in this case, on the second
-but in this paper our focus will be in the second aspect: we will try to
-design a decentralized system that, at the same time, is able to use
-all available resources for finding the solution of an evolutionary algorithm. This design will be done incrementally by
-changing client and the server and measuring its
-impact on the overall performance: time and evaluations needed to find
-the solution. Eventually, we want to find a system that, whatever the
+experiment. The challenge is to design a system that, whatever the
 number of users available to perform the experiment, is able to
-maximize their contribution to the evolutionary algorithm, at the same
-time that the evolutionary algorithm itself makes the most of those
-contributions and is able to find
-the solution to the problem in a minimum time, with the least
-number of contributions.
+maximize their contribution to the evolutionary algorithm. One obvious
+way of achieving that is trying to find as many users as possible. But
+another approach, and the one that we will be taking in this paper, is
+to delve into the socio-technical nature of the system, using
+gamification to improve the amount of time every user will lend to the
+system. Gamification has to include some way of identifying the user,
+which is a hurdle that might take users away; however, the open design
+of the system we use means that we can actually mix anonymous-non
+gamified and identified-gamified users, with only the small overhead
+of adding the mechanism for gamification.
 

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Back to 10 pages

@@
 
-This latest system offloads evaluation to browsers so the actual
-algorithm runs in some other place; in general, most researchers do
-not try to assess the performance of the algorithm 
-itself in the precise conditions they find in a volunteer computing
-system, although papers such as the one by Laredo et
-al. \cite{churn08:ijhpsa}, using models, try to find out the resulting
-performance when the users are changing continuously according to a
-Weibull distribution. In the case of Klein and Spector \cite{unwitting-ec}, the algorithm
-is actually run on the server, using the browser mainly for fitness
-evaluation, and thus no actual contribution to the evaluation of the
-evolutionary algorithm itself is made.
-
-There is another factor that should be taken into account: since the
-user has control of the browser, there is a limited amount of
-interaction with it, namely, the fact that by reloading the webpage they
-can apply a kind of hypermutation, killing the current population and
-generating new individuals some of which will make their way to the
-common pool via migration. In that sense, volunteer computing is also
-a way of {\em human computation} \cite{quinn2011human}, a concept that
-has also been applied to evolutionary algorithms \cite{972056}, in
-this case extensively and with all operators. It is quite difficult to
-find out how many times this happens and what is their effect on the
-overall algorithm, but this only reaffirms the fact that there is a
-long list of issues with volunteer evolutionary computation, and that
-the volunteer him or herself is at the center of many of them.
-
-In this paper we will make contributions towards the design of
-efficient volunteer evolutionary algorithms by studying the effect of
-several design decisions on performance. In the next Section we will
-present the general framework and the initial setup.
-
-In volunteer distributed computing the BOINC projects applies a gamification technique 
+As indicated in the introduction, the human is an integral part of the system, which can be considered {\em human computation} \cite{quinn2011human}.
+Giving the user not only some more control, by including registration, but also a measure to compare with other users might have positive effects.
+In fact, the BOINC projects applies a gamification technique
 by having a web page to present the ``Top 100 multi-project BOINC participants'' where
@@
 The last thing a user wants to see is yet another registration form.
-Then, the first design decision for this version is that registration is optional
-and not promoted by the interface. There is only a button with a ``Login/Register'' label,
-using the graphical interface shown in Figure \ref{fig:login}. 
-The registration form is very basic, having only the user-name and 
-password (with out confirmation) as the required fields. Email is optional
-and is not verified. The problem of users that forgot their password
-happened, and it was resolved by sending an email to the administrators
-or some users just created another account.
+Then, the first design decision for this version is that registration is optional and simple as shown in Figure \ref{fig:login}.
 
@@
 The function returns a normalized value that is multiplied by the maximum level 
-in this case 100. The variables $a$=100 and $b$=7360 give the slope of the function as
-seen in Figure~\ref{fig:s}. The function was set to give users a rapid increase of 
+in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of 
 levels at the beginning.
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=2.5in]{img/s.png}
-    \caption{$S$ function used to assign the user's level.
-    }
-    \label{fig:s}
-\end{figure*}
+% \begin{figure*}[htbp]
+%     \centering
+%         \includegraphics[width=2.5in]{img/s.png}
+%     \caption{$S$ function used to assign the user's level.
+%     }
+%     \label{fig:s}
+% \end{figure*}
 
@@
 \subsection{Problem}
-In the browser, each page visiting the experiment loads an HTML Worker
+In the browser, each page visiting the experiment loads an WebWorker
 that runs a local island of an evolutionary algorithm to solve a
@@
 the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
-bits. As seen in Figure \ref{fig:trap} the fitness falls into a {\em trap} 
+bits, this makes the  the fitness fall into a {\em trap} 
 as the number of bits is increased, decreasing linearly until a change in slope 
@@
 \]
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=2.5in]{img/trap.png}
-    \caption{Trap function.
-    }
-    \label{fig:trap}
-\end{figure*}
+% \begin{figure*}[htbp]
+%     \centering
+%         \includegraphics[width=2.5in]{img/trap.png}
+%     \caption{Trap function.
+%     }
+%     \label{fig:trap}
+% \end{figure*}
 %
-\begin{table}[htbp]
+
+Each local GA had the following parameters, the initial population was randomly generated 
+with sizes between 128 and 256 individuals, the period to send individuals to the server
+was set at 100 local generations, the parameters for the trap function are $l$ = 4,
+$a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.
+
+We made a call to participation on November 23th, 
+2016 through the authors' social networks: ``Asking again for your help, we are conducting a computational experiment
+that requires computer power. Can we borrow some of your CPU? just visit the web page (link)
+and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
+The experiment will run until November 27th, Thanks!.'' We also included links to register and a deadline.
+Users visiting the pageafter the dealine were presented with a thank you message and a static leader board.
+%
+
+\subsection{Results}
+\label{sec:results}
+
+\begin{table}[h!tbp]
   \small
@@
 \end{table}
-
-Each local GA had the following parameters, the initial population was randomly generated 
-with sizes between 128 and 256 individuals, the period to send individuals to the server
-was set at 100 local generations, the parameters for the trap function are $l$ = 4,
-$a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.
-
-In order to begin the experiment a call to participation was issued on November 23th, 
-2016 through the authors social networks:
-``Asking again for your help, we are conducting a computational experiment
-that requires computer power. Can we borrow some of your CPU? just visit the web page (link)
-and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
-The experiment will run until November 27th, Thanks!''. In the message, the registration
-to the application is presented, and also a deadline of the experiment. When the
-experiment was over, users visiting the page were presented with a thank you message 
-and a static leader board. 
 %
-\begin{figure*}[htbp]
-    \centering
-    \subfigure  [Registered users]
-    {
-        \includegraphics[width=4in]{img/puts_user.png}
-    }
-    \subfigure  [Anonymous users]
-    {
-        \includegraphics[width=4in]{img/puts_ip.png}
-    }
-
-    \caption{
-        Registered vs Anonymous users.  Users are ranked by
-        the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis 
-        number or PUTs in a logarithmic scale.    }
-    \label{fig:puts}
-\end{figure*}
-%
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=4in]{img/puts_box.png}
-    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
-     In the \emph{y} axis are number or PUTs in a logarithmic scale.
-    }
-    \label{fig:box}
-\end{figure*}
-%
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=5in]{img/workers_put_ip.png}
-    \caption{ Participation of registered users. Each circle represents a user relating  
-        in the \emph{x} axis the number of workers used and in the \emph{y} axis 
-        number or PUTs in a logarithmic scale. The area of the circle is the number 
-        of unique IPs used by the user.
-    }
-    \label{fig:worker-put-ips}
-\end{figure*}
-%
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=5in]{img/puts_by_time.png}
-    \caption{Number of PUTs by User or IP in 15 minute slots.
-    }
-    \label{fig:puts-time}
-\end{figure*}
-\begin{figure*}[htbp]
-    \centering
-        \includegraphics[width=5in]{img/workers_best_user.png}
-    \caption{ Number of workers used by the top ranked user in 15 minute slots.
-    }
-    \label{fig:top-user}
-\end{figure*}
-
-\subsection{Results}
-\label{sec:results}
-At the end of the experiment the resulting log file contained 933,513 contributions both 
+At the end of the experiment the resulting log file (whose fields are shown in Table \ref{tab:record}) contained 933,513 contributions both 
 registered users and anonymous volunteers. Registered users where responsible of around
@@
 
+\begin{figure}[h!tbp]
+    \centering
+        \includegraphics[width=2.3in]{img/puts_user.png}
+        \includegraphics[width=2.3in]{img/puts_ip.png}
+    \caption{Usage by registered (left) and anonymous (right) users.  Users are ranked by
+        the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis 
+        number or PUTs in a logarithmic scale.}
+    \label{fig:puts}
+\end{figure}
+%
 Contributions are expressed as the number of HTTP PUT requests sent to the server, 
@@
 
-The box plot of the amount of participation between the two groups, highlights
+%
+\begin{figure*}[htbp]
+    \centering
+        \includegraphics[width=4in]{img/puts_box.png}
+    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
+     In the \emph{y} axis are number or PUTs in a logarithmic scale.
+    }
+    \label{fig:box}
+\end{figure*}
+%
+The box plot \ref{fig:box} of the amount of participation between the two groups, highlights
 the difference and also the number of outliers in the groups. Two registered 
@@
 
+%
+\begin{figure*}[htbp]
+    \centering
+        \includegraphics[width=5in]{img/workers_put_ip.png}
+    \caption{ Participation of registered users. Each circle represents a user relating  
+        in the \emph{x} axis the number of workers used and in the \emph{y} axis 
+        number or PUTs in a logarithmic scale. The area of the circle is the number 
+        of unique IPs used by the user.
+    }
+    \label{fig:worker-put-ips}
+\end{figure*}
+%
 Figure~\ref{fig:worker-put-ips} gives an interesting view on the amount of
@@
 
+%
+\begin{figure*}[htbp]
+    \centering
+        \includegraphics[width=4in]{img/puts_by_time.png}
+    \caption{Number of PUTs by User or IP in 15 minute slots.
+    }
+    \label{fig:puts-time}
+\end{figure*}
+%
+\begin{figure*}[htbp]
+    \centering
+        \includegraphics[width=5in]{img/workers_best_user.png}
+    \caption{ Number of workers used by the top ranked user in 15 minute slots.
+    }
+    \label{fig:top-user}
+\end{figure*}
+%
 The participation of the top ranked user is presented in 15 seconds time slots in

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Adds gamification references

@@
 even compete with each other when issuing their commands.
-
-Most
-of these papers focused on testing the system by itself as proofs of
-concepts, more than on its performance or their actual influence on
-the result of an evolutionary algorithm. In fact, some of them do not
-even work in as a distributed computing system
-\cite{EvoStar2014:jsEO}, describing the expressive capabilities of
-JavaScript for evolutionary computation.
+This approach has been called {\em social cloud} \cite{6404452}, enphasizing the fact that user participation is gamified in order to obtain the most from it \cite{7027564}.
 
@@
 find out  how many users join the effort and how many the system is
-able to support. In fact, systems described in
+able to support. In fact, systems such as the one described in
 \cite{gecco07:workshop:dcor} had serious scaling issues; some of them
@@
 with single-user single-computer implementations of the same
-algorithm. However, the latest systems try to emphasize the seamless
-integration of peer and volunteer computing systems \cite{leclerc2016seamless}.
+algorithm.
+Lately, researchers have tried to integrate volunteer as a part of a larger distributed computint effort \cite{leclerc2016seamless}.
+In fact, exclusive volunteer systems exhibit a certain amount of unpredictability and might be best used combined with available computing power.
 
@@
 
-\section{Gamification Technique}
+Next we will present the gamification technique used in this experiment.
+
+\section{Gamification}
 \label{sec:gamification}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Adds conclusions

@@
 
+\section{Conclusions, discussion and future work}
+
+In this paper we have added gamification to a volunteer computing
+system via a leaderboard of contributions of registered users. In
+fact, we have proved that, although only a minority of the users opt
+for registration, their individual contribution is on average much
+higher than for non-registered/non-gamified users. The fact that less
+than one fifth of users opt for registering would discard the
+exclusive use of this feature; however, it is an interesting addition
+to the volunteer system that in fact enhances its computing power.
+
 \section*{Acknowledgments}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Final version submitted

@@
 
-\begin{figure*}[htbp]
+\begin{figure*}[htb]
     \centering
@@
 levels at the beginning.
-% \begin{figure*}[htbp]
+% \begin{figure*}[htb]
 %     \centering
@@
 \]
-% \begin{figure*}[htbp]
+% \begin{figure*}[htb]
 %     \centering
@@
 %
-\begin{figure*}[htbp]
+\begin{figure*}[htb]
     \centering
@@
 %
-The box plot \ref{fig:box} of the amount of participation between the two groups, highlights
+The box plot featured in Figure \ref{fig:box} shows the amount of participation between the two groups, highlights
 the difference and also the number of outliers in the groups. Two registered 
@@
 
-User engagement is related to the amount of time a user spends in the 
-application. Figure~\ref{fig:puts-time} shows to important aspects of
-engagement, the overall time spent and the amount of resources shared
-in that time. Registered users where more engaged during the experiment.
-They contribute more resources during longer periods of time.
-
-%
-\begin{figure*}[htbp]
+\begin{figure*}[htb]
     \centering
@@
 %
-\begin{figure*}[htbp]
+User engagement is related to the amount of time a user spends in the 
+application. Figure~\ref{fig:puts-time} shows to important aspects of
+engagement, the overall time spent and the amount of resources shared
+in that time. Registered users where more engaged during the experiment.
+They contribute more resources during longer periods of time.
+
+\begin{figure*}[htb]
     \centering
@@
 
+The fact that gamification actually work highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation.
+That is why future lines of work will look for intermediate ways of including
+gamification without registration, as well as some ways of making the user increase control over the algorithm that would go beyond simply staying for longer or leaving the page.
+Since it is a sociotechnical system, including social features in it and studying emerging social networks might be an interesting line of work too.
+
+Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
+
 \section*{Acknowledgments}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Conclusion added
@JJ Please check if there is time,

--- 2017-02-17 2017-EUROPAR/Gamification/paper.tex
@@
 
+\section{Conclusions}
+\label{sec:conclusions}
+
+The objective of this work was to test the kind of impact 
+applying a gamification technique has on user engagement in 
+the {\sf NodIO} volunteer computing framework; In concordance with 
+the results obtained in other 
+browser-based volunteer systems, after 
+applying the gamification techniques, user participation was increased.
+
+One of the interesting future lines of work would be to look a bit
+more closely at the behavior of users as they are participating  
+in the web system.
+
+Another line of work would be to study the possible negative effects of using  
+gamification techniques to improve engagement, like cheating or
+literally {\em gaming} the system to defeat competition. We already
+found some hints of this behavior, but more subtle effect could be taking place. 
+Finally, the refinement of the proposed framework will need
+more case studies and further multi-disciplinary research.
+
 \section*{Acknowledgments}

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Spelling checked
And caught quite a few... Phew!

@@
 provide their own computing resources or storage to contribute to a common effort.
-By runnning a script in a web page, collaboration is straightforward, but also ephemeral.
-Resources depend on the amount of time a user lends, whicn means that 
+By running a script in a web page, collaboration is straightforward, but also ephemeral.
+Resources depend on the amount of time a user lends, which means that 
 the user has to be kept engaged to obtain as many computing cycles as
@@
 participation thus increasing the overall computing power. We present the results of
-an experiment where a gammification technique is applied by adding a leaderboard 
+an experiment where a gamification technique is applied by adding a leader-board 
 showing the top scores achieved by registered contributors. In the NodIO system volunteers can
@@
 if the need to register would have a negative impact on user participation. 
-The experiment results show that even if only a small percentege of users created an account,
+The experiment results show that even if only a small percentage of users created an account,
 those participating in the competition provided around 90\%.
@@
 even compete with each other when issuing their commands.
-This approach has been called {\em social cloud} \cite{6404452}, enphasizing the fact that user participation is gamified in order to obtain the most from it \cite{7027564}.
+This approach has been called {\em social cloud} \cite{6404452}, emphasizing the fact that user participation is gamified in order to obtain the most from it \cite{7027564}.
 
@@
 algorithm.
-Lately, researchers have tried to integrate volunteer as a part of a larger distributed computint effort \cite{leclerc2016seamless}.
+Lately, researchers have tried to integrate volunteer as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
 In fact, exclusive volunteer systems exhibit a certain amount of unpredictability and might be best used combined with available computing power.
@@
 The experiment will run until November 27th, Thanks!.'' We also included links to register and a deadline.
-Users visiting the pageafter the dealine were presented with a thank you message and a static leader board.
+Users visiting the page after the deadline were presented with a thank you message and a static leader board.
 %

### 2017-04-05 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Authors

@@
 system \cite{eiben2015interactive}.
-They are usually applied in problems where the fitness function is not known or simply
-does not exist, and the result of optimization should fit a certain human need or
-desire such as an aesthetic ideal.  
-That is why their use cases include the evolution of objects with subjective characteristics,
-such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
-considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
-or creativity \cite{yu2011cooks}. In the cases when 
-human interaction is responsible of other 
-aspects of the evolutionary process, these IEC methods are classified by some authors  
-as human-based evolutionary computation \cite{972056} 
-or as human-based computation \cite{quinn2011human}.
+% They are usually applied in problems where the fitness function is not known or simply
+% does not exist, and the result of optimization should fit a certain human need or
+% desire such as an aesthetic ideal.  
+% That is why their use cases include the evolution of objects with subjective characteristics,
+% such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
+% considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
+% or creativity \cite{yu2011cooks}. In the cases when 
+% human interaction is responsible of other 
+% aspects of the evolutionary process, these IEC methods are classified by some authors  
+% as human-based evolutionary computation \cite{972056} 
+% or as human-based computation \cite{quinn2011human}.
 
 % Human Based Computation
-IEC systems are an interesting venue of research, since they have demonstrated 
-their ability for effectively 
-producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-as well as other types of artifacts in many other domains \cite{ie1}. 
+% IEC systems are an interesting venue of research, since they have demonstrated 
+% their ability for effectively 
+% producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
+% as well as other types of artifacts in many other domains \cite{ie1}. 
 % Problems
@@
 many of which are not interesting or are very similar to each other.
-% C-IEAs
-Moreover, the performance of these systems effectively depends on the number of users
-they are able to include; to reach more users,
-IEC systems are some times developed as web applications depending on visitors to help
-with the search, using both anonymous and registered users. Some systems 
-employ a collaborative technique, where several users participate in 
-the evaluation, this method is called Collaborative-IEC (C-IEC)
-\cite{picbreeder,seyama2016development,wagy2014collective}.
+% % C-IEAs
+% Moreover, the performance of these systems effectively depends on the number of users
+% they are able to include; to reach more users,
+% IEC systems are some times developed as web applications depending on visitors to help
+% with the search, using both anonymous and registered users. Some systems 
+% employ a collaborative technique, where several users participate in 
+% the evaluation, this method is called Collaborative-IEC (C-IEC)
+% \cite{picbreeder,seyama2016development,wagy2014collective}.
 % Volunteers 
@@
 
-For example, in a C-IEC application fitness assignment depends on the
-actions of a social network of users.  These actions are triggered when 
-they tag, share, rate, store or delete a phenotype. 
-Then, the selection of parents could depend on the previous actions, leveraging information, 
-such as the fact that they both have the same tag, or were shared by
-similar users.
+% For example, in a C-IEC application fitness assignment depends on the
+% actions of a social network of users.  These actions are triggered when 
+% they tag, share, rate, store or delete a phenotype. 
+% Then, the selection of parents could depend on the previous actions, leveraging information, 
+% such as the fact that they both have the same tag, or were shared by
+% similar users.
 
@@
 
-It was shown how the estimation of fitness of candidate solutions
-can take advantage of a graph model representation. Also, the graph
-offers another view of the system that it would have been
-difficult to replicate with a relational database. 
-% More conclusions and some way of relating the three case studies
-% would be something that is missing and should definitely go here - JJ
-% No longer needed as they where removed - Mario
-
 One of the interesting future lines of work would be to look a bit

### 2017-04-05 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Authors

@@
  Evolutionary Computation}
-%\titlenote{Produces the permission block, and
-%  copyright information}
-\subtitle{Anonymized Version}
-%\subtitlenote{The full version of the author's guide is available as
-%  \texttt{acmart.pdf} document}
 
-\author{Ben Trovato}
-%\authornote{Dr.~Trovato insisted his name be first.}
-\orcid{1234-5678-9012}
+\author{Mario Garc\'ia-Valdez,\\ Jos\'e-C  Romero}
 \affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin} 
-  \state{Ohio} 
-  \postcode{43017-6221}
+  \institution{Dept. of Graduate Studies,\\ Instituto
+Tecnol\'ogico de Tijuana, Tijuana, M\'exico}
   }
-\email{trovato@corporation.com}
+\email{(mario|jcromero)tectijuana.edu.mx}
 
-\author{G.K.M. Tobin}
-%\authornote{The secretary disavows any knowledge of this author's actions.}
+\author{Juan-J.~Merelo}
 \affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin} 
-  \state{Ohio} 
-  \postcode{43017-6221}
+  \institution{Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain}
   }
-\email{webmaster@marysville-ohio.com}
+\email{jmerelo@ugr.es}
 
-\author{Lars Th{\o}rv{\"a}ld}
-%\authornote{This author is the
-%  one who did all the really hard work.}
+\author{Francisco Fern\'andez-de-Vega}
 \affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla} 
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Lawrence P. Leipuner}
-\affiliation{
-  \institution{Brookhaven Laboratories}
-  \streetaddress{P.O. Box 5000}}
-\email{lleipuner@researchlabs.org}
+  \institution{Grupo de Evoluci\'on Artificial,\\
+Universidad de Extremadura, Spain}
+  }
+\email{fcofdez@unex.es}
 
@@
 development of a collaborative evolutionary interactive system. 
-Both conceptual and implementation details are provided where the   
-technique is used to measure and increase user engagement and participation.
-Our experiments show that the model can be successfully applied in a
-gamification technique developed to increase user engagement, which
-implies that this technique can successfully be used to 
-decrease user fatigue and boredom, and thus increase the {\em performance} of the
-interactive system.
+% Both conceptual and implementation details are provided where the   
+% technique is used to measure and increase user engagement and participation.
+% Our experiments show that the model can be successfully applied in a
+% gamification technique developed to increase user engagement, which
+% implies that this technique can successfully be used to 
+% decrease user fatigue and boredom, and thus increase the {\em performance} of the
+% interactive system.
 \end{abstract}

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
First draft

@@
 
-\author{Mario Garc\'ia-Valdez,\\ Jos\'e-C  Romero}
+\author{Mario Garc\'ia-Valdez, Jos\'e-C Romero}
 \affiliation{%
@@
   }
-\email{(mario|jcromero)tectijuana.edu.mx}
+\email{mariosky@gmail.com}
+\email{josec.romero@tectijuana.edu.mx}
 
@@
 
-\begin{abstract}
-The necessary intervention of humans in interactive evolutionary
-computational systems has inherent drawbacks arising from the very nature of 
-the algorithms, namely, the human fatigue caused by the interaction, and
-the boredom arising when users evaluate a large number of artifacts.
-To tackle these issues, in this paper we
-propose a human-centered framework to model complex interactions on these systems.
-A case study is presented where the model is applied in the 
-development of a collaborative evolutionary interactive system. 
-% Both conceptual and implementation details are provided where the   
-% technique is used to measure and increase user engagement and participation.
-% Our experiments show that the model can be successfully applied in a
-% gamification technique developed to increase user engagement, which
-% implies that this technique can successfully be used to 
-% decrease user fatigue and boredom, and thus increase the {\em performance} of the
-% interactive system.
+\begin{abstract} The necessary intervention of humans in interactive
+evolutionary computational systems has inherent drawbacks arising from the
+very nature of the algorithms, namely, the human fatigue caused by the
+interaction, and the boredom arising when users evaluate a large number of
+artifacts. To tackle these issues, in this paper we propose a human-centered
+framework that can be used to  increase volunteer participation in C-IEC 
+systems by using gamification techniques. A case study is
+presented where the model is applied in the development of a collaborative
+evolutionary interactive system.
 \end{abstract}
@@
 <concept>
-<concept_id>10010147.10010178.10010205.10010206</concept_id>
-<concept_desc>Computing methodologies~Heuristic function construction</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10010147.10010178.10010205.10010207</concept_id>
-<concept_desc>Computing methodologies~Bio Inspired Algorithms</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
 </ccs2012>
@@
 \ccsdesc[500]{Human-centered computing~Collaborative and social computing systems and tools}
-\ccsdesc[300]{Computing methodologies~Heuristic function construction}
-% We no longer use \terms command
-%\terms{Theory}
-
 \keywords{Interactive evolutionary computation, Human Centered Computing}

### 2017-04-06 2017-EUROPAR/Gamification/paper.tex

@@
 CPU they where using.
+
+\section{Conclusions, discussion and future work}
+
+In this paper we have added gamification to a volunteer computing
+system via a leaderboard of contributions of registered users. In
+fact, we have proved that, although only a minority of the users opt
+for registration, their individual contribution is on average much
+higher than for non-registered/non-gamified users. The fact that less
+than one fifth of users opt for registering would discard the
+exclusive use of this feature; however, it is an interesting addition
+to the volunteer system that in fact enhances its computing power.
+
+The fact that gamification actually work highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation.
+That is why future lines of work will look for intermediate ways of including
+gamification without registration, as well as some ways of making the user increase control over the algorithm that would go beyond simply staying for longer or leaving the page.
+Since it is a sociotechnical system, including social features in it and studying emerging social networks might be an interesting line of work too.
+
+Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
 

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Spell-check and references fixed

@@
 \section{Introduction}
-% Main Ideas:
 
-% IEAs
-% Human Based Computation
-% C-IEAs
-% Problems
-% Volunteers 
-% Human Centered
-% C-IEA Complex Interactions
-  % User -> Human
-  % Social Network
-  % Devices IoT
-  % Activities
-  % Engagement
-
-% Why an HC Framework (Objectives)  
-% Presentation
-
-% IEAs
-Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
-whose fitness evaluations are performed by humans through an interactive 
-system \cite{eiben2015interactive}.
-% They are usually applied in problems where the fitness function is not known or simply
-% does not exist, and the result of optimization should fit a certain human need or
-% desire such as an aesthetic ideal.  
-% That is why their use cases include the evolution of objects with subjective characteristics,
-% such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
-% considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
-% or creativity \cite{yu2011cooks}. In the cases when 
-% human interaction is responsible of other 
-% aspects of the evolutionary process, these IEC methods are classified by some authors  
-% as human-based evolutionary computation \cite{972056} 
-% or as human-based computation \cite{quinn2011human}.
-
-% Human Based Computation
-% IEC systems are an interesting venue of research, since they have demonstrated 
-% their ability for effectively 
-% producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
-% as well as other types of artifacts in many other domains \cite{ie1}. 
-% Problems
-However, the necessary intervention of humans brings certain challenges 
-to designers of IEC methods; namely, human evaluations are scarce, slow and expensive, there is
-human fatigue caused by the interaction \cite{ie1}, and also boredom arises
-when users evaluate a large number of phenotypes, 
-many of which are not interesting or are very similar to each other.
-% % C-IEAs
-% Moreover, the performance of these systems effectively depends on the number of users
-% they are able to include; to reach more users,
-% IEC systems are some times developed as web applications depending on visitors to help
-% with the search, using both anonymous and registered users. Some systems 
-% employ a collaborative technique, where several users participate in 
-% the evaluation, this method is called Collaborative-IEC (C-IEC)
-% \cite{picbreeder,seyama2016development,wagy2014collective}.
-% Volunteers 
-Including an C-IEC in a volunteer system can lower the requirements for
-participants in the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
-system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
-volunteer's lack of accountability, and the need to build trust between participants and project
-owners. Other issues of interest for project owners are also the difficulty of predicting 
-the amount of time and resources a volunteer is willing to spend on the system, 
-and how they decide if they participate or not \cite{JJ:2016}.
-
-% Human Centered
-In order to increase volunteer participation and to tackle some of the issues mentioned above,  
-we proposed a software framework following a human centered design \cite{greenhouse2012human},
-giving extensive attention to volunteers, not only because their
-explicit evaluation is essential, but also because the context of the 
-interaction affects the system as a whole.
-
-% For example, in a C-IEC application fitness assignment depends on the
-% actions of a social network of users.  These actions are triggered when 
-% they tag, share, rate, store or delete a phenotype. 
-% Then, the selection of parents could depend on the previous actions, leveraging information, 
-% such as the fact that they both have the same tag, or were shared by
-% similar users.
-
-Data available from the interaction is also used to increase the engagement of 
-users in the system by applying  gamification techniques. Gamification
-is a technique defined by 
-Deterding et al. \cite{deterding2011game} as
-\begin{quote}
-  the use of game design elements in non-game contexts.
-\end{quote}  
+The intervention of humans brings particular challenges
+to designers of Interactive evolutionary computation (IEC) systems.
+Namely, human evaluations are scarce, slow and expensive, there is human 
+fatigue caused by the interaction \cite{ie1}, and
+also boredom arises when users evaluate a large number of phenotypes.
+The general goal of this research is to develop a  human-centered 
+\cite{gasson2003human} software framework that can be used to 
+increase volunteer participation in C-IEC systems by using gamification techniques.
+Gamification is the use of game design elements in non-game contexts \cite{deterding2011game}. 
 The gamification element employed in this work is a rewarding mechanism
-\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
-with score points, levels and leader boards. Points are awarded to users in response of 
-the accomplishment of certain activities that need to be encouraged. Levels depend
-on the score and certain features of the game are only available to gamers when 
-they reach a giving level.
-
-The development of a web based C-IEC application 
-for evolving artistic drawings is described to give the reader details about 
-the utilization of the proposed framework. Then, a rewarding mechanism is 
-implemented in the same application to study the changes in participation
-when applying a gamification technique. Finally, an experiment is conducted
-comparing three versions of the application: one with out gamification and
-two employing a different gamification technique.
-
-The rest of this paper is organized as follows.
-Section \ref{sec:related} presents related work on the topic 
-of collaborative interactive evolution. Then, Section \ref{sec:framework} 
-presents the human centered framework
-for C-IEC applications which the main proposal of this work. Implementation details of the data model
-are presented in Section \ref{sec:implementation}. 
-Next the EvoDrawings application case study is presented in
-Section \ref{sec:evodraw}. The experimental set-up is described in Section \ref{sec:experiment},
-an the results are discussed in Section \ref{sec:results}. Finally, concluding remarks are provided 
-in Section \ref{sec:conclusions}.
-
-\section{Related Work}
-\label{sec:related}
-
-An early example of C-IEC is the Galapagos Project \cite{sims1997interactivity},
-an exhibit in the Tokyo Multimedia Museum (1997--2000) 
-were visitors interacted with images presented in 
-twelve displays by selecting those they found most aesthetically interesting by standing on
-step sensors in front of them. Web based systems were introduced later, with  
-Langdon's system \cite{langdon:2004} which evolved fractal representations of virtual creatures. 
-Similarly, Secretan et al. \cite{picbreeder} and Clune and Lipson \cite{forms} 
-use web-based IEAs to evolve artistic artifacts using a generative encoding.
-
-Some C-IEC systems promote user engagement by presenting interesting information to 
-users, for instance the genetic lineage of each phenotype or the most popular or 
-best rated solution \cite{picbreeder,forms}. An example is the recent work by 
-Wagy \& Bongard \cite{wagy2014collective} where user interaction 
-is needed for evaluating fitness and developing
-new designs of robot locomotion. Collaboration is encouraged by gamifying the system 
-using the maximum distance indicator to inspire the user to try and ``beat'' previous designs. 
-In any case, using gamification techniques imply dealing with IEC
-systems as socio-technical constructs, where the social aspects are
-essential to understand its dynamics. In this sense, conclusions
-reached with other systems such as NodEO \cite{DBLP:conf/gecco/MereloCGCRV16}
-can also be applied to these systems; and applying social 
-network techniques such as graph analysis
-to their study will allow us to understand them more thoroughly.
-
-% C-IEAS
-Given the human fatigue limitation when applying IEAs, some authors 
-have tried to mitigate the problem by allowing the algorithm to 
-collaborate with the user, so that sometimes 
-users perform the evaluation, but also specific measures are included 
-into the algorithm to perform the
-evaluation of some features automatically. For instance, Reis et al. added
-some terrain measures (such as accessibility and edge length) 
-to a standard  IEC in \cite{DBLP:journals/soco/FradeVC12}. 
-This way the algorithm was capable of providing terrains that would otherwise have needed the
-users' evaluation for these specific features. Seyama and Munetomo \cite{seyama2016development}
-also propose the reduction of user fatigue by using 
-a collaborative filtering algorithm to show only the information utilized by similar users as 
-they collaborate with a large number of users for the interactive modeling of 3D glasses.
-
-Promoting user engagement and reducing user fatigue requires the management of the complex 
-interactions found in these systems. Our approach is proposed next.
-% Volunteer Based
-% HBC
-% Engagement Techniques
-% Models
-% Is that left for later? No space, I'm afraid - JJ
-\begin{figure*}[!t]
-    \centering
-        \includegraphics[width=7in]{img/framework.eps}
-    \caption{IEC Human-centered framework.}
-    \label{fig:hc_framework}
-\end{figure*}
+\cite{dubois2013understanding}. An example of the use of this technique is a recent
+work by Wagy \& Bongard \cite{wagy2014collective} where user interaction is
+needed for developing new designs of robot locomotion.
+Collaboration is encouraged by gamifying the system using the maximum distance
+indicator to inspire the user to try and ``beat'' previous designs. Also Seyama and
+Munetomo \cite{seyama2016development} propose the reduction of user
+fatigue by using a collaborative filtering algorithm to show only the
+information utilized by similar users. In Section \ref{sec:framework}
+we present our proposed framework. Next the EvoDrawings application case study 
+is presented in Section \ref{sec:evodraw} and the
+results are discussed in Section \ref{sec:results}. Finally, some concluding
+remarks are provided in Section \ref{sec:conclusions}.
 
@@
 \label{sec:framework} 
-% You have to clearly express the objectives of the paper from tine
-% introduction. If the objective is to reduce fatigue via social
-% engagement, you have to frame the results that way.
-% Ok, added - Mario 
-The general goal of this research is to develop a human-centered \cite{gasson2003human} 
-software framework that can be used to increase volunteer participation 
-in C-IEC systems. A framework is defined as a reusable architectural design together
-with an implementation \cite{campbell1991choices}, in this case 
-providing generalized components to developers and researchers of C-IEC systems. 
-The proposed framework includes components that can be refined to increase
-participation and also to minimize the amount of evaluations needed for the evolutionary 
-process in a given IEC application. Software frameworks often have 
-a vision \cite{carneiro2010introducing} guiding their
-design. Before diving into details the main design considerations of the
-framework are explained next:
-
+The main design considerations of the framework are explained next: 
 \begin{itemize}
-\item {\bf Users are human.} 
-  The framework follows the approach of human centered computing \cite{sebe2010human},
-  in which the context, environment, interfaces, preferences, accessibility, human relations,
-  cognitive limitations, culture, creativity and other human aspects are an integral part 
-  of the system. Humans are the computing resources of the system, having unique characteristics
-  as those identified by Sun \& Dance \cite{Sun2013}:
-\begin{quote}
-  (1) humans can solve computer hard problems; (2) humans are very good at exception handling,
-  (3) humans have creativity, (4) humans have cognitive load limitation, (5) humans are
-  vulnerable to psychological manipulation, (6) humans are prone to errors,
-  especially for reflective tasks.
-\end{quote}
-
 \item {\bf Users are volunteers.} Users donate their computing resources, so they are 
@@
   \cite{ahuja1993network}. 
-  A graph model could enable researchers to find other ways of identifying leaders of 
-  opinion or measuring the similarity between user's preferences. 
-  These measures can then be used by recommender algorithms selecting 
-  phenotypes according to their preferences.
-
-\item {\bf Context of use matters.}
-  Fischer \cite{fischer2012context}
-  defines context as
-  \begin{quote}
-  the interaction between humans and
-  computers in socio-technical systems that takes place in a certain
-  context referring to the physical and social situation in which
-  computational devices and environments are embedded.
-\end{quote}   
-  Fischer also identifies the important aspects to consider when the context is used: how is
-  contextual data obtained, how is context represented and what
-  goals and purposes the context has when it is used in a particular
-  application. An IEC system will  be used within a certain range 
-  of technical, physical and social or
-  organizational environments \cite{maguire2001context} that may affect its use.
-
 \item {\bf Interaction is a stream of actions.}
@@
   captured by sensors, or directly captured as user input. For example, social networks encourage
-  users to publish their interactions with other users, media objects and places. Users of 
-  social networks (for instance the Facebook Graph) are accustomed to express these 
-  complex relationships in sentences such as: ``John and Ann eating breakfast at Tony's''. 
-  As this is gaining accptence, there are iniciatives like the W3C Activity Streams 2.0 
-  specification, used for representing common activities in social 
-  web applications \cite{json:streams}. 
+  users to publish their interactions with other users, media objects and places.
+  there are initiatives like the W3C Activity Streams 2.0  specification, used for 
+  representing common activities in social web applications \cite{json:streams}. 
 \end{itemize}
 
-The above considerations have guided the design of the framework, and throughout they have been
-treated as application constraints. In order to satisfy theses requirements the
-human-centered C-IEC architecture consists of three high level components depicted
-in figure \ref{fig:hc_framework}:
+A graph is proposed for modeling the social network of users and their interactions 
+with candidate solutions, and the relationships between them in the population.
+The graph database system used in the implementation is Neo4J, which is a scalable solution 
+\cite{miller2013graph,holzschuher2013performance}. 
+This graph is also used to increase engagement through gamification, as explained in the next section.
 
-\begin{enumerate}
-  \item {\bf Interactive System.} 
-  This is the real world system that we are going to represent in the data model, 
-  it consist of human users and their interactions with one or more phenotypes
-  from the population. There are many ways in which humans could interact 
-  with these phenotypes. The interaction consists of a set of actions and 
-  takes place in a certain context; for example, through a mobile device  
-  or by interacting with real world objects 
-  \cite{de2014artists,de2013unplugging}. 
-  There is also the possibility that fitness or even part of the search 
-  is done by devices lent by humans \cite{DBLP:conf/gecco/MereloCGCRV16}.
-  The information gathered through the interaction is the primary focus
-  as it will guide the search.
-
-  \item {\bf Data Model.}
-  A data model is used to describe the IEC system prior to a physical 
-  implementation.  Depending on the domain several models can be used, these
-  options and their physical implememtations are explained in Section \ref{sec:implementation}.
-
-  \item {\bf Human Interface.}
-  The appropriate human interface will depend on the application domain, 
-  the current framework implementation employs a web based application. Developers 
-  can use different templates, to present one or more phenotypes 
-  at a time, and the type of rating system:  based on ``likes'' or in a rating from one to five (stars). 
-  Other implementations do not need a graphical interface at all, using sensors and actuators.
-
-  \item {\bf Evolutionary Algorithm.}
-  The Evolutionary Algorithm (EA) algorithm interacts with the data model.     
-  The EA is decoupled from the data model, the algorithm could be implemented internally, 
-  use an external library, or even be human based evolution, 
-  where volunteers select the parents of the next generation and upload the 
-  new individuals as in the XYZ project \cite{de2014artists}.
-
-  \item {\bf Analytics.} 
-  Researchers and deveopers will need to query the data model in order to analyzing the 
-  system's behaviour.
-
-\end{enumerate}
-
-\section{Data Model Implementation}
-\label{sec:implementation}
-
-In this section, the core component of the framework is presented.
-Three database models are employed to store different elements of the system, next 
-these are described including the technology selected for their physical implementation:
-
-\begin{itemize}
-\item {\bf EvoSpace-Redis.} 
-EvoSpace is a population store \cite{Evospace}  for the development of 
-evolutionary algorithms that are intended to run on a cloud computing model. 
-The population is decoupled from any particular evolutionary algorithm. 
-Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
-processed and replaced using a specified set of methods \cite{GValdez2015}. The population
-is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
-database management system because it provides a hash based implementation of sets and 
-queues which are natural data structures for the EvoSpace model. Basically, a sample of 
-candidate solutions are retrieved from the server, evaluated and then sent back. 
-The same operations are used to evolve the population. In EvoSpace individuals replaced 
-in the population are stored indefinitely, to permit users to store permalinks to them.
-Implementation details are presented in \cite{garcia2013evospace}.
-
-\item {\bf Graph-Neo4J.} 
- Collaborative IEC (C-IEC) systems need to store highly connected data, as it is common 
- in current applications like social networks. In order to deal with large datasets of connected 
- data found in these systems, graph databases \cite{angles2012comparison} have been proposed 
- as an alternative to relational databases which have performance limitations when dealing with 
- highly connected data \cite{holzschuher2013performance}.
- A graph is proposed for modeling the social network of users, their interaction with 
- candidate solutions, and the relationships between them in the population.
- The graph database system used in the implementation is Neo4J, which is
- a scalable solution \cite{miller2013graph,holzschuher2013performance}, well 
- supported and documented in PaaS infrastructures like Heroku. The Cypher query 
- language it used to retrieve views from the graph.
- An example query is shown in Figure \ref{fig:cypher} where the relations 
- between users and solutions are presented.
-
-  \begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/gui-neo.png}
-    \caption{Example query in Cypher.}
-    \label{fig:cypher}
-  \end{figure}
-
-\item {\bf Relational Objects-PostgreSQL.}
-  The PostgreSQL relational database system is also employed because user 
-  sessions and authentication,
-  as well as dynamic web pages are handled directly by the Django web
-  framework \cite{garcia2013evospace}. % these implementations details
-                                % could go if needed.
-
-\end{itemize}
-
-\section{Case Study: EvoDrawings}
+\section{Case Study: EvoDrawings Gamification}
 \label{sec:evodraw}
 As a case study, a C-IEC application was developed by extending the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
-A brief description of the application is presented next.
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.5in]{img/interface.png}
-    \caption{User interface of the EvoDrawings application. (ANONIMIZED IN THIS VERSION)}
-    \label{fig:web}
-\end{figure}
-
-\subsection{Collaboration}
-\label{sec:col}
-Users need to authenticate themselves to the system using their
-Facebook account. In that sense, some users might not be interested
-either because they do not want to give that information or simply because
-they do not use that social network. Even if we might lose
-some users that way, the additional information we obtain for scoring
-phenotypes more than balances that. 
-After logging in,
-users can collaborate with their Facebook friends, 
-sharing the phenotypes they like, or by taking phenotypes
-from their friend's collections by using the web interface depicted 
-in Figure \ref{fig:web}.
-At the top left corner a list of Facebook friends is presented
-to encourage users to interact with the system. In the central 
-\emph{ Wall } area, a phenotype sampled from the population that is
-being evolved via the evolutionary algorithm 
-is shown to the user.
-Here, the user can interact with the system in two ways.
-First, he can assign a rating to the phenotype or choose to add an image to one of their \emph{Collections}.
-A collection is a special folder that stores those phenotypes a user likes and wishes
-to save. After the user finishes interacting with the phenotype
-on the Wall, he can choose to retrieve a new one from the population.
-At the left hand side of Figure \ref{fig:web}, the web page shows the \emph{Collections} section.
-The user can create several collections, to group and organize his favorite 
-artifacts. Moreover, users can browse the content of each collection and from
-there share images through their social network. This makes the assignment 
-of fitness through the rating system a {\em social} activity, 
-pursuing the objective of this work, which is to increase user engagement.
-
-\subsection{Graph Model} 
-The graph model for the EvoDrawings application has the following types of
-nodes: {\bf User}, {\bf Phenotype} and {\bf Collection}. The Collection node represents
-a collection of drawings belonging to a user. One collection can contain many phenotypes 
-or be empty. A single phenotype could be shared by many collections. The interaction 
-between these entities are represented by the following edges:
-
-\begin{itemize}
-\item {\bf Likes} This relation describes the interaction between a user and
-a phenotype in which a rating value is assigned.
-
-\item {\bf Knows} The relation connects two users that know each
-other in the Facebook social graph.
-
-\item {\bf Parent} Describes what phenotype is the parent of a new
-  phenotype.
-
-\item {\bf Has} The relation describes an ownership relation between users and
-those collections they own. 
-\end{itemize}
-% Once again, these implementation details could go - JJ
-
-\subsection{Gamification}
-The rewarding mechanism as it is applied in EvoDrawings gives more importance 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. 
+The open source code is hosted in GitHub at
+\url{https://github.com/mariosky/evo-drawings}. The rewarding 
+mechanism as it is applied in EvoDrawings gives more importance 
 to the preference of those users with higher reputation
@@
 the wall to a collection, save a phenotypes from a friend's collection, and
-explore collections of other friends.
+explore collections of other friends. Each of these actions are stored in
+the Neo4J graph. Two variables are used to determine the weight of a user's 
+preference:
 
-Two variables are used to determine the weight of a user's preference:
 \begin{itemize}
@@
 between 0 and 100. A new user starts at zero, and the experience increases until
-it reaches 100 actions. It is assumed for this case, that once a user
-reaches this value, it has enough experience on using the
-application. % What happens before that? - JJ  
-% explained - Mario
+it reaches 100 actions.
+
 \item {\bf Participation}: This variable is simply the degree of the user node 
@@
 
-\section{Experimental Setup}
-\label{sec:experiment}
-
-Three versions of EvoDrawings were compared:
-\begin{itemize}
-\item Base (B): All users have the same weight.
-\item Non Graph Gamification (G): Only experience is considered.
-\item Graph Gamification (GG): Both experience and participation are considered.
-\end{itemize}
-When gamification was employed, all score values known where presented to users
-and a ranking of users by weight was shown in a window. Table \ref{tab:params} shows 
-the parameters used for the evolutionary algorithm.
-
-\begin{table}
-  \small
-  \caption{ Parameters for experiments.  }
-  \label{tab:params} 
-  \centering
-  \small
-  \begin{tabular}{l  c}
-    \hline\noalign{\smallskip}
-     Parameter & Value \\
-    \noalign{\smallskip}\hline\noalign{\smallskip}
-    Initial Population Size   & 80 \\ \hline
-    Sample Size & 1 \\ \hline
-    Step Size & 8 Samples \\ \hline
-    Mutation &  \\ \hline
-    Selection & Tournament \\ \hline
-    Tournament Size &  6 \\ \hline
-  \end{tabular}
-\end{table}
-
-% Anuncements
-At the start of every experiment, a call to participation was issued
-through social networks.  The link used in the call for participation
-was a shortened Google URL, that provides meta-data and analytics for the
-users that click on it. 
- %Will we comment on that? - JJ
- % To do.
-  In Table \ref{tab:urls} the URL for each 
-deployment in the Heroku platform is shown,
-along with the short URL and the analytics link. In the same table a link to the GitHub 
-application repository used to deploy to Heroku is also listed (ANONIMIZED IN THIS VERSION). Only data 
-for the first week of deployment was considered for the experiments, and they where conducted 
-between January and May of 2016.
+Three versions of EvoDrawings were compared: Base (B): All users have the same weight,
+Non Graph Gamification (G): Only experience is considered and Graph Gamification (GG): 
+Both experience and participation are considered.
 
@@
 \label{sec:results} 
-Before release, each deployment was first tried with a few beta testers. 
+Before release, the deployment was first tried with a few beta testers. 
 When applying the leader board gamification technique for the first time a 
@@
     \hline\noalign{\smallskip}
-     Deployment &  Users &  Nodes &  Edges & URL \\
+     Deployment &  Users &  Nodes &  Edges \\
     \noalign{\smallskip}\hline\noalign{\smallskip}
-    B   & 53 &  595   & 2220  & goo.gl/jLis4Q.info \\ \hline
-    G   & 54 &  648   & 2596  & goo.gl/jqjNy5.info \\ \hline
-    GG  & 68 &  932   & 3594  & goo.gl/J8TCe1.info \\ \hline
+    B   & 53 &  595   & 2220   \\ \hline
+    G   & 54 &  648   & 2596   \\ \hline
+    GG  & 68 &  932   & 3594   \\ \hline
     \end{tabular}
@@
 Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
-in the graph after each experiment. Moreover, the total number of evaluated 
-phenotypes for each volunteer is presented in figure 
-\ref{fig:top-ranked-participation} where users are ranked by the 
-number of phenotypes they rated. In the \emph{x} axis is the rank and in the \emph{y} axis 
-the number of phenotypes evaluated using a logarithmic scale. Results show that when 
-considering deployments B and G, the difference came with users with a medium level of participation.
-When comparing all the experiments the deployment GG had the higher
+in the graph after each experiment. 
+% Moreover, the total number of evaluated 
+% phenotypes for each volunteer is presented in figure 
+% \ref{fig:top-ranked-participation} where users are ranked by the 
+% number of phenotypes they rated. 
+% This figure is missing. Maybe we should publish it in ArXiV or something
+When comparing all the experiments the GG deployment had the higher
 number of participation, besides attracting also the higher number of users.
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.3in]{img/comparison.png}
-    \caption{Users ranked by the number of phenotypes they 
-    rated vs. the number of phenotypes evaluated using a logarithmic scale. }
-    \label{fig:top-ranked-participation}
-\end{figure}
-
-Moreover, modeling the performance of an C-IEC system
-involves understanding its dynamics. Previous works on browser-based
-volunteer computing have used basic metrics such as
-the number of users or the time spent by every one in the
-computation \cite{DBLP:journals/gpem/LaredoBGVAGF14, 2016arXiv160101607M}. 
-While works on other platforms such as SETI@home
-\cite{javadi2009mining} have found that the Weibull, log-normal, and
-Gamma distributions where viable models of the  
-the availability of computing resources in several clusters, which  is 
-in concordance with the results obtained in \cite{agajaj}, and also
-browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
-The shape of those distributions is a skewed bell
-with more resources in the {\em low} areas than in the high areas; i.e., 
-there are many users that give a small amount of cycles, while there
-are just a few that give many cycles. In order to assess gamification techniques alongside C-IEC follows the
-same pattern, participation was fitted to a 
-Weibull distribution,
-and shown in Figures \ref{fig:w1} to \ref{fig:w3}, 
-confirming this a model for user interaction, although with different fitted
-values in each case. A graph query was used to compare the number of volunteers in deployments
-B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) showing 
-additional patterns on both social networks.
-
-%
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.4in]{img/weibull_1.png}
-    \caption{Deployment B. Weibull Fit.}
-    \label{fig:w1}
-\end{figure}
-%
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.4in]{img/weibull_2.png}
-    \caption{Deployment G. Weibull Fit.}
-    \label{fig:w2}
-\end{figure}
-
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.4in]{img/weibull_3.png}
-    \caption{Deployment GG. Weibull Fit.}
-    \label{fig:w3}
-\end{figure}
-
-% This has nothing to do with the rest of the paper. You are not
-% modeling anything. It's just another application of the
-% framework. Could be mentioned in the state of the art, that's
-% all. We should devote more space to the study of user interactions.
-
-% Once again, no model, no relation to the rest of the paper. Either
-% you talk about a model of interactions with humans and their
-% similarity with the first case study, or you eliminate it. How do
-% you tackle boredom? How do you increase engagement? Those are the
-% paper's topics, you can't just drop it here because it's interactive
-% - JJ
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.2in]{img/user_known_2.png}
-    \caption{User network view of the Base deployment (B). (ANONIMIZED IN THIS VERSION) }
-    \label{fig:B-network}
-\end{figure}
-\begin{figure}[!t]
-    \centering
-        \includegraphics[width=3.2in]{img/user_known_3.png}
-    \caption{User network view of the Graph Gamification deployment(GG) (ANONIMIZED IN THIS VERSION) .}
-    \label{fig:GG-network}
-  \end{figure}
 
@@
 
-Volunteer-based and C-IEC systems involve the dynamic 
-interaction of many entities and artifacts. Employing a human-centered approach 
-will allow researchers to understand and visualize this kind of
-systems better. % But you would have to provide some proof of this - JJ
-
-In this paper a human-centered software framework was proposed, and was validated
-through the implementation and refinement of a C-IEC application.
-This framework enabled the implementation of a gamification technique 
-to improve engagement in a case study which provides a common arena where users are aware of the
-activities of other users in the social neighborhood.
-
 In concordance with the results obtained in other 
 browser-based volunteer systems, after 
-applying the gamification techniques, user participation follows the
-same pattern and it was fitted to a Weibull distribution.
+applying the gamification techniques, user participation was increased.
 
@@
 for instance \cite{Merelo2016:repomining}.
-The dynamics of this kind of system are fundamentally different, and our future research will
-include exploring these aspects of the system.
 
-Another line of work would be to study the possible negative effects of using  
+One of the interesting future lines of work would be to study the possible negative effects of using  
 gamification techniques to improve engagement, like cheating or
-literally {\em gaming} the system to defeat competition. We already
-found some hints of this behavior at the beginning of the release of
-this system, but more subtle effect could be taking place. 
+literally {\em gaming} the system to defeat competition. 
 Finally, the refinement of the proposed Human-Centered framework will need
@@
 \begin{acks}
-  Reserved\\
-  Space for acks.
+This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P (UGR-EPHEMECH) and by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590.
 

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex

--- 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
   users to publish their interactions with other users, media objects and places.
-  there are initiatives like the W3C Activity Streams 2.0  specification, used for 
+  there are iniciatives like the W3C Activity Streams 2.0  specification, used for 
   representing common activities in social web applications \cite{json:streams}. 
@@
 \cite{miller2013graph,holzschuher2013performance}. 
-This graph is also used to increase engagement through gamification, as explained in the next section.
+This graph is also used to increase engagement through gamification, as expleained in the next section.
 
@@
 Three versions of EvoDrawings were compared: Base (B): All users have the same weight,
-Non Graph Gamification (G): Only experience is considered and Graph Gamification (GG): 
+Non Graph Gamification (G): Only experience is consideredm and Graph Gamification (GG): 
 Both experience and participation are considered.
@@
 Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
-in the graph after each experiment. 
-% Moreover, the total number of evaluated 
-% phenotypes for each volunteer is presented in figure 
-% \ref{fig:top-ranked-participation} where users are ranked by the 
-% number of phenotypes they rated. 
-% This figure is missing. Maybe we should publish it in ArXiV or something
-When comparing all the experiments the GG deployment had the higher
+in the graph after each experiment. Moreover, the total number of evaluated 
+phenotypes for each volunteer is presented in figure 
+\ref{fig:top-ranked-participation} where users are ranked by the 
+number of phenotypes they rated. When comparing all the experiments the deployment GG had the higher
 number of participation, besides attracting also the higher number of users.

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Small changes in abs

@@
 very nature of the algorithms, namely, the human fatigue caused by the
-interaction, and the boredom arising when users evaluate a large number of
+interaction and the boredom arising when users evaluate a large number of
 artifacts. To tackle these issues, in this paper we propose a human-centered
-framework that can be used to  increase volunteer participation in C-IEC 
+framework that can be used to increase volunteer participation in collaborative interactive evolutionary computational (C-IEC)
 systems by using gamification techniques. A case study is

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-front.tex
Two page files
Small changes in abs

--- 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
@@
 
-\begin{abstract} The necessary intervention of humans in interactive
-evolutionary computational systems has inherent drawbacks arising from the
-very nature of the algorithms, namely, the human fatigue caused by the
-interaction, and the boredom arising when users evaluate a large number of
-artifacts. To tackle these issues, in this paper we propose a human-centered
-framework that can be used to  increase volunteer participation in C-IEC 
-systems by using gamification techniques. A case study is
-presented where the model is applied in the development of a collaborative
-evolutionary interactive system.
+\begin{abstract} 
+The necessary intervention of humans in interactive evolutionary
+computational systems has inherent drawbacks arising from the very
+nature of the algorithms, namely, the human fatigue caused by the
+interaction and the boredom arising when users evaluate a large number
+of artifacts. To tackle these issues, in this paper we propose a
+human-centered framework that can be used to increase volunteer
+participation in Collaborative Interactive Evolutionary Computation
+(C-IEC) systems by using gamification techniques. A case study is
+presented where the model is applied in the development of a
+collaborative evolutionary interactive system.
 \end{abstract}
@@
 
-\input{paper-body}
+\input{paper-short-body}
 

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
Spell check and elimination of non-referenced figures here too
Modifications to intro

--- 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
@@
 
-The intervention of humans brings particular challenges
-to designers of Interactive evolutionary computation (IEC) systems.
-Namely, human evaluations are scarce, slow and expensive, there is human 
-fatigue caused by the interaction \cite{ie1}, and
-also boredom arises when users evaluate a large number of phenotypes.
-The general goal of this research is to develop a  human-centered 
+Interactive evolutionary computation (IEC) systems include persons in
+a computational system performing different kind of tasks, which brings particular challenges
+to their designers: human {\em computational cycles} are scarce, slow
+and expensive to acquire, there is human 
+fatigue caused by the interaction \cite{ie1}; besides, boredom arises
+when users evaluate a large number of phenotypes. In this paper we
+intend to solve these issues by developing a  human-centered 
 \cite{gasson2003human} software framework that can be used to 
-increase volunteer participation in C-IEC systems by using gamification techniques.
+increase volunteer participation in C-IEC systems by using
+gamification techniques, which are used to {\em human computational
+  cycles} more abundant and possibly decrease boredom.
+
 Gamification is the use of game design elements in non-game contexts \cite{deterding2011game}. 
@@
 % number of phenotypes they rated. 
-% This figure is missing. Maybe we should publish it in ArXiV or something
+% Same as in the other, figure missing - JJ
 When comparing all the experiments the GG deployment had the higher

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
Revises conclusions

@@
 browser-based volunteer systems, after 
-applying the gamification techniques, user participation was increased.
+applying the gamification techniques, user participation was
+increased. By the nature of the experiment, we will have to repeat it
+several times and check causal correlation, but in general all
+measurements indicate that is the case.
 
@@
 to reach a critical state, as has been found in software repositories,
-for instance \cite{Merelo2016:repomining}.
-
-One of the interesting future lines of work would be to study the possible negative effects of using  
+for instance \cite{Merelo2016:repomining}. It would be also
+interesting to study the possible negative effects of using   
 gamification techniques to improve engagement, like cheating or

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
Paco's  suggestions

@@
 as given by their score points and experience levels.  
-Each time a user does on of these actions their score is incremented by one:
+Each time a user does one of these actions their score is incremented by one:
 start a session, rate a phenotype, create a collection, save a phenotype of 
-the wall to a collection, save a phenotypes from a friend's collection, and
+the wall to a collection, save a phenotype from a friend's collection, and
 explore collections of other friends. Each of these actions are stored in
@@
 
-Three versions of EvoDrawings were compared: Base (B): All users have the same weight,
-Non Graph Gamification (G): Only experience is considered and Graph Gamification (GG): 
+Three versions of EvoDrawings were compared: Base (B): All users have the same weight;
+Non Graph Gamification (G): Only experience is considered; and Graph Gamification (GG): 
 Both experience and participation are considered.
@@
 \begin{acks}
-This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P (UGR-EPHEMECH) and by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590.
+This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-{1,2,3}-P (UGR-EPHEMECH), by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590, and by Junta de Extremadura FEDER, project GR15068
 

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
Project codes

@@
 \begin{acks}
-This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-{1,2,3}-P (UGR-EPHEMECH), by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590, and by Junta de Extremadura FEDER, project GR15068
+This work has been supported in part by: de Ministerio espa\~{n}ol de Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-\{1,2,3\}-P (UGR-EPHEMECH), by Consejo Nacional de Ciencia y Tecnolog\'{\i}a under project 220590, and by Junta de Extremadura FEDER, project GR15068
 

### 2017-04-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
Github URL as footnote

@@
 As a case study, a C-IEC application was developed by extending the 
-EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}. 
-The open source code is hosted in GitHub at
-\url{https://github.com/mariosky/evo-drawings}. The rewarding 
+EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}, 
+with the open source code hosted in GitHub\footnote{\url{https://github.com/mariosky/evo-drawings}}.
+The rewarding 
 mechanism as it is applied in EvoDrawings gives more importance 

### 2017-11-03 2017-EUROPAR/Gamification/paper.tex
Revising abstract for EvoPar

@@
 
-One way of creating distributed computing system is to use volunteers who
+Distributed computing systems can be created using volunteers who
 provide their own computing resources or storage to contribute to a common effort.
-By running a script in a web page, collaboration is straightforward, but also ephemeral.
-Resources depend on the amount of time a user lends, which means that 
-the user has to be kept engaged to obtain as many computing cycles as
+They can, for instance,  run a script embedded in a web page, collaboration is straightforward, but also ephemeral, with resources depending on the amount of time a user lends; this  implies that
+the user has to be kept engaged so as to obtain as many computing cycles as
 possible. In this paper, we analyze a volunteer-based evolutionary computing system called

### 2017-11-03 2017-EUROPAR/Gamification/paper.tex
Enlarging and hiding some stuff for double blind
Anonymous Authors

@@
 
-\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}
+\author{anonymous author name \inst{1} \and anonymous author name\inst{2} \and  anonymous author name \inst{1}}
 
-\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
+\institute{anonymous institution, some city, country
 \and
-Universidad de Granada, Granada, Spain
-\email{mario@tectijuana.edu.mx}\\
-\email{jmerelo@geneura.ugr.es}}
+anonymous institution, some city, country
+\email{example@example.edu}\\
+\email{example2@example.edu}}
 
-\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }
+%\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}
+
+%\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
+%\and
+%Universidad de Granada, Granada, Spain
+%\email{mario@tectijuana.edu.mx}\\
+%\email{jmerelo@geneura.ugr.es}}
+
+%\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }
 
@@
     \centering
-        \includegraphics[width=2.3in]{img/puts_user.png}
-        \includegraphics[width=2.3in]{img/puts_ip.png}
+        \includegraphics[width=0.9\linewidth]{img/puts_user.png}
+        \includegraphics[width=0.9\linewidth]{img/puts_ip.png}
     \caption{Usage by registered (left) and anonymous (right) users.  Users are ranked by
@@
 
-This work has been supported in part by: Ministerio espa\~{n}ol de
-Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
-(UGR-EPHEMECH).
+% This work has been supported in part by: Ministerio espa\~{n}ol de
+% Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
+% (UGR-EPHEMECH).
+Hidden for double blind.
 

### 2017-11-06 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-short-body.tex
old uncommitted changes

@@
 
-Interactive evolutionary computation (IEC) systems include persons in
+Interactive Evolutionary Computation (IEC) systems include persons in
 a computational system performing different kind of tasks, which brings particular challenges
@@
 \cite{dubois2013understanding}. An example of the use of this technique is a recent
-work by Wagy \& Bongard \cite{wagy2014collective} where user interaction is
+work by Wagy \& Bongard \cite{wagy2014collective} in which interaction is
 needed for developing new designs of robot locomotion.
 Collaboration is encouraged by gamifying the system using the maximum distance
-indicator to inspire the user to try and ``beat'' previous designs. Also Seyama and
-Munetomo \cite{seyama2016development} propose the reduction of user
+indicator to inspire the user to try and ``beat'' previous designs. Also 
+Seyama \& Munetomo \cite{seyama2016development} propose the reduction of user
 fatigue by using a collaborative filtering algorithm to show only the
 information utilized by similar users. In Section \ref{sec:framework}
-we present our proposed framework. Next the EvoDrawings application case study 
+we present our proposed framework. Next, the EvoDrawings application case study 
 is presented in Section \ref{sec:evodraw} and the
@@
 A graph is proposed for modeling the social network of users and their interactions 
-with candidate solutions, and the relationships between them in the population.
+with candidate solutions and the relationships between them in the population.
 The graph database system used in the implementation is Neo4J, which is a scalable solution 
@@
 Before release, the deployment was first tried with a few beta testers. 
-When applying the leader board gamification technique for the first time a 
+When applying the leader board gamification technique for the first time, a 
 problem was found: some users were cheating by giving a
 rating to an animation even before it was returned from the server, this was done by just
-constantly clicking the mouse button. This is a common problem found in systems using leader
+constantly clicking the mouse button. Cheating is a common problem found in systems using leader
 boards because by making the scores visible to other players they are encouraged 
-to compete \cite{hickman2010total}. The version used in experiments disabled the button until 
+to compete against each other \cite{hickman2010total}. The version used in experiments disabled the button until 
 the drawing animation was over. The results of each of the three experiments in 

### 2017-11-06 2017-EUROPAR/Gamification/paper.tex
Anonymous
Some typos
Some changes and corrections.
Also the atom editor deleted all trailing spaces.

@@
 \and
-anonymous institution, some city, country
+anonymous institution, some city, country\\
 \email{example@example.edu}\\
@@
 The experiment results show that even if only a small percentage of users created an account,
-those participating in the competition provided around 90\%.
+those participating in the competition provided around 90\% of the work.
 
@@
 
-The World Wide Web is an increasingly reliable and high-performance
+The world wide web is an increasingly reliable and high-performance
 operating system for running distributed applications. Besides the
@@
 factors: the JavaScript virtual machine embedded in every browser
-and the REST interface, a {\em de facto} standard interface for
+and the REST interface, providing a {\em de facto} standard interface for
 communication. A distributed system based on the browser can be
@@
 
-We are mainly interested in evolutionary algorithms \cite{jj-ppsn98},
-which are population-based search and optimization methods inspired by
-the theory of Evolution and its molecular basis. The fact that they
-are population-based makes easy its distribution in different clients
-and since their asynchronous nature has no impact on their performance
-and might even enhance it, makes them ideal candidates for volunteer
-computing setups such as the one presented in this paper.
+We are mainly interested in evolutionary algorithms \cite{jj-ppsn98},, population-based search and optimization methods inspired by
+the theory of Evolution and its molecular basis.
+The fact that they are population-based makes them suitable for a straightforward distribution of work to different clients, and since they have an asynchronous nature, there is no impact on their performance.
+All these properties make them ideal candidates for volunteer computing setups such as the one presented in this paper.
 
-In this line of research that uses volunteer computing for
-evolutionary algorithms, there are several pending issues, of which
-the most important is approaching volunteer computing as a 
-socio-technical system \cite{vespignani2009predicting} which integrates
-user decisions and behavioral patterns in the system model; this
+There are several open issues in the research regarding the application of volunteer computing techniques to the development of  distributed evolutionary algorithms of which the most important is approaching volunteer computing as a socio-technical system \cite{vespignani2009predicting} which integrates user decisions and behavioral patterns in the system model; this
 includes trying to optimize the number of users in a particular
@@
 way of achieving that is trying to find as many users as possible. But
-another approach, and the one that we will be taking in this paper, is
+another approach, and the one that we will be following in this paper, is
 to delve into the socio-technical nature of the system, using
-gamification to improve the amount of time every user will lend to the
-system. Gamification has to include some way of identifying the user,
+gamification techniques to improve the amount of time every user will lend to the system. Gamification has to include some way of user authentification,
 which is a hurdle that might take users away; however, the open design
 of the system we use means that we can actually mix anonymous-non
-gamified and identified-gamified users, with only the small overhead
-of adding the mechanism for gamification.
+gamified and authentified-gamified users, with only the small overhead
+of adding the gamification mechanism.
 
@@
 Volunteer computing involves users deciding to run a program that acts
-as a client or as a peer in an experiment
+as a client or as a peer in a distributed computing experiment
 and, as such, has been deployed in many different ways from the
@@
 authors, including more recent efforts
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}. In fact, this last paper \cite{pan2015gray} performs
-an analysis of what it calls {\em Gray computing} doing an analysis of
-feasibility, cost-effectiveness, change in the user experience and
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}.
+In fact, this last paper \cite{pan2015gray} performs an analysis of what is called {\em Gray computing} covering aspects of
+feasibility, cost-effectiveness, change in the user experience and the
 architectural optimization needed, concluding that the computing power
-available can be vast and it that it can be cost-effective to use it.
+available is vast and it can be cost-effective to use it.
 
-Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness 
-evaluations across a heterogeneous pool of cloud compute nodes and volunteer
-peer computers running a browser. In SPACE peers establish a bi-directional 
-communication with the server using the \texttt{Socket.io} JavaScript library.  
-As an experiment they provide an \texttt{ASM.js} compiled version of the RoboGen 
-software platform, showing that JavaScript can be used in a broad 
+Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness evaluations across a heterogeneous pool of cloud compute nodes and volunteer peer computers running a browser.
+In SPACE, peers establish a bi-directional communication with the server using the \texttt{Socket.io} JavaScript library.
+As an experiment they provide an \texttt{ASM.js} compiled version of the RoboGen software platform, showing that JavaScript can be used in a broad
 spectrum of applications.
@@
 Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose
-training simulated robots for the grounding of language symbols. They do this by using human 
-observers that issue arbitrary commands to these robots via the web, 
-providing positive or negative reinforcement in response to the
-robot\'s resulting actions. For interaction, they use the Twitch 
-video streaming platform where users can issue commands through the
+training simulated robots for the grounding of language symbols.
+They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement in response to the
+robot\'s resulting actions. For interaction, they use the Twitch streaming platform where users can issue commands through the
 platform's integrated chat service. Subjects were incentivized to
 interact with the system by GUI features that provided participants
-with a sense of involvement with the simulation. This kind of systems
+with a sense of involvement with the simulation. This kind of system
 could add additional dynamics, because users can collaborate or
@@
 
-The proofs of concept described above do not go further than trying to
-find out  how many users join the effort and how many the system is
-able to support. In fact, systems such as the one described in
+The proof-of-concept efforts described above, do not go any further than to trying to find out how many users join the effort, and how many of them the system is able to support.
+In fact, systems such as the one described in
 \cite{gecco07:workshop:dcor} had serious scaling issues; some of them
@@
 or, alternatively, how many users would be needed to be competitive
-with single-user single-computer implementations of the same
+against single-user single-computer implementations of the same
 algorithm.
-Lately, researchers have tried to integrate volunteer as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
-In fact, exclusive volunteer systems exhibit a certain amount of unpredictability and might be best used combined with available computing power.
+Lately, researchers have tried to integrate volunteer computing techinques as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
+In fact, exclusive volunteer-based systems exhibit a certain amount of unpredictability and might be best used in combination with ready-available computing power.
 
-As indicated in the introduction, the human is an integral part of the system, which can be considered {\em human computation} \cite{quinn2011human}.
-Giving the user not only some more control, by including registration, but also a measure to compare with other users might have positive effects.
+As indicated in the introduction, the human is an integral part of the system, which can be considered as {\em human computation} \cite{quinn2011human}.
+Giving the user not only more control, by including registration, but also a measure to compare with other users might have positive effects.
 In fact, the BOINC projects applies a gamification technique
@@
         \includegraphics[width=5in]{img/login.png}
-    \caption{ User interface: Landing page, and Login and registration dialogs.  
+    \caption{ User interface: Landing page, and Log-in and registration dialogs.
     }

### 2017-11-07 2017-EUROPAR/Gamification/paper.tex
Hidden last identifying url
Closes #5

Also adds some figures that had been edited out since we have enough space.

@@
 in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of
-levels at the beginning.
-% \begin{figure*}[htb]
-%     \centering
-%         \includegraphics[width=2.5in]{img/s.png}
-%     \caption{$S$ function used to assign the user's level.
-%     }
-%     \label{fig:s}
-% \end{figure*}
+levels at the beginning; this function is represented in Figure \ref{fig:s}.
+%
+\begin{figure*}[htb]
+    \centering
+        \includegraphics[width=2.5in]{img/s.png}
+    \caption{$S$ function used to assign the user's level.
+    }
+    \label{fig:s}
+\end{figure*}
 
@@
 In the front page a link to the open source code of the experiment was also
-available \url{https://github.com/lucero21/login-master}
+available \url{https://undisclos.ed}
+% \url{https://github.com/lucero21/login-master}
 
@@
 that runs a local island of an evolutionary algorithm to solve a
-multi-modal problem called {\em l-trap}, which has been used extensively
+multi-modal problem called {\em l-trap}, represented in figure \ref{fig:trap}, which has been used extensively
 as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}.
@@
  \]
-% \begin{figure*}[htb]
-%     \centering
-%         \includegraphics[width=2.5in]{img/trap.png}
-%     \caption{Trap function.
-%     }
-%     \label{fig:trap}
-% \end{figure*}
  %
+\begin{figure*}[htb]
+    \centering
+        \includegraphics[width=2.5in]{img/trap.png}
+    \caption{Trap function.
+    }
+    \label{fig:trap}
+\end{figure*}
 

### 2017-11-07 2017-EUROPAR/Gamification/paper.tex
A new paragraph appended to Intro
 @JJ  please have a look. I´m  trying to state the question addressed
in the paper.

@@
 the theory of Evolution and its molecular basis.
-The fact that they are population-based makes them suitable for a straightforward distribution of work to different clients, and since they have an asynchronous nature, there is no impact on their performance.
+The fact that they are population-based makes them suitable for a straightforward distribution of work to different clients, and since they have an asynchronous nature, there is not a considerable impact on their performance.
 All these properties make them ideal candidates for volunteer computing setups such as the one presented in this paper.
 
-There are several open issues in the research regarding the application of volunteer computing techniques to the development of  distributed evolutionary algorithms of which the most important is approaching volunteer computing as a socio-technical system \cite{vespignani2009predicting} which integrates user decisions and behavioral patterns in the system model; this
+There are several open issues in the research regarding the application of volunteer computing techniques to the development of distributed evolutionary algorithms of which the most important is approaching the problem as a socio-technical system \cite{vespignani2009predicting}.
+which integrates user decisions and behavioral patterns in the model; this
 includes trying to optimize the number of users in a particular
@@
 of adding the gamification mechanism.
+
+There is a trade-off between ease of use, where users just need to visit a web page, versus personlaization, where users need to register and subsecuently authenticate in order to participate.
+Registration could be seen as a barrier to volunteers trying to participate in an ephemeral experiment, so the question is what is the impact on user participation.
 
@@
 
-The proof-of-concept efforts described above, do not go any further than to trying to find out how many users join the effort, and how many of them the system is able to support.
+The proof-of-concept efforts described above, do not go any further than  trying to find out how many users join the effort, and how many of them the system is able to support.
 In fact, systems such as the one described in

### 2017-11-07 2017-EUROPAR/Gamification/paper.tex
Gamification technique
Typos spotted by reviewer #2

@@
 They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement in response to the
-robot\'s resulting actions. For interaction, they use the Twitch streaming platform where users can issue commands through the
+robot resulting actions. For interaction, they use the Twitch streaming platform where users can issue commands through the
 platform's integrated chat service. Subjects were incentivized to
@@
 providing leader boards are ways of encouraging players to compete \cite{hickman2010total}.
+In this work, a leader board was implemented in order to promote competition, presenting only the all-time top five participating users.
 Competition can also have problematic consequences, which can result in
@@
 this functionality was not desired because it was seen as a barrier
-for participation. After all the advantage of a browser based volunteer system
+for participation. After all, the advantage of a browser based volunteer system
 is precisely the minimum amount of user intervention needed to start.
@@
 \subsection{Problem}
-In the browser, each page visiting the experiment loads an WebWorker
+In the browser, each page visiting the experiment loads an HTTP Web Worker
 that runs a local island of an evolutionary algorithm to solve a
@@
 %
-Contributions are expressed as the number of HTTP PUT requests sent to the server,
-in Figure~\ref{fig:puts} a comparison between registered vs anonymous users is
+Contributions are expressed as the number of HTTP PUT requests sent to the server. In Figure~\ref{fig:puts} a comparison between the contribution of registered vs anonymous users is
 presented. There where only 18 registered users, and the exact number of anonymous
@@
 User engagement is related to the amount of time a user spends in the
-application. Figure~\ref{fig:puts-time} shows to important aspects of
+application. Figure~\ref{fig:puts-time} shows two important aspects of
 engagement, the overall time spent and the amount of resources shared

### 2017-11-08 2017-EUROPAR/Gamification/paper.tex
Anonymizes  system name
Refs #5 which was already closed, but..

@@
 
-Distributed computing systems can be created using volunteers who
-provide their own computing resources or storage to contribute to a common effort.
+Distributed computing systems can be created using volunteers who decide to
+provide their own  resources or storage to contribute to a common effort.
 They can, for instance, run a script embedded in a web page, collaboration is straightforward, but also ephemeral, with resources depending on the amount of time a user lends; this  implies that
@@
 possible. In this paper, we analyze a volunteer-based evolutionary computing system called
-NodIO with the objective of discovering rules that encourage volunteer
+% NodIO
+Anonymous with the objective of discovering rules that encourage volunteer
 participation thus increasing the overall computing power. We present the results of
 an experiment where a gamification technique is applied by adding a leader-board
-showing the top scores achieved by registered contributors. In the NodIO system volunteers can
+showing the top scores achieved by registered contributors. In the
+% NodIO
+Anonymous system volunteers can
 participate without the need to create an account, so the question was
@@
 In this experiment a gamification technique is applied to
-the {\sf NodIO} volunteer computing framework, using the
+the Anonymous
+% {\sf NodIO}
+volunteer computing framework, using the
 particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
@@
 functionality is available to all users and is the same as the previous
-{\sf NodIO} version, showing the state of the current algorithm.
+%{\sf NodIO}
+Anonymous version, showing the state of the current algorithm.
 In the front page a link to the open source code of the experiment was also
@@
 applying a gamification technique has on user engagement in
-the {\sf NodIO} volunteer computing framework; In concordance with
+the Anonymous
+% {\sf NodIO}
+volunteer computing framework; In concordance with
 the results obtained in other

### 2017-11-08 2017-EUROPAR/Gamification/paper.tex
Rewriting the introductory paragraph
And also adding a bit. Please @mariosky check it out.

@@
 
-There is a trade-off between ease of use, where users just need to visit a web page, versus personlaization, where users need to register and subsecuently authenticate in order to participate.
-Registration could be seen as a barrier to volunteers trying to participate in an ephemeral experiment, so the question is what is the impact on user participation.
+There is a trade-off between ease of use, allowing users to
+participate by just visiting a web page, and personalization, making
+registration compulsory and subsequently authenticate in order to
+participate. In a sociotechnical system such as this one, every
+decision the user has to take has an influence on the overall
+performance, and registration could be seen as a hurdle to volunteers
+trying to participate in an ephemeral experiment, so it is a challenge
+to analyze what actually happens in this case; this is one of the main
+questions addressed in this paper.
 

### 2017-11-08 2017-EUROPAR/Gamification/paper.tex
Adds bib for socio-technical systems

@@
 
-\keywords{Distributed Evolutionary Algorithms, Volunteer Computing}
+\keywords{Distributed Evolutionary Algorithms, Volunteer Computing,
+  Socio-Technical Systems}
 \end{abstract}
@@
 
-There are several open issues in the research regarding the application of volunteer computing techniques to the development of distributed evolutionary algorithms of which the most important is approaching the problem as a socio-technical system \cite{vespignani2009predicting}.
-which integrates user decisions and behavioral patterns in the model; this
+There are several open issues in the research regarding the
+application of volunteer computing techniques to the development of
+distributed evolutionary algorithms of which the most important is
+approaching the problem as a socio-technical system
+\cite{vespignani2009predicting,merelo2015designing}. which integrates
+user decisions and behavioral patterns in the model; this 
 includes trying to optimize the number of users in a particular
@@
 to delve into the socio-technical nature of the system, using
-gamification techniques to improve the amount of time every user will lend to the system. Gamification has to include some way of user authentification,
+gamification techniques to improve the amount of time every user will
+lend to the system. Gamification has to include some way of user authentification,
 which is a hurdle that might take users away; however, the open design
@@
 
+The objective of this work was to test the kind of impact
+applying a gamification technique has on user engagement in
+the Anonymous
+% {\sf NodIO}
+volunteer computing framework; In concordance with
+the results obtained in other
+browser-based volunteer systems, after
+applying the gamification techniques, user participation was increased.
+
 In this paper we have added gamification to a volunteer computing
@@
 That is why future lines of work will look for intermediate ways of including
-gamification without registration, as well as some ways of making the user increase control over the algorithm that would go beyond simply staying for longer or leaving the page.
-Since it is a sociotechnical system, including social features in it and studying emerging social networks might be an interesting line of work too.
+gamification without registration, as well as some ways of making the
+user increase control over the algorithm that would go beyond simply
+staying for longer or leaving the page. Since it is a sociotechnical
+system, including social features in it and studying emerging social
+networks is an interesting line of work too.
 
 Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-The objective of this work was to test the kind of impact
-applying a gamification technique has on user engagement in
-the Anonymous
-% {\sf NodIO}
-volunteer computing framework; In concordance with
-the results obtained in other
-browser-based volunteer systems, after
-applying the gamification techniques, user participation was increased.
 

### 2017-11-08 2017-EUROPAR/Gamification/paper.tex
Some edits

--- 2017-11-08 2017-EUROPAR/Gamification/paper.tex
@@
 
-Distributed computing systems can be created using volunteers who decide to
-provide their own  resources or storage to contribute to a common effort.
+Distributed computing systems can be created using volunteers who
+provide their own computing resources or storage to contribute to a common effort.
 They can, for instance,  run a script embedded in a web page, collaboration is straightforward, but also ephemeral, with resources depending on the amount of time a user lends; this  implies that
@@
 possible. In this paper, we analyze a volunteer-based evolutionary computing system called
-% NodIO
-Anonymous with the objective of discovering rules that encourage volunteer
+NodIO with the objective of discovering rules that encourage volunteer
 participation thus increasing the overall computing power. We present the results of
 an experiment where a gamification technique is applied by adding a leader-board
-showing the top scores achieved by registered contributors. In the
-% NodIO
-Anonymous system volunteers can
+showing the top scores achieved by registered contributors. In the NodIO system volunteers can
 participate without the need to create an account, so the question was
@@
 
-There are several open issues in the research regarding the application of volunteer computing techniques to the development of distributed evolutionary algorithms of which the most important is approaching the problem as a socio-technical system \cite{vespignani2009predicting}.
-which integrates user decisions and behavioral patterns in the model; this
+There are several open issues in the research regarding the application of volunteer computing techniques to the development of distributed evolutionary algorithms of which the most important is approaching the problem as a socio-technical system \cite{vespignani2009predicting} which integrates user decisions and behavioral patterns in the model; this
 includes trying to optimize the number of users in a particular
 experiment. The challenge is to design a system that, whatever the
-number of users available to perform the experiment, is able to
+number of available users to perform the experiment, it is able to
 maximize their contribution to the evolutionary algorithm. One obvious
@@
 
-There is a trade-off between ease of use, where users just need to visit a web page, versus personlaization, where users need to register and subsecuently authenticate in order to participate.
-Registration could be seen as a barrier to volunteers trying to participate in an ephemeral experiment, so the question is what is the impact on user participation.
+There is a trade-off between ease of use, where users just need to visit a web page, versus personalization, where users need to register and subsequently authenticate to participate.
+Registration can be seen as a barrier to volunteers trying to join an ephemeral experiment, so the question is: what is the impact on user participation.
 
@@
 computation (EC). Section \ref{sec:gamification} will describe the
-framework and problem used in the experiments, which are publicly
-available under a free license. We will present the results of the
-different steps in the incremental design in Section
+framework and the problem used in the experiments, which are publicly
+available under a free license. The results of each of the steps in the incremental design are presented in Section
 \ref{sec:experiments}, to finally wrap up with the conclusions.
@@
 
-Volunteer computing involves users deciding to run a program that acts
+Volunteer computing involves users who decide to run a program that acts
 as a client or as a peer in a distributed computing experiment
@@
 systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
-require higher availability of volunteer resources, achieving this
-with downloadable clients which in some cases, like HTCondor, need
-administration access to the resources.
+requires higher availability of volunteer resources.
+In order to increase availability, systems rely on downloadable clients, and in some cases like HTCondor, they also require administrative access to certain resources.
 
@@
 \cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
-evolutionary computation and it has been used ever since by several
-authors, including more recent efforts
+evolutionary computation and it has been used by several
+authors ever since. More recent efforts include
 \cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}.
 In fact, this last paper \cite{pan2015gray} performs an analysis of what is called {\em Gray computing} covering aspects of
-feasibility, cost-effectiveness, change in the user experience and the
+feasibility, cost-effectiveness, change in the user\'s experience and the
 architectural optimization needed, concluding that the computing power
@@
 
-Using a peer to peer approach, the SPACE framework \cite{leclerc2016seamless} distributes fitness evaluations across a heterogeneous pool of cloud compute nodes and volunteer peer computers running a browser.
+The SPACE framework \cite{leclerc2016seamless}, using a peer to peer approach, distributes fitness evaluations across a heterogeneous pool of cloud compute nodes and volunteer computers running a browser.
 In SPACE, peers establish a bi-directional communication with the server using the \texttt{Socket.io} JavaScript library.
@@
 training simulated robots for the grounding of language symbols.
-They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement in response to the
-robot resulting actions. For interaction, they use the Twitch streaming platform where users can issue commands through the
-platform's integrated chat service. Subjects were incentivized to
+They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement to the resulting actions of the robot.
+For interaction, they use the Twitch streaming platform where users can issue commands through the
+platform\'s integrated chat service. Subjects were incentivized to
 interact with the system by GUI features that provided participants
@@
 
-The proof-of-concept efforts described above, do not go any further than  trying to find out how many users join the effort, and how many of them the system is able to support.
+The proof-of-concept efforts described above, do not go any further than trying to find out how many users join the effort, and how many of them the system can support.
 In fact, systems such as the one described in
-\cite{gecco07:workshop:dcor} had serious scaling issues; some of them
+\cite{gecco07:workshop:dcor} had severe scaling issues; some of them
 also tried to find out how much time was needed to find the solution
@@
 Lately, researchers have tried to integrate volunteer computing techinques as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
-In fact, exclusive volunteer-based systems exhibit a certain amount of unpredictability and might be best used in combination with ready-available computing power.
+In fact, systems using volunteers exclusively, exhibit a certain amount of unpredictability, and they might be better used in combination with ready-available computing power.
 
-As indicated in the introduction, the human is an integral part of the system, which can be considered as {\em human computation} \cite{quinn2011human}.
-Giving the user not only more control, by including registration, but also a measure to compare with other users might have positive effects.
-In fact, the BOINC projects applies a gamification technique
-by having a web page to present the ``Top 100 multi-project BOINC participants'' where
-the name of the volunteer, the number of projects, GFLOPS, country and team are presented
+As indicated in the introduction, the human is an integral part of the system, which we can consider a human computation {\em human computation} \cite{quinn2011human}.
+Giving the user more control, and adding visibility to their participation might have positive effects.
+In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' where the name of the volunteer, the number of projects, GFLOPS, country and team are displayed
 (\url{https://boinc.berkeley.edu/chart_list.php}).
@@
 A definition given by Huotari  \cite{huotari2012defining} is ``Gamification is
-the process by which gaming concepts are brought to the real world tasks associated with
+the process by which gaming concepts are brought to real world tasks associated with
 real people''. Gamification uses game design elements out of the domain of games
-with the objective of enhancing the user's experience, engagement, productivity,
+with the objective of enhancing the user\'s experience, engagement, productivity,
 learning, among others. Deterding et al. proposes the following definition:
@@
 negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
-such as women \cite{kumar2013gamification}. Another techniques to gamification
+such as women \cite{kumar2013gamification}. Another gamification techinque
 is to make existing tasks feel more like games \cite{deterding2010just}.
@@
 In this experiment a gamification technique is applied to
-the Anonymous
-% {\sf NodIO}
-volunteer computing framework, using the
+the {\sf NodIO} volunteer computing framework, using the
 particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
@@
 functionality is available to all users and is the same as the previous
-%{\sf NodIO}
-Anonymous version, showing the state of the current algorithm.
+{\sf NodIO} version, showing the state of the current algorithm.
 In the front page a link to the open source code of the experiment was also
@@
 that runs a local island of an evolutionary algorithm to solve a
-multi-modal problem called {\em l-trap}, represented in figure \ref{fig:trap}, which has been used extensively
-as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}.
+multi-modal problem called {\em l-trap} which has been used extensively
+as a benchmark for evolutionary algorithms and is represented in figure \ref{fig:trap},  \cite{fernandes2009using,nijssen2003analysis}.
 This function counts the number of bits in a sequence $l$ and assigns
 the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
-bits, this makes the  the fitness fall into a {\em trap}
+bits, this makes the fitness fall into a {\em trap}
 as the number of bits is increased, decreasing linearly until a change in slope
-is reached at point $z$, adding deceptive component for evolutionary algorithms.
-To increase the difficulty trap functions can be concatenated.
-In our case we have used $40$ concatenated traps. The trap function  defined as:
+is reached at point $z$, adding a deceptive component for evolutionary algorithms.
+Trap functions can be concatenated in order to increase the difficulty .
+In our case we have used $40$ concatenated traps.
+The trap function is defined as:
 \[ f(u)=
@@
 $a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.
+
+The experiment was planed as a usibility test, employing a relativly small sample of users.
+
+A/B testing or split testing is an experimental approach to web desig
 
@@
 applying a gamification technique has on user engagement in
-the Anonymous
-% {\sf NodIO}
-volunteer computing framework; In concordance with
+the {\sf NodIO} volunteer computing framework; In concordance with
 the results obtained in other

### 2017-11-08 2017-EUROPAR/Gamification/paper.tex

@@
 
-There are several open issues in the research regarding the application of volunteer computing techniques to the development of distributed evolutionary algorithms of which the most important is approaching the problem as a socio-technical system \cite{vespignani2009predicting} which integrates user decisions and behavioral patterns in the model; this
+There are several open issues in the research regarding the
+application of volunteer computing techniques to the development of
+distributed evolutionary algorithms of which the most important is
+approaching the problem as a socio-technical system
+\cite{vespignani2009predicting,merelo2015designing}. which integrates
+user decisions and behavioral patterns in the model; this
 includes trying to optimize the number of users in a particular
@@
 
-There is a trade-off between ease of use, where users just need to visit a web page, versus personalization, where users need to register and subsequently authenticate to participate.
-Registration can be seen as a barrier to volunteers trying to join an ephemeral experiment, so the question is: what is the impact on user participation.
+There is a trade-off between ease of use, allowing users to
+participate by just visiting a web page, and personalization, making
+registration compulsory and subsequently authenticate in order to
+participate. In a sociotechnical system such as this one, every
+decision the user has to take has an influence on the overall
+performance, and registration could be seen as a hurdle to volunteers
+trying to participate in an ephemeral experiment, so it is a challenge
+to analyze what actually happens in this case; this is one of the main
+questions addressed in this paper.
 
@@
 That is why future lines of work will look for intermediate ways of including
-gamification without registration, as well as some ways of making the user increase control over the algorithm that would go beyond simply staying for longer or leaving the page.
-Since it is a sociotechnical system, including social features in it and studying emerging social networks might be an interesting line of work too.
+gamification without registration, as well as some ways of making the
+user increase control over the algorithm that would go beyond simply
+staying for longer or leaving the page. Since it is a sociotechnical
+system, including social features in it and studying emerging social
+networks is an interesting line of work too.
 
 Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-The objective of this work was to test the kind of impact
-applying a gamification technique has on user engagement in
-the {\sf NodIO} volunteer computing framework; In concordance with
-the results obtained in other
-browser-based volunteer systems, after
-applying the gamification techniques, user participation was increased.
 

### 2017-11-09 2017-EUROPAR/Gamification/paper.tex
Removed a paragraph I was starting to write
Corrections and edits.

@@
 participation thus increasing the overall computing power. We present the results of
-an experiment where a gamification technique is applied by adding a leader-board
+an experiment in which a gamification technique is applied by adding a leader-board
 showing the top scores achieved by registered contributors. In the
@@
 gamification techniques to improve the amount of time every user will
-lend to the system. Gamification has to include some way of user authentification,
+lend to the system. Gamification has to include some way of user authentication,
 which is a hurdle that might take users away; however, the open design
 of the system we use means that we can actually mix anonymous-non
-gamified and authentified-gamified users, with only the small overhead
+gamified and authenticated-gamified users, with only the small overhead
 of adding the gamification mechanism.
@@
 They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement to the resulting actions of the robot.
-For interaction, they use the Twitch streaming platform where users can issue commands through the
+For interaction, they use the Twitch streaming platform at which users can issue commands through the
 platform\'s integrated chat service. Subjects were incentivized to
@@
 Giving the user more control, and adding visibility to their participation might have positive effects.
-In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' where the name of the volunteer, the number of projects, GFLOPS, country and team are displayed
+In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed
 (\url{https://boinc.berkeley.edu/chart_list.php}).
@@
 negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
-such as women \cite{kumar2013gamification}. Another gamification techinque
+such as women \cite{kumar2013gamification}. Another gamification technique
 is to make existing tasks feel more like games \cite{deterding2010just}.
@@
     \centering
-        \includegraphics[width=2.5in]{img/s.png}
+        \includegraphics[width=3in]{img/s.png}
     \caption{$S$ function used to assign the user's level.
@@
 After login, registered users are welcomed by their name, and they can see
-their score and level. All users can see in a modal dialog the leader board.
+their current score and level. All users can see in a modal dialog the leader board.
 While the modal is opened, the scores are refreshed every second. All other
@@
 This function counts the number of bits in a sequence $l$ and assigns
-the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
+the local maximum $a$ if it has 0 bits and the global maximum $b$ if it has $l$
 bits, this makes the fitness fall into a {\em trap}
-as the number of bits is increased, decreasing linearly until a change in slope
+as the number of bits increases, and decreasing linearly until a change in slope
 is reached at point $z$, adding a deceptive component for evolutionary algorithms.
-Trap functions can be concatenated in order to increase the difficulty .
-In our case we have used $40$ concatenated traps.
+In order to increase difficulty, trap functions can be concatenated, in our case we have used $40$ concatenated traps.
 The trap function is defined as:
@@
     \centering
-        \includegraphics[width=2.5in]{img/trap.png}
+        \includegraphics[width=3in]{img/trap.png}
     \caption{Trap function.
@@
 Each local GA had the following parameters, the initial population was randomly generated
-with sizes between 128 and 256 individuals, the period to send individuals to the server
+with a random size between 128 and 256 individuals, the period to send individuals to the server
 was set at 100 local generations, the parameters for the trap function are $l$ = 4,
-$a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.
+$a$=1, $b$ = 2, $z$=3 with a chromosome length of 160 bits.
 
-The experiment was planed as a usibility test, employing a relativly small sample of users.
-
-A/B testing or split testing is an experimental approach to web desig
+%%%% Exp
 
@@
 The experiment will run until November 27th, Thanks!.'' We also included links to register and a deadline.
-Users visiting the page after the deadline were presented with a thank you message and a static leader board.
-%
+Users visiting the page after the deadline were presented with a thank you message and a final leader board.
+Considering the number of friends, followers and the organic sharing of the posts, we estimate the post was viewed by around 2000 users.
+Relatively small samples of users ($10\pm$) are sufficient for discovering 80\% of usability problems \cite{Schmettow2012}.
 
@@
 At the end of the experiment the resulting log file (whose fields are shown in Table \ref{tab:record}) contained 933,513 contributions both
-registered users and anonymous volunteers. Registered users where responsible of around
+registered users and anonymous volunteers. Registered users were responsible of around
 90\% of the contributions. The JSON data recorded for each contribution is  presented
@@
 Contributions are expressed as the number of HTTP PUT requests sent to the server. In Figure~\ref{fig:puts} a comparison between the contribution of registered vs anonymous users is
-presented. There where only 18 registered users, and the exact number of anonymous
+presented. There were only 18 registered users, and the exact number of anonymous
 is not known as we only recorded the IP of the request, also a registered users
@@
 about 16\% of users attending the call decided to participate as named users.
+
 There is a notable difference in the amount of participation between the two
@@
     \centering
-        \includegraphics[width=4in]{img/puts_box.png}
+        \includegraphics[width=5in]{img/puts_box.png}
     \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
@@
 %
-The box plot featured in Figure \ref{fig:box} shows the amount of participation between the two groups, highlights
+The box plot featured in Figure \ref{fig:box} shows the amount of participation between the two groups, it highlights
 the difference and also the number of outliers in the groups. Two registered
-outliers where very important participants maybe competing between them. But
-there where more outliers between anonymous IPs, and some of them could be related
-to the same users. The lower quartile in in the users box-plot has more
+outliers were very important participants, maybe competing between them. But
+there were more outliers between anonymous IPs, and some of them could be related to the same users. The lower quartile in the users box-plot has more
 spread, indicating a participation similar to the median of anonymous IPs.
@@
     \centering
-        \includegraphics[width=5in]{img/workers_put_ip.png}
+        \includegraphics[width=5.3in]{img/workers_put_ip.png}
     \caption{ Participation of registered users. Each circle represents a user relating
@@
     \centering
-        \includegraphics[width=4in]{img/puts_by_time.png}
-    \caption{Number of PUTs by User or IP in 15 minute slots.
+        \includegraphics[width=5.3in]{img/puts_by_time.png}
+    \caption{Number of PUTs by User (green dots) or IP (blue triangles) in 15 minute slots.
     }
@@
 engagement, the overall time spent and the amount of resources shared
-in that time. Registered users where more engaged during the experiment.
+in that time. Registered users were more engaged during the experiment.
 They contribute more resources during longer periods of time.
@@
 %
-The participation of the top ranked user is presented in 15 seconds time slots in
+The amount of participation of the top ranked user is presented in 15 seconds time slots in
 Figure~\ref{fig:top-user}. In the beginning this user was using more than 30
-workers, this means that more than 30 tabs of the page where open at the
+workers, this means that more than 30 tabs of the page were open at the
 same time, perhaps using several computing devices. Some users reported that
 they wanted to test the limits of their own systems, checking the percentage of
-CPU they where using.
+CPU they were using.
 
@@
 for registration, their individual contribution is on average much
-higher than for non-registered/non-gamified users. The fact that less
+higher than for non-authenticated/non-gamified users. The fact that less
 than one fifth of users opt for registering would discard the

### 2017-11-09 2017-EUROPAR/Gamification/paper.tex
Needs a sentence.
@JJ please help with comment
Improved the paragraph and added a few comments
Check it out @mariosky to see if it's correct.

@@
 
-%%%% Exp
-
 We made a call to participation on November 23th,
@@
 and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
-The experiment will run until November 27th, Thanks!.'' We also included links to register and a deadline.
-Users visiting the page after the deadline were presented with a thank you message and a final leader board.
-Considering the number of friends, followers and the organic sharing of the posts, we estimate the post was viewed by around 2000 users.
-Relatively small samples of users ($10\pm$) are sufficient for discovering 80\% of usability problems \cite{Schmettow2012}.
+The experiment will run until November 27th, Thanks!.'' As you can
+see, the message mentioned a deadline; a link that could be used to
+register was also included. Users visiting the page after the deadline
+were presented with a thank you message and a final leader
+board. Considering the number of friends, followers and the organic
+sharing of the posts, we estimate the post was visited by around 2000
+users. Only some of them registered, but in fact relatively small samples of users ($10\pm$) are sufficient for
+discovering 80\% of usability problems \cite{Schmettow2012}. In fact,
+we were not so much interested in making a precise model but on how
+the behavior of the registered users was different from non-registered
+ones; in non-parametric comparisons, using 15 samples is in general
+good enough, as seems to be the case in this particular
+experiment, where we were also interested in computing the number of
+users who actually registered, proving that, in fact, for most users
+the fact that they have to register represents a bigger effort they
+seem to be able to muster for this kind of experiments.
 

### 2017-11-09 2017-EUROPAR/Gamification/paper.tex
Please add this data
@mariosky I think it's important that we enphasize that the registered
users contribute, per person, much more than non-registereed ones.
Check out the last comment
Basically use also statistical measures to compare both, @mariosky. Do
you have the data handy?

@@
 At the end of the experiment the resulting log file (whose fields are shown in Table \ref{tab:record}) contained 933,513 contributions both
-registered users and anonymous volunteers. Registered users were responsible of around
-90\% of the contributions. The JSON data recorded for each contribution is  presented
+registered users and anonymous volunteers. Registered users were responsible for about
+90\% of the contributions. % While being only x% of the total number
+                           % of unique users - JJ
+The JSON data recorded for each contribution is  presented
 in Table~\ref{tab:record}.
 
-\begin{figure}[h!tbp]
+\begin{figure}[h!tb]
     \centering
@@
         \includegraphics[width=0.9\linewidth]{img/puts_ip.png}
-    \caption{Usage by registered (left) and anonymous (right) users.  Users are ranked by
+    \caption{Usage by registered (top) and anonymous (bottom) users.  Users are ranked by
         the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis
@@
       \label{fig:puts}
+% Maybe we should scale dividing by number of users and put them
+% together. - JJ
 \end{figure}
 %
-Contributions are expressed as the number of HTTP PUT requests sent to the server. In Figure~\ref{fig:puts} a comparison between the contribution of registered vs anonymous users is
-presented. There were only 18 registered users, and the exact number of anonymous
-is not known as we only recorded the IP of the request, also a registered users
-could sometimes be anonymous too, this could be because the first time they visit
-the application they automatically begin to work as anonymous. Nevertheless,
-there were 91 distinct IPs that had `anonymous' as the user. In the plot of
+Contributions are expressed as the number of HTTP {\tt PUT} requests
+sent to the server. In Figure~\ref{fig:puts} a comparison between the
+contribution of registered vs anonymous users is 
+presented.
+
+The first observation arises from the length of the $x$ axis. There
+were only 18 registered users vs 91 distinct IPs; the 
+exact number of anonymous users is upper-bounded by that amount but is
+not known precisely as we only recorded the IP of the request, we did
+not use cookies or other means of identification since it was not
+actually needed for the experiment. Besides, a registered user
+could sometimes be anonymous too, which could be caused by the fact that the first time they visited
+the application they automatically begin to work as anonymous, so the
+number of non-anonymous users, 91, is actually the amount of unique
+IPs that had `anonymous' as the user. In the plot of 
 anonymous users, some of the last in the rank could in fact be registered users,
 momentarily participating as anonymous. Even if this is not considered, only
-about 16\% of users attending the call decided to participate as named users.
+about 16\% of users attending the call decided to participate as named
+users and thus enter the game dynamics.
 
-There is a notable difference in the amount of participation between the two
-groups, and also in the slope of the plot.
+The second observation is apparent in the $y$ axis of Figure
+\ref{fig:puts}. There is a notable difference in the amount of
+participation between the two groups, and also in the slope of the
+plot. %Maybe a statistical comparison? Median, for instance - JJ
 

### 2017-11-09 2017-EUROPAR/Gamification/paper.tex
Minor clarification

@@
 the behavior of the registered users was different from non-registered
-ones; in non-parametric comparisons, using 15 samples is in general
+ones; in non-parametric comparisons, using 15 sample units is in general
 good enough, as seems to be the case in this particular

### 2017-11-09 2017-EUROPAR/Gamification/log_code/by_time_slot.py
Code

@@ -1,0 +1,44 @@
+import  sqlite3
+import time
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# logs table: timestamp, IP , level , user , worker
+c.execute(
+    """ SELECT datetime((strftime('%s', timestamp) / 900) * 900, 'unixepoch') interval, count(*)
+        FROM logs
+        WHERE user <> 'anonimo'
+        GROUP BY interval
+        ORDER BY interval
+""")
+
+users = [( time.mktime(time.strptime(r[0],'%Y-%m-%d %H:%M:%S' )),r[1]) for r in c]
+
+
+c.execute(
+    """ SELECT datetime((strftime('%s', timestamp) / 900) * 900, 'unixepoch') interval, count(*)
+        FROM logs
+        WHERE user = 'anonimo'
+        GROUP BY interval
+        ORDER BY interval
+""")
+
+IPs = [( time.mktime(time.strptime(r[0],'%Y-%m-%d %H:%M:%S' )),r[1]) for r in c]
+
+
+conn.close()
+
+
+import matplotlib.pyplot as plt
+from matplotlib.ticker import ScalarFormatter, MaxNLocator
+
+
+fig , ax =plt.subplots()
+
+ax.plot( [ u[0] for u in users] , [ u[1] for u in users],'g.')
+ax.plot( [ i[0] for i in IPs] , [ i[1] for i in IPs],'b^')
+ax.set_ylabel('PUTs')
+ax.set_xlabel('15 minute slots (unix epoch)')
+ax.grid(True)
+
+plt.show()

### 2017-11-09 2017-EUROPAR/Gamification/log_code/log2sqlite.py

@@ -1,0 +1,26 @@
+import json, sqlite3
+
+# Create DB
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# Create table
+c.execute('''CREATE TABLE logs
+             (timestamp text, IP text, level text, user text, worker text)''')
+
+
+
+file = open("nodio-2016-11-23-0-json.txt")
+
+records = [(dict_entry["timestamp"],dict_entry["IP"],dict_entry["level"],dict_entry["user"],dict_entry["worker_uuid"] )
+           for dict_entry in [ json.loads(line) for line in file]]
+print records[:5]
+
+c.executemany("INSERT INTO logs VALUES (?,?,?,?,?)", records)
+
+# Save (commit) the changes
+conn.commit()
+
+# We can also close the connection if we are done with it.
+# Just be sure any changes have been committed or they will be lost.
+conn.close()

### 2017-11-09 2017-EUROPAR/Gamification/log_code/user_puts.py

@@ -1,0 +1,73 @@
+
+import sqlite3
+
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# logs table: timestamp, IP , level , user , worker
+c.execute(
+    """ SELECT user , count(*)
+        FROM logs
+        WHERE user <> 'anonimo'
+        GROUP BY user
+        ORDER BY 2 DESC
+""")
+
+users = [r[1]  for r in c]
+
+
+c.execute(
+        """ SELECT IP , count(*)
+            FROM logs
+            WHERE user == 'anonimo'
+            GROUP BY IP
+            ORDER BY 2 DESC
+    """)
+
+IPs = [r[1] for r in c]
+
+
+# Save (commit) the changes
+conn.commit()
+
+# We can also close the connection if we are done with it.
+# Just be sure any changes have been committed or they will be lost.
+conn.close()
+
+import matplotlib.pyplot as plt
+from matplotlib.ticker import ScalarFormatter, MaxNLocator
+
+x_users = range(len(users))
+print len(users)
+# automatically update ylim of ax2 when ylim of ax1 changes.
+fig , ax =plt.subplots()
+ax.plot(x_users,users,'.')
+ax.set_yscale('log')
+ax.axis([1, 20, 1, 100000])
+for axis in [ax.xaxis, ax.yaxis]:
+    axis.set_major_formatter(ScalarFormatter())
+
+ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+ax.set_ylabel('PUTs')
+ax.set_xlabel('users')
+ax.grid(True)
+
+plt.show()
+
+
+x_ips = range(len(IPs))
+print len(IPs)
+fig , ax =plt.subplots()
+print IPs
+ax.plot(x_ips,IPs,'.')
+ax.set_yscale('log')
+ax.axis([1, 100, 1, 100000])
+for axis in [ax.xaxis, ax.yaxis]:
+    axis.set_major_formatter(ScalarFormatter())
+
+ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+ax.set_ylabel('PUTs')
+ax.set_xlabel('IPs')
+ax.grid(True)
+plt.show()
+

### 2017-11-09 2017-EUROPAR/Gamification/log_code/user_puts_box.py

@@ -1,0 +1,52 @@
+
+import  sqlite3
+
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# logs table: timestamp, IP , level , user , worker
+c.execute(
+    """ SELECT user , count(*)
+        FROM logs
+        WHERE user <> 'anonimo'
+        GROUP BY user
+        ORDER BY 2 DESC
+""")
+
+users = [r[1] for r in c]
+
+
+c.execute(
+        """ SELECT IP , count(*)
+            FROM logs
+            WHERE user == 'anonimo'
+            GROUP BY IP
+            ORDER BY 2 DESC
+    """)
+
+IPs = [r[1] for r in c]
+
+
+# Save (commit) the changes
+conn.commit()
+
+# We can also close the connection if we are done with it.
+# Just be sure any changes have been committed or they will be lost.
+conn.close()
+
+
+import matplotlib.pyplot as plt
+from matplotlib.ticker import ScalarFormatter, MaxNLocator
+
+x_users = range(len(users))
+# automatically update ylim of ax2 when ylim of ax1 changes.
+fig , ax =plt.subplots()
+ax.boxplot([users,IPs], vert=True,patch_artist=True)
+ax.set_yscale('log')
+# add x-tick labels
+plt.setp([ax], xticklabels=['Users', 'IPs'])
+ax.set_ylabel('PUTs')
+ax.grid(True)
+plt.show()
+
+

### 2017-11-09 2017-EUROPAR/Gamification/log_code/workers_per_user.py

@@ -1,0 +1,53 @@
+import  sqlite3
+import time
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# logs table: timestamp, IP , level , user , worker
+c.execute(
+    """ SELECT user, count(distinct IP), count(distinct worker), count(*)
+        FROM logs
+        WHERE user <> 'anonimo'
+        GROUP BY user
+        ORDER BY 2 desc, user
+""")
+
+users = [ r for r in c]
+conn.close()
+print users[0]
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import FuncFormatter
+
+def log_10_product(x, pos):
+    """The two args are the value and tick position.
+    Label ticks with the product of the exponentiation"""
+    return '%1i' % (x)
+
+colors = np.random.rand(len(users))
+
+fig , ax = plt.subplots()
+
+ax.set_yscale('log')
+ax.set_xscale('log')
+ax.set_ylabel('PUTs')
+ax.set_xlabel('Workers')
+ax.set_ylim(1e-1, 1e7)
+ax.set_xlim(1e-1, 1e5)
+formatter = FuncFormatter(log_10_product)
+ax.xaxis.set_major_formatter(formatter)
+ax.yaxis.set_major_formatter(formatter)
+
+ax.grid(True)
+
+
+
+
+ax.scatter([ u[2] for u in users], [ u[3] for u in users], s=[ u[1]*50 for u in users], c=colors, alpha=0.5)
+
+plt.show()
+
+
+
+

### 2017-11-09 2017-EUROPAR/Gamification/log_code/workers_per_user_time_slot_best.py

@@ -1,0 +1,45 @@
+import  sqlite3
+import time
+conn = sqlite3.connect('log.db')
+c = conn.cursor()
+
+# logs table: timestamp, IP , level , user , worker
+c.execute(
+    """ SELECT datetime((strftime('%s', timestamp) / 15) * 15, 'unixepoch') interval, count(distinct worker)
+        FROM logs
+        WHERE user = 'd625621a78a8362a0b39d2306a433944c7b987df'
+        GROUP BY interval
+        ORDER BY interval
+""")
+
+users = [( time.mktime(time.strptime(r[0],'%Y-%m-%d %H:%M:%S' )),r[1], r[0]) for r in c]
+conn.close()
+
+
+ticks =[(1479994980.0, 0, u'05:43:00'),
+    (1480043640.0, 2880, u'19:14:00'),
+    (1480058790.0, 3840, u'23:26:30'),
+    (1480074180.0, 4320, u'03:43:00'),
+    (1480089930.0, 4800, u'08:05:30'),
+    (1480104345.0, 5760, u'12:05:45'),
+    (1480153065.0, 6720, u'01:37:45'),
+    (1480260390.0, 7200, u'07:26:30'),
+    (1480322565.0, 7680, u'00:42:45')]
+
+
+
+
+import matplotlib.pyplot as plt
+from matplotlib.ticker import ScalarFormatter, MaxNLocator
+
+
+fig , ax =plt.subplots()
+
+ax.plot( [ u[0] for u in users] , [ u[1] for u in users],'g.')
+
+ax.set_ylabel('Workers')
+ax.set_xlabel('15 second slots')
+ax.grid(True)
+
+plt.xticks([t[0] for t in ticks] , [t[2] for t in ticks] , rotation=30)
+plt.show()

### 2017-11-09 2017-EUROPAR/Gamification/log_code/stats.py
stats

@@ -1,5 +1,8 @@
-
 import  sqlite3
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+from scipy.stats import ttest_ind
 
 conn = sqlite3.connect('log.db')
 c = conn.cursor()
@@ -13,7 +16,7 @@ c.execute(
         ORDER BY 2 DESC
 """)
 
-users = [r[1] for r in c]
+users =  [ (_id, _count,'user') for (_id, _count) in c]
 
 
 c.execute(
@@ -24,33 +27,21 @@ c.execute(
             ORDER BY 2 DESC
     """)
 
-IPs = [r[1] for r in c]
+IPs = [(_id, _count,'IP') for (_id, _count) in c]
 
 
+
+df_users = pd.DataFrame(users, columns= ['id','count','category'])
+df_IPs = pd.DataFrame(IPs, columns= ['id','count','category'])
+
+df_all = pd.concat([df_users, df_IPs])
+
+print df_all.groupby('category').describe()
+
+print ttest_ind(df_users['count'], df_IPs['count'])
 # Save (commit) the changes
 #conn.commit()
 
 # We can also close the connection if we are done with it.
 # Just be sure any changes have been committed or they will be lost.
 conn.close()
-
-print len(IPs)
-print len(users)
-
-import matplotlib.pyplot as plt
-from matplotlib.ticker import ScalarFormatter, MaxNLocator
-
-
-
-x_users = range(len(users))
-# automatically update ylim of ax2 when ylim of ax1 changes.
-fig , ax =plt.subplots()
-ax.boxplot([users,IPs], vert=True,patch_artist=True)
-ax.set_yscale('log')
-# add x-tick labels
-plt.setp([ax], xticklabels=['Users', 'IPs'])
-ax.set_ylabel('PUTs')
-ax.grid(True)
-plt.show()
-
-

### 2017-11-09 2017-EUROPAR/Gamification/paper.tex
Assuming IPs are unique users
Comparison of user participation

@@
 registered users and anonymous volunteers. Registered users were responsible for about
-90\% of the contributions. % While being only x% of the total number
-                           % of unique users - JJ
+90\% of the contributions, while being only 16\% of the total number of unique users (assuming each IP is from a single user).
 The JSON data recorded for each contribution is  presented
@@
 participation between the two groups, and also in the slope of the
-plot. %Maybe a statistical comparison? Median, for instance - JJ
+plot.
+An independent-samples t-test was conducted to compare participation, and there was a significant difference in the scores for anonymous users  (M=1057.97, SD= 3669.86) and non-anonymous users (M=46513.55, SD=93537.20); t-test= 4.70, p = 7.5e-06.
 
+%Maybe a statistical comparison? Median, for instance - JJ
+%count          mean           std  min      25%      50%      75%   
+%category                                                                     
+%IP        91.0   1057.978022   3669.862398  1.0     1.50     18.0    173.0   
+%user      18.0  46513.555556  93537.207948  7.0  3962.75  13034.5  46993.5
+
+%      max  
+%category            
+%IP         26593.0  
+%user      396081.0  
+%Ttest_indResult(statistic=4.7070787365897289, pvalue=7.565505602664807e-06)
 %
+
 \begin{figure*}[htb]

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
@mariosky check out title change

@@
 
-\title{Increasing User Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System}
-\titlerunning{Increasing User Engagement in Volunteer Evolutionary Computing}
+\title{Increasing Performance via Gamification in a Volunteer-Based Evolutionary Computation System}
+\titlerunning{Increasing Performance via Gamification in a Volunteer
+  EC System}
 

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Revising abstract

@@
 
-Distributed computing systems can be created using volunteers who decide to
-provide their own  resources or storage to contribute to a common effort.
-They can, for instance, run a script embedded in a web page, collaboration is straightforward, but also ephemeral, with resources depending on the amount of time a user lends; this  implies that
+Distributed computing systems can be created using volunteers; these
+are users who spontaneously, after receiving an invitation, decide to
+provide their own resources or storage to contribute to a common
+effort. They can, for instance, run a script embedded in a web page;
+thus, 
+collaboration is straightforward, but also ephemeral, with resources
+depending on the amount of time the user decides to lend. This  implies that 
 the user has to be kept engaged so as to obtain as many computing cycles as
@@
 % NodIO
-Anonymous with the objective of discovering rules that encourage volunteer
-participation thus increasing the overall computing power. We present the results of
+Anonymous with the objective of discovering design decisions that encourage volunteer
+participation, thus increasing the overall computing power. We present the results of
 an experiment in which a gamification technique is applied by adding a leader-board
@@
 Anonymous system volunteers can
-participate without the need to create an account, so the question was
+participate without the need to create an account, so one of the
+questions we wanted to address was
 if the need to register would have a negative impact on user participation.
 The experiment results show that even if only a small percentage of users created an account,
-those participating in the competition provided around 90\% of the work.
+those participating in the competition provided around 90\% of the
+work, thus effectively increasing the performance of the overall
+system.
 

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Responding to one of the reviewers
Saying that we are more interested in the real amount of time to see
if it is worth the while. Please @mariosky check it out.

@@
 \cite{sarmenta2001volunteer} and it dates back, in different shapes
-and underlying mechanisms, to the origin of the web
+and underlying mechanisms, to the origin of the Internet
 \cite{david-seti:home}.
 
-We are mainly interested in evolutionary algorithms \cite{jj-ppsn98},, population-based search and optimization methods inspired by
-the theory of Evolution and its molecular basis.
-The fact that they are population-based makes them suitable for a straightforward distribution of work to different clients, and since they have an asynchronous nature, there is not a considerable impact on their performance.
-All these properties make them ideal candidates for volunteer computing setups such as the one presented in this paper.
+In this context, we are mainly interested in evolutionary algorithms
+\cite{milani2004online,sherry2012flex,jj-ppsn98}%,
+% population-based search and optimization methods inspired by
+% the theory of Evolution and its molecular basis.
+. The fact that they are population-based makes them suitable for a
+straightforward distribution of work to different clients, and since
+they have an asynchronous nature, there is no considerable impact on
+their performance, and might even be beneficial for a number of
+reasons, including increased diversity
+\cite{cantu-paz:migration-policies}. All these properties make them
+ideal candidates for volunteer computing setups such as the one
+presented in this paper, at least if some minimal changes are made to
+the usual sequential or parallel algorithms that rely on a static
+setup.
 
-There are several open issues in the research regarding the
+But even if adaptation is straightforward, there are several open issues in the research regarding the
 application of volunteer computing techniques to the development of
@@
 approaching the problem as a socio-technical system
-\cite{vespignani2009predicting,merelo2015designing}. which integrates
+\cite{vespignani2009predicting,merelo2015designing}, which integrates
 user decisions and behavioral patterns in the model; this
-includes trying to optimize the number of users in a particular
+includes trying to optimize the number of users or the overall time
+they contribute in a particular
 experiment. The challenge is to design a system that, whatever the
-number of available users to perform the experiment, it is able to
+number of users available and willing to perform the experiment, it is able to
 maximize their contribution to the evolutionary algorithm. One obvious
@@
 to delve into the socio-technical nature of the system, using
-gamification techniques to improve the amount of time every user will
-lend to the system. Gamification has to include some way of user authentication,
+gamification techniques to try and improve the amount of time every user will
+lend to the system. However, there is a tradeoff with gamification,
+since users have to be identified in some way to participate, and thus
+it has to include some way of user authentication,
 which is a hurdle that might take users away; however, the open design
@@
 
-There is a trade-off between ease of use, allowing users to
+This trade-off between ease of use, allowing users to
 participate by just visiting a web page, and personalization, making
 registration compulsory and subsequently authenticate in order to
-participate. In a sociotechnical system such as this one, every
+participate, is what we will analyze in this paper. In a sociotechnical system such as this one, every
 decision the user has to take has an influence on the overall
@@
 trying to participate in an ephemeral experiment, so it is a challenge
-to analyze what actually happens in this case; this is one of the main
-questions addressed in this paper.
+to analyze what actually happens in this case; quite clearly, the
+inclusion of gamification techniques has an impact on page loading and
+complexity of server-side programming; if eigher the number of users
+or the amount of time they contribute to the experiment is not enough,
+it would be better to revert to the purely anonymous system.
 

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
More anon stuff refs #5

@@
 The rest of the paper is organized as follows: Next we will briefly
-present the state of the art in opportunistic distributed evolutionary
+present the state of the art in opportunistic/volunteer distributed evolutionary
 computation (EC). Section \ref{sec:gamification} will describe the
@@
 as a client or as a peer in a distributed computing experiment
-and, as such, has been deployed in many different ways from the
+and, as such, has been deployed in many different ways since the
 beginning of the Internet, starting with the SETI@home framework for
-processing extraterrestrial signals \cite{david-seti:home}, or a
+processing extraterrestrial signals \cite{david-seti:home}, which used
+a client that acted as an screensaver, or a
 high-throughput queuing system such as HTCondor \cite{HTCondor}.
@@
 frameworks such as the one we are using here, and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15}. Systems based on the
+was described in \cite{DBLP:conf/gecco/GuervosG15anom}. Systems based on the
 JavaScript/browser combination

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Deleted typo

--- 2017-11-10 2017-EUROPAR/Gamification/paper.tex
@@
 inclusion of gamification techniques has an impact on page loading and
-complexity of server-side programming; if eigher the number of users
+complexity of server-side programming; if the number of users
 or the amount of time they contribute to the experiment is not enough,

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Adding to related work refs #2

--- 2017-11-10 2017-EUROPAR/Gamification/paper.tex
@@
 made this combination the most popular for volunteer computing
-frameworks such as the one we are using here, and whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15anom}. Systems based on the
+frameworks in what is generally described as BBVC or Browser-Based
+Volunteer Computing \cite{fabisiak2017browser}. Systems based on the
 JavaScript/browser combination
 emphasize the ephemerality, ease of use, and universality, while
-systems such as HTCondor or BOINC might be more adequate for work that % citation for BOINC (?)
-requires higher availability of volunteer resources.
-In order to increase availability, systems rely on downloadable clients, and in some cases like HTCondor, they also require administrative access to certain resources.
+systems such as HTCondor or BOINC \cite{boinc_grid04} might be more adequate for work that 
+requires higher availability of volunteer resources. In order to
+increase availability, systems rely on downloadable clients, and in
+some cases like HTCondor, they also require administrative access to
+certain resources. In fact, Fabisiak \cite{fabisiak2017browser}
+analyzes the different capabilities of Internet and browser-based
+systems, including their ability to handle large sets of input data
+and time-consuming problems, concluding that BBVC is mostly apt for
+short, data-light tasks. This recent survey does not analyze the
+impact of gamification as we do in this paper, which might help in
+this area.
 
@@
 authors ever since. More recent efforts include
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}.
-In fact, this last paper \cite{pan2015gray} performs an analysis of what is called {\em Gray computing} covering aspects of
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}. In
+fact, this last paper \cite{pan2015gray} performs an analysis of what
+is called {\em Gray computing} covering aspects of 
 feasibility, cost-effectiveness, change in the user\'s experience and the
@@
 
-Next we will present the gamification technique used in this experiment.
+Next we will present the gamification technique used in this
+experiment, using a system whose first version
+was described in \cite{DBLP:conf/gecco/GuervosG15anom}.
 

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Improves bibliography

@@
 frameworks in what is generally described as BBVC or Browser-Based
-Volunteer Computing \cite{fabisiak2017browser}. Systems based on the
+Volunteer Computing \cite{fabisiak2017browser}.
+
+Systems based on the
 JavaScript/browser combination
@@
 systems such as HTCondor or BOINC \cite{boinc_grid04} might be more adequate for work that 
-requires higher availability of volunteer resources. In order to
-increase availability, systems rely on downloadable clients, and in
+requires higher availability of volunteer resources.
+In order to attract volunteers, systems rely on downloadable clients, and in
 some cases like HTCondor, they also require administrative access to
@@
 authors ever since. More recent efforts include
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray}. In
-fact, this last paper \cite{pan2015gray} performs an analysis of what
+\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray,desell2017large,rivas2017time,leclerc2016seamless,merelo2017performance}. In
+fact, \cite{pan2015gray} performs an analysis of what
 is called {\em Gray computing} covering aspects of 

### 2017-11-10 2017-EUROPAR/Gamification/paper.tex
Comments on user behavior
Maybe @mariosky knows them better and can check whether this is true.

@@
         number or PUTs in a logarithmic scale. The area of the circle is the number
-        of unique IPs used by the user.
+        of unique IPs associated to every user. 
     }
@@
 resources shared by registered users. The area of the circle is proportional
-to the number of unique IPs used by each user, the user with more participation
-used a total of 66 different IPs, this number could be related with the number
-of devices used during the participation. There are some users that using less
+to the number of unique IPs used by each user; the user with the most participation
+used a total of 66 different IPs. this number could be related with the number
+of different devices used during the participation. There are some users that using less
 devices participated more, this could mean a more powerful device, faster
-Internet connection or simply more time spent participating.
+Internet connection or simply more time spent participating. At any
+rate, the chart seems to indicate that the main mechanism registered
+users employ to increase their rank seems to be using all devices
+available, and maxing out their CPUs by increasing the number of
+workers. This indicates that, in fact, registered users who have
+incurred in additional time expenses by performing registration and
+login are ready to work even more by manually deploying more workers
+and using more computers, which is another proof of the
+socio-technical nature of volunteer computing systems.
 
@@
 
-The objective of this work was to test the kind of impact
+In this paper we have added gamification to a volunteer computing
+system via a leaderboard of contributions of registered users. The objective of this work was to test the kind of impact
 applying a gamification technique has on user engagement in
@@
 % {\sf NodIO}
-volunteer computing framework; In concordance with
+volunteer computing framework; in concordance with
 the results obtained in other
 browser-based volunteer systems, after
-applying the gamification techniques, user participation was increased.
+applying the gamification techniques, registered users participated much more intensely than anonymous ones.
 
-In this paper we have added gamification to a volunteer computing
-system via a leaderboard of contributions of registered users. In
-fact, we have proved that, although only a minority of the users opt
-for registration, their individual contribution is on average much
+In fact, we have proved that, although only a minority of the users opt
+for registration, their individual contributions are on average much
 higher than for non-authenticated/non-gamified users. The fact that less
-than one fifth of users opt for registering would discard the
+than one fifth of users actually registering would discard the
 exclusive use of this feature; however, it is an interesting addition

### 2018-01-06 2018-IPMU/Gamification/paper.tex
Changes names for new conference

--- 2017-11-10 2017-EUROPAR/Gamification/paper.tex
@@
 effort. They can, for instance, run a script embedded in a web page;
-thus, 
-collaboration is straightforward, but also ephemeral, with resources
-depending on the amount of time the user decides to lend. This  implies that 
-the user has to be kept engaged so as to obtain as many computing cycles as
-possible. In this paper, we analyze a volunteer-based evolutionary computing system called
-% NodIO
-Anonymous with the objective of discovering design decisions that encourage volunteer
-participation, thus increasing the overall computing power. We present the results of
-an experiment in which a gamification technique is applied by adding a leader-board
-showing the top scores achieved by registered contributors. In the
-% NodIO
-Anonymous system volunteers can
-participate without the need to create an account, so one of the
-questions we wanted to address was
-if the need to register would have a negative impact on user participation.
-The experiment results show that even if only a small percentage of users created an account,
-those participating in the competition provided around 90\% of the
-work, thus effectively increasing the performance of the overall
-system.
+thus, collaboration is straightforward, but also ephemeral, with
+resources depending on the amount of time the user decides to
+lend. This implies that the user has to be kept engaged so as to
+obtain as many computing cycles as possible. In this paper, we analyze
+a volunteer-based evolutionary computing system called % NodIO
+Anonymous with the objective of discovering design decisions that
+encourage volunteer participation, thus increasing the overall
+computing power. We present the results of an experiment in which a
+gamification technique is applied by adding a leader-board showing the
+top scores achieved by registered contributors. In the % NodIO
+Anonymous system volunteers can participate without the need to create
+an account, so one of the questions we wanted to address was if the
+need to register would have a negative impact on user participation.
+The experiment results show that even if only a small percentage of
+users created an account, those participating in the competition
+provided around 90\% of the work, thus effectively increasing the
+performance of the overall system.
 
@@
 gamification techniques to try and improve the amount of time every user will
-lend to the system. However, there is a tradeoff with gamification,
+lend to the system. However, there is a trade-off with gamification,
 since users have to be identified in some way to participate, and thus
@@
 algorithm.
-Lately, researchers have tried to integrate volunteer computing techinques as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
+Lately, researchers have tried to integrate volunteer computing techniques as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
 In fact, systems using volunteers exclusively, exhibit a certain amount of unpredictability, and they might be better used in combination with ready-available computing power.
@@
 90\% of the contributions, while being only 16\% of the total number of unique users (assuming each IP is from a single user).
-The JSON data recorded for each contribution is  presented
+The data recorded in JSON format for every contribution is  presented
 in Table~\ref{tab:record}.
@@
 %
-Contributions are expressed as the number of HTTP {\tt PUT} requests
+Contributions are counted by the number of HTTP {\tt PUT} requests
 sent to the server. In Figure~\ref{fig:puts} a comparison between the 
-contribution of registered vs anonymous users is 
-presented.
+contribution of registered vs anonymous users is presented.
 
@@
 participation between the two groups, and also in the slope of the
-plot.
-An independent-samples t-test was conducted to compare participation, and there was a significant difference in the scores for anonymous users  (M=1057.97, SD= 3669.86) and non-anonymous users (M=46513.55, SD=93537.20); t-test= 4.70, p = 7.5e-06.
-
+plot.  An independent-samples t-test was conducted to compare
+participation, and there was a significant difference in the scores
+for anonymous users  (M=1057.97, SD= 3669.86) and non-anonymous users
+(M=46513.55, SD=93537.20); t-test= 4.70, p = 7.5e-06; that is, the
+contribution of registered users was two orders of magnitude bigger
+than for anonymous users. This difference is highlighted in 
+%
 %Maybe a statistical comparison? Median, for instance - JJ
@@
 %user      18.0  46513.555556  93537.207948  7.0  3962.75  13034.5  46993.5   
-
+%
+%           
 %      max  
@@
 %
-
+%
 \begin{figure*}[htb]
@@
         \includegraphics[width=5in]{img/puts_box.png}
-    \caption{ Box-plot of the number of requests sent by registered users and anonymous IPs.
+    \caption{Box-plot of the number of requests sent by registered users and anonymous IPs.
      In the \emph{y} axis are number or PUTs in a logarithmic scale.
@@
 %
-The box plot featured in Figure \ref{fig:box} shows the amount of participation between the two groups, it highlights
-the difference and also the number of outliers in the groups. Two registered
-outliers were very important participants, maybe competing between them. But
-there were more outliers between anonymous IPs, and some of them could be related to the same users. The lower quartile in the users box-plot has more
-spread, indicating a participation similar to the median of anonymous IPs.
+ Figure \ref{fig:box}, which is a boxplot of the  participation by two
+ groups, showing the big difference between registered users (shown as
+ ``users'') and anonymous ones (shown as ``IPs''. This chart also
+ shows the number of outliers in the groups, which is an obvious
+ consequence of the ephemeral and spontaneous nature of the
+ collaboration; in fact, two registered
+outliers were very important participants, maybe competing between
+them, an expected consequence of the gamification, in fact, although
+not in this precise form and quantity. But
+there were more outliers between anonymous IPs, and some of them could
+be related to the same users. The lower quartile in the users box-plot
+has more 
+spread, indicating a participation similar to the median of anonymous
+IPs, indicating that the difference in distribution is mainly due to a
+few highly engaged and registered users.
 

### 2018-01-06 2018-IPMU/Gamification/paper.tex
De-anonymizes
Revises and reduces related work, closes #6

@@
 
-\author{anonymous author name \inst{1} \and anonymous author name\inst{2} \and  anonymous author name \inst{1}}
+\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}
 
-\institute{anonymous institution, some city, country
+\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
 \and
-anonymous institution, some city, country\\
-\email{example@example.edu}\\
-\email{example2@example.edu}}
+Universidad de Granada, Granada, Spain
+\email{mario@tectijuana.edu.mx}\\
+\email{jmerelo@geneura.ugr.es}}
 
-%\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}
-
-%\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
-%\and
-%Universidad de Granada, Granada, Spain
-%\email{mario@tectijuana.edu.mx}\\
-%\email{jmerelo@geneura.ugr.es}}
-
-%\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }
+\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }
 
@@
 
-Distributed computing systems can be created using volunteers; these
-are users who spontaneously, after receiving an invitation, decide to
+Distributed computing systems can be created using volunteers, users who spontaneously, after receiving an invitation, decide to
 provide their own resources or storage to contribute to a common
@@
 obtain as many computing cycles as possible. In this paper, we analyze
-a volunteer-based evolutionary computing system called % NodIO
-Anonymous with the objective of discovering design decisions that
+a volunteer-based evolutionary computing system called  NodIO
+ with the objective of discovering design decisions that
 encourage volunteer participation, thus increasing the overall
@@
 gamification technique is applied by adding a leader-board showing the
-top scores achieved by registered contributors. In the % NodIO
-Anonymous system volunteers can participate without the need to create
+top scores achieved by registered contributors. In the  NodIO
+system volunteers can participate without the need to create
 an account, so one of the questions we wanted to address was if the
@@
 
-The SPACE framework \cite{leclerc2016seamless}, using a peer to peer approach, distributes fitness evaluations across a heterogeneous pool of cloud compute nodes and volunteer computers running a browser.
-In SPACE, peers establish a bi-directional communication with the server using the \texttt{Socket.io} JavaScript library.
-As an experiment they provide an \texttt{ASM.js} compiled version of the RoboGen software platform, showing that JavaScript can be used in a broad
-spectrum of applications.
-
 Recent works have been using crowdsourcing in order to train robots
@@
 Giving the user more control, and adding visibility to their participation might have positive effects.
-In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed
-(\url{https://boinc.berkeley.edu/chart_list.php}).
-
-Next we will present the gamification technique used in this
-experiment, using a system whose first version
-was described in \cite{DBLP:conf/gecco/GuervosG15anom}.
+In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed.
 

### 2018-01-06 2018-IPMU/Gamification/paper.tex
Revises conclusions closes #7

@@
 applying a gamification technique has on user engagement in
-the Anonymous
-% {\sf NodIO}
+the {\sf NodIO}
 volunteer computing framework; in concordance with
@@
 
-Other factors, such as the language used, the way the experiment is announced, are also very important in this social context, which is why they will be included in future lines of work too.
-
-One of the interesting future lines of work would be to look a bit
-more closely at the behavior of users as they are participating
+Other factors, such as the language used, the way the experiment is
+announced, are also very important in this social context, which is
+why they will be included in future lines of work too, along with the
+study of the behavior of users as they are participating
 in the web system.
@@
 
-% This work has been supported in part by: Ministerio espa\~{n}ol de
-% Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
-% (UGR-EPHEMECH).
-Hidden for double blind.
+This work has been supported in part by: Ministerio espa\~{n}ol de
+ Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
+ (UGR-EPHEMECH).
 

### 2018-01-06 2018-IPMU/Gamification/paper.tex
Fixes typos closes #9

@@
 
-As indicated in the introduction, the human is an integral part of the system, which we can consider a human computation {\em human computation} \cite{quinn2011human}.
+As indicated in the introduction, the human is an integral part of the system, which we can consider a {\em human computation} \cite{quinn2011human}.
 Giving the user more control, and adding visibility to their participation might have positive effects.
@@
 IPs that had `anonymous' as the user. In the plot of 
-anonymous users, some of the last in the rank could in fact be registered users,
+anonymous users, some of the lasts in the rank could in fact be registered users,
 momentarily participating as anonymous. Even if this is not considered, only
@@
  Figure \ref{fig:box}, which is a boxplot of the  participation by two
- groups, showing the big difference between registered users (shown as
- ``users'') and anonymous ones (shown as ``IPs''. This chart also
+ groups, shows the big difference between registered users (shown as
+ ``users'') and anonymous ones (shown as ``IPs''). This chart also
  shows the number of outliers in the groups, which is an obvious

### 2018-01-06 2018-IPMU/Gamification/paper.tex
Reduced to 13 pages

@@
 
-Systems based on the
-JavaScript/browser combination
-emphasize the ephemerality, ease of use, and universality, while
-systems such as HTCondor or BOINC \cite{boinc_grid04} might be more adequate for work that 
-requires higher availability of volunteer resources.
+% ystems based on the
+% JavaScript/browser combination
+% emphasize the ephemerality, ease of use, and universality, while
+% systems such as HTCondor or BOINC \cite{boinc_grid04} might be more adequate for work that 
+% requires higher availability of volunteer resources.
+
 In order to attract volunteers, systems rely on downloadable clients, and in
@@
 and time-consuming problems, concluding that BBVC is mostly apt for
-short, data-light tasks. This recent survey does not analyze the
-impact of gamification as we do in this paper, which might help in
-this area.
+short, data-light tasks.
 
@@
 evolutionary computation and it has been used by several
-authors ever since. More recent efforts include
-\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO,martinez2015capataz,pan2015gray,desell2017large,rivas2017time,leclerc2016seamless,merelo2017performance}. In
+authors ever since. In
 fact, \cite{pan2015gray} performs an analysis of what
@@
 negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
-such as women \cite{kumar2013gamification}. Another gamification technique
-is to make existing tasks feel more like games \cite{deterding2010just}.
-Some techniques used in this approach include adding meaningful choice,
-on-boarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.
+such as women \cite{kumar2013gamification}.
 
-\begin{figure*}[htb]
-    \centering
-        \includegraphics[width=5in]{img/login.png}
-    \caption{ User interface: Landing page, and Log-in and registration dialogs.
-    }
-    \label{fig:login}
-\end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%         \includegraphics[width=5in]{img/login.png}
+%     \caption{ User interface: Landing page, and Log-in and registration dialogs.
+%     }
+%     \label{fig:login}
+% \end{figure*}
 
@@
 In this experiment a gamification technique is applied to
-the Anonymous
-% {\sf NodIO}
+the {\sf NodIO}
 volunteer computing framework, using the
@@
 The last thing a user wants to see is yet another registration form.
-Then, the first design decision for this version is that registration is optional and simple as shown in Figure \ref{fig:login}.
+Then, the first design decision for this version is that registration
+is optional and simple.
+%as shown in Figure \ref{fig:login}.
 
@@
 in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of
-levels at the beginning; this function is represented in Figure \ref{fig:s}.
+levels at the beginning.%; this function is represented in Figure \ref{fig:s}.
 %
-\begin{figure*}[htb]
-    \centering
-        \includegraphics[width=3in]{img/s.png}
-    \caption{$S$ function used to assign the user's level.
-    }
-    \label{fig:s}
-\end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%         \includegraphics[width=3in]{img/s.png}
+%     \caption{$S$ function used to assign the user's level.
+%     }
+%     \label{fig:s}
+% \end{figure*}
 
@@
 While the modal is opened, the scores are refreshed every second. All other
-functionality is available to all users and is the same as the previous
-%{\sf NodIO}
-Anonymous version, showing the state of the current algorithm.
+functionality is available to all users and is the same as the
+previous {\sf NodIO} version, showing the state of the current algorithm.
 In the front page a link to the open source code of the experiment was also
-available \url{https://undisclos.ed}
-% \url{https://github.com/lucero21/login-master}
+available \url{https://github.com/lucero21/login-master}.
 
@@
 multi-modal problem called {\em l-trap} which has been used extensively
-as a benchmark for evolutionary algorithms and is represented in figure \ref{fig:trap},  \cite{fernandes2009using,nijssen2003analysis}.
+as a benchmark for evolutionary algorithms % and is represented in
+% figure \ref{fig:trap}
+,  \cite{nijssen2003analysis}.
 This function counts the number of bits in a sequence $l$ and assigns
@@
  %
-\begin{figure*}[htb]
-    \centering
-        \includegraphics[width=3in]{img/trap.png}
-    \caption{Trap function.
-    }
-    \label{fig:trap}
-\end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%         \includegraphics[width=3in]{img/trap.png}
+%     \caption{Trap function.
+%     }
+%     \label{fig:trap}
+% \end{figure*}
 
@@
 
-\begin{table}[h!tbp]
-  \small
-  \caption{ Log file record structure}
-  \label{tab:record}
-  \centering
-  \small
-  \begin{tabular}{l  l}
-    \hline\noalign{\smallskip}
-    Attribute & Value \\
-    \noalign{\smallskip}\hline\noalign{\smallskip}
-    chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
-    fitness & fitness as an integer.  \\ \hline
-    IP & IP address of the participant (later anonymized) .\\ \hline
-    user & user-name or the string `anonymous'.  \\ \hline
-    worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
-    level &  level of user or the string `info' \\ \hline
-    message & Only PUT messages were recorded in this experiment. \\ \hline
-    timestamp & A time-stamp based on the Unix epoch\\ \hline
-  \end{tabular}
-\end{table}
+% \begin{table}[h!tbp]
+%   \small
+%   \caption{ Log file record structure}
+%   \label{tab:record}
+%   \centering
+%   \small
+%   \begin{tabular}{l  l}
+%     \hline\noalign{\smallskip}
+%     Attribute & Value \\
+%     \noalign{\smallskip}\hline\noalign{\smallskip}
+%     chromosome   & The individual sent to the server as a string of 160 bits.  \\ \hline
+%     fitness & fitness as an integer.  \\ \hline
+%     IP & IP address of the participant (later anonymized) .\\ \hline
+%     user & user-name or the string `anonymous'.  \\ \hline
+%     worker\_uuid & unique identifier of the HTTP Worker that sent the participation.   \\ \hline
+%     level &  level of user or the string `info' \\ \hline
+%     message & Only PUT messages were recorded in this experiment. \\ \hline
+%     timestamp & A time-stamp based on the Unix epoch\\ \hline
+%   \end{tabular}
+% \end{table}
 %
-At the end of the experiment the resulting log file (whose fields are shown in Table \ref{tab:record}) contained 933,513 contributions both
+At the end of the experiment the resulting log file
+%(whose fields are shown in Table \ref{tab:record})
+contained 933,513 contributions both
 registered users and anonymous volunteers. Registered users were responsible for about

### 2018-01-09 2018-IPMU/Gamification/paper.tex
@jj please check the figure
Single Figure 1 as requested in issue #10.

@@
     \centering
-        \includegraphics[width=0.9\linewidth]{img/puts_user.png}
-        \includegraphics[width=0.9\linewidth]{img/puts_ip.png}
-    \caption{Usage by registered (top) and anonymous (bottom) users.  Users are ranked by
+        \includegraphics[width=0.9\linewidth]{img/puts.png}
+
+    \caption{Usage by registered (green triangles) and anonymous (red dots) users.  Users are ranked by
         the number of requests sent, in the \emph{x} axis is the rank, and in the \emph{y} axis
-        number or PUTs in a logarithmic scale.}
+        number of PUTs in a logarithmic scale.}
       \label{fig:puts}

### 2018-01-09 2018-IPMU/Gamification/log_code/user_puts.py

@@ -38,28 +38,13 @@ import matplotlib.pyplot as plt
 from matplotlib.ticker import ScalarFormatter, MaxNLocator
 
 x_users = range(len(users))
+x_ips = range(len(IPs))
+
 print len(users)
 # automatically update ylim of ax2 when ylim of ax1 changes.
 fig , ax =plt.subplots()
-ax.plot(x_users,users,'.')
-ax.set_yscale('log')
-ax.axis([1, 20, 1, 100000])
-for axis in [ax.xaxis, ax.yaxis]:
-    axis.set_major_formatter(ScalarFormatter())
-
-ax.xaxis.set_major_locator(MaxNLocator(integer=True))
-ax.set_ylabel('PUTs')
-ax.set_xlabel('users')
-ax.grid(True)
-
-plt.show()
-
-
-x_ips = range(len(IPs))
-print len(IPs)
-fig , ax =plt.subplots()
-print IPs
-ax.plot(x_ips,IPs,'.')
+ax.plot(x_users,users,'g^')
+ax.plot(x_ips,IPs,'r.')
 ax.set_yscale('log')
 ax.axis([1, 100, 1, 100000])
 for axis in [ax.xaxis, ax.yaxis]:
@@ -67,7 +52,25 @@ for axis in [ax.xaxis, ax.yaxis]:
 
 ax.xaxis.set_major_locator(MaxNLocator(integer=True))
 ax.set_ylabel('PUTs')
-ax.set_xlabel('IPs')
+ax.set_xlabel('Users')
 ax.grid(True)
+
 plt.show()
+
+
+#x_ips = range(len(IPs))
+#print len(IPs)
+#fig , ax =plt.subplots()
+#print IPs
+#ax.plot(x_ips,IPs,'.')
+#ax.set_yscale('log')
+#ax.axis([1, 100, 1, 100000])
+#for axis in [ax.xaxis, ax.yaxis]:
+#    axis.set_major_formatter(ScalarFormatter())
+
+#ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+#ax.set_ylabel('PUTs')
+#ax.set_xlabel('IPs')
+#ax.grid(True)
+#plt.show()
 

### 2018-01-09 2018-IPMU/Gamification/paper.tex
Reviewed @mariosky: it looks fine
Also reduced a bit more everything.

@@
 processing extraterrestrial signals \cite{david-seti:home}, which used
-a client that acted as an screensaver, or a
-high-throughput queuing system such as HTCondor \cite{HTCondor}.
+a client that acted as an screensaver%, or a
+%high-throughput queuing system such as HTCondor \cite{HTCondor}.
 However,
@@
 
-In order to attract volunteers, systems rely on downloadable clients, and in
-some cases like HTCondor, they also require administrative access to
-certain resources. In fact, Fabisiak \cite{fabisiak2017browser}
+In order to attract volunteers, many systems rely on downloadable clients;
+% , and in
+% some cases like HTCondor, they also require administrative access to
+% certain resources
+Fabisiak \cite{fabisiak2017browser}
 analyzes the different capabilities of Internet and browser-based
@@
 
-Recent works have been using crowdsourcing in order to train robots
-for human interaction, a representative work of this trend is that of
-Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose
-training simulated robots for the grounding of language symbols.
+% Recent works have been using crowdsourcing in order to train robots
+% for human interaction, a representative work of this trend is that of
+% Anetsberger \& Bongard \cite{anetsbergerrobots} in which they propose
+% training simulated robots for the grounding of language symbols.
 They do this by using human observers that issue arbitrary commands to these robots via the web, providing positive or negative reinforcement to the resulting actions of the robot.
 For interaction, they use the Twitch streaming platform at which users can issue commands through the
-platform\'s integrated chat service. Subjects were incentivized to
-interact with the system by GUI features that provided participants
-with a sense of involvement with the simulation. This kind of system
-could add additional dynamics, because users can collaborate or
-even compete with each other when issuing their commands.
+% platform\'s integrated chat service. Subjects were incentivized to
+% interact with the system by GUI features that provided participants
+% with a sense of involvement with the simulation. This kind of system
+% could add additional dynamics, because users can collaborate or
+% even compete with each other when issuing their commands.
 This approach has been called {\em social cloud} \cite{6404452}, emphasizing the fact that user participation is gamified in order to obtain the most from it \cite{7027564}.
 
-The proof-of-concept efforts described above, do not go any further than trying to find out how many users join the effort, and how many of them the system can support.
-In fact, systems such as the one described in
+Many systems described above do not go any further than trying to find out how many users join the effort, and how many of them the system can support. In fact, systems such as the one described in
 \cite{gecco07:workshop:dcor} had severe scaling issues; some of them
@@
 against single-user single-computer implementations of the same
-algorithm.
-Lately, researchers have tried to integrate volunteer computing techniques as a part of a larger distributed computing effort \cite{leclerc2016seamless}.
-In fact, systems using volunteers exclusively, exhibit a certain amount of unpredictability, and they might be better used in combination with ready-available computing power.
+algorithm. Lately, researchers have tried to integrate volunteer
+computing techniques as a part of a larger distributed computing
+effort \cite{leclerc2016seamless}.  In fact, systems using volunteers
+exclusively, exhibit a certain amount of unpredictability, and they
+might be better used in combination with ready-available computing
+power.
 
@@
 90\% of the contributions, while being only 16\% of the total number of unique users (assuming each IP is from a single user).
-The data recorded in JSON format for every contribution is  presented
-in Table~\ref{tab:record}.
+% The data recorded in JSON format for every contribution is  presented
+% in Table~\ref{tab:record}.
 

### 2018-01-09 2018-IPMU/Gamification/paper.tex
Down to 11 pages
Suppressed some paragraphs and citations. Closes #8
Also @mariosky regot figure 1 in a single chart, so closes #10

@@
 given non-game context application \cite{deterding2011game,hamari2014does}.
-Some works give a form of reward to users, these include
-points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework},
-the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
+% Some works give a form of reward to users, these include
+% points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework},
+% the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
 By Making the rewards for  tasks achievements visible to other players or
-providing leader boards are ways of encouraging players to compete \cite{hickman2010total}.
-In this work, a leader board was implemented in order to promote competition, presenting only the all-time top five participating users.
-Competition can also have problematic consequences, which can result in
-negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
-such as women \cite{kumar2013gamification}.
+providing leader boards are ways of encouraging players to compete
+\cite{hickman2010total}. In this work, a leader board was implemented
+in order to promote competition, presenting only the all-time top five
+participating users.
+% Competition can also have problematic consequences, which can result in
+% negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
+% such as women \cite{kumar2013gamification}.
 
@@
 This work has been supported in part by: Ministerio espa\~{n}ol de
- Econom\'{\i}a y Competitividad under project TIN2014-56494-C4-3-P
- (UGR-EPHEMECH).
+ Econom\'{\i}a y Competitividad under projects TIN2014-56494-C4-3-P
+ (UGR-EPHEMECH) and TIN2017-85727-C4-2-P (UGR-DeepBio).
 

### 2018-01-11 2018-IPMU/Gamification/paper.tex
Intro

@@
 gamification technique is applied by adding a leader-board showing the
-top scores achieved by registered contributors. In the  NodIO
-system volunteers can participate without the need to create
-an account, so one of the questions we wanted to address was if the
+top scores achieved by registered contributors. In NodIO,
+volunteers can participate without creating an account, so one of the questions we wanted to address was if the
 need to register would have a negative impact on user participation.
@@
 \section{Introduction}
-
 The world wide web is an increasingly reliable and high-performance
@@
 communication. A distributed system based on the browser can be
-created easily using these mechanisms, which have lately been enhanced
-by WebSockets for lower latency; you can gather users by just
+created easily using these mechanisms, which lately have been enhanced
+by WebSockets for lower latency; you can start to gather users just by
 announcing the URL. This approach for creating distributed experiments is called {\em
@@
 their performance, and might even be beneficial for a number of
-reasons, including increased diversity
+reasons, including an increased diversity
 \cite{cantu-paz:migration-policies}. All these properties make them
 ideal candidates for volunteer computing setups such as the one
-presented in this paper, at least if some minimal changes are made to
-the usual sequential or parallel algorithms that rely on a static
-setup.
+presented in this paper.
+%, at least if some minimal changes are made to
+%the usual sequential or parallel algorithms that rely on a static
+%setup.  -- Not needed, Mario
 
-But even if adaptation is straightforward, there are several open issues in the research regarding the
-application of volunteer computing techniques to the development of
-distributed evolutionary algorithms of which the most important is
+But even if adaptation is straightforward, the application of volunteer computing techniques to the development of distributed evolutionary algorithms 
+still has several open issues. An important research aspect is 
 approaching the problem as a socio-technical system
@@
 which is a hurdle that might take users away; however, the open design
-of the system we use means that we can actually mix anonymous-non
-gamified and authenticated-gamified users, with only the small overhead
+of the system we use means that we can actually mix anonymous (with no
+gamification) and authenticated (gamified) users, with only the small overhead
 of adding the gamification mechanism.
@@
 participate by just visiting a web page, and personalization, making
-registration compulsory and subsequently authenticate in order to
-participate, is what we will analyze in this paper. In a sociotechnical system such as this one, every
+registration a requirement in order to % I just simplified the sentence -Mario
+participate, is what we will analyze in this paper. In a socio-technical system such as this one, every
 decision the user has to take has an influence on the overall
@@
 or the amount of time they contribute to the experiment is not enough,
-it would be better to revert to the purely anonymous system.
+it would be better to revert NodIO to a purely anonymous system.
 
-The rest of the paper is organized as follows: Next we will briefly
-present the state of the art in opportunistic/volunteer distributed evolutionary
-computation (EC). Section \ref{sec:gamification} will describe the
+The rest of the paper is organized as follows: The state of the art in opportunistic/volunteer distributed evolutionary
+computation (EC) is presented next. Section \ref{sec:gamification} will describe the
 framework and the problem used in the experiments, which are publicly

### 2018-01-11 2018-IPMU/Gamification/paper.tex
Problem

@@
 by WebSockets for lower latency; you can start to gather users just by
-announcing the URL. This approach for creating distributed experiments is called {\em
-  volunteer}, {\em cycle-scavenging}, or {\em opportunistic} computing
+announcing the URL. This approach for creating distributed 
+experiments is called {\em volunteer}, {\em cycle-scavenging}, 
+or {\em opportunistic} computing
 \cite{sarmenta2001volunteer} and it dates back, in different shapes
@@
 
-But even if adaptation is straightforward, the application of volunteer computing techniques to the development of distributed evolutionary algorithms 
+But even if adaptation is straightforward, the application of volunteer
+computing techniques to the development of distributed evolutionary algorithms 
 still has several open issues. An important research aspect is 
@@
 processing extraterrestrial signals \cite{david-seti:home}, which used
-a client that acted as an screensaver%, or a
+a client that acted as an screensaver.%, or a
 %high-throughput queuing system such as HTCondor \cite{HTCondor}.
@@
 \cite{unwitting-ec,boldrin2007distributed,apolonia2012enhancing} or volunteer
-\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
-evolutionary computation and it has been used by several
+\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed EC and it has been used by several
 authors ever since. In
@@
 algorithm. Lately, researchers have tried to integrate volunteer
-computing techniques as a part of a larger distributed computing
+computing techniques as a part of a larger distributed evolutionary computing 
 effort \cite{leclerc2016seamless}.  In fact, systems using volunteers
@@
 After login, registered users are welcomed by their name, and they can see
-their current score and level. All users can see in a modal dialog the leader board.
+their current score and level. All users can see the 
+leader board in a modal dialog.
 While the modal is opened, the scores are refreshed every second. All other
@@
 multi-modal problem called {\em l-trap} which has been used extensively
-as a benchmark for evolutionary algorithms % and is represented in
-% figure \ref{fig:trap}
-,  \cite{nijssen2003analysis}.
+as a benchmark for evolutionary algorithms 
+% and is represented in figure \ref{fig:trap},  
+\cite{nijssen2003analysis}.
 This function counts the number of bits in a sequence $l$ and assigns
@@
 as the number of bits increases, and decreasing linearly until a change in slope
-is reached at point $z$, adding a deceptive component for evolutionary algorithms.
-In order to increase difficulty, trap functions can be concatenated, in our case we have used $40$ concatenated traps.
-The trap function is defined as:
+is reached at point $z$, adding a deceptive component for evolutionary
+algorithms. In order to increase difficulty, trap functions can be
+concatenated, in our case we have used $40$ concatenated traps. The trap
+function is defined as:
 \[ f(u)=
@@
 
-Each local GA had the following parameters, the initial population was randomly generated
-with a random size between 128 and 256 individuals, the period to send individuals to the server
-was set at 100 local generations, the parameters for the trap function are $l$ = 4,
+Each local GA had the following parameters, the initial population was randomly 
+generated with a size between 128 and 256 individuals, the period to send individuals to the server was set at 100 local generations, the parameters for the trap function are fixed at $l$ = 4,
 $a$=1, $b$ = 2, $z$=3 with a chromosome length of 160 bits.

### 2018-01-11 2018-IPMU/Gamification/paper.tex
results
Finished spell-check and revision
All that's left is @fergunet revision (and adding himself as an author)

@@
 This trade-off between ease of use, allowing users to
-participate by just visiting a web page, and personalization, making
+participate by just visiting a web page, and customization, making
 registration a requirement in order to % I just simplified the sentence -Mario
@@
 %(whose fields are shown in Table \ref{tab:record})
-contained 933,513 contributions both
-registered users and anonymous volunteers. Registered users were responsible for about
+contained 933,513 contributions considering both
+registered and anonymous volunteers. Registered users were responsible for about
 90\% of the contributions, while being only 16\% of the total number of unique users (assuming each IP is from a single user).

### 2018-01-11 2018-IPMU/Gamification/paper.tex
Added some sugestions and questions. Please fix the summary of the paper in introduction, as it links to incorrect sections (also Gamification section should be merged). Closes #11

@@
 
-\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1}}
+\author{Mario Garc\'ia-Valdez\inst{1} \and Juan J. Merelo Guerv\'os\inst{2} \and  Lucero Lara \inst{1} \and Pablo Garc\'ia-S\'anchez \inst{3}}
 
@@
 Universidad de Granada, Granada, Spain
+\and
+Universidad de C\'adiz, C\'adiz, Spain
 \email{mario@tectijuana.edu.mx}\\
@@
 
-\authorrunning{Garc\'ia-Valdez, Merelo, \& Lara }
+\authorrunning{Garc\'ia-Valdez, Merelo, Lara \& Garc\'ia-S\'anchez}
 
@@
 provided around 90\% of the work, thus effectively increasing the
-performance of the overall system.
+performance of the overall system. %Pablo: how the performance is compared with respect to the previous version? Less time?
 
@@
 communication. A distributed system based on the browser can be
-created easily using these mechanisms, which lately have been enhanced
+easily created using these mechanisms, which lately have been enhanced
 by WebSockets for lower latency; you can start to gather users just by
@@
 it has to include some way of user authentication,
-which is a hurdle that might take users away; however, the open design
+which is a hurdle that might take users away; however, the open design %Pablo: use colon instead semicolon?
 of the system we use means that we can actually mix anonymous (with no
@@
 \ref{sec:experiments}, to finally wrap up with the conclusions.
+%Pablo: check out this paragraph!
+Section 3 is not describing the problem and framework, it is only discussing on related work on gamification!
+I suggest combine section Gamification into Section 2, and separate Section 4.2 (Results) as a new Section before Conclusions.
 
@@
 In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed.
+%Pablo: I suggest to move this paragraph above the previous one, as it works as a better introduction to this kind of systems and can be linked to the next paragraph.
 
@@
 
-\section{Experiment}
+\section{Experiment} %Pablo: change this to Experimental Setup, as it also describes the system
 \label{sec:experiments}
@@
 the accomplishment of certain activities that need to be encouraged. In these case
-a point is awarded for each PUT sent to the server. Levels are a long
+a point is awarded for each HTTP {\tt PUT} Request sent to the server. Levels are a long
 term achievement, in this case the level depends on the score:
@@
 
-After login, registered users are welcomed by their name, and they can see
+After login, %Pablo: after login where? Add a link to the page or explain that the system was run in our servers
+registered users are welcomed by their name, and they can see
 their current score and level. All users can see the 
@@
 In the front page a link to the open source code of the experiment was also
-available \url{https://github.com/lucero21/login-master}.
+available \url{https://github.com/lucero21/login-master}. %Pablo: I suggest to move the link to the introduction, just after the first time it is mentioned as free software
 
@@
 
+%Pablo: this should be a new subsection, as we are not describing the problem from here
 We made a call to participation on November 23th,
@@
 Contributions are counted by the number of HTTP {\tt PUT} requests
-sent to the server. In Figure~\ref{fig:puts} a comparison between the 
+sent to the server. %Pablo: the PUT request concept was mentioned before the level scores description. This explanation should be moved there. 
+In Figure~\ref{fig:puts} a comparison between the 
 contribution of registered vs anonymous users is presented.

### 2018-01-11 2018-IPMU/Gamification/paper.tex
changes proposed my pablo

@@
 performance of the overall system. %Pablo: how the performance is compared with respect to the previous version? Less time?
+% In this case is more work, more indivudals put in to the
+% pool or more generations - Mario
 
@@
 The rest of the paper is organized as follows: The state of the art in opportunistic/volunteer distributed evolutionary
-computation (EC) is presented next. Section \ref{sec:gamification} will describe the
+computation (EC) and Gamification Techinques as presented next. Section \ref{sec:experiments} will describe the
 framework and the problem used in the experiments, which are publicly
 available under a free license. The results of each of the steps in the incremental design are presented in Section
-\ref{sec:experiments}, to finally wrap up with the conclusions.
+\label{sec:results}, to finally wrap up with the conclusions.
 %Pablo: check out this paragraph!
@@
 I suggest combine section Gamification into Section 2, and separate Section 4.2 (Results) as a new Section before Conclusions.
+% Done - Mario
 
@@
 short, data-light tasks.
+
+As indicated in the introduction, the human is an integral part of the system, which we can consider a {\em human computation} \cite{quinn2011human}.
+Giving the user more control, and adding visibility to their participation might have positive effects.
+In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed.
+%Pablo: I suggest to move this paragraph above the previous one, as it works as a better introduction to this kind of systems and can be linked to the next paragraph.
+% Moved - Mario
 
@@
 
-As indicated in the introduction, the human is an integral part of the system, which we can consider a {\em human computation} \cite{quinn2011human}.
-Giving the user more control, and adding visibility to their participation might have positive effects.
-In fact, the BOINC project applies this technique by having a web page presenting the ``Top 100 multi-project BOINC participants'' in which the name of the volunteer, the number of projects, GFLOPS, country and team are displayed.
-%Pablo: I suggest to move this paragraph above the previous one, as it works as a better introduction to this kind of systems and can be linked to the next paragraph.
-
 \section{Gamification}
@@
 
-\section{Experiment} %Pablo: change this to Experimental Setup, as it also describes the system
+\section{Experimental Setup and Execution} %Pablo: change this to Experimental Setup, as it also describes the system
+% OK - Mario
 \label{sec:experiments}
@@
 
-After login, %Pablo: after login where? Add a link to the page or explain that the system was run in our servers
+After login to the NodIO web application, %Pablo: after login where? Add a link to the page or explain that the system was run in our servers
 registered users are welcomed by their name, and they can see
@@
 %Pablo: this should be a new subsection, as we are not describing the problem from here
+% Ok -Mario
+\subsection{Experiment Execution}
 We made a call to participation on November 23th,
@@
 the {\sf NodIO}
-volunteer computing framework; in concordance with
+volunteer evolutionary computing framework; in concordance with
 the results obtained in other
@@
 
-The fact that gamification actually work highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation.
+The fact that applying a gamification techinque improved user participation highlights the social nature of volunteer computing systems, piling on the fact that successful {\em social clouds} must embed games and other social mechanisms to enhance participation.
 That is why future lines of work will look for intermediate ways of including

### 2018-03-08 2018-IPMU/Gamification/paper.tex
REcoloca correos
Edit to address some the comments from  reviewers.
Edit to address some the comments from  reviewers.  Added a new
proposal section.

@@
 
-\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico
+\institute{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico; 
+  \email{mario@tectijuana.edu.mx}
 \and
-Universidad de Granada, Granada, Spain
+Universidad de Granada, Granada, Spain; 
+\email{jmerelo@geneura.ugr.es}
 \and
 Universidad de C\'adiz, C\'adiz, Spain
-\email{mario@tectijuana.edu.mx}\\
-\email{jmerelo@geneura.ugr.es}}
+}
 
@@
 \section{Introduction}
-The world wide web is an increasingly reliable and high-performance
+The World Wide Web is an increasingly reliable and high-performance
 operating system for running distributed applications. Besides the
@@
 it has to include some way of user authentication,
-which is a hurdle that might take users away; however, the open design %Pablo: use colon instead semicolon?
+which is a hurdle that might take users away. However, the open design %Pablo: use colon instead semicolon?
 of the system we use means that we can actually mix anonymous (with no
@@
 The rest of the paper is organized as follows: The state of the art in opportunistic/volunteer distributed evolutionary
-computation (EC) and Gamification Techinques as presented next. Section \ref{sec:experiments} will describe the
+computation (EC) is presented next. The proposed Gamification Techinque is described in 
+Section~\ref{sec:gamification}. Section~\ref{sec:experiments} will describe the
 framework and the problem used in the experiments, which are publicly
-available under a free license. The results of each of the steps in the incremental design are presented in Section
-\label{sec:results}, to finally wrap up with the conclusions.
+available under a free license. The results of each of the steps in the incremental design are presented in Section~\ref{sec:results}, to finally wrap up with the conclusions. 
 %Pablo: check out this paragraph!
@@
 
-Many systems described above do not go any further than trying to find out how many users join the effort, and how many of them the system can support. In fact, systems such as the one described in
+Many of the systems described above do not go any further than trying to find out how many users join the effort, and how many of them the system can support.
+In fact, systems such as the one described in
 \cite{gecco07:workshop:dcor} had severe scaling issues; some of them
@@
 
-\section{Gamification}
+\section{Proposed Gamification Technique}
 \label{sec:gamification}
-A definition given by Huotari \cite{huotari2012defining} is ``Gamification is
+A definition of Gamification given by Huotari \cite{huotari2012defining} is ``Gamification is
 the process by which gaming concepts are brought to real world tasks associated with
@@
 providing leader boards are ways of encouraging players to compete
-\cite{hickman2010total}. In this work, a leader board was implemented
+\cite{hickman2010total}.
+
+In this work, a leader board was implemented
 in order to promote competition, presenting only the all-time top five
 participating users.
+
+Another gamification technique employed in this work is based on a rewarding mechanism
+\cite{dubois2013understanding}. In general rewards  consist of a reputation system
+with score points, levels and leader boards. Points are awarded to users in response of
+the accomplishment of certain activities that need to be encouraged. In these case
+a point is awarded for each HTTP {\tt PUT} Request sent to the server. Levels are a long
+term achievement, in this case the level depends on the score:
+\[ \text{level}(\text{score};a,b)=
+    \begin{cases}
+      0,                                    &  \text{score}\leq a\\
+      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
+      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
+      1,                                    & \text{score}\geq b
+   \end{cases}
+\]
+The function returns a normalized value that is multiplied by the maximum level
+in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of
+levels at the beginning. The effect $a$ and $b$ have on the results was not 
+determined, and other techinques could be used.
+
+ %; this function is represented in Figure \ref{fig:s}.
+
 % Competition can also have problematic consequences, which can result in
@@
 
-In this experiment a gamification technique is applied to
-the {\sf NodIO}
-volunteer computing framework, using the
-particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
+For this experiment the gamification technique is applied to
+the {\sf NodIO} volunteer computing framework, using the
+particular version described in detail in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
 The objective of the experiment is to test the kind of impact
@@
 
-The gamification technique employed in this work is based on a rewarding mechanism
-\cite{dubois2013understanding}. In general rewards  consist of a reputation system
-with score points, levels and leader boards. Points are awarded to users in response of
-the accomplishment of certain activities that need to be encouraged. In these case
-a point is awarded for each HTTP {\tt PUT} Request sent to the server. Levels are a long
-term achievement, in this case the level depends on the score:
-\[ \text{level}(\text{score};a,b)=
-    \begin{cases}
-      0,                                    &  \text{score}\leq a\\
-      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
-      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
-      1,                                    & \text{score}\geq b
-   \end{cases}
-\]
-The function returns a normalized value that is multiplied by the maximum level
-in this case 100. The variables $a$=100 and $b$=7360 are set to give users a rapid increase of
-levels at the beginning.%; this function is represented in Figure \ref{fig:s}.
 %
@@
 In the front page a link to the open source code of the experiment was also
-available \url{https://github.com/lucero21/login-master}. %Pablo: I suggest to move the link to the introduction, just after the first time it is mentioned as free software
+available in \url{https://github.com/lucero21/login-master}. %Pablo: I suggest to move the link to the introduction, just after the first time it is mentioned as free software
 
@@
 
-\subsection{Results}
+\section{Results}
 \label{sec:results}
@@
 
-The second observation is apparent in the $y$ axis of Figure
-\ref{fig:puts}. There is a notable difference in the amount of
+The second observation is apparent in the $y$ axis of Figure~\ref{fig:puts}.
+There is a notable difference in the amount of
 participation between the two groups, and also in the slope of the
@@
 %
- Figure \ref{fig:box}, which is a boxplot of the  participation by two
+ Figure~\ref{fig:box}, which is a boxplot of the  participation by two
  groups, shows the big difference between registered users (shown as
